### 2020-08-06 paper_introduction.Rmd
draft paper introduction methods

---
title: "How many researchers use 'boilerplate' statistical analysis sections?"
author: "Nicole White, Richi Nayak, Adrian Barnett"
date: "21/05/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
link-citations: yes
bibliography: references.bib
subtitle: An observational study of papers published in _PLOS ONE_.
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(pander)
library(ggplot2)
library(DiagrammeR) 
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('keep.line.breaks', TRUE)
panderOptions('table.alignment.default','left')
```

An ideal statistical analysis will use appropriate methods to create insights from the data and inform the research questions. Unfortunately many current statistical analyses are far from ideal, with many researchers using the wrong methods, misinterpreting the results, or failing to adequately check their assumptions [@Goodman2008;@Leek2017]. Some researchers take a "mechanistic" approach to statistics, copying the few methods they know regardless of their appropriateness, and then going through the motions of the analysis [@Stark2018].

Many researchers may not have received adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
However, using the wrong statistical methods can cause real harm [@Altman1994] and bad statistical practices are being to used abet weak science [@Stark2018].
Statistical mistakes are a key source of waste in research and partly explain the current reproducibility crisis in science [@Allison2016]. Even when the correct methods are used, many researchers fail to describe them adequately, making it difficult to reproduce the results [@Ernst2017;@Zhou2018].

The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019]. Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked.
A recent survey of editors found that only 23% of health and medical journals used expert statistical review for all articles [@Hardwicke2020], which was little different from a survey from 22 years ago [@Goodman1998].

As statisticians we have heard researchers admit that they sometimes copy-and-paste their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to estimate the extent that researchers are using 'boilerplate' statistical methods sections.
Use of these methods sections indicates that little thought has gone into the statistical analysis.

# Methods

## Data source

We chose _PLOS ONE_ because it is a large multi-disciplinary open access journal, and because the journal laudibly make the full text of papers easily available for analysis via an API. This enabled us to isolate the statistical analysis section of the paper, and also collect descriptive data on the paper (e.g., field of research).

<!--- https://journals.plos.org/plosone/s/editorial-and-peer-review-process --->
<!--- https://journals.plos.org/plosone/s/criteria-for-publication --->
The criteria for publication include, "Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail".
_PLOS ONE_ does not use a separate statistical peer reviewer. Articles are handled by an academic editor who selects peer reviewers. 
_PLOS ONE_ expect authors to use reporting guidelines from the EQUATOR network, which were designed in part because of the long history of poor statistical application and reporting in health and medical journals [@Altman2016].

We used the 'rplos' package to read the data from the API into R [@rplos].

Questions:

* What questions are asked of PLOS ONE Reviewers?

- Is the manuscript technically sound, and do the data support the conclusions?
- Has the statistical analysis been performed appropriately and rigorously?
- Does the manuscript adhere to the PLOS Data Policy?
- Is the manuscript presented in an intelligible fashioin and written in standard English?

## Search strategy
Figure \ref{fig:search-strategy} outlines the main steps of our search strategy. Searches were completed using the 'rplos' package [@rplos]. All full-text research articles published in _PLOS ONE_ since 2006 were eligible for inclusion (last accessed: 5 June 2020).

Records were first identified using combinations of analysis-related search terms that could appear anywhere in the full-text. Metadata for each record consisted of the digital objected identifier (DOI), article title, journal volume, subject area(s) and total article views since online publication. For each DOI returned, Full text XML was extracted and organsied by listed section and subsection headings. In order to consistently identify descriptions of statistical analyses undertaken, subsection headings were filtered using common titles (e.g. 'Data analysis'; 'Statisical methods'). Records that returned more than one heading at the filtering stage were combined (?). Filtering resulted were cleaned using the 'textclean' package [@textclean] (see Supplementary File 1 for full details)

```{r echo=F,fig.cap='\\label{fig:search-strategy}Search strategy',fig.align='center',fig.height=5}
knitr::include_graphics("figures/search_strategy.png")

```

# Results
```{r source, message=F,echo=F}
load('data/search_results_n.RData')
```

* total records found: 178654; 131847 unique DOIs
* total stats sections after filtering: 111731 (85%)
* Plot over time. Currently double counting if more than one search query for same doi.

```{r echo=F,fig.cap="\\label{fig:search-results-n}Figure caption here"}
plot_dat_n = n_by_searchterm_volume %>% group_by(volume) %>% 
  summarise(n_matches = sum(n_unique)) %>% right_join(.,n_by_volume %>% select(-n_records),by='volume') %>% 
  gather(variable,value,-volume)

plot_dat_n %>% ggplot(.,aes(x=volume,y=value,group=variable))+
  geom_bar(stat='identity',position='dodge',aes(fill=variable))

```

# Discussion

The first line in many statistical analysis sections was the software used, implying that the software is the most important detail. As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994]. This is not the case, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a relatively minor detail compared with detailing what methods were used and why.

Despite the extensive array of tests available, many authors are reporting the same few methods.

## Limitations

We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods section would be fine.

We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section.

We only examined one journal and hence our results may not be generalisable to all journals, especially those that use a statistical reviewer for all papers.

# References

### 2020-08-11 paper_introduction.Rmd
update search strategy text

@@
 library(ggplot2)
-library(DiagrammeR) 
 panderOptions('table.emphasize.rownames', FALSE)
@@
 
+* Reference updates to stats reporting guidelines; e.g see:
+
+- https://everyone.plos.org/2019/09/26/new-plos-one-statistical-reporting-guidelines/.
+- https://web.archive.org/web/20190607174803/https://journals.plos.org/plosone/s/submission-guidelines
+- https://web.archive.org/web/20150507175314/https://journals.plos.org/plosone/s/submission-guidelines
+
 ## Search strategy
-Figure \ref{fig:search-strategy} outlines the main steps of our search strategy.
-Searches were completed using the 'rplos' package [@rplos].
-All full-text research articles published in _PLOS ONE_ since 2006 were eligible for inclusion (last accessed: 5 June 2020).
+Our search strategy comprised targeted searches of the PLOS API and filtering of full-text articles based on section headings (Figure \ref{fig:search-strategy}).
+All full-text Research Articles published in _PLOS ONE_ since 2006 were eligible for inclusion.
 
-Records were first identified using combinations of analysis-related search terms that could appear anywhere in the full-text.
-Metadata for each record consisted of the digital objected identifier (DOI), article title, journal volume, subject area(s) and total article views since online publication.
-For each DOI returned, Full text XML was extracted and organsied by listed section and subsection headings.
-In order to consistently identify descriptions of statistical analyses undertaken, subsection headings were filtered using common titles (e.g.
-'Data analysis'; 'Statisical methods').
-Records that returned more than one heading at the filtering stage were combined (?).
-Filtering resulted were cleaned using the 'textclean' package [@textclean] (see Supplementary File 1 for full details)
+Initial searches of the PLOS API were completed using the ‘rplos’ package [@rplos].
+Submitted search queries focussed on the presence of analysis-related terms anywhere within a full-text article.
+Individual terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
+Results terms were intended to be broad and, by allowing terms to appear anywhere within the article, we accounted for the placement of relevant text in different sections, for example, in the Material and Methods section versus Supplementary Information.
+Search results were indexed by Digital Object Identifier (DOI), which enabled the removal of duplicate records.
+Metadata collected for each DOI included journal volume, PLOS subject classification(s) and number of article views since publication.
+
+Full-text XML data were then extracted for all search results for further processing.
+Data were extracted by DOI and organised by individual section and subsection headings.
+In order to consistently identify statistical methods sections, partial matching on commonly used sections headings was done to filter results (see Figure \ref{fig:search-strategy}).
+The final dataset was cleaned prior to analysis using the ‘textclean’ package [@textclean].
+Cleaning steps included conversion of all text to lower case, replacement of common symbol (e.g % to ‘percent’; < to ‘less than’) and removal of punctuation and excess whitespace.
+Commonly used statistical methods were standardised across all records, which included changing terms from plural to singular (e.g.
+tukey’s to tukey) as well as hyphenated terms (e.g.
+“kruskal-wallis” to “kruskal wallis”).
+The full list of standardised terms is provided as a Supplementary File.
+
+## Clustering algorithm
 
@@
 # Results
+
 ```{r source, message=F,echo=F}

### 2020-08-12 5_process_stats_section.R
update text cleaning code

@@ -3,9 +3,9 @@ library(stringr)
 library(textclean)
 library(tm)
 library(spelling)
-library(xlsx)
+library(readxl)
 
-load('./data/stats_section_info.rda')
+load('stats_section_info.rda')
 
 stats_section = bind_rows(stats_section)
 
@@ -51,51 +51,55 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
 #remove () including text within brackets 
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*\\([^\\)]+\\)",""))
 
-#remove punctutation
-#stats_section = stats_section %>% mutate(text_data_clean = strip(text_data_clean,char.keep = c("~~", "."),
-#                                                                 apostrophe.remove=T,digit.remove = F))
+#remove punctuation
 stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = c("~~","."),apostrophe.remove=T,digit.remove=F)
-
 
 #replace other non-ascii characters
 stats_section = stats_section %>% mutate(text_data_clean = replace_non_ascii(text_data_clean))
 
 #make common statistical terms and methods consistent
-common_stat_tests = read.xlsx('./data/methods_dictionary.xlsx',sheetName = 'common_stat_tests')
-other_terms = read.xlsx('./data/methods_dictionary.xlsx',sheetName = 'other')
+common_stat_tests = read_xlsx('methods_dictionary.xlsx',sheet = 'common_stat_tests')
+other_terms = read_xlsx('methods_dictionary.xlsx',sheet = 'other')
 
 #for each common method, create combined and unique plural terms
-common_stat_tests = common_stat_tests %>% mutate(combined_term = str_remove_all(term,'-'))
-unique_terms = unique(unlist(str_split(common_stat_tests$term,'-')))
+common_stat_tests = common_stat_tests %>% mutate(combined_term = str_remove_all(term,' '))
+unique_terms = unique(unlist(str_split(common_stat_tests$term,' ')))
 plural_terms  = paste0(unique_terms,'s')
 
 #str_c
-combined_terms_all = str_c(common_stat_tests$combined_term,collapse="|")
-plural_terms_all = str_c(plural_terms,collapse="|")
-other_terms_all = str_c(other_terms$term,collapse='|')
[…]

### 2020-08-13 6_process_meta.R
add script 6_process_meta.R

@@ -1,0 +1,38 @@
+#6_process_meta
+#explore meta data for final set of plos records
+library(tidyverse)
+#load meta data for all records found
+load('plos_meta_data.rda')
+
+#load final stats_section
+load('stats_section_info.rda')
+
+#bind_rows
+stats_section = bind_rows(stats_section)
+
+#pull doi from stats_section; mutate doi
+doi_list = stats_section %>% distinct(doi) %>% mutate(doi = str_replace_all(doi,'.*10.1371.','10.1371/'))
+
+
+#filter meta data
+meta_stats_section = meta_dat_allrecords %>% filter(doi %in% doi_list[['doi']])
+
+#split subject_level_1; gather/filter
+stats_section_subject = separate(meta_stats_section %>% select(doi,subject_level_1), subject_level_1, into = paste0("subject", 1:10), sep = ',')
+stats_section_subject = gather(stats_section_subject,subject_level_number,subject_level_name,-doi) %>% filter(!is.na(subject_level_name))
+
+#str_remove
+stats_section_subject = stats_section_subject %>% mutate(subject_level_number = str_remove_all(subject_level_number,'subject'))
+
+#count total subjects per doi
+stats_section_subject %>% count(subject_level_number)
+
+#count subject_level_name
+stats_section_subject %>% count(subject_level_name)
+
+#create indicator matrix for subjects by doi
+doi_subject_matrix = stats_section_subject %>% group_by(doi) %>% count(subject_level_name) %>% 
+  ungroup() %>% spread(.,subject_level_name,n,fill=0) 
+
+#number of dois per subject area
+doi_subject_matrix %>% summarise_at(vars("Biology and life sciences":"Social sciences"),sum)

### 2020-08-13 6_process_meta.R
add subject dendrogram

@@ -36,3 +36,9 @@ doi_subject_matrix = stats_section_subject %>% group_by(doi) %>% count(subject_l
 
 #number of dois per subject area
 doi_subject_matrix %>% summarise_at(vars("Biology and life sciences":"Social sciences"),sum)
+
+#distinct patterns
+subject_patterns = distinct(doi_subject_matrix %>% select(-doi))
+
+hc = hclust(dist(subject_patterns))
+plot(hc)

### 2020-08-18 5_process_stats_section.R
update text cleaning strings

@@ -9,91 +9,92 @@ load('stats_section_info.rda')
 
 stats_section = bind_rows(stats_section)
 
-#remove all unicode characters
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data,"\\s*<U\\+\\w+>\\s*|<U+\\w+>$"," "))
+#1. remove formatting/special characters
 
-#remove equations
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*(\n)(.*)(\n)\\s*"," "))
+#unicode characters (e.g. hair space <U+200A>)
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data,pattern="\\s*<U\\+\\w+>\\s*|\\<\\S\\S\\S\\S\\S\\S\\>",replacement=" "))
 
-#replace >,<,= with text
-#\\s* - zero or more whitespaces
+#centered equations/other
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(\n)(.*)(\n)\\s*",replacement = " "))
 
-#less than
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*(<)\\s*"," less than "))
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[<]"," less than "))
+#numbered references eg [23]
+#stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*\\[\\d+\\]\\s*",""))
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\[\\S{1,3}\\]",replacement = ""))
 
-#greater than
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*(>)\\s*"," greater than "))
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[>]"," greater than "))
+#remove () including text within brackets  "\\s*\\([^\\)]+\\)"
+#if want to keep text inside brackets, use "[()]"
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[()]",""))
 
-#equal to
[…]

### 2020-08-20 5_extract_unicodes.R
extract unique unicodes before text cleaning

@@ -1,0 +1,25 @@
+library(tidyverse)
+library(tidytext)
+library(stringr)
+library(textclean)
+library(tm)
+library(spelling)
+library(readxl)
+
+load('stats_section_info.rda')
+
+stats_section = bind_rows(stats_section)
+
+#change to native encoding
+stats_section = stats_section %>% mutate(text_data_clean = enc2native(text_data))
+
+#1. remove formatting/special characters
+all_words = stats_section %>% unnest(text_data_clean) %>% 
+  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
+
+#find all unicode characters
+unicode_all = grep('\\s*<U\\+\\w+>\\s*|"\\s*\\<\\S\\S\\S\\S\\S\\S\\>\\s*"',all_words,value=T)
+unicode_unique = str_extract_all(string=unicode_all,pattern=regex("<U\\+\\w+>")) %>% unlist() %>% unique()
+
+#write
+save(unicode_unique,file='unicode_characters.rda')

### 2020-08-20 5_extract_unicodes.R
extract unicode characters for conversion

@@ -22,4 +22,4 @@ unicode_all = grep('\\s*<U\\+\\w+>\\s*|"\\s*\\<\\S\\S\\S\\S\\S\\S\\>\\s*"',all_w
 unicode_unique = str_extract_all(string=unicode_all,pattern=regex("<U\\+\\w+>")) %>% unlist() %>% unique()
 
 #write
-save(unicode_unique,file='unicode_characters.rda')
+write.csv(unicode_unique,file='unicode_characters.csv',row.names=F)

### 2020-08-20 5_extract_unicodes.R
udpate script convert to symbols

@@ -1,10 +1,6 @@
 library(tidyverse)
-library(tidytext)
 library(stringr)
-library(textclean)
-library(tm)
-library(spelling)
-library(readxl)
+library(stringi)
 
 load('stats_section_info.rda')
 
@@ -18,8 +14,23 @@ all_words = stats_section %>% unnest(text_data_clean) %>%
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
 
 #find all unicode characters
-unicode_all = grep('\\s*<U\\+\\w+>\\s*|"\\s*\\<\\S\\S\\S\\S\\S\\S\\>\\s*"',all_words,value=T)
-unicode_unique = str_extract_all(string=unicode_all,pattern=regex("<U\\+\\w+>")) %>% unlist() %>% unique()
+unicode_unique = str_extract_all(string=all_words,pattern=regex("<U\\+\\w+>")) %>% 
+  unlist() %>%
+  as_tibble() %>% count(value)
+
+#mutate, rename, arrange
+unicode_lookup = unicode_unique %>% mutate(value = gsub('000','',value)) %>%
+  rename('unicode'=value) %>%
+  arrange(-n)
+
+#convert to uft8
+unicode_lookup = unicode_lookup %>% mutate(utf8 = gsub("<U\\+(\\w+)>", "\\\\u\\1", unicode))
+
+#add symbols
+unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
+
+#select
+unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
 
 #write
-write.csv(unicode_unique,file='unicode_characters.csv',row.names=F)
+save(unicode_lookup,file='unicode_characters.rda')

### 2020-08-20 code/1_rplos_search.R
organise files into folders

@@ -1,0 +1,49 @@
+library(rplos)
+library(rentrez)
+library(stringr)
+library(XML)
+library(tidyverse)
+
+
+#search all full-text research articles in PLoS One and return meta data only
+results_all  =  searchplos("*:*",
+             fl=c('id','volume','subject','subject_level_1'),
+             fq=list('journal_key:PLoSONE','article_type:"Research Article"','doc_type:full'),
+             limit=0)
+
+#number of records by volume (current volume = 15). meta data only
+q = paste0("volume:",1:15)
+results_by_volume = lapply(q, function(x) searchplos(q=paste(x),
+                                                     fq=list('journal_key:PLoSONE','article_type:"Research Article"','doc_type:full'),
+                                                     limit=0)) %>%
+  setNames(q)
+
+save(results_by_volume,file='data/results_by_volume_info.RData')
+
+#by analysis-related search term
+#search terms
+#'method' picks up 'methods'
+#'modelling' and 'modelling' return same number of records. assume UK spelling
+prefix = c('data','statistical')
+suffix = c('analysis','analyses','method','methodology','modelling')
+search_terms = apply(expand.grid(prefix,suffix),1,paste,collapse=' ')
+q = paste0(search_terms,'~1')
+
+results_by_searchterm = lapply(q, function(x) searchplos(q=paste(x),
+                                                     fl=c('id','title','volume','counter_total_all','subject','subject_level_1'),
+                                                     fq=list('journal_key:PLoSONE','article_type:"Research Article"','doc_type:full'),
+                                                     limit=90000)) %>%
+  setNames(q)
+
+save(results_by_searchterm,file='data/results_by_searchterm_fulltext_info.RData')
+
+#meta-data
+meta_dat = lapply(results_by_searchterm,function(x) x$data %>% select(id,volume,counter_total_all,subject,subject_level_1))
+meta_dat = do.call(rbind.data.frame,meta_dat)
+meta_dat = distinct(meta_dat)                            
+
+meta_dat = meta_dat %>% rename('doi'=id,
+                               'citations'=counter_total_all)
+
+save(meta_dat,file='data/plos_searchresults_metadata.RData')
+

### 2020-08-20 code/2_split_meta_dat.R

@@ -1,0 +1,23 @@
+#split meta_dat
+#splits meta_dat into equal sized datasets for QUT HPC batches
+library(tidyverse)
+#plos full text
+meta_dat = readRDS('./data/plos_searchresults_metadata.RDS')
+
+num_groups = 5
+meta_dat = meta_dat %>% mutate(bin = (row_number()-1) %/% (n()/num_groups)) 
+
+meta_dat_list = split(meta_dat,meta_dat$bin)
+
+#store as batches and save
+meta_dat_1 = meta_dat_list[[1]]
+meta_dat_2 = meta_dat_list[[2]]
+meta_dat_3 = meta_dat_list[[3]]
+meta_dat_4 = meta_dat_list[[4]]
+meta_dat_5 = meta_dat_list[[5]]
+
+saveRDS(meta_dat_1,file='data/plos_meta_1.RDS')
+saveRDS(meta_dat_2,file='data/plos_meta_2.RDS')
+saveRDS(meta_dat_3,file='data/plos_meta_3.RDS')
+saveRDS(meta_dat_4,file='data/plos_meta_4.RDS')
+saveRDS(meta_dat_5,file='data/plos_meta_5.RDS')

### 2020-08-20 code/3_extract_fulltext_xml.R

@@ -1,0 +1,40 @@
+library(rplos)
+library(XML)
+library(tidyverse)
+
+#plos full text - e,g, batch 1
+meta_dat = readRDS('./data/plos_meta_1.RDS')
+
+doi_list = meta_dat %>% pull(doi)
+
+for(i in 1:length(doi_list)){
+
+  full_text <- plos_fulltext(doi_list[i])
+  full_text_xml <- xmlParse(full_text,asText=T,useInternalNodes = T,encoding='UTF-8')
+  methods_text = list()
+  node_info = getNodeSet(full_text_xml,"//sec")
+  
+  methods_text[['text_heading']] = sapply(node_info,xpathSApply,"./title",xmlValue)
+  methods_text[['text_heading']] =  unlist(as.character(methods_text[['text_heading']]))
+
+  text_out = sapply(node_info,xpathSApply,"./p",xmlValue)
+  methods_text[['text_data']] = sapply(text_out,function(s) tolower(paste(s,collapse='')))
+  
+  if (!is.null(methods_text[['text_heading']]) & !is.null(methods_text[['text_data']])){
+    dat = do.call(cbind.data.frame,methods_text)
+    rownames(dat) <- NULL
+    dat = as_tibble(dat) %>% mutate_if(is.factor,as.character)
+    
+    #remove empty text fields
+    dat = dat %>% filter(text_data!="")
+    
+    #save as text file. save non-empty tibbles only
+    if (nrow(dat)>0){
+      fileOut = gsub('/','.',doi_list[i])
+      save(dat,file=paste0('./data/full text records/batch_1/doi_',fileOut,'.rda'))
+    }
+  }
+}
+
+
+

### 2020-08-20 code/5_process_stats_section.R

@@ -1,0 +1,133 @@
+library(tidyverse)
+library(stringr)
+library(textclean)
+library(tm)
+library(spelling)
+library(readxl)
+
+load('stats_section_info.rda')
+
+stats_section = bind_rows(stats_section)
+
+#1. remove formatting/special characters
+
+#unicode characters (e.g. hair space <U+200A>)
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data,pattern="\\s*<U\\+\\w+>\\s*|\\<\\S\\S\\S\\S\\S\\S\\>",replacement=" "))
+
+#centered equations/other
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(\n)(.*)(\n)\\s*",replacement = " "))
+
+#numbered references eg [23]
+#stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*\\[\\d+\\]\\s*",""))
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\[\\S{1,3}\\]",replacement = ""))
+
+#remove () including text within brackets  "\\s*\\([^\\)]+\\)"
+#if want to keep text inside brackets, use "[()]"
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[()]",""))
+
+#2. replace/standardise common symbols
+#standardise dashes 
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(–+)\\s*",replacement = "-"))
+
+#p value (doesn't resolve p less-than, p equal to)
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp value\\b","p-value"))
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp less than\\b|\\bp less-than\\b","p <"))
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp equal to\\b|\\bp equal-to\\b","p ="))
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp greater than\\b|\\bp greater-than\\b","p >"))
+
+#standardise spacing for >,<,= 
[…]

### 2020-08-20 code/6_process_meta.R

@@ -1,0 +1,44 @@
+#6_process_meta
+#explore meta data for final set of plos records
+library(tidyverse)
+#load meta data for all records found
+load('plos_meta_data.rda')
+
+#load final stats_section
+load('stats_section_info.rda')
+
+#bind_rows
+stats_section = bind_rows(stats_section)
+
+#pull doi from stats_section; mutate doi
+doi_list = stats_section %>% distinct(doi) %>% mutate(doi = str_replace_all(doi,'.*10.1371.','10.1371/'))
+
+
+#filter meta data
+meta_stats_section = meta_dat_allrecords %>% filter(doi %in% doi_list[['doi']])
+
+#split subject_level_1; gather/filter
+stats_section_subject = separate(meta_stats_section %>% select(doi,subject_level_1), subject_level_1, into = paste0("subject", 1:10), sep = ',')
+stats_section_subject = gather(stats_section_subject,subject_level_number,subject_level_name,-doi) %>% filter(!is.na(subject_level_name))
+
+#str_remove
+stats_section_subject = stats_section_subject %>% mutate(subject_level_number = str_remove_all(subject_level_number,'subject'))
+
+#count total subjects per doi
+stats_section_subject %>% count(subject_level_number)
+
+#count subject_level_name
+stats_section_subject %>% count(subject_level_name)
+
+#create indicator matrix for subjects by doi
+doi_subject_matrix = stats_section_subject %>% group_by(doi) %>% count(subject_level_name) %>% 
+  ungroup() %>% spread(.,subject_level_name,n,fill=0) 
+
+#number of dois per subject area
+doi_subject_matrix %>% summarise_at(vars("Biology and life sciences":"Social sciences"),sum)
+
+#distinct patterns
+subject_patterns = distinct(doi_subject_matrix %>% select(-doi))
+
+hc = hclust(dist(subject_patterns))
+plot(hc)

### 2020-08-21 3_extract_fulltext_xml.R
'WIP'

@@ -1,40 +1,0 @@
-library(rplos)
-library(XML)
-library(tidyverse)
-
-#plos full text - e,g, batch 1
-meta_dat = readRDS('./data/plos_meta_1.RDS')
-
-doi_list = meta_dat %>% pull(doi)
-
-for(i in 1:length(doi_list)){
-
-  full_text <- plos_fulltext(doi_list[i])
-  full_text_xml <- xmlParse(full_text,asText=T,useInternalNodes = T,encoding='UTF-8')
-  methods_text = list()
-  node_info = getNodeSet(full_text_xml,"//sec")
-  
-  methods_text[['text_heading']] = sapply(node_info,xpathSApply,"./title",xmlValue)
-  methods_text[['text_heading']] =  unlist(as.character(methods_text[['text_heading']]))
-
-  text_out = sapply(node_info,xpathSApply,"./p",xmlValue)
-  methods_text[['text_data']] = sapply(text_out,function(s) tolower(paste(s,collapse='')))
-  
-  if (!is.null(methods_text[['text_heading']]) & !is.null(methods_text[['text_data']])){
-    dat = do.call(cbind.data.frame,methods_text)
-    rownames(dat) <- NULL
-    dat = as_tibble(dat) %>% mutate_if(is.factor,as.character)
-    
-    #remove empty text fields
-    dat = dat %>% filter(text_data!="")
-    
-    #save as text file. save non-empty tibbles only
-    if (nrow(dat)>0){
-      fileOut = gsub('/','.',doi_list[i])
-      save(dat,file=paste0('./data/full text records/batch_1/doi_',fileOut,'.rda'))
-    }
-  }
-}
-
-
-

### 2020-08-21 4_combine_batch_info.R

@@ -1,25 +1,0 @@
-#combine_batch_info
-library(tidyverse)
-
-#filter: identify dois with statitical analyses subsection based on text_heading (partial string matching)
-analysis_keywords = c('statistical analysis','statistical analyses',
-                      'statistical method','statistics',
-                      'data analysis', 'data analyses')
-#loop over batches
-batch = 1:5
-stats_section = list()
-for (b in batch){
-  folder_name = paste0('data/full text records/batch_',b)
-  results_by_doi = list.files(folder_name)
-  doi_list  = gsub('.rda','',results_by_doi)
-  text_results = lapply(results_by_doi,
-                        function(x) {load(paste(folder_name,x,sep='/'))
-                          return(dat %>% mutate_if(is.factor,as.character))}
-  )
-  names(text_results) <- doi_list
-  text_results = bind_rows(text_results,.id='doi')
-  stats_section[[b]] = text_results %>% filter(grepl(paste(analysis_keywords,collapse='|'),text_heading,
-                                       ignore.case=T,fixed=F))
-}
-
-save(stats_section,file='data/stats_section_info.rda')

### 2020-08-21 5_process_stats_section.R

@@ -1,133 +1,0 @@
-library(tidyverse)
-library(stringr)
-library(textclean)
-library(tm)
-library(spelling)
-library(readxl)
-
-load('stats_section_info.rda')
-
-stats_section = bind_rows(stats_section)
-
-#1. remove formatting/special characters
-
-#unicode characters (e.g. hair space <U+200A>)
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data,pattern="\\s*<U\\+\\w+>\\s*|\\<\\S\\S\\S\\S\\S\\S\\>",replacement=" "))
-
-#centered equations/other
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(\n)(.*)(\n)\\s*",replacement = " "))
-
-#numbered references eg [23]
-#stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*\\[\\d+\\]\\s*",""))
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\[\\S{1,3}\\]",replacement = ""))
-
-#remove () including text within brackets  "\\s*\\([^\\)]+\\)"
-#if want to keep text inside brackets, use "[()]"
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[()]",""))
-
-#2. replace/standardise common symbols
-#standardise dashes 
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(–+)\\s*",replacement = "-"))
-
-#p value (doesn't resolve p less-than, p equal to)
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp value\\b","p-value"))
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp less than\\b|\\bp less-than\\b","p <"))
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp equal to\\b|\\bp equal-to\\b","p ="))
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp greater than\\b|\\bp greater-than\\b","p >"))
-
-#standardise spacing for >,<,= 
[…]

### 2020-08-21 6_process_meta.R

@@ -1,44 +1,0 @@
-#6_process_meta
-#explore meta data for final set of plos records
-library(tidyverse)
-#load meta data for all records found
-load('plos_meta_data.rda')
-
-#load final stats_section
-load('stats_section_info.rda')
-
-#bind_rows
-stats_section = bind_rows(stats_section)
-
-#pull doi from stats_section; mutate doi
-doi_list = stats_section %>% distinct(doi) %>% mutate(doi = str_replace_all(doi,'.*10.1371.','10.1371/'))
-
-
-#filter meta data
-meta_stats_section = meta_dat_allrecords %>% filter(doi %in% doi_list[['doi']])
-
-#split subject_level_1; gather/filter
-stats_section_subject = separate(meta_stats_section %>% select(doi,subject_level_1), subject_level_1, into = paste0("subject", 1:10), sep = ',')
-stats_section_subject = gather(stats_section_subject,subject_level_number,subject_level_name,-doi) %>% filter(!is.na(subject_level_name))
-
-#str_remove
-stats_section_subject = stats_section_subject %>% mutate(subject_level_number = str_remove_all(subject_level_number,'subject'))
-
-#count total subjects per doi
-stats_section_subject %>% count(subject_level_number)
-
-#count subject_level_name
-stats_section_subject %>% count(subject_level_name)
-
-#create indicator matrix for subjects by doi
-doi_subject_matrix = stats_section_subject %>% group_by(doi) %>% count(subject_level_name) %>% 
-  ungroup() %>% spread(.,subject_level_name,n,fill=0) 
-
-#number of dois per subject area
-doi_subject_matrix %>% summarise_at(vars("Biology and life sciences":"Social sciences"),sum)
-
-#distinct patterns
-subject_patterns = distinct(doi_subject_matrix %>% select(-doi))
-
-hc = hclust(dist(subject_patterns))
-plot(hc)

### 2020-08-21 code/5_extract_unicodes.R
organise files into folders

@@ -1,0 +1,36 @@
+library(tidyverse)
+library(stringr)
+library(stringi)
+
+load('stats_section_info.rda')
+
+stats_section = bind_rows(stats_section)
+
+#change to native encoding
+stats_section = stats_section %>% mutate(text_data_clean = enc2native(text_data))
+
+#1. remove formatting/special characters
+all_words = stats_section %>% unnest(text_data_clean) %>% 
+  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
+
+#find all unicode characters
+unicode_unique = str_extract_all(string=all_words,pattern=regex("<U\\+\\w+>")) %>% 
+  unlist() %>%
+  as_tibble() %>% count(value)
+
+#mutate, rename, arrange
+unicode_lookup = unicode_unique %>% mutate(value = gsub('000','',value)) %>%
+  rename('unicode'=value) %>%
+  arrange(-n)
+
+#convert to uft8
+unicode_lookup = unicode_lookup %>% mutate(utf8 = gsub("<U\\+(\\w+)>", "\\\\u\\1", unicode))
+
+#add symbols
+unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
+
+#select
+unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
+
+#write
+save(unicode_lookup,file='unicode_characters.rda')

### 2020-08-21 5_extract_unicodes.R
delete duplicate files

@@ -1,36 +1,0 @@
-library(tidyverse)
-library(stringr)
-library(stringi)
-
-load('stats_section_info.rda')
-
-stats_section = bind_rows(stats_section)
-
-#change to native encoding
-stats_section = stats_section %>% mutate(text_data_clean = enc2native(text_data))
-
-#1. remove formatting/special characters
-all_words = stats_section %>% unnest(text_data_clean) %>% 
-  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
-
-#find all unicode characters
-unicode_unique = str_extract_all(string=all_words,pattern=regex("<U\\+\\w+>")) %>% 
-  unlist() %>%
-  as_tibble() %>% count(value)
-
-#mutate, rename, arrange
-unicode_lookup = unicode_unique %>% mutate(value = gsub('000','',value)) %>%
-  rename('unicode'=value) %>%
-  arrange(-n)
-
-#convert to uft8
-unicode_lookup = unicode_lookup %>% mutate(utf8 = gsub("<U\\+(\\w+)>", "\\\\u\\1", unicode))
-
-#add symbols
-unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
-
-#select
-unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
-
-#write
-save(unicode_lookup,file='unicode_characters.rda')

### 2020-08-21 code/5_extract_unicodes.R
update lookup table

@@ -2,7 +2,7 @@ library(tidyverse)
 library(stringr)
 library(stringi)
 
-load('stats_section_info.rda')
+load('./data/stats_section_info.rda')
 
 stats_section = bind_rows(stats_section)
 
@@ -14,12 +14,12 @@ all_words = stats_section %>% unnest(text_data_clean) %>%
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
 
 #find all unicode characters
-unicode_unique = str_extract_all(string=all_words,pattern=regex("<U\\+\\w+>")) %>% 
+unicode_lookup = str_extract_all(string=all_words,pattern=regex("<U\\+\\w+>")) %>% 
   unlist() %>%
   as_tibble() %>% count(value)
 
 #mutate, rename, arrange
-unicode_lookup = unicode_unique %>% mutate(value = gsub('000','',value)) %>%
+unicode_lookup = unicode_lookup %>% mutate(value = gsub('000','',value)) %>%
   rename('unicode'=value) %>%
   arrange(-n)
 
@@ -33,4 +33,4 @@ unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
 unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
 
 #write
-save(unicode_lookup,file='unicode_characters.rda')
+save(unicode_lookup,file='./data/unicode_characters.rda')

### 2020-08-21 code/5_process_stats_section.R
update filepaths

@@ -119,7 +119,7 @@ out = right_join(meta_dat,stats_section,by='doi') %>%
   select(-text_data)
 
 
-#not run
+#not run - save on git repo?
 #split into 5 batches to reduce file size
 # ngrps = 5
 # out = out %>% mutate(batch= (row_number()-1) %/% (n()/ngrps)) 

### 2020-08-21 code/6_process_meta.R
update filepaths

@@ -2,7 +2,7 @@
 #explore meta data for final set of plos records
 library(tidyverse)
 #load meta data for all records found
-load('plos_meta_data.rda')
+load('./data/plos_meta_data.rda')
 
 #load final stats_section
 load('stats_section_info.rda')

### 2020-08-21 code/5_extract_unicodes.R
add comments to code

@@ -9,7 +9,7 @@ stats_section = bind_rows(stats_section)
 #change to native encoding
 stats_section = stats_section %>% mutate(text_data_clean = enc2native(text_data))
 
-#1. remove formatting/special characters
+#unnest to individual words
 all_words = stats_section %>% unnest(text_data_clean) %>% 
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
 

### 2020-08-21 code/5_find_stopwords.R
script to find most common words

@@ -1,0 +1,17 @@
+#5_find_stopwords.R
+library(tidyverse)
+library(stringr)
+library(stringi)
+
+load('./data/stats_section_info.rda')
+
+stats_section = bind_rows(stats_section)
+
+#change to native encoding
+stats_section = stats_section %>% mutate(text_data_clean = enc2native(text_data))
+#remove unicodes
+stats_section = stats_section %>% mutate(text_data_clean = str_remove_all(text_data_clean,
+                                                                          pattern="\\s*<U\\+\\w+>\\s*|\\<\\S\\S\\S\\S\\S\\S\\>"))
+#unnest to individual words
+all_words = stats_section %>% unnest(text_data_clean) %>% 
+  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()

### 2020-08-21 code/5_find_stopwords.R
add word frequencies

@@ -15,3 +15,6 @@ stats_section = stats_section %>% mutate(text_data_clean = str_remove_all(text_d
 #unnest to individual words
 all_words = stats_section %>% unnest(text_data_clean) %>% 
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
+
+#word frequencies
+word_freq = as_tibble(all_words) %>% count(value) %>% arrange(-n)

### 2020-08-24 code/5_extract_unicodes.R
add unicode text labels

@@ -1,6 +1,7 @@
 library(tidyverse)
-library(stringr)
 library(stringi)
+library(Unicode)
+library(readr)
 
 load('./data/stats_section_info.rda')
 
@@ -23,14 +24,20 @@ unicode_lookup = unicode_lookup %>% mutate(value = gsub('000','',value)) %>%
   rename('unicode'=value) %>%
   arrange(-n)
 
+#u_char_name for text label
+unicode_lookup = unicode_lookup %>% mutate(label = u_char_name(gsub("<(.*)>",'\\1',unicode)))
+
+
 #convert to uft8
 unicode_lookup = unicode_lookup %>% mutate(utf8 = gsub("<U\\+(\\w+)>", "\\\\u\\1", unicode))
 
 #add symbols
 unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
 
+
 #select
 unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
 
 #write
 save(unicode_lookup,file='./data/unicode_characters.rda')
+write_excel_csv(unicode_lookup,path='./data/unicode_characters.csv')

### 2020-08-24 code/5_extract_unicodes.R
update code to seperate greek letters and symbols

@@ -26,7 +26,10 @@ unicode_lookup = unicode_lookup %>% mutate(value = gsub('000','',value)) %>%
 
 #u_char_name for text label
 unicode_lookup = unicode_lookup %>% mutate(label = u_char_name(gsub("<(.*)>",'\\1',unicode)))
-
+#mutate label
+unicode_lookup = unicode_lookup %>% mutate(label_clean = gsub(' ','-',label) %>% tolower())
+#update 'small-letters' e.g greek-small-letter-chi to chi TODO confirm 
+unicode_lookup = unicode_lookup %>% mutate(label_clean = gsub('.*small-letter-|greek-|slanted-','',label_clean))
 
 #convert to uft8
 unicode_lookup = unicode_lookup %>% mutate(utf8 = gsub("<U\\+(\\w+)>", "\\\\u\\1", unicode))
@@ -36,7 +39,7 @@ unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
 
 
 #select
-unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
+unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,label,label_clean,n)
 
 #write
 save(unicode_lookup,file='./data/unicode_characters.rda')

### 2020-08-24 code/5_extract_unicodes.R
update unicode text cleaning

@@ -26,7 +26,10 @@ unicode_lookup = unicode_lookup %>% mutate(value = gsub('000','',value)) %>%
 
 #u_char_name for text label
 unicode_lookup = unicode_lookup %>% mutate(label = u_char_name(gsub("<(.*)>",'\\1',unicode)))
-
+#mutate label
+unicode_lookup = unicode_lookup %>% mutate(label_clean = gsub(' ','-',label) %>% tolower())
+#update 'small-letters' e.g greek-small-letter-chi to chi
+unicode_lookup = unicode_lookup %>% mutate(label_clean = gsub('.*small-letter-|greek-|-symbol|slanted-','',label_clean))
 
 #convert to uft8
 unicode_lookup = unicode_lookup %>% mutate(utf8 = gsub("<U\\+(\\w+)>", "\\\\u\\1", unicode))
@@ -36,7 +39,7 @@ unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(utf8))
 
 
 #select
-unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,n)
+unicode_lookup = unicode_lookup %>% select(unicode,utf8,symbol,label,label_clean,n)
 
 #write
 save(unicode_lookup,file='./data/unicode_characters.rda')

### 2020-08-24 code/5_process_stats_section.R
update unicode conversion

@@ -6,16 +6,16 @@ library(spelling)
 library(readxl)
 
 load('./data/stats_section_info.rda')
+load('./data/unicode_characters.rda')
+
 
 stats_section = bind_rows(stats_section)
 
 #1. remove formatting/special characters
 
-#unicode characters (e.g. hair space <U+200A>)
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data,pattern="\\s*<U\\+\\w+>\\s*|\\<\\S\\S\\S\\S\\S\\S\\>",replacement=" "))
-
-#centered equations/other
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(\n)(.*)(\n)\\s*",replacement = " "))
+#change to native encoding
+stats_section = stats_section %>% mutate(text_data = enc2native(text_data))
+stats_section = stats_section %>% mutate(text_data_clean = text_data)
 
 #numbered references eg [23]
 #stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*\\[\\d+\\]\\s*",""))
@@ -25,20 +25,50 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
 #if want to keep text inside brackets, use "[()]"
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[()]",""))
 
+#centered equations/other
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(\n)(.*)(\n)\\s*",replacement = " "))
+
+
 #2. replace/standardise common symbols
+#unicode
+#format unicode characters (e.g. hair space <U+200A>)
+#remove general punctuation unicodes U+20xxx (no dashes identified)
+unicode_spaces = unicode_lookup %>% filter(grepl("U\\+20(\\w+)",unicode)) %>% pull(unicode) 
+unicode_spaces = gsub("U\\+(\\w+)", "\\U\\\\+\\1",unicode_spaces) %>% str_c(.,collapse='|')
+
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern=unicode_spaces,replacement=" "))
+
+#change unicode with label_clean
+unicode_set = gsub("U\\+(\\w+)", "\\U\\\\+\\1",unicode_lookup[['unicode']])
+unicode_set = str_c(unicode_set,collapse='|')
+
+#add white space around unicode labels
[…]

### 2020-08-24 code/5_process_stats_section.R
remove remaining non-asci

@@ -78,8 +78,8 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
 stats_section = stats_section %>% mutate(text_data_clean = replace_symbol(text_data_clean))
 
 #3. remove any remaining non-ascii characters, curly quotes
-#stats_section = stats_section %>% mutate(text_data_clean = replace_non_ascii(text_data_clean)) #NOT RUN: will remove common greek symbols,<,>,etcs
-#stats_section = stats_section %>% mutate(text_data_clean = replace_curly_quote(text_data_clean))
+stats_section = stats_section %>% mutate(text_data_clean = replace_non_ascii(text_data_clean)) 
+stats_section = stats_section %>% mutate(text_data_clean = replace_curly_quote(text_data_clean))
 
 #remove punctuation except for '.','-'
 stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = c("~~",".","-"),apostrophe.remove=T,digit.remove=F)

### 2020-08-25 code/5_find_stopwords.R
amended to filter out tm stop words

@@ -1,8 +1,7 @@
 #5_find_stopwords.R
 library(tidyverse)
-library(stringr)
 library(stringi)
-
+library(tm)
 load('./data/stats_section_info.rda')
 
 stats_section = bind_rows(stats_section)
@@ -18,3 +17,7 @@ all_words = stats_section %>% unnest(text_data_clean) %>%
 
 #word frequencies
 word_freq = as_tibble(all_words) %>% count(value) %>% arrange(-n)
+
+#remove stop words
+word_freq = word_freq %>% filter(!value %in% stopwords('en'))
+View(word_freq)

### 2020-08-25 code/5_process_stats_section.R
update str_replace_all for common stats terms

@@ -1,5 +1,4 @@
 library(tidyverse)
-library(stringr)
 library(textclean)
 library(tm)
 library(spelling)
@@ -55,9 +54,6 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
 #standardise dashes 
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\s*(–+)\\s*",replacement = "-"))
 
-#p value (doesn't resolve p less-than, p equal to)
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bp value\\b","p-value"))
-
 #standardise text and spacing for >,<,= 
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*[<]\\s*"," less-than "))
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*[>]\\s*"," greater-than "))
@@ -93,12 +89,16 @@ common_stat_terms = common_stat_terms %>% mutate(combined_term = str_remove_all(
 plural_terms  = unique(paste0(c(common_stat_terms$term,common_stat_terms$combined_term,common_stat_terms$update),'s'))
 
 #str_c
-stats_terms_all = str_c("\\b",c(common_stat_terms$term,common_stat_terms$combined_term),"\\b",collapse="|")
+stats_terms_all = str_c("\\b",common_stat_terms[['term']],"\\b",collapse="|")
+stats_terms_combined = str_c("\\b",common_stat_terms[['combined_term']],"\\b",collapse="|")
 plural_terms_all = str_c("\\b",plural_terms,"\\b",collapse="|")
-other_terms_all = str_c("\\b",other_terms$term,"\\b",collapse='|')
+other_terms_all = str_c("\\b",other_terms[['term']],"\\b",collapse='|')
 
 change_stats_terms = function(input){
-  common_stat_terms %>% filter(combined_term==input|term==input) %>% pull(update)
+  common_stat_terms %>% filter(term==input) %>% pull(update)
+}
+change_stats_combined = function(input){
+    common_stat_terms %>% filter(combined_term==input) %>% pull(update)
 }
 
 change_other = function(input){
@@ -115,24 +115,26 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
                                                                            plural_terms_all,
[…]

### 2020-08-25 manuscript/paper_introduction.Rmd
organise files into folders
update paper draft, methods section

--- 2020-08-11 paper_introduction.Rmd
@@
 ## Search strategy
-Our search strategy comprised targeted searches of the PLOS API and filtering of full-text articles based on section headings (Figure \ref{fig:search-strategy}).
+Our search strategy comprised targeted searches of the PLOS API and filtering of full-text articles based on section headings.
 All full-text Research Articles published in _PLOS ONE_ since 2006 were eligible for inclusion.
@@
 Results terms were intended to be broad and, by allowing terms to appear anywhere within the article, we accounted for the placement of relevant text in different sections, for example, in the Material and Methods section versus Supplementary Information.
-Search results were indexed by Digital Object Identifier (DOI), which enabled the removal of duplicate records.
+Search results were indexed by Digital Object Identifier (DOI), which allowed for removal of duplicate records.
 Metadata collected for each DOI included journal volume, PLOS subject classification(s) and number of article views since publication.
@@
 Data were extracted by DOI and organised by individual section and subsection headings.
-In order to consistently identify statistical methods sections, partial matching on commonly used sections headings was done to filter results (see Figure \ref{fig:search-strategy}).
-The final dataset was cleaned prior to analysis using the ‘textclean’ package [@textclean].
-Cleaning steps included conversion of all text to lower case, replacement of common symbol (e.g % to ‘percent’; < to ‘less than’) and removal of punctuation and excess whitespace.
-Commonly used statistical methods were standardised across all records, which included changing terms from plural to singular (e.g.
-tukey’s to tukey) as well as hyphenated terms (e.g.
-“kruskal-wallis” to “kruskal wallis”).
-The full list of standardised terms is provided as a Supplementary File.
+In order to consistently identify statistical methods sections, partial matching on commonly used sections headings was done to filter results.
+[TODO add more details of headings/partial matching here.
+strings were: 'statistical analysis','statistical analyses','statistical method','statistics','data analysis', 'data analyses'].
+
+Text cleaning focussed on standardising notation and references to statistical terminology, whilst minimising major changes to article style and formatting.
+Details of text cleaning steps, including R syntax, are provided in Supplementary File 1.
+
+Mathematical notation including Greek letters was converted from Unicode characters to plain text (Supplementary File 2).
+For example, instances of the code ‘<U+03B8>’, representing the Greek letter $\theta$, were replaced with ‘theta’.
+Common symbols not captured by Unicode blocks were also converted into plain text, using functions available in the 'textclean' package (e.g.
+% to ‘percent’, < to ‘less-than’) [@textclean].
+General punctuation, in-text references (e.g.
+[23]) centred equations and remaining non-ascii characters were removed.
+Text contained inside brackets was retained in the dataset to maximise content for analysis, with brackets removed.
+
+We compiled an extensive list of common methods and related terminology as a way of standardising descriptions of statistical analyses undertaken.
+An initial list was compiled by calculating individual word frequencies in the data and identifying relevant terms that appeared at least 100 times.
+Additional terms were added based on index searches of statistical textbooks [TODO need to mention this here?
+If so, add references] and spell checks (e.g.
+‘fischer’ instead of ‘fisher’).
+Using the final list, we performed multiple searches to identify possible variants of each term and replaced these with a standardised spelling.
+For example, instances ‘chi square’ and ‘chi squared’ were replaced with ‘chi-square’.
+Other variants of terminology considered plurals (e.g.
+chi squares), combinations of characters and numbers (e.g.
+chi 2) and combined text (e.g.
+chisquare).
+
+As a final step, common stop words including pronouns, contractions and selected prepositions were removed from the data.
+Stop words were identified from lists available from the ‘tm’ R package (Supplementary File 2).
+We chose to keep some prepositions included in the available list to minimise potential changes to the context of statistical method(s) being described (e.g.
+‘between’, ‘against’).
+
+<!-- ```{r echo=F,fig.cap='\\label{fig:search-strategy}Search strategy',fig.align='center',fig.height=5} -->
+<!-- knitr::include_graphics("search_strategy.png") -->
+
+<!-- ``` -->
 
@@
 
-```{r echo=F,fig.cap='\\label{fig:search-strategy}Search strategy',fig.align='center',fig.height=5}
-knitr::include_graphics("figures/search_strategy.png")
+<!-- # Results -->
 
-```
+<!-- ```{r source, message=F,echo=F} -->
+<!-- load('data/search_results_n.RData') -->
+<!-- ``` -->
 
-# Results
+<!-- * total records found: 178654; 131847 unique DOIs -->
+<!-- * total stats sections after filtering: 111731 (85%) -->
+<!-- * Plot over time. Currently double counting if more than one search query for same doi. -->
 
-```{r source, message=F,echo=F}
-load('data/search_results_n.RData')
-```
+<!-- ```{r echo=F,fig.cap="\\label{fig:search-results-n}Figure caption here"} -->
+<!-- plot_dat_n = n_by_searchterm_volume %>% group_by(volume) %>%  -->
+<!--   summarise(n_matches = sum(n_unique)) %>% right_join(.,n_by_volume %>% select(-n_records),by='volume') %>%  -->
+<!--   gather(variable,value,-volume) -->
 
-* total records found: 178654; 131847 unique DOIs
-* total stats sections after filtering: 111731 (85%)
-* Plot over time. Currently double counting if more than one search query for same doi.
+<!-- plot_dat_n %>% ggplot(.,aes(x=volume,y=value,group=variable))+ -->
+<!--   geom_bar(stat='identity',position='dodge',aes(fill=variable)) -->
 
-```{r echo=F,fig.cap="\\label{fig:search-results-n}Figure caption here"}
-plot_dat_n = n_by_searchterm_volume %>% group_by(volume) %>% 
-  summarise(n_matches = sum(n_unique)) %>% right_join(.,n_by_volume %>% select(-n_records),by='volume') %>% 
-  gather(variable,value,-volume)
-
-plot_dat_n %>% ggplot(.,aes(x=volume,y=value,group=variable))+
-  geom_bar(stat='identity',position='dodge',aes(fill=variable))
-
-```
+<!-- ``` -->
 

### 2020-08-25 code/5_process_stats_section.R
update code to save output

@@ -125,12 +125,11 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
                                                                            stats_terms_combined,
                                                                            change_stats_combined))
 
-#choose random 1% sample for check progress of text cleaning
-test_data= sample_frac(stats_section, 0.01)
-View(test_data)
 
 #remove excess whitespace
 stats_section = stats_section %>% mutate(text_data_clean = replace_white(text_data_clean))
+
+#final spell check TODO
 
 #remove stop words? TODO
 #remove_stopwords = str_c("\\b",stopwords('en'),"\\b",collapse='|')
@@ -140,16 +139,32 @@ stats_section = stats_section %>% mutate(text_data_clean = replace_white(text_da
 stats_section = stats_section %>% mutate(doi = str_replace_all(doi,'doi_',''))
 stats_section = stats_section %>% mutate(doi = str_replace_all(doi,'.journal','/journal'))
 
-save(stats_section,file='data/stats_section_cleaned.rda')
+#choose 100 dois to check cleaning
+sample_dois = stats_section %>% distinct(doi) %>% sample_n(.,100)
+
+sample_data = stats_section %>% filter(doi %in% sample_dois[['doi']])
+
+#split stats_section into batches to reduce file size
+
+
+#save(stats_section,file='data/stats_section_cleaned.rda')
+
+ngrps = 5
+out = stats_section %>% mutate(batch= (row_number()-1) %/% (n()/ngrps)) 
+out_list = split(out,out$batch)
+
+lapply(1:ngrps,function(b){
+  output = out_list[[b]] %>% select(-batch)
+  save(output,file=paste0('data/stats_section_cleaned_',b,'.rda'))
+})
 #write.table(stats_section %>% select(-text_data),file='data/stats_section_cleaned.txt',sep='\t',row.names = F)
 
 #merge with meta-data
 #meta_dat = readRDS('./data/plos_searchresults_metadata.rda')
 
-out = right_join(meta_dat,stats_section,by='doi') %>%
-  rename('counter_total_all' = citations) %>%
-  select(-text_data)
-
+# out = right_join(meta_dat,stats_section,by='doi') %>%
+#   rename('counter_total_all' = citations) %>%
+#   select(-text_data)
 
 #not run - save on git repo?
 #split into 5 batches to reduce file size

### 2020-08-26 code/5_process_stats_section.R
update functions for common stats methods

@@ -7,6 +7,10 @@ library(readxl)
 load('./data/stats_section_info.rda')
 load('./data/unicode_characters.rda')
 
+stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
+stat_terms_singular = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'single_terms')
+other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
+
 
 stats_section = bind_rows(stats_section)
 
@@ -65,10 +69,8 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bequal to\\b","equal-to"))
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\bgreater than\\b","greater-than"))
 
-
 #plus or minus 
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*(±)\\s*|\\s*(\\+/-+)\\s*"," plus-or-minus "))
-
 
 #common symbols covered by text clean ($,%,#,@,&,w/)
 stats_section = stats_section %>% mutate(text_data_clean = replace_symbol(text_data_clean))
@@ -81,49 +83,103 @@ stats_section = stats_section %>% mutate(text_data_clean = replace_curly_quote(t
 stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = c("~~",".","-"),apostrophe.remove=T,digit.remove=F)
 
 #4. make common statistical terms and methods consistent
-common_stat_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'stats_terms')
-other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
 
-#for each common method, create combined and unique plural terms
-common_stat_terms = common_stat_terms %>% mutate(combined_term = str_remove_all(term,' '))
-plural_terms  = unique(paste0(c(common_stat_terms$term,common_stat_terms$combined_term,common_stat_terms$update),'s'))
+#unnest to individual words
+all_words = stats_section %>% unnest(text_data_clean) %>% 
+  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist() 
+#word frequencies
[…]

### 2020-08-27 code/5_process_stats_section.R
update code for methods text find and replace

@@ -8,9 +8,8 @@ load('./data/stats_section_info.rda')
 load('./data/unicode_characters.rda')
 
 stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
-stat_terms_singular = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'single_terms')
+stat_terms_model = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'models')
 other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
-
 
 stats_section = bind_rows(stats_section)
 
@@ -41,7 +40,7 @@ unicode_spaces = gsub("U\\+(\\w+)", "\\U\\\\+\\1",unicode_spaces) %>% str_c(.,co
 
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern=unicode_spaces,replacement=" "))
 
-#change unicode with label_clean
+#repalce unicode with plain text description (label_clean)
 unicode_set = gsub("U\\+(\\w+)", "\\U\\\\+\\1",unicode_lookup[['unicode']])
 unicode_set = str_c(unicode_set,collapse='|')
 
@@ -84,42 +83,22 @@ stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = 
 
 #4. make common statistical terms and methods consistent
 
-#unnest to individual words
-all_words = stats_section %>% unnest(text_data_clean) %>% 
-  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist() 
-#word frequencies
-word_freq = tibble::enframe(all_words) %>% count(value) %>% arrange(-n)
-
 #for each hyphenated term, create combined and unique plural terms
+#plural includes entries from stat_terms_modell eg regressions to regression
 stat_terms_hyphen = stat_terms_hyphen %>% mutate(combined_term = str_remove_all(term,' '))
-plural_terms  = unique(paste0(c(stat_terms_hyphen[['term']],stat_terms_hyphen[['combined_term']],stat_terms_hyphen[['update']]),'s'))
-
-#filter word_freq to find variants of common methods
-combined_found = word_freq %>% filter(value %in% stat_terms_hyphen[['combined_term']])
-hyphen_found = word_freq %>% filter(value %in% stat_terms_hyphen[['update']])
[…]

### 2020-08-27 code/5_process_stats_section.R
add save for text files

@@ -157,38 +157,30 @@ stats_section = stats_section %>% mutate(doi = str_replace_all(doi,'.journal','/
 
 #choose 100 dois to check cleaning
 sample_dois = stats_section %>% distinct(doi) %>% sample_n(.,100)
-
 sample_data = stats_section %>% filter(doi %in% sample_dois[['doi']])
 
-#split stats_section into batches to reduce file size
-#save(stats_section,file='data/stats_section_cleaned.rda')
+write_rds(sample_data,path='data/stats_section_sample100.rds',"xz", compression = 9L)
+write_rds(stats_section,path='data/stats_section_cleaned.rds',"xz", compression = 9L)
 
-ngrps = 5
-out = stats_section %>% mutate(batch= (row_number()-1) %/% (n()/ngrps)) 
-out_list = split(out,out$batch)
+# save as text files for richi (5 batches)
+stats_section_txt = stats_section %>% select(-text_data)
+#split stats_section into batches to reduce file size
+ ngrps = 5
+ out = stats_section_txt %>% mutate(batch= (row_number()-1) %/% (n()/ngrps)) 
+ out_list = split(out,out$batch)
 
 lapply(1:ngrps,function(b){
   output = out_list[[b]] %>% select(-batch)
-  save(output,file=paste0('data/stats_section_cleaned_',b,'.rda'))
+  write.table(output,file=paste0('data/stats_section_cleaned_',b,'.txt'),sep='\t',row.names = F)
 })
-#write.table(stats_section %>% select(-text_data),file='data/stats_section_cleaned.txt',sep='\t',row.names = F)
 
-#merge with meta-data
-#meta_dat = readRDS('./data/plos_searchresults_metadata.rda')
+
+
+
 
 # out = right_join(meta_dat,stats_section,by='doi') %>%
 #   rename('counter_total_all' = citations) %>%
 #   select(-text_data)
-
-#not run - save on git repo?
-#split into 5 batches to reduce file size
-# ngrps = 5
-# out = out %>% mutate(batch= (row_number()-1) %/% (n()/ngrps)) 
-# out_list = split(out,out$batch)
-# 
-# lapply(1:5,function(b) write.table(out_list[[b]] %>% 
-#                                      select(-batch),file=paste0('data/stats_section_cleaned_',b,'.txt'),
-#                                    sep='\t',row.names=F))
 # 
 # 
 # 

### 2020-08-28 code/5_process_stats_section.R
add spell check code

@@ -9,7 +9,6 @@ load('./data/unicode_characters.rda')
 
 stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
 stat_terms_model = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'models')
-other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
 
 stats_section = bind_rows(stats_section)
 
@@ -24,7 +23,7 @@ stats_section = stats_section %>% mutate(text_data_clean = text_data)
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\[\\S{1,3}\\]",replacement = ""))
 
 #remove () including text within brackets  "\\s*\\([^\\)]+\\)"
-#if want to keep text inside brackets, use "[()]"
+#option to keep text inside brackets is "[()]"
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[()]",""))
 
 #centered equations/other
@@ -82,18 +81,16 @@ stats_section = stats_section %>% mutate(text_data_clean = replace_curly_quote(t
 stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = c("~~",".","-"),apostrophe.remove=T,digit.remove=F)
 
 #4. make common statistical terms and methods consistent
-
 #for each hyphenated term, create combined and unique plural terms
-#plural includes entries from stat_terms_modell eg regressions to regression
+#plurals includes entries from stat_terms_model eg regressions to regression
 stat_terms_hyphen = stat_terms_hyphen %>% mutate(combined_term = str_remove_all(term,' '))
 plural_terms  = unique(c(paste0(c(stat_terms_hyphen[['term']],stat_terms_hyphen[['combined_term']],
                                 stat_terms_hyphen[['update']]),'s'),stat_terms_model[['term']]))
 
-#str_c
+#str_c: define search strings
 plural_terms_all = str_c("\\b",plural_terms,"\\b",collapse="|") #to turn plural to singular
 stats_terms_all = str_c("\\b",stat_terms_hyphen[['term']],"\\b",collapse="|") #to join method words by hyphen
 stats_combined_all = str_c("\\b",stat_terms_hyphen[['combined_term']],"\\b",collapse="|") #to split method words by hyphen
-other_terms_all = str_c("\\b",other_terms[['term']],"\\b",collapse='|') #incorrect spellings identified
 
[…]

### 2020-08-28 code/5_process_stats_section.R
add spell check

@@ -9,7 +9,7 @@ load('./data/unicode_characters.rda')
 
 stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
 stat_terms_model = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'models')
-
+other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
 stats_section = bind_rows(stats_section)
 
 #1. remove formatting/special characters
@@ -114,6 +114,16 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
                                                                            stats_combined_all,
                                                                            change_stats_combined))
 
+#update common US to GB spelling
+other_terms_all = str_c(other_terms[['term']],collapse='|') #incorrect spellings identified
+change_other = function(input){
+  other_terms %>% filter(term==input) %>% pull(update)
+}
+
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,
+                                                                            other_terms_all,
+                                                                            change_other))
+
 #check for remaining instances
 all_words = stats_section %>% unnest(text_data_clean) %>% 
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist() 
@@ -159,5 +169,4 @@ lapply(1:ngrps,function(b){
    group_by(word) %>% summarise(n=sum(as.numeric(found))) %>%
    arrange(-n)
  save(spelling_errors,file='./data/common_spelling_errors_gb.rda')
- write.table(spelling_errors,file='./data/common_spelling_errors_gb.txt',sep='\t',row.names=F)
 

### 2020-09-01 code/5_process_stats_section.R
expand list of plural terms

@@ -8,6 +8,7 @@ load('./data/stats_section_info.rda')
 load('./data/unicode_characters.rda')
 
 stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
+stat_terms_single = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'single_terms')
 stat_terms_model = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'models')
 other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
 stats_section = bind_rows(stats_section)
@@ -84,13 +85,17 @@ stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = 
 #for each hyphenated term, create combined and unique plural terms
 #plurals includes entries from stat_terms_model eg regressions to regression
 stat_terms_hyphen = stat_terms_hyphen %>% mutate(combined_term = str_remove_all(term,' '))
-plural_terms  = unique(c(paste0(c(stat_terms_hyphen[['term']],stat_terms_hyphen[['combined_term']],
-                                stat_terms_hyphen[['update']]),'s'),stat_terms_model[['term']]))
+plural_terms  = unique(c(paste0(c(stat_terms_hyphen[['term']],
+                                  stat_terms_hyphen[['combined_term']],
+                                  stat_terms_hyphen[['update']]),'s'),
+                                  stat_terms_model[['term']],
+                                  stat_terms_single[['term']]))
 
 #str_c: define search strings
 plural_terms_all = str_c("\\b",plural_terms,"\\b",collapse="|") #to turn plural to singular
 stats_terms_all = str_c("\\b",stat_terms_hyphen[['term']],"\\b",collapse="|") #to join method words by hyphen
 stats_combined_all = str_c("\\b",stat_terms_hyphen[['combined_term']],"\\b",collapse="|") #to split method words by hyphen
+stats_terms_single = str_c("\\b",stat_terms_single[['term']],"\\b",collapse="|")
 
 #change plural terms to singular
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,

### 2020-09-02 manuscript/draft_manuscript.Rmd
update draft manuscript

--- 2020-08-25 manuscript/paper_introduction.Rmd
@@
 
-We compiled an extensive list of common methods and related terminology as a way of standardising descriptions of statistical analyses undertaken.
-An initial list was compiled by calculating individual word frequencies in the data and identifying relevant terms that appeared at least 100 times.
+We compiled an extensive list of common methods and statistical terminology to standardise descriptions of statistical analyses undertaken.
+An initial list was compiled by calculating individual word frequencies in the data and identifying relevant terms.
 Additional terms were added based on index searches of statistical textbooks [TODO need to mention this here?
-If so, add references] and spell checks (e.g.
+If so, add references] and spelling checks (e.g.
 ‘fischer’ instead of ‘fisher’).
-Using the final list, we performed multiple searches to identify possible variants of each term and replaced these with a standardised spelling.
+Multiple searches were then performed to identify possible variants of list terms, which were replaced with a standardised spelling.
 For example, instances ‘chi square’ and ‘chi squared’ were replaced with ‘chi-square’.
@@
 As a final step, common stop words including pronouns, contractions and selected prepositions were removed from the data.
-Stop words were identified from lists available from the ‘tm’ R package (Supplementary File 2).
-We chose to keep some prepositions included in the available list to minimise potential changes to the context of statistical method(s) being described (e.g.
+We chose to retain selected stop words that, if excluded, may have changed the context of statistical method(s) being described (e.g.
 ‘between’, ‘against’).
-
-<!-- ```{r echo=F,fig.cap='\\label{fig:search-strategy}Search strategy',fig.align='center',fig.height=5} -->
-<!-- knitr::include_graphics("search_strategy.png") -->
-
-<!-- ``` -->
 

### 2020-09-03 code/5_process_stats_section.R
add code to remove list of stopwords

@@ -129,6 +129,14 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
                                                                             other_terms_all,
                                                                             change_other))
 
+#remove stopwords (adpated from list availble in tm package)
+#keep select stop words in text
+add_stop = c("use","used","using")
+keep_words = c('against','between','before','after','over','under','above','below')
+stop_words_all = str_c("\\b",c(setdiff(stopwords("en"),keep_words),add_stop),"\\b",collapse='|')
+
+stats_section = stats_section %>% mutate(text_data_clean = str_remove_all(text_data_clean,stop_words_all)) 
+
 #check for remaining instances
 all_words = stats_section %>% unnest(text_data_clean) %>% 
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist() 
@@ -143,6 +151,7 @@ word_freq %>% filter(value %in% stat_terms_hyphen[['update']])
 
 word_freq %>% filter(value %in% other_terms[['term']])
 word_freq %>% filter(value %in% other_terms[['update']])
+word_freq %>% filter(value %in% c(setdiff(stopwords("en"),keep_words),add_stop))
 
 
 #remove excess whitespace
@@ -157,6 +166,7 @@ sample_data = stats_section %>% filter(doi %in% sample_dois[['doi']])
 
 write_rds(sample_data,path='data/stats_section_sample100.rds',"xz", compression = 9L)
 write_rds(stats_section,path='data/stats_section_cleaned.rds',"xz", compression = 9L)
+
 
 # save as text files (5 batches)
 stats_section_txt = stats_section %>% select(-text_data)

### 2020-09-03 code/5_process_stats_section.R
fix errors in stopword removal

@@ -97,6 +97,18 @@ stats_terms_all = str_c("\\b",stat_terms_hyphen[['term']],"\\b",collapse="|") #t
 stats_combined_all = str_c("\\b",stat_terms_hyphen[['combined_term']],"\\b",collapse="|") #to split method words by hyphen
 stats_terms_single = str_c("\\b",stat_terms_single[['term']],"\\b",collapse="|")
 
+
+#update common US to GB spelling
+other_terms_all = str_c(other_terms[['term']],collapse='|') #incorrect spellings identified
+change_other = function(input){
+  other_terms %>% filter(term==input) %>% pull(update)
+}
+
+stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,
+                                                                           other_terms_all,
+                                                                           change_other))
+
+
 #change plural terms to singular
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,
                                                                            plural_terms_all,
@@ -119,23 +131,21 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
                                                                            stats_combined_all,
                                                                            change_stats_combined))
 
-#update common US to GB spelling
-other_terms_all = str_c(other_terms[['term']],collapse='|') #incorrect spellings identified
-change_other = function(input){
-  other_terms %>% filter(term==input) %>% pull(update)
+
+
+#remove stopwords (adapted from list available in tm package)
+rm_words <- function(string, words) {
+  stopifnot(is.character(string), is.character(words))
+  spltted <- strsplit(string, " ", fixed = TRUE) # fixed = TRUE for speedup
+  vapply(spltted, function(x) paste(x[!tolower(x) %in% words], collapse = " "), character(1))
 }
 
-stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,
-                                                                            other_terms_all,
-                                                                            change_other))
-
-#remove stopwords (adpated from list availble in tm package)
 #keep select stop words in text
 add_stop = c("use","used","using")
 keep_words = c('against','between','before','after','over','under','above','below')
-stop_words_all = str_c("\\b",c(setdiff(stopwords("en"),keep_words),add_stop),"\\b",collapse='|')
+stop_words_all = c(setdiff(stopwords("en"),keep_words),add_stop)
 
[…]

### 2021-03-02 code/5_extract_unicodes.R
upload anzctr data

@@ -15,7 +15,7 @@ all_words = stats_section %>% unnest(text_data_clean) %>%
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
 
 #find all unicode characters
-unicode_lookup = str_extract_all(string=all_words,pattern=regex("<U\\+\\w+>")) %>% 
+unicode_lookup = str_extract_all(string=all_words,pattern=("\\\\u")) %>% 
   unlist() %>%
   as_tibble() %>% count(value)
 

### 2021-03-02 code/5_process_stats_section_actrn.R

@@ -10,7 +10,7 @@ library(readxl)
 library(stringi)
 library(Unicode)
 load('./data/StatsSections.RData')
-#load('./data/unicode_characters.rda')
+load('./data/unicode_characters.rda')
 
 stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
 stat_terms_single = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'single_terms')
@@ -19,20 +19,22 @@ other_terms = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'other')
 stats_section = bind_rows(studies) #changed to studies
 
 #change column headings
-stats_section = stats_section %>% rename('text_data' = stats_section)
+stats_section = studies %>% rename('text_data' = stats_section)
+#change to native encoding
+stats_section = stats_section %>% mutate(text_data = enc2native(text_data))
+stats_section = stats_section %>% mutate(text_data_clean = text_data)
 
 
 #1. remove formatting/special characters
 
-#change to native encoding, escape unicoes
-stats_section = stats_section %>% mutate(text_data = enc2native(text_data))
+#change to native encoding, escape unicodes
 stats_section = stats_section %>% mutate(text_data_clean = stri_escape_unicode(text_data))
 all_words = stats_section %>% unnest(text_data_clean) %>% 
   mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
 
 
 #find all unicode characters
-unicode_lookup = str_extract_all(all_words,pattern=regex("\\\\u\\w{4}|<U(.*)>")) %>% 
+unicode_lookup = str_extract_all(all_words,pattern=regex("\\\\u\\w{4}")) %>% 
   unlist() %>%
   as_tibble() %>% count(value) %>%
   rename('unicode'=value) %>%

### 2021-03-03 code/5_process_stats_section_actrn.R
text cleaning code for anzctr

@@ -10,7 +10,7 @@ library(readxl)
 library(stringi)
 library(Unicode)
 load('./data/StatsSections.RData')
-load('./data/unicode_characters.rda')
+#load('./data/unicode_characters.rda')
 
 stat_terms_hyphen = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'hyphen_terms')
 stat_terms_single = read_xlsx('./data/methods_dictionary.xlsx',sheet = 'single_terms')
@@ -27,28 +27,6 @@ stats_section = stats_section %>% mutate(text_data_clean = text_data)
 
 #1. remove formatting/special characters
 
-#change to native encoding, escape unicodes
-stats_section = stats_section %>% mutate(text_data_clean = stri_escape_unicode(text_data))
-all_words = stats_section %>% unnest(text_data_clean) %>% 
-  mutate(y=strsplit(text_data_clean,' ')) %>% pull(y) %>% unlist()
-
-
-#find all unicode characters
-unicode_lookup = str_extract_all(all_words,pattern=regex("\\\\u\\w{4}")) %>% 
-  unlist() %>%
-  as_tibble() %>% count(value) %>%
-  rename('unicode'=value) %>%
-  arrange(-n)
-unicode_lookup = unicode_lookup %>% mutate(label = u_char_name(gsub('\\\\u','',unicode)))
-#mutate label
-unicode_lookup = unicode_lookup %>% mutate(label_clean = gsub(' ','-',label) %>% tolower())
-
-#update 'small-letters' e.g greek-small-letter-chi to chi TODO confirm 
-unicode_lookup = unicode_lookup %>% mutate(label_clean = gsub('.*small-letter-|greek-|slanted-','',label_clean))
-unicode_lookup = unicode_lookup %>% mutate(symbol = stri_unescape_unicode(unicode))
-
-
-
 #numbered references eg [23]
 #stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"\\s*\\[\\d+\\]\\s*",""))
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,pattern="\\[\\S{1,3}\\]",replacement = ""))
@@ -57,35 +35,16 @@ stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_
 #option to keep text inside brackets is "[()]"
 stats_section = stats_section %>% mutate(text_data_clean = str_replace_all(text_data_clean,"[()]",""))
 
-#remove carriage returns
[…]

### 2021-03-03 manuscript/draft_manuscript.Rmd
methods section updated

@@
 
-## Data source
+## Data sources
 
-We chose _PLOS ONE_ because it is a large multi-disciplinary open access journal, and because the journal laudibly make the full text of papers easily available for analysis via an API.
-This enabled us to isolate the statistical analysis section of the paper, and also collect descriptive data on the paper (e.g., field of research).
+### Public Library of Science (PLOS ONE)
 
-<!--- https://journals.plos.org/plosone/s/editorial-and-peer-review-process --->
-<!--- https://journals.plos.org/plosone/s/criteria-for-publication --->
-The criteria for publication include, "Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail".
-_PLOS ONE_ does not use a separate statistical peer reviewer. Articles are handled by an academic editor who selects peer reviewers. 
-_PLOS ONE_ expect authors to use reporting guidelines from the EQUATOR network, which were designed in part because of the long history of poor statistical application and reporting in health and medical journals [@Altman2016].
+PLOS ONE is a large open access journal that publishes original research across the fundamental sciences.
+Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise.
+Submissions do not undergo formal statistical review.
+Instead, reviewers are required to assess submissions against several publication criteria, including whether “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”.
+Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
+For example, the EQUATOR network has developed reporting guidelines for study designs commonly used in health research, in part because of the long history of poor statistical application and reporting in health and medical journals [@Altman2016].
 
-We used the 'rplos' package to read the data from the API into R [@rplos].
+All articles published in PLOS ONE are freely accessible via the PLOS Application Programming Interface (API).
+This functionality enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
+Our search strategy comprised of targeted API searches followed by article filtering based on available section headings.
 
-Questions:
+_Step 1_: Targeted API searches.
+API searches were completed using the R package ‘rplos’ [@rplos].
+Search queries targeted the presence of analysis-related terms anywhere within a full-text article.
+Individual search terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
+Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2).
+By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the Material and Methods section versus Supplementary Information.
+Search results were indexed by a unique Digital Object Identifier (DOI).
+Attribute data collected per DOI included journal volume, subject classification(s) and total article views since publication.
 
-* What questions are asked of PLOS ONE Reviewers?
+_Step 2_: Partial matching on section headings.
+Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
+Since PLOS ONE does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
+To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]
 
-- Is the manuscript technically sound, and do the data support the conclusions?
-- Has the statistical analysis been performed appropriately and rigorously?
-- Does the manuscript adhere to the PLOS Data Policy?
-- Is the manuscript presented in an intelligible fashioin and written in standard English?
+### Australia and New Zealand Clinical Trials Registry (ANZCTR)
 
-* Reference updates to stats reporting guidelines; e.g see:
+ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting.
+Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment.
+Details required for registration follow a standardised template (reference or supp file), which covers important details about participant eligibility, intervention(s) being evaluated, study design and outcomes.
+As part of study design, researchers are required to provide a summary of statistical methods and planned analyses.
+Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications.
+All trials registered on AZNCTR are publicly available and can be searched for via an online portal.
+[TODO: details of how stat.section text was extracted for analysis]
 
-- https://everyone.plos.org/2019/09/26/new-plos-one-statistical-reporting-guidelines/.
-- https://web.archive.org/web/20190607174803/https://journals.plos.org/plosone/s/submission-guidelines
-- https://web.archive.org/web/20150507175314/https://journals.plos.org/plosone/s/submission-guidelines
+## Full-text processing
+Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
+Full details of text cleaning steps undertaken and corresponding R code are provided in Supplementary File X.
 
-## Search strategy
-Our search strategy comprised targeted searches of the PLOS API and filtering of full-text articles based on section headings.
-All full-text Research Articles published in _PLOS ONE_ since 2006 were eligible for inclusion.
-
-Initial searches of the PLOS API were completed using the ‘rplos’ package [@rplos].
-Submitted search queries focussed on the presence of analysis-related terms anywhere within a full-text article.
-Individual terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
-Results terms were intended to be broad and, by allowing terms to appear anywhere within the article, we accounted for the placement of relevant text in different sections, for example, in the Material and Methods section versus Supplementary Information.
-Search results were indexed by Digital Object Identifier (DOI), which allowed for removal of duplicate records.
-Metadata collected for each DOI included journal volume, PLOS subject classification(s) and number of article views since publication.
-
-Full-text XML data were then extracted for all search results for further processing.
-Data were extracted by DOI and organised by individual section and subsection headings.
-In order to consistently identify statistical methods sections, partial matching on commonly used sections headings was done to filter results.
-[TODO add more details of headings/partial matching here.
-strings were: 'statistical analysis','statistical analyses','statistical method','statistics','data analysis', 'data analyses'].
-
-Text cleaning focussed on standardising notation and references to statistical terminology, whilst minimising major changes to article style and formatting.
-Details of text cleaning steps, including R syntax, are provided in Supplementary File 1.
-
-Mathematical notation including Greek letters was converted from Unicode characters to plain text (Supplementary File 2).
-For example, instances of the code ‘<U+03B8>’, representing the Greek letter $\theta$, were replaced with ‘theta’.
-Common symbols not captured by Unicode blocks were also converted into plain text, using functions available in the 'textclean' package (e.g.
-% to ‘percent’, < to ‘less-than’) [@textclean].
-General punctuation, in-text references (e.g.
-[23]) centred equations and remaining non-ascii characters were removed.
+Mathematical notation including Greek letters was converted from Unicode characters to plain text.
+For example, the Unicode characters corresponding to $\theta$ (<U+03B8>) was replaced with ‘theta’.
+Similarly, common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were converted into plain text, using functions available in the ‘textclean’ package [@textclean].
+General formatting including carriage returns, punctuation marks, in-text references (e.g.
+[42]) centred equations and other non-ascii characters were removed.
 Text contained inside brackets was retained in the dataset to maximise content for analysis, with brackets removed.
 
-We compiled an extensive list of common methods and statistical terminology to standardise descriptions of statistical analyses undertaken.
-An initial list was compiled by calculating individual word frequencies in the data and identifying relevant terms.
-Additional terms were added based on index searches of statistical textbooks [TODO need to mention this here?
-If so, add references] and spelling checks (e.g.
-‘fischer’ instead of ‘fisher’).
-Multiple searches were then performed to identify possible variants of list terms, which were replaced with a standardised spelling.
-For example, instances ‘chi square’ and ‘chi squared’ were replaced with ‘chi-square’.
-Other variants of terminology considered plurals (e.g.
-chi squares), combinations of characters and numbers (e.g.
-chi 2) and combined text (e.g.
-chisquare).
+We compiled an extensive list of statistical terms to standardise descriptions of statistical methods reported across included record in both datasets.
+An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times.
+Further terms were sourced from index searches of statistics reference textbooks [ref].
+The final list is provided as Supplementary Material.
+Possible variants including plurals (e.g.
+‘chi-squares’) unhyphenated (e.g chi square) and combined (e.g.
+chisquare) terms were transformed to singular, hyphenated form (e.g chi-square).
+Common statistical tests were also hyphenated (e.g.
+hosmer lemeshow to hosem-lemeshow).
 
-As a final step, common stop words including pronouns, contractions and selected prepositions were removed from the data.
-We chose to retain selected stop words that, if excluded, may have changed the context of statistical method(s) being described (e.g.
+As a final step, common stop words including pronouns, contractions and selected prepositions were removed.
+We chose to keep certain stop words that, if excluded, may have changed the context of statistical method(s) being described (e.g.
 ‘between’, ‘against’).
+
+<!-- Questions: -->
+
+<!-- * What questions are asked of PLOS ONE Reviewers? -->
+
+<!-- - Is the manuscript technically sound, and do the data support the conclusions? -->
+<!-- - Has the statistical analysis been performed appropriately and rigorously? -->
+<!-- - Does the manuscript adhere to the PLOS Data Policy? -->
+<!-- - Is the manuscript presented in an intelligible fashioin and written in standard English? -->
+
+<!-- * Reference updates to stats reporting guidelines; e.g see: -->
+
+<!-- - https://everyone.plos.org/2019/09/26/new-plos-one-statistical-reporting-guidelines/. -->
+<!-- - https://web.archive.org/web/20190607174803/https://journals.plos.org/plosone/s/submission-guidelines -->
+<!-- - https://web.archive.org/web/20150507175314/https://journals.plos.org/plosone/s/submission-guidelines -->
 
@@
 We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section.
+Reiterate outcomes of random sample checking here.
 

### 2021-03-03 manuscript/draft_manuscript.Rmd
results outline

@@
 
+# Results
+
+## PLOS ONE
+
+* Final sample size
+* Breakdown of included articles by year of publication (did/did not include stats section). Final sample broken down by subject classification(s) [most are biology and life sciences]
+* Figure showing change in gof measure as function of number of clusters (use to justify choice of number of clusters)
+
+* Desribe topic clusters + examples
+
+## ANZCTR
+
+* Final sample size
+
 # Discussion

### 2021-03-08 code/6_process_meta.R
update code and data files

@@ -1,11 +1,20 @@
 #6_process_meta
 #explore meta data for final set of plos records
 library(tidyverse)
+library(ggalluvial)
 #load meta data for all records found
 load('./data/plos_meta_data.rda')
 
+g.theme = theme_bw() + theme(legend.title = element_blank(),legend.position = 'top',
+                             legend.direction = 'horizontal',axis.text = element_text(size=12),
+                             axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
+
+#number of unique API search records
+n_api = meta_dat_allrecords %>% distinct(doi,.keep_all = F) %>% nrow()
+
+
 #load final stats_section
-load('stats_section_info.rda')
+load('./data/stats_section_info.rda')
 
 #bind_rows
 stats_section = bind_rows(stats_section)
@@ -16,6 +25,39 @@ doi_list = stats_section %>% distinct(doi) %>% mutate(doi = str_replace_all(doi,
 
 #filter meta data
 meta_stats_section = meta_dat_allrecords %>% filter(doi %in% doi_list[['doi']])
+
+n_stats.section = meta_stats_section %>% distinct(doi,.keep_all = F) %>% nrow()
+
+#compare % of records that passed filter by volume
+to_plot_1 = meta_dat_allrecords %>% distinct(doi,.keep_all = T) %>% count(volume)
+to_plot_2 = meta_stats_section %>% distinct(doi,.keep_all = T) %>% count(volume)
+
+to_plot = bind_rows('API' = to_plot_1,'Included in analysis' = to_plot_2,.id='stage') %>%
+  mutate(stage = factor(stage,levels=c('API','Included in analysis')))
+
+g = ggplot(to_plot,aes(x=volume,y=n,fill=stage)) + geom_bar(stat='identity',position='dodge') + 
+  scale_fill_grey(guide = guide_legend(reverse = TRUE) ) + 
+  expand_limits(y=c(0,20000))+
+  scale_y_continuous('Total records',breaks = seq(0,20000,2500)) + 
+  scale_x_continuous('Volume',breaks = 1:15) + g.theme
+jpeg('manuscript/figures/searchresults.plos.jpg',width=600,height=480,units='px')
+g
+invisible(dev.off())
+png('manuscript/figures/searchresults.plos.png',width=600,height=480)
+g
+invisible(dev.off())
+
+#find peak
+ttab_byvolume  = to_plot %>% spread(stage,n) %>% arrange(-API)
+#percentage of records that met partial matching filters
[…]

### 2021-03-08 code/7_summarise_stats_sections.R

@@ -1,0 +1,24 @@
+#7_summarise_stats_sections.R
+library(tidyverse)
+library(tidytext)
+dat = readRDS('data/stats_section_cleaned.rds')
+
+tidy_dat = dat %>%
+  unnest_tokens(word,text_data_clean)
+
+wordcounts <- tidy_dat %>%
+  group_by(doi) %>%
+  summarise(words = n(),.groups='drop')
+
+#join with metadata
+#load meta data for all records found
+load('./data/plos_meta_data.rda')
+
+wordcounts = left_join(wordcounts,meta_dat_allrecords,by='doi')
+wordcounts %>% group_by(volume) %>% summarise(med=median(words),
+                                              Min=min(words),Max=max(words),
+                                              q1=quantile(words,.25),q3=quantile(words,.75),.groups='drop')
+
+
+ggplot(wordcounts,aes(y=log(words),x=volume,group=volume))+geom_boxplot()  
+?geom_boxplot

### 2021-03-08 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
-Many researchers may not have received adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
+Many researchers lack adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
 However, using the wrong statistical methods can cause real harm [@Altman1994] and bad statistical practices are being to used abet weak science [@Stark2018].
@@
 
-As statisticians we have heard researchers admit that they sometimes copy-and-paste their statistical methods sections from other papers, regardless of whether they are appropriate. 
-The aim of this paper is to use text-mining methods to estimate the extent that researchers are using 'boilerplate' statistical methods sections.
+Two statisticians on this paper (AB and NW) have heard researchers admit that they sometimes copy-and-paste their statistical methods sections from other papers, regardless of whether they are appropriate.
+The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
 Use of these methods sections indicates that little thought has gone into the statistical analysis.
@@
 
+We used two openly available data sources to find statistical methods sections.
+
 ### Public Library of Science (PLOS ONE)
 
-PLOS ONE is a large open access journal that publishes original research across the fundamental sciences.
+PLOS ONE is a large open access journal that publishes original research across a wide range of scientific fields.
 Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise.
 Submissions do not undergo formal statistical review.
-Instead, reviewers are required to assess submissions against several publication criteria, including whether “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”.
+Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”.
 Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
-For example, the EQUATOR network has developed reporting guidelines for study designs commonly used in health research, in part because of the long history of poor statistical application and reporting in health and medical journals [@Altman2016].
+Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].
 
@@
 This functionality enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
-Our search strategy comprised of targeted API searches followed by article filtering based on available section headings.
+To find papers with a statistical methods section we used targeted API searches followed by article filtering based on available section headings.
 
@@
 API searches were completed using the R package ‘rplos’ [@rplos].
-Search queries targeted the presence of analysis-related terms anywhere within a full-text article.
+Search queries targeted the presence of analysis-related terms anywhere in a full-text article.
 Individual search terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
@@
 
-ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting.
+The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting.
 Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment.
@@
 As part of study design, researchers are required to provide a summary of statistical methods and planned analyses.
-Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications.
-All trials registered on AZNCTR are publicly available and can be searched for via an online portal.
+<!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
+All trials registered on ANZCTR are publicly available and can be searched for via an online portal.
 [TODO: details of how stat.section text was extracted for analysis]

### 2021-03-08 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 All trials registered on ANZCTR are publicly available and can be searched for via an online portal.
-[TODO: details of how stat.section text was extracted for analysis]
+
+All studies on ANZCTR were downloaded on 1~February~2020 in XML format. Studies were excluded if the statistical methods section was empty or labelled as "Not applicable", "Nil" or "None".
 
 ## Full-text processing
+
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
@@
 Similarly, common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were converted into plain text, using functions available in the ‘textclean’ package [@textclean].
-General formatting including carriage returns, punctuation marks, in-text references (e.g.
-[42]) centred equations and other non-ascii characters were removed.
+General formatting was removed, this included carriage returns, punctuation marks, in-text references (e.g.
+"[42]") centred equations, and other non-ascii characters.
 Text contained inside brackets was retained in the dataset to maximise content for analysis, with brackets removed.
 
-We compiled an extensive list of statistical terms to standardise descriptions of statistical methods reported across included record in both datasets.
+We compiled an extensive list of statistical terms to standardise descriptions of statistical methods reported across both datasets.
 An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times.
@@
 Possible variants including plurals (e.g.
-‘chi-squares’) unhyphenated (e.g chi square) and combined (e.g.
-chisquare) terms were transformed to singular, hyphenated form (e.g chi-square).
+‘chi-squares’) unhyphenated (e.g ‘chi square’) and combined (e.g.
+‘chisquare’) terms were transformed to singular, hyphenated form (e.g ‘chi-square’).
 Common statistical tests were also hyphenated (e.g.
-hosmer lemeshow to hosem-lemeshow).
+‘hosmer lemeshow’ to ‘hosmer-lemeshow’).
 
 As a final step, common stop words including pronouns, contractions and selected prepositions were removed.
-We chose to keep certain stop words that, if excluded, may have changed the context of statistical method(s) being described (e.g.
-‘between’, ‘against’).
+We retained selected stop words that, if excluded, may have changed the context of statistical method(s) being described, for example ‘between’ and ‘against’.
 
@@
 
-Despite the extensive array of tests available, many authors are reporting the same few methods.
+Despite the extensive array of statistical tests available, many authors are reporting the same few methods.
 
@@
 
-We only examined one journal and hence our results may not be generalisable to all journals, especially those that use a statistical reviewer for all papers.
+We only examined one journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that use a statistical reviewer.
 

### 2021-03-08 manuscript/draft_manuscript.Rmd
add results to manuscript

@@
 title: "How many researchers use 'boilerplate' statistical analysis sections?"
-author: "Nicole White, Richi Nayak, Adrian Barnett"
+author: "Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett"
 date: "21/05/2020"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 library(dplyr)
@@
 library(ggplot2)
+library(ggpubr)
 panderOptions('table.emphasize.rownames', FALSE)
@@
 panderOptions('table.alignment.default','left')
+
+g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',
+                           axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
 ```
@@
 
-<!-- # Results -->
-
-<!-- ```{r source, message=F,echo=F} -->
-<!-- load('data/search_results_n.RData') -->
-<!-- ``` -->
-
-<!-- * total records found: 178654; 131847 unique DOIs -->
-<!-- * total stats sections after filtering: 111731 (85%) -->
-<!-- * Plot over time. Currently double counting if more than one search query for same doi. -->
-
-<!-- ```{r echo=F,fig.cap="\\label{fig:search-results-n}Figure caption here"} -->
-<!-- plot_dat_n = n_by_searchterm_volume %>% group_by(volume) %>%  -->
-<!--   summarise(n_matches = sum(n_unique)) %>% right_join(.,n_by_volume %>% select(-n_records),by='volume') %>%  -->
-<!--   gather(variable,value,-volume) -->
-
-<!-- plot_dat_n %>% ggplot(.,aes(x=volume,y=value,group=variable))+ -->
-<!--   geom_bar(stat='identity',position='dodge',aes(fill=variable)) -->
-
-<!-- ``` -->
-
 # Results
@@
 
-* Final sample size
-* Breakdown of included articles by year of publication (did/did not include stats section). Final sample broken down by subject classification(s) [most are biology and life sciences]
-* Figure showing change in gof measure as function of number of clusters (use to justify choice of number of clusters)
+```{r}
+load('total.records.plos.rda')
+n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
+n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
+perc.include = round(100*n.include/n.api)
+min.perc.stats = total.records[[2]] %>% filter(percentage==min(percentage)) %>% select(volume,percentage)
+max.perc.stats = total.records[[2]] %>% filter(percentage==max(percentage)) %>% select(volume,percentage)
 
-* Desribe topic clusters + examples
+```
+
+```{r}
+load('wordcount.plos.rda')
+n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
+```
+
+```{r,fig.width=10,fig.height=12,fig.cap='\\label{fig:plos-n}Search results by PLOS ONE volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
+to_plot = total.records[[2]] %>% select(-percentage) %>% gather(variable,value,-volume) %>%
+  mutate(variable=factor(variable,levels=c('API','Included in analysis')))
+
+g1 = ggplot(to_plot,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
+  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())
+
+g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
+  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count (log10 transformed)',breaks=seq(0,4,0.5))+g.theme
+
+load('stats_section_subject.classifications.rda')
+stats_section_subject = mutate(stats_section_subject,
+top3 = ifelse(subject_level_number %in% 1:3,1,0))
+to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')
+
+to_plot.bar = to_plot %>% arrange(-n) %>% select(-n) %>% gather('Top 3 subject classification',value,-subject_level_name) %>%
+mutate(subject_level_name=factor(subject_level_name,levels=rev(unique(subject_level_name))),
+'Top 3 subject classification'=factor(`Top 3 subject classification`,levels = c('No','Yes')))
+
+g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = 'top',
+legend.direction = 'horizontal',axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
+
+ggarrange(g1,g2,g3,nrow=3,labels=LETTERS[1:3])
+
+n.top3 = to_plot %>% filter(n==Yes)
+n.biology = to_plot %>% filter(subject_level_name=='Biology and life sciences') %>% pull(n)
+n.earth = to_plot %>% filter(subject_level_name=='Earth sciences') %>% pull(n)
+n.computer = to_plot %>% filter(subject_level_name=='Computer and information sciences') %>% pull(n)
+
+```
+
+API searches returned `r n.api` unique records, of which `r n.include` (`r perc.include`%) included a statistical methods section based on our search criteria.
+Search results varied by journal volume (Figure 1A).
+The total number of API search results peaked at volumes 8 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`) and 9 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`), corresponding to years 2013 and 2014.
+This trend aligned with the total number of papers published in PLOS ONE over the same timeframe.
+The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).
+
+The median length of statistical methods sections after text cleaning was `r n.words[['med']]` per DOI (IQR: `r n.words[['q1']]` to `r n.words[['q3']]`) (Figure 1B).
+
+Across all records included in analysis, all included Biology and life sciences (n = `r n.biology`), Earth sciences
+(n = `r n.earth`) and/or Computer and information sciences (n = `r n.computer`) within their top 3 subject classifications
+(Figure 1C).
 

### 2021-03-08 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 Details required for registration follow a standardised template (reference or supp file), which covers important details about participant eligibility, intervention(s) being evaluated, study design and outcomes.
-As part of study design, researchers are required to provide a summary of statistical methods and planned analyses.
+As part of study design, researchers are required to provide a "brief description" of the sample size calculations, statistical methods and planned analyses [@ANZCTR].
 <!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
@@
 
-All studies on ANZCTR were downloaded on 1~February~2020 in XML format. Studies were excluded if the statistical methods section was empty or labelled as "Not applicable", "Nil" or "None".
+All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. Studies were excluded if the statistical methods section was empty or labelled as "Not applicable", "Nil" or "None".
 
@@
 
+There were 28,008 downloaded studies, and 9,632 (34\%) had a statistical methods section.
+
 * Final sample size

### 2021-03-09 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
+[Figure here or as supplementary - 5 topic solution]
+
+Results of topic modelling assuming five clusters showed clear separation between statistical software, hypothesis testing and linear modelling (Figure).
+The largest cluster, representing 48% of all included studies (Topic 2, n = 53,491).
+
+Two topics (26% of study sample) were dominated by terms related statistical software, notably Prism/GraphPad (Topic 3, n = 10,940) and SPSS (Topic 5, n = 18,029).
+Remaining topics focussed on hypothesis testing (Topic, n = 12,531) and analysis of variance/multiple comparisons (Topic 4, n = 16,740)
+
+[Figure here or as supplementary - 10 topic solution]
+
 ## ANZCTR

### 2021-03-09 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
+Top matches (Topic 5)
+* 10.1371/journal.pone.0120046: Statistical analysis was performed using the SPSS version 18.0 (SPSS Inc., Chicago, IL, USA).
+Mann-Whitney test was used for the analysis of continuous variables.
+Chi-square test, Fisher’s exact test or one-way ANOVA test was employed for the analysis of categorical variables.
+P < 0.05 was considered statistically significant.
+
+* 10.1371/journal.pone.0204950: Statistical analyses were performed using SPSS version 20.0 for Windows (SPSS Inc., Chicago, IL).
+Categorical and continuous variables were evaluated with the non-parametric chi-square test, the Kruskal-Wallis test, and the Mann-Whitney U test.
+All statistical tests were two-sided, and a p value <0.05 was considered statistically significant.
+
+* 10.1371/journal.pone.0062685: SPSS version 21.0 for Windows (SPSS, Inc., Chicago, IL, USA) was used for statistical analysis.
+Categorical variables were analyzed by a chi-square test or Fisher’s exact test.
+Continuous variables were analyzed by independent t-test or Mann-Whitney test.
+Logistic regression analysis was performed to evaluate the effects of independent variables on clinical outcomes.
+A p-value of <0.05 by the two-tailed test was considered statistically significant.
+
+* 10.1371/journal.pone.0133783: Data are expressed as means ± SD or medians (with interquartile ranges).
+Statistical analyses were performed using SPSS software version 19.0 (SPSS, Inc., Chicago, IL, USA).
+Statistical significance for intergroup differences was assessed by the χ2 test or Fisher’s exact test for categorical variables, and by the Student’s t-test, or Mann–Whitney U test for continuous variables.
+Correlations between variables were tested with Spearman’s correlation analysis.
+A P-value <0.05 was considered significant for all tests.
+
+* 10.1371/journal.pone.0098797: Univariate analysis was performed using Mann–Whitney U-tests for continuous variables and Pearson's chi-square or Fisher's exact tests for categorical variables.
+Statistical significance was set at a P-value <0.05.
+SPSS statistical software version 19 (SPSS, Chicago, IL, USA) was used to perform the statistical analyses.
+
 [Figure here or as supplementary - 10 topic solution]

### 2021-03-09 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
+Top matches (Topic 3)
+
+* 10.1371/journal.pone.0043519: All statistical analyses were performed using GraphPad Prism (GraphPad Software Inc., San Diego, CA).
+* 10.1371/journal.pone.0045453: GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.
+* 10.1371/journal.pone.0170640: All statistical analysis of the data was performed using GraphPad Prism software (GraphPad Prism Software Inc., La Jolla, Ca).
+* 10.1371/journal.pone.0036750: All statistical analysis was performed using Graphpad Prism software.
+* 10.1371/journal.pone.0058148: All statistical analysis was performed using the Graphpad Prism software.
+
 Top matches (Topic 5)

### 2021-03-11 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 There were 28,008 downloaded studies, and 9,632 (34\%) had a statistical methods section.
+Of the 12,700 eligible applications, 9523 (75%) had a statistic section
+
+Variable & Odds ratio & 95% CI
+Study type = Observational & 0.78 & 0.69, 0.89
+Date (per year) & 0.90 & 0.88, 0.91
+Number of funders & 0.80 & 0.74, 0.86
+Target sample size (per doubling) & 0.90 & 0.88, 0.92
 
@@
 
+One reason these inadequate sections get published is that most journals do not use statistical reviewers, despite empirical evidence showing they improve manuscript quality [@Hardwicke2020].
+
 ## Limitations

### 2021-03-11 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting.
+All trials registered on ANZCTR are publicly available and can be searched for via an online portal.
 Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment.
 Details required for registration follow a standardised template (reference or supp file), which covers important details about participant eligibility, intervention(s) being evaluated, study design and outcomes.
-As part of study design, researchers are required to provide a "brief description" of the sample size calculations, statistical methods and planned analyses [@ANZCTR].
+Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, but this section is not compulsory [@ANZCTR].
 <!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
-All trials registered on ANZCTR are publicly available and can be searched for via an online portal.
 
-All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. Studies were excluded if the statistical methods section was empty or labelled as "Not applicable", "Nil" or "None".
+All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. 
+Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
 
@@
 ## Clustering algorithm
+
+## Missing statistical methods sections
+
+The statistical methods section for the ANZCTR data was somtimes missing and so we examined if there were particular studies where this section was more likely to be missing. 
+We used a logistic regression model fitted using a Bayesian paradigm. A small number of sections were labelled as "Not applicable", "Nil" or "None" and we changed these to missing.
 
@@
 
-There were 28,008 downloaded studies, and 9,632 (34\%) had a statistical methods section.
-Of the 12,700 eligible applications, 9523 (75%) had a statistic section
+There were 28,008 downloaded studies, and 9,632 (34%) had a statistical methods section. 
+The numbers of excluded studies are shown in Figure~X. 
+Of the 12,700 eligible applications, 9523 (75%) had a statistic section.
+
+The odds ratios and 95% credible intervals are shown in Table~X.
 
@@
 Target sample size (per doubling) & 0.90 & 0.88, 0.92
+
+Some of the sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS". The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
 

### 2021-03-11 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
-There were 28,008 downloaded studies, and 9,632 (34%) had a statistical methods section. 
-The numbers of excluded studies are shown in Figure~X. 
+There were 28,008 downloaded studies. The numbers of excluded studies are shown in Figure~X. 
 Of the 12,700 eligible applications, 9523 (75%) had a statistic section.
 
-The odds ratios and 95% credible intervals are shown in Table~X.
+We examined if four study characteristics predicted a missing statistics section.
+The odds ratios and 95% credible intervals are in Table~X.
+Observational studies were less likely to have a missing methods section compared with interventional studies.
+Missing sections become less likely over time.
+Studies with more funders and a larger target sample size were less likely to have a missing methods section.
 
@@
 
-Some of the sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS". The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+Some of the statistics sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS". The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
 

### 2021-03-11 code/1_analysis_anzctr_stats_section.R
Files for the missing logistic model

@@ -1,0 +1,83 @@
+# 1_analysis_anzctr_stats_section.R
+# analysis of whether there's a stats section or not
+# March 2021
+library(dplyr)
+library(INLA)
+library(stringr)
+library(inlatools)
+library(flextable)
+
+### section 1 ###
+# get the data and process ready for the model
+load('data/AnalysisReady_Stats_Section.RData') # 0_read_data_anzctr_stats_section.R
+
+# count sections as null if they are virtually empty
+null_stats = c('NA', 'N/A', '\n    ', 'Nil', 'Pending', 'NIl ', 'None', 'None.')
+studies = filter(studies, 
+                 !str_detect(number, '^NCT')) %>% # none of these have stats section
+  mutate( # binary outcome:
+           stats_section = ifelse(stats_section %in% null_stats, NA, stats_section),
+           missing = as.numeric(is.na(stats_section)),
+           # length outcome (number of characters):
+           length = nchar(stats_section),
+           length = ifelse(missing==TRUE, 0, length),
+           l_length = log(length+1), # log-transformed
+          # number of words
+          stats_section = str_replace_all(string=stats_section, pattern='  ', replacement = ' '),
+          stats_section = str_replace_all(string=stats_section, pattern='\n', replacement = ' '),
+          words = str_count(string=stats_section, pattern=' ') + 1,
+           # time
+           year = as.numeric(format(submitted, '%Y')),
+           date = as.numeric(submitted - as.Date('2015-01-01')) / (1*365.25)) # standardised to one year
+# check
+filter(studies, words==10) %>% sample_n(3) %>%pull(stats_section)
+
+# table of counts by year
+tab = group_by(studies, year, missing) %>%
+  tally() %>%
+  group_by(year) %>%
+  mutate(p = round(prop.table(n)*100)) %>%
+  ungroup() %>%
+  filter(missing==1)
+tab
+
+# it is clear that stats section was only available from 2013 onwards - confirmed on wayback machine that it was included then -- no earlier documentation
+studies = filter(studies, 
+                 year >= 2013)
+
+### section 2 ###
+## Bayesian regression model of word count with selected variables ##
+count_model <- inla(words ~ study_type + date + n_funding + log2(samplesize_target), family='poisson', data=studies, 
+                  control.compute=list(dic=TRUE))
+summary(count_model)
+# add residuals
+studies = mutate(studies, 
+            f= fitted(count_model),
[…]

### 2021-03-11 code/1_excluded_anzctr.R

@@ -1,0 +1,52 @@
+# 1_excluded_anzctr.R
+# graph of excluded studies for ANZCTR (for appendix)
+# March 2021
+library(dplyr)
+library(stringr)
+library(diagram)
+
+# get the data and process ready for the model
+load('data/AnalysisReady_Stats_Section.RData') # 0_read_data_anzctr_stats_section.R
+
+N_start = nrow(excluded) + nrow(studies)
+counts = group_by(excluded, reason) %>%
+  tally()
+
+# exclude NCT, stats section all missing
+n0 = nrow(studies)
+studies = filter(studies, !str_detect(number, '^NCT'))
+n1 = n0 - nrow(studies) # excluded number
+
+# exclude studies prior to 2013
+n0 = nrow(studies)
+studies = mutate(studies,
+                 year = as.numeric(format(submitted, '%Y'))) %>%
+  filter(year >= 2013)
+n2 = n0 - nrow(studies)
+
+## function to make diagram ANZCTR
+make_diagram_anzctr = function(){
+  par(mai=c(0,0,0,0))
+  labels = c(paste('Downloaded\n n = ', format(N_start, big.mark = ','), sep=''),
+             paste('Excluded\n- ', counts$reason[1], ', n = ', counts$n[1],
+                   '\n- ', counts$reason[2], ', n = ', counts$n[2],
+                   '\n- ', 'clinicaltrials.gov', ', n = ', n1,
+                   '\n- ', 'Pre 2013', ', n = ', n2, sep=''),
+             paste('Analysed\n n =', format(nrow(studies), big.mark = ',')))
+  n_labels = length(labels)
+  M = matrix(nrow=n_labels, ncol=n_labels)
+  M[3,1] = "' '" 
+  pos = matrix(data=c(0.29,0.8,
+                      0.7,0.5,
+                      0.29,0.2), ncol=2, byrow=TRUE)
+  sizes=c(1.5,2.3,1.5) / 10
+  props = c(0.4,0.5,0.4) # narrower for first and last
+  plotmat(M, name=labels, pos=pos, box.type = 'rect', box.size=sizes, box.prop = props, curve = 0, arr.pos=0.85)
+  shape::Arrows(x0=0.29, x1=0.45, y0=0.5, y1=0.5, arr.width=0.2, arr.length=0.22, arr.type='triangle')
+}
+make_diagram_anzctr()
+# export
[…]

### 2021-03-12 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
-PLOS ONE is a large open access journal that publishes original research across a wide range of scientific fields.
+_PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields.
 Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise.
@@
 
-All articles published in PLOS ONE are freely accessible via the PLOS Application Programming Interface (API).
+All _PLOS ONE_ articles are freely accessible via the PLOS Application Programming Interface (API).
 This functionality enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
-To find papers with a statistical methods section we used targeted API searches followed by article filtering based on available section headings.
+To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings.
+The data were downloaded on xx-xxxx-xxx.
 
@@
 API searches were completed using the R package ‘rplos’ [@rplos].
-Search queries targeted the presence of analysis-related terms anywhere in a full-text article.
-Individual search terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
+Search queries targeted the presence of analysis-related terms anywhere in the article.
+Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
 Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2).
@@
 Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
-Since PLOS ONE does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
+Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
 To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]
@@
 
-The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting.
-All trials registered on ANZCTR are publicly available and can be searched for via an online portal.
-Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment.
-Details required for registration follow a standardised template (reference or supp file), which covers important details about participant eligibility, intervention(s) being evaluated, study design and outcomes.
-Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, but this section is not compulsory [@ANZCTR].
+The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered.
+All studies registered on ANZCTR are publicly available and can be searched for via an online portal.
+<!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
+Details required for registration follow a standardised template (reference or supp file), which covers participant eligibility, intervention(s) being evaluated, study design and outcomes.
+Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR].
 <!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
@@
 All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. 
+We used all the text available in the "Statistical methods" section.
+We also collated basic information about the study including the study type (interventional or observational), submission date, number of funders and target sample size.
+These variables were chosen as we believed they might influence the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.
 Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
+Some studies were originally registered on the trial database _clinicaltrials.gov_ and essentially copied over to ANZCTR.
+We excluded these studies because they almost all had no completed statistical methods section as this is not a field in _clinicaltrials.gov_.
 

### 2021-03-12 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 _PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields.
+Articles must be in English.
 Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise.
@@
 Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2).
-By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the Material and Methods section versus Supplementary Information.
+By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_.
 Search results were indexed by a unique Digital Object Identifier (DOI).
@@
 Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR].
+All information must be in English.
+Studies are reviewed by ANZCTR staff, but are not peer reviewed.
 <!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
@@
 
-We only examined one journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that use a statistical reviewer.
+We only examined one journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.
+
+We searched the full text of _PLOS ONE_ papers but not the supporting information which may contain statistical methods sections for some papers.
+The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x%).
+We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.
 

### 2021-03-12 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered.
-All studies registered on ANZCTR are publicly available and can be searched for via an online portal.
+All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au/).
 <!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
-Details required for registration follow a standardised template (reference or supp file), which covers participant eligibility, intervention(s) being evaluated, study design and outcomes.
+Details required for registration follow a standardised template (reference or supp file), which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes.
 Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR].
-All information must be in English.
-Studies are reviewed by ANZCTR staff, but are not peer reviewed.
+The information provided must be in English.
+Studies are reviewed by ANZCTR staff for completeness of key information, which does not include the completeness of the statistical methods sections.
+Studies are not peer reviewed.
 <!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
@@
 We also collated basic information about the study including the study type (interventional or observational), submission date, number of funders and target sample size.
-These variables were chosen as we believed they might influence the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.
+These variables were chosen as we believed they might influence the completeness of the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.
 Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
-Some studies were originally registered on the trial database _clinicaltrials.gov_ and essentially copied over to ANZCTR.
-We excluded these studies because they almost all had no completed statistical methods section as this is not a field in _clinicaltrials.gov_.
+Some studies were first registered on the alternative trial database _clinicaltrials.gov_ and then also posted to ANZCTR.
+We excluded these studies because they almost all had no completed statistical methods section as this requested by _clinicaltrials.gov_.
 
@@
 
-There were 28,008 downloaded studies. The numbers of excluded studies are shown in Figure~X. 
-Of the 12,700 eligible applications, 9523 (75%) had a statistic section.
+We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
+Of the 12,700 included studies, 9,523 had a statistic section which is 75% (95% CI 74% to 76%).
 
-We examined if four study characteristics predicted a missing statistics section.
+We examined if four study characteristics were associated with a missing statistics section.
 The odds ratios and 95% credible intervals are in Table~X.
 Observational studies were less likely to have a missing methods section compared with interventional studies.
-Missing sections become less likely over time.
+Missing sections became less likely over time.
 Studies with more funders and a larger target sample size were less likely to have a missing methods section.
@@
 
-Some of the statistics sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS". The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+Some of the non-missing statistics sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".
+The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
 
@@
 
-The first line in many statistical analysis sections was the software used, implying that the software is the most important detail.
+The first line in many statistical analysis sections in _PLOS ONE_ was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail.
 As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994].
-This is not the case, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a relatively minor detail compared with detailing what methods were used and why.
+This is not the case, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.
 

### 2021-03-12 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
+Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects.
+This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018].
+This is concerning because using the wrong statistical methods can reduce the potential of study, or worse, invalidate the entire study.
+Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].
+
 The first line in many statistical analysis sections in _PLOS ONE_ was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail.

### 2021-03-12 manuscript/draft_manuscript.Rmd
update draft manuscript

@@
 n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
+p.extremes = wordcounts %>% summarise(morethan500 = round(100*mean(words>500)),lessthan50=round(100*mean(words<50)))
+
 ```
@@
 
-The median length of statistical methods sections after text cleaning was `r n.words[['med']]` per DOI (IQR: `r n.words[['q1']]` to `r n.words[['q3']]`) (Figure 1B).
+The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure 1B).
+`r n.extremes[['morethan500']]` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
+`r n.extremes[['lessthan50']]` articles (`r p.extremes[['lessthan50']]`%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.
 
-Across all records included in analysis, all included Biology and life sciences (n = `r n.biology`), Earth sciences
+All papers included Biology and life sciences (n = `r n.biology`), Earth sciences
 (n = `r n.earth`) and/or Computer and information sciences (n = `r n.computer`) within their top 3 subject classifications
 (Figure 1C).
+
+```{r,fig.width=10,fig.cap="\\label{fig:plot-5-topics}}
+
+include_graphics{'figures/plosone.5topics.jpg'}
+
+```
+
+```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}}
+
+include_graphics{'figures/plosone.10topics.jpg'}
+
+```
 

### 2021-03-12 manuscript/draft_manuscript.Rmd
update manuscript results

@@
 output:
-  pdf_document: default
+  pdf_document:
+    latex_engine: xelatex
   html_document:
@@
 library(ggpubr)
+library(knitr)
 panderOptions('table.emphasize.rownames', FALSE)
@@
 p.extremes = wordcounts %>% summarise(morethan500 = round(100*mean(words>500)),lessthan50=round(100*mean(words<50)))
+n.extremes = wordcounts %>% summarise(morethan500 = sum(words>500),lessthan50=sum(words<50))
 
@@
 API searches returned `r n.api` unique records, of which `r n.include` (`r perc.include`%) included a statistical methods section based on our search criteria.
+
 Search results varied by journal volume (Figure 1A).
@@
 The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).
+[TODO: breakdown of number of records that had stats methods in supp material, other observations about those excluded at second stage filter]
 
@@
 
-```{r,fig.width=10,fig.cap="\\label{fig:plot-5-topics}}
+```{r,fig.width=10,fig.cap="\\label{fig:plot-5-topics}Five topic solution from clustering algorithm"}
 
-include_graphics{'figures/plosone.5topics.jpg'}
+include_graphics("figures/plosone.5topics.jpg")
 
@@
 
-```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}}
+```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Ten topic solution from clustering algorithm"}
 
-include_graphics{'figures/plosone.10topics.jpg'}
+include_graphics("figures/plosone.10topics.jpg")
 
@@
 
-[Figure here or as supplementary - 5 topic solution]
+Increasing the number of clusters decreases the cluster quality.
+This indicated that the data comprised of one large cluster and multiple small clusters.
+Word cloud representing the clustering solutions based on five and ten topics are presented in Figures \@ref(fig:plot-5-topics) and \@ref(fig:plot-10-topics).
+Common topics reflected statistical software, hypothesis testing by student's t-test and one-way analysis of variance.
 
-Results of topic modelling assuming five clusters showed clear separation between statistical software, hypothesis testing and linear modelling (Figure).
-The largest cluster, representing 48% of all included studies (Topic 2, n = 53,491).
-
-Two topics (26% of study sample) were dominated by terms related statistical software, notably Prism/GraphPad (Topic 3, n = 10,940) and SPSS (Topic 5, n = 18,029).
-Remaining topics focussed on hypothesis testing (Topic, n = 12,531) and analysis of variance/multiple comparisons (Topic 4, n = 16,740)
-
-Top matches (Topic 3)
+Two topics concentrated on the use of Prism/GraphPad (Topic 3, n = 10,940; 9.8%) and SPSS (Topic 5, n = 18,029; 16%) software.
+A review of records with greatest alignment to Topic 3 showed clear overlaps in text used:
 
@@
 
-Top matches (Topic 5)
+Boilerplate text to describe statistical methods for between-group comparisons was identified in Topic 1 (n = 12,531; 11%) and Topic 4 (n = 16,740; 15%)
+Topic 1 had the smallest number of unique terms in both five (#terms = 54,780) and ten (#terms = 24,686) topic solutions.
+Statistical methods sections that were strong matches to Topic 1 focused on hypothesis testing using Student's t-test and definitions of statistical significance:
+
+* 10.1371/journal.pone.0077872: Student’s t-test was used for statistical analysis. A p value of <0.05 was considered statistically significant.
+* 10.1371/journal.pone.0144950: Student’s t-test was used for statistical analysis. A p value < 0.05 was considered to be statistically significant.
+* 10.1371/journal.pone.0019412: The Student t-test was used for all statistical analyses, and a P value less than 0.05 was considered statistically significant.
+* 10.1371/journal.pone.0055941: Statistical analysis was performed using a two-tailed Student’s t-test, and p-values <0.05 were considered statistically significant.
+* 10.1371/journal.pone.0101993: Statistical analysis was performed using Student's t-test. P-values <0.05 were considered statistically significant.
+
+Top matches for Topic 4 followed a template of summarising data by means with standard errors, formal analysis by one-way analysis of variance and post hoc mulitple comparisons for testing statistical significance:
+
+* 10.1371/journal.pone.0098207: Data are presented as the mean ±SEM.
+The data were analyzed using the Student's t-test for comparison of two groups, or by one-way analysis of variance (ANOVA) followed by Tukey's post-hoc test for multiple comparisons.
+Comparisons with p-values less than 0.05 were considered significant.
+
+* 10.1371/journal.pone.0051180: Results are presented as mean ± SEM.
+Comparison between two groups were analyzed using Student’s t test.
+Comparisons among multiple groups were performed using one-way analysis of variance (ANOVA), followed by Tukey’s test.
+
+* 10.1371/journal.pone.0207844: Values are presented as mean ± SEM.
+Differences were analyzed between groups using one-way analysis of variance (ANOVA) followed by a post hoc Tukey’s multiple comparison to determine statistical significance.
+
+* 10.1371/journal.pone.0046301: All data were presented as mean ± SEM.
+Statistical significance was assessed with one-way analysis of variance (ANOVA) followed by a post hoc (Bonferroni) test for multiple group comparison.
+Differences with p-value less than 0.05 were considered statistically significant.
+
+* 10.1371/journal.pone.0145237: Results were presented as Mean ±SEM.
+Mann-Whitney’s U test was used to compare two groups.
+For multiple group comparisons, one-way ANOVA tests were performed followed up by Tukey’s post-hoc multiple comparisons test.
+P< 0.05 was considered statistically significant.
+
+Top matches (Topic 5):
+
 * 10.1371/journal.pone.0120046: Statistical analysis was performed using the SPSS version 18.0 (SPSS Inc., Chicago, IL, USA).
@@
 SPSS statistical software version 19 (SPSS, Chicago, IL, USA) was used to perform the statistical analyses.
-
-[Figure here or as supplementary - 10 topic solution]
 

### 2021-03-15 code/7_summarise_stats_sections.R
update manuscript results plos

@@ -1,24 +1,39 @@
 #7_summarise_stats_sections.R
 library(tidyverse)
 library(tidytext)
+library(openxlsx)
 dat = readRDS('data/stats_section_cleaned.rds')
+#ten topics
+matches = read.xlsx('manuscript/top.matches.10topics.xlsx')
+matches = left_join(matches,dat,by='doi')
+write.xlsx(matches,file='manuscript/top.matches.10topics.text.xlsx',colNames=T)
 
-tidy_dat = dat %>%
-  unnest_tokens(word,text_data_clean)
-
-wordcounts <- tidy_dat %>%
-  group_by(doi) %>%
-  summarise(words = n(),.groups='drop')
-
-#join with metadata
-#load meta data for all records found
-load('./data/plos_meta_data.rda')
-
-wordcounts = left_join(wordcounts,meta_dat_allrecords,by='doi')
-wordcounts %>% group_by(volume) %>% summarise(med=median(words),
-                                              Min=min(words),Max=max(words),
-                                              q1=quantile(words,.25),q3=quantile(words,.75),.groups='drop')
+#five topics
+matches = read.xlsx('manuscript/top.matches.5topics.xlsx')
+matches = left_join(matches,dat,by='doi')
+write.xlsx(matches,file='manuscript/top.matches.5topics.text.xlsx',colNames=T)
 
 
-ggplot(wordcounts,aes(y=log(words),x=volume,group=volume))+geom_boxplot()  
-?geom_boxplot
+
+#topic 1
+filter(matches,topic==1) %>% select(doi,text_data) %>% View()
+
+# not run:
+# tidy_dat = dat %>%
+#   unnest_tokens(word,text_data_clean)
+# 
+# wordcounts <- tidy_dat %>%
+#   group_by(doi) %>%
+#   summarise(words = n(),.groups='drop')
+# 
+# #join with metadata
+# #load meta data for all records found
+# load('./data/plos_meta_data.rda')
+# 
+# wordcounts = left_join(wordcounts,meta_dat_allrecords,by='doi')
+# wordcounts %>% group_by(volume) %>% summarise(med=median(words),
+#                                               Min=min(words),Max=max(words),
+#                                               q1=quantile(words,.25),q3=quantile(words,.75),.groups='drop')
+# 
+# 
+# ggplot(wordcounts,aes(y=log(words),x=volume,group=volume))+geom_boxplot()  

### 2021-03-15 manuscript/draft_manuscript.Rmd
update manuscript results plos
update manuscript results

@@
 To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings.
-The data were downloaded on xx-xxxx-xxx.
+The data were downloaded on 3 July 2020.
 
@@
 
-Increasing the number of clusters decreases the cluster quality.
-This indicated that the data comprised of one large cluster and multiple small clusters.
-Word cloud representing the clustering solutions based on five and ten topics are presented in Figures \@ref(fig:plot-5-topics) and \@ref(fig:plot-10-topics).
-Common topics reflected statistical software, hypothesis testing by student's t-test and one-way analysis of variance.
+Classifying statistical methods sections into two clusters showed a general split between means-based testing and model-based analysis.
+Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure - Silhouette + between- & within-cluster dispersions).
+This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.
+To examine the effect of increasing the number of clusters on cluster content, we examined themes present among smaller clusters under five-topic (\@ref(fig:plot-5-topics)) and ten-topic (\@ref(fig:plot-10-topics) solutions.
+Examples of content among top matches for each topic are presented in Table X.
 
-Two topics concentrated on the use of Prism/GraphPad (Topic 3, n = 10,940; 9.8%) and SPSS (Topic 5, n = 18,029; 16%) software.
-A review of records with greatest alignment to Topic 3 showed clear overlaps in text used:
+Topic 1 had the smallest number of unique terms under both five (n = 12,531; 54,780 unique terms) and ten (n = 3,775; 24,686 unique terms) topic solutions.
+Sections assigned to Topic 1 focused on the use of Student's t-test followed by statements of statistical significance.
+Under the ten topic solution, Topic 1 was split into two clusters based on the use of 5% (Topic 1: n = 3,775) versus multiple thresholds for declaring statistical significance (Topic 9: n = 6,104).
 
-* 10.1371/journal.pone.0043519: All statistical analyses were performed using GraphPad Prism (GraphPad Software Inc., San Diego, CA).
-* 10.1371/journal.pone.0045453: GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.
-* 10.1371/journal.pone.0170640: All statistical analysis of the data was performed using GraphPad Prism software (GraphPad Prism Software Inc., La Jolla, Ca).
-* 10.1371/journal.pone.0036750: All statistical analysis was performed using Graphpad Prism software.
-* 10.1371/journal.pone.0058148: All statistical analysis was performed using the Graphpad Prism software.
+Two topics concentrated on statistical software representing 26% and 17% of all statistical methods sections under five and ten topic solutions.
+Review of top matching articles assigned to Topic 3 (Prism GraphPad) revealed evidence of boilerplate text, with many omitting details about statistical methods used.
+For sections citing the use of SPSS, increasing the number of clusters appeared to isolate references to software only versus articles that provided further details on statistical methods used.
 
-Boilerplate text to describe statistical methods for between-group comparisons was identified in Topic 1 (n = 12,531; 11%) and Topic 4 (n = 16,740; 15%)
-Topic 1 had the smallest number of unique terms in both five (#terms = 54,780) and ten (#terms = 24,686) topic solutions.
-Statistical methods sections that were strong matches to Topic 1 focused on hypothesis testing using Student's t-test and definitions of statistical significance:
+Topic 4 was characterised by the use descriptive statistics followed by between-group testing including post-hoc multiple comparisons.
+Descriptive statistics were commonly reported as means with standard deviations or standard errors.
+Related topics focused on analysis of experimental designs (Topic 10: n = 4,664) and two-group analyses only (Topic 6: n = 4,746).
 
-* 10.1371/journal.pone.0077872: Student’s t-test was used for statistical analysis. A p value of <0.05 was considered statistically significant.
-* 10.1371/journal.pone.0144950: Student’s t-test was used for statistical analysis. A p value < 0.05 was considered to be statistically significant.
-* 10.1371/journal.pone.0019412: The Student t-test was used for all statistical analyses, and a P value less than 0.05 was considered statistically significant.
-* 10.1371/journal.pone.0055941: Statistical analysis was performed using a two-tailed Student’s t-test, and p-values <0.05 were considered statistically significant.
-* 10.1371/journal.pone.0101993: Statistical analysis was performed using Student's t-test. P-values <0.05 were considered statistically significant.
+Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression/survival analysis (Topic 2: n = 19,478) and larger sections that combined data processing and analysis (Topic 8: n = 38,183) [TODO: more work needed on this]
 
-Top matches for Topic 4 followed a template of summarising data by means with standard errors, formal analysis by one-way analysis of variance and post hoc mulitple comparisons for testing statistical significance:
-
-* 10.1371/journal.pone.0098207: Data are presented as the mean ±SEM.
-The data were analyzed using the Student's t-test for comparison of two groups, or by one-way analysis of variance (ANOVA) followed by Tukey's post-hoc test for multiple comparisons.
-Comparisons with p-values less than 0.05 were considered significant.
-
-* 10.1371/journal.pone.0051180: Results are presented as mean ± SEM.
-Comparison between two groups were analyzed using Student’s t test.
-Comparisons among multiple groups were performed using one-way analysis of variance (ANOVA), followed by Tukey’s test.
-
-* 10.1371/journal.pone.0207844: Values are presented as mean ± SEM.
-Differences were analyzed between groups using one-way analysis of variance (ANOVA) followed by a post hoc Tukey’s multiple comparison to determine statistical significance.
-
-* 10.1371/journal.pone.0046301: All data were presented as mean ± SEM.
-Statistical significance was assessed with one-way analysis of variance (ANOVA) followed by a post hoc (Bonferroni) test for multiple group comparison.
-Differences with p-value less than 0.05 were considered statistically significant.
-
-* 10.1371/journal.pone.0145237: Results were presented as Mean ±SEM.
-Mann-Whitney’s U test was used to compare two groups.
-For multiple group comparisons, one-way ANOVA tests were performed followed up by Tukey’s post-hoc multiple comparisons test.
-P< 0.05 was considered statistically significant.
-
-Top matches (Topic 5):
-
-* 10.1371/journal.pone.0120046: Statistical analysis was performed using the SPSS version 18.0 (SPSS Inc., Chicago, IL, USA).
-Mann-Whitney test was used for the analysis of continuous variables.
-Chi-square test, Fisher’s exact test or one-way ANOVA test was employed for the analysis of categorical variables.
-P < 0.05 was considered statistically significant.
-
-* 10.1371/journal.pone.0204950: Statistical analyses were performed using SPSS version 20.0 for Windows (SPSS Inc., Chicago, IL).
-Categorical and continuous variables were evaluated with the non-parametric chi-square test, the Kruskal-Wallis test, and the Mann-Whitney U test.
-All statistical tests were two-sided, and a p value <0.05 was considered statistically significant.
-
-* 10.1371/journal.pone.0062685: SPSS version 21.0 for Windows (SPSS, Inc., Chicago, IL, USA) was used for statistical analysis.
-Categorical variables were analyzed by a chi-square test or Fisher’s exact test.
-Continuous variables were analyzed by independent t-test or Mann-Whitney test.
-Logistic regression analysis was performed to evaluate the effects of independent variables on clinical outcomes.
-A p-value of <0.05 by the two-tailed test was considered statistically significant.
-
-* 10.1371/journal.pone.0133783: Data are expressed as means ± SD or medians (with interquartile ranges).
-Statistical analyses were performed using SPSS software version 19.0 (SPSS, Inc., Chicago, IL, USA).
-Statistical significance for intergroup differences was assessed by the χ2 test or Fisher’s exact test for categorical variables, and by the Student’s t-test, or Mann–Whitney U test for continuous variables.
-Correlations between variables were tested with Spearman’s correlation analysis.
-A P-value <0.05 was considered significant for all tests.
-
-* 10.1371/journal.pone.0098797: Univariate analysis was performed using Mann–Whitney U-tests for continuous variables and Pearson's chi-square or Fisher's exact tests for categorical variables.
-Statistical significance was set at a P-value <0.05.
-SPSS statistical software version 19 (SPSS, Chicago, IL, USA) was used to perform the statistical analyses.
+and  Larger clusters focused on more advanced statistical methods
 

### 2021-03-15 manuscript/draft_manuscript.Rmd
revise plos results to 10 clusters

@@
 date: "21/05/2020"
-output:
-  pdf_document:
-    latex_engine: xelatex
-  html_document:
-    df_print: paged
+output: bookdown::pdf_document2
+toc: false
 link-citations: yes
@@
 
-Search results varied by journal volume (Figure 1A).
+Search results varied by journal volume (Figure \@ref(fig:plos-n)A).
 The total number of API search results peaked at volumes 8 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`) and 9 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`), corresponding to years 2013 and 2014.
@@
 
-The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure 1B).
+The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \@ref(fig:plos-n)B).
 `r n.extremes[['morethan500']]` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
@@
 (n = `r n.earth`) and/or Computer and information sciences (n = `r n.computer`) within their top 3 subject classifications
-(Figure 1C).
+(Figure \@ref(fig:plos-n)C).
 
-```{r,fig.width=10,fig.cap="\\label{fig:plot-5-topics}Five topic solution from clustering algorithm"}
-
-include_graphics("figures/plosone.5topics.jpg")
-
-```
-
-```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Ten topic solution from clustering algorithm"}
+```{r,fig.width=10,fig.align="center",fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in PLOS ONE statistical methods sections assuming ten clusters"}
 
@@
 
-Classifying statistical methods sections into two clusters showed a general split between means-based testing and model-based analysis.
-Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure - Silhouette + between- & within-cluster dispersions).
+We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
+Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters.
 This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.
-To examine the effect of increasing the number of clusters on cluster content, we examined themes present among smaller clusters under five-topic (\@ref(fig:plot-5-topics)) and ten-topic (\@ref(fig:plot-10-topics) solutions.
-Examples of content among top matches for each topic are presented in Table X.
 
-Topic 1 had the smallest number of unique terms under both five (n = 12,531; 54,780 unique terms) and ten (n = 3,775; 24,686 unique terms) topic solutions.
-Sections assigned to Topic 1 focused on the use of Student's t-test followed by statements of statistical significance.
-Under the ten topic solution, Topic 1 was split into two clusters based on the use of 5% (Topic 1: n = 3,775) versus multiple thresholds for declaring statistical significance (Topic 9: n = 6,104).
+Figure \@ref(fig:plot-10-topics) displays topic clouds based on ten clusters.
+General themes across topics reflected the use of statistical software, between-group hypothesis testing, definitions of statistical significance and more advanced statistical methods.
 
-Two topics concentrated on statistical software representing 26% and 17% of all statistical methods sections under five and ten topic solutions.
-Review of top matching articles assigned to Topic 3 (Prism GraphPad) revealed evidence of boilerplate text, with many omitting details about statistical methods used.
-For sections citing the use of SPSS, increasing the number of clusters appeared to isolate references to software only versus articles that provided further details on statistical methods used.
+Two topics reflected the use of statistical software: Prism GraphPad (Topic 3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
+Manual review of top matching sections assigned to Topic 3 showed evidence of boilerplate text.
+Out of the top 10 matches, 9 did not specify which statistical methods were used.
 
-Topic 4 was characterised by the use descriptive statistics followed by between-group testing including post-hoc multiple comparisons.
+___
+**Box 1: Examples of boilerplate text (Statistical software)**
+
+Topic 3
+
+* *graphpad prism (graphpad software, san diego, ca) was used for all analyses.*
+* *all statistical analysis was performed using the graphpad prism software.*
+* *statistical analysis were done using graphpad prism software (graphpad software, inc., la jolla, ca).*
+
+Topic 5
+
+* *all statistical analyses were conducted in spss 19 (spss inc., chicago, il, usa).*
+* *all statistical analyses were performed using the spss version 15 for windows (spss inc., chicago, il, usa). a p-value less than 0.05 was considered a significant difference.*
+* *statistical analysis was performed using the spss 17.0 software package (spss, chicago, il, usa). all data were analyzed using t tests, values of p < 0.05 were considered significant.*
+
+___
+
+Evidence of boilerplate text was identified among topics that focused two-group hypothesis testing as the only statistical method used (Box 2).
+Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance (n = 3,775).
+Similar methods sections were identified under Topic 9 (n = 6,104) but were differentiated by multiple thresholds for declaring statistical significance: \*p<0.05, \*\*p<0.01 and \*\*\*p<0.001.
+
+___
+**Box 2: Examples of boilerplate text (Between-group hypothesis testing, two groups)**
+
+Topic 1
+
+* *student’s t-test was used for statistical analysis. a p value of <0.05 was considered statistically significant.*
+* *statistical analysis was performed using student`s t-test. a p-value <0.05 was considered to be statistically significant.*
+* *statistical analysis was performed using a two-tailed student’s t-test, and p-values <0.05 were considered statistically significant.*
+
+Topic 9
+
+* *statistical significance was determined by students t-tests.* \**p<0.05,* \*\* *p = <0.01,* \*\*\* *p<0.001.*
+* *statistical analyses were performed using the nonparametric mann-whitney test. differences were considered significant for* \**p<0.05,* \*\* *p<0.01, \*\*\* *p<0.001.*
+* *comparisons between two groups were performed by student’s test. statistical significance was defined as \**p<0.05;* \*\* *p<0.01;* \*\*\* *p<0.001.*
+
+___
+
+Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3).
+[TODO: link to topics 1 and 9 here]Statistical methods sections associated with Topic 4 (n = 10,163) combined descriptive statistics followed by Student's t-test and/or one-way analysis of variance with post-hoc multiple comparisons.
 Descriptive statistics were commonly reported as means with standard deviations or standard errors.
-Related topics focused on analysis of experimental designs (Topic 10: n = 4,664) and two-group analyses only (Topic 6: n = 4,746).
+Topic 6 (n = 4,746) followed a similar template minus multiple comparisons.
+Topic 10 (Topic 10: n = 4,664) concentrated on between-group analysis applied to experimental replicates.
 
 Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression/survival analysis (Topic 2: n = 19,478) and larger sections that combined data processing and analysis (Topic 8: n = 38,183) [TODO: more work needed on this]
-
-and  Larger clusters focused on more advanced statistical methods
 

### 2021-03-15 manuscript/draft_manuscript.Rmd
update plos results for 10 clusters

@@
 Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3).
-[TODO: link to topics 1 and 9 here]Statistical methods sections associated with Topic 4 (n = 10,163) combined descriptive statistics followed by Student's t-test and/or one-way analysis of variance with post-hoc multiple comparisons.
+Topics appeared to follow the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared.
+Sections strongly associated with Topic 6 (n = 4,746) combined descriptive statistics with follow-up testing of group differences by Student's t-test, Mann Whitney U or one-way analysis of variance.
 Descriptive statistics were commonly reported as means with standard deviations or standard errors.
-Topic 6 (n = 4,746) followed a similar template minus multiple comparisons.
-Topic 10 (Topic 10: n = 4,664) concentrated on between-group analysis applied to experimental replicates.
+Topic 10 (n = 4,664) reported similar methods applied to replicates from experimental designs.
+Variations to descriptions of one-way analysis of variance were identified under Topic 4 (n = 10,163) which emphasised the use of post-hoc multiple comparisons.
 
-Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression/survival analysis (Topic 2: n = 19,478) and larger sections that combined data processing and analysis (Topic 8: n = 38,183) [TODO: more work needed on this]
+Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression/survival analysis (Topic 2: n = 19,478) and larger sections that combined data processing and analysis (Topic 8: n = 38,183).
+Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, followed by details of multivariable regression modelling.
+Examples of multivariable models included (stepwise) logistic regression and cox proportional hazards regression.
 

### 2021-03-15 manuscript/draft_manuscript.Rmd
update plos results for 10 clusters

--- 2021-03-15 manuscript/draft_manuscript.Rmd
@@
 Figure \@ref(fig:plot-10-topics) displays topic clouds based on ten clusters.
-General themes across topics reflected the use of statistical software, between-group hypothesis testing, definitions of statistical significance and more advanced statistical methods.
+General themes reflected the use of statistical software, between-group hypothesis testing, definitions of statistical significance and regression modelling.
 
@@
 
-Evidence of boilerplate text was identified among topics that focused two-group hypothesis testing as the only statistical method used (Box 2).
+Evidence of boilerplate text was identified among topics that focused on two-group hypothesis testing as the only statistical method used (Box 2).
 Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance (n = 3,775).
@@
 
-Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3).
-[TODO: link to topics 1 and 9 here]Statistical methods sections associated with Topic 4 (n = 10,163) combined descriptive statistics followed by Student's t-test and/or one-way analysis of variance with post-hoc multiple comparisons.
+Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3 - include?).
+Sections assigned to these topics appeared to follow the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared.
+Sections strongly associated with Topic 6 (n = 4,746) combined descriptive statistics with follow-up testing of group differences by Student's t-test, Mann Whitney U or one-way analysis of variance.
 Descriptive statistics were commonly reported as means with standard deviations or standard errors.
-Topic 6 (n = 4,746) followed a similar template minus multiple comparisons.
-Topic 10 (Topic 10: n = 4,664) concentrated on between-group analysis applied to experimental replicates.
+Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates obtained under experimental designs.
+Variations to descriptions of one-way analysis of variance were identified under Topic 4 (n = 10,163), with an emphasis on post-hoc multiple comparisons.
 
-Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression/survival analysis (Topic 2: n = 19,478) and larger sections that combined data processing and analysis (Topic 8: n = 38,183) [TODO: more work needed on this]
+Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183).
+Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling.
+Examples of multivariable models were (stepwise) logistic regression and cox proportional hazards regression.
 

### 2021-03-15 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 bibliography: references.bib
-subtitle: An observational study of papers published in _PLOS ONE_.
+subtitle: An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.
 always_allow_html: yes
@@
 Even when the correct methods are used, many researchers fail to describe them adequately, making it difficult to reproduce the results [@Ernst2017;@Zhou2018].
-
-The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019].
-Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked.
+Poor statistical methods might not be caught by reviewers, as they may not be qualified to judge the statistics.
 A recent survey of editors found that only 23% of health and medical journals used expert statistical review for all articles [@Hardwicke2020], which was little different from a survey from 22 years ago [@Goodman1998].
 
-Two statisticians on this paper (AB and NW) have heard researchers admit that they sometimes copy-and-paste their statistical methods sections from other papers, regardless of whether they are appropriate.
+There is guideance for researchers on how to write up their statistical methods and results.
+The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019].
+More detailed guideance is given by the SAMPL and EQUATOR guidelines [@Lang2013;@Altman2016] with the latter covering all apsects of the paper.
+Both of these guidelines were led by Doug Altman, who spoke often and for many years about the need for better statistical reporting.
+The awareness of these guidelines could likely be improved.
+There were 256 Google Scholar citations to the SAMPL paper (as at 15~March~2021) which is a good citation statistic for most papers, but feels low considering the millions of papers that use statistical analysis.
+<!--- Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked. --->
+
+Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate.
 The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
@@
 Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”.
-Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
-Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].
+Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
+<!--- Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].
+--->
 
 All _PLOS ONE_ articles are freely accessible via the PLOS Application Programming Interface (API).
-This functionality enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
+This enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
 To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings.
@@
 The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered.
-All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au/).
+All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au).
 <!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
-Details required for registration follow a standardised template (reference or supp file), which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes.
+Details required for registration follow a standardised template [@ANZCTR], which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes.
 Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR].
@@
 Some studies were first registered on the alternative trial database _clinicaltrials.gov_ and then also posted to ANZCTR.
-We excluded these studies because they almost all had no completed statistical methods section as this requested by _clinicaltrials.gov_.
+We excluded these studies because they almost all had no completed statistical methods section as this section is not included in _clinicaltrials.gov_.
 
@@
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
-Full details of text cleaning steps undertaken and corresponding R code are provided in Supplementary File X.
+Full details of the text cleaning steps  and corresponding _R_ code are provided in Supplementary File X.
 
 Mathematical notation including Greek letters was converted from Unicode characters to plain text.
-For example, the Unicode characters corresponding to $\theta$ (<U+03B8>) was replaced with ‘theta’.
+For example, the Unicode characters corresponding to $\theta$ (<U+03B8>) were replaced with ‘theta’.
 Similarly, common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were converted into plain text, using functions available in the ‘textclean’ package [@textclean].
@@
 An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times.
-Further terms were sourced from index searches of statistics reference textbooks [ref].
+Further terms were sourced from index searches of three statistics textbooks [@Dobson2018,Diggle2013,Bland2015].
 The final list is provided as Supplementary Material.
-Possible variants including plurals (e.g.
-‘chi-squares’) unhyphenated (e.g ‘chi square’) and combined (e.g.
-‘chisquare’) terms were transformed to singular, hyphenated form (e.g ‘chi-square’).
-Common statistical tests were also hyphenated (e.g.
-‘hosmer lemeshow’ to ‘hosmer-lemeshow’).
+Plurals (e.g., ‘chi-squares’) unhyphenated (e.g., ‘chi square’) and combined (e.g.
+‘chisquare’) terms were transformed to singular, hyphenated form (e.g., ‘chi-square’).
+Common statistical tests were also hyphenated (e.g., ‘hosmer lemeshow’ to ‘hosmer-lemeshow’).
 

### 2021-03-15 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”.
+All reviewers are asked the question: "Has the statistical analysis been performed appropriately and rigorously?", with the possible responses of "Yes", "No" and "I don't know".
+
 Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
@@
 
-<!-- * What questions are asked of PLOS ONE Reviewers? -->
-
-<!-- - Is the manuscript technically sound, and do the data support the conclusions? -->
-<!-- - Has the statistical analysis been performed appropriately and rigorously? -->
-<!-- - Does the manuscript adhere to the PLOS Data Policy? -->
-<!-- - Is the manuscript presented in an intelligible fashioin and written in standard English? -->
-
 <!-- * Reference updates to stats reporting guidelines; e.g see: -->
@@
 
-The statistical methods section for the ANZCTR data was somtimes missing and so we examined if there were particular studies where this section was more likely to be missing. 
+The statistical methods section for the ANZCTR data was missing for some studies and we examined if there were particular studies where this section was more likely to be missing. 
+We used four independent variables of date, study type (observational or interventional), number of funders and target sample size. 
 We used a logistic regression model fitted using a Bayesian paradigm. A small number of sections were labelled as "Not applicable", "Nil" or "None" and we changed these to missing.
@@
 
-## PLOS ONE
+## _PLOS ONE_
 
@@
 max.perc.stats = total.records[[2]] %>% filter(percentage==max(percentage)) %>% select(volume,percentage)
-
 ```
@@
 n.extremes = wordcounts %>% summarise(morethan500 = sum(words>500),lessthan50=sum(words<50))
-
 ```
 
-```{r,fig.width=10,fig.height=12,fig.cap='\\label{fig:plos-n}Search results by PLOS ONE volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
+```{r,fig.width=10,fig.height=12,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
 to_plot = total.records[[2]] %>% select(-percentage) %>% gather(variable,value,-volume) %>%
@@
 
-API searches returned `r n.api` unique records, of which `r n.include` (`r perc.include`%) included a statistical methods section based on our search criteria.
+API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search criteria.
 
 Search results varied by journal volume (Figure \@ref(fig:plos-n)A).
-The total number of API search results peaked at volumes 8 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`) and 9 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`), corresponding to years 2013 and 2014.
-This trend aligned with the total number of papers published in PLOS ONE over the same timeframe.
+The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`), corresponding to years 2013 and 2014.
+This trend aligned with the total number of papers published in _PLOS ONE_ over the same timeframe.
 The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).
@@
 The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \@ref(fig:plos-n)B).
-`r n.extremes[['morethan500']]` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
-`r n.extremes[['lessthan50']]` articles (`r p.extremes[['lessthan50']]`%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.
+`r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
+`r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.
 
-All papers included Biology and life sciences (n = `r n.biology`), Earth sciences
-(n = `r n.earth`) and/or Computer and information sciences (n = `r n.computer`) within their top 3 subject classifications
+All papers included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
+(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications
 (Figure \@ref(fig:plos-n)C).
 
-```{r,fig.width=10,fig.align="center",fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in PLOS ONE statistical methods sections assuming ten clusters"}
+```{r,fig.width=10,fig.align="center",fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}
 
@@
 Two topics reflected the use of statistical software: Prism GraphPad (Topic 3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
-Manual review of top matching sections assigned to Topic 3 showed evidence of boilerplate text.
-Out of the top 10 matches, 9 did not specify which statistical methods were used.
+Manual review of top matching sections assigned to Topic 3 showed strong evidence of boilerplate text.
+Out of the top 10 matches, 9 did not specify which statistical methods were used, but did give the software.
 
@@
 
-Evidence of boilerplate text was identified among topics that focused on two-group hypothesis testing as the only statistical method used (Box 2).
+Boilerplate text was identified among topics that focused on two-group hypothesis testing as the only statistical method used (Box 2).
 Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance (n = 3,775).

### 2021-03-15 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
+Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [wikipedia].
 Use of these methods sections indicates that little thought has gone into the statistical analysis.
@@
 Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3 - include?).
-Sections assigned to these topics appeared to follow the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared.
+Sections assigned to these topics followed the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared.
 Sections strongly associated with Topic 6 (n = 4,746) combined descriptive statistics with follow-up testing of group differences by Student's t-test, Mann Whitney U or one-way analysis of variance.
@@
 Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling.
-Examples of multivariable models were (stepwise) logistic regression and cox proportional hazards regression.
+Examples of multivariable models were (stepwise) logistic regression and Cox proportional hazards regression.
 
@@
 We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
-Of the 12,700 included studies, 9,523 had a statistic section which is 75% (95% CI 74% to 76%).
+Of the 12,700 included studies, 9,523 had a statistical methods section, 75% (95% CI 74% to 76%).
 
@@
 
+```{r}
+table =  read.table(header=TRUE, sep='&', text='
 Variable & Odds ratio & 95% CI
@@
 Number of funders & 0.80 & 0.74, 0.86
-Target sample size (per doubling) & 0.90 & 0.88, 0.92
+Target sample size (per doubling) & 0.90 & 0.88, 0.92')
+ftab = flextable(table) %>%
+theme_box()
+autofit()
+ftab
+```
 
@@
 The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+
+The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5) and repeated measures ANOVA (topic 10).
+There were cases where the exact same method section had been ...
 

### 2021-03-15 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 ___
-**Box 1: Examples of boilerplate text (Statistical software)**
+**Box 1: Examples of boilerplate text from _PLOS ONE_ (Statistical software)**
 
@@
 ___
-**Box 2: Examples of boilerplate text (Between-group hypothesis testing, two groups)**
+**Box 2: Examples of boilerplate text from _PLOS ONE_ (Between-group hypothesis testing, two groups)**
 
@@
 
-The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5) and repeated measures ANOVA (topic 10).
+The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
 There were cases where the exact same method section had been ...
+
+We also found evidence of 'boilerplate' sections clustered as topic 3, example text
+
+___
+**Box x: Examples of boilerplate text from ANZCTR**
+
+Topic 3
+
+* "Shapiro Wilk test was used as normality test.
+Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
+Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test.
+Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed"
+* "Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant."
+___
 
@@
 As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994].
-This is not the case, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.
+This is far from the truth, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.
+
+A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5% level.
+This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [mcshane] and the common misinterpretations of p-values [goodman].
 

### 2021-03-16 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 Many researchers lack adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
-However, using the wrong statistical methods can cause real harm [@Altman1994] and bad statistical practices are being to used abet weak science [@Stark2018].
+However, using the wrong statistical methods can cause real harm [@Altman1994;@Brown2018] and bad statistical practices are being to used abet weak science [@Stark2018].
 Statistical mistakes are a key source of waste in research and partly explain the current reproducibility crisis in science [@Allison2016].

### 2021-03-16 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 
+A related paper has criticised vague statistical methods sections because they deprive readers and reviewers for the opportunity to confirm that the appropriate methods were used [@Weissgerber2018].
+These authors checked hundreds of papers using ANOVA and found that 95\% did not contain the information needed to determine what type of ANOVA was performed.
+This lack of information could well be because the authors used a boilerplate statistical methods section that was missing key details.
+
+If authors shared their code then this would provide an alternative route for checking what statistical methods were used.
+This is not a perfect solution, as we still want authors to accurately report their methods, but it does increase transparency.
+However, a recent paper found that code sharing was very low in biomedical papers, with just 2\% of a sample of over 6,000 papers sharing code [@Serghiou2021].
+
 ## Limitations

### 2021-03-22 manuscript/draft_manuscript.Rmd
revised plos one results

@@
 library(knitr)
+library(flextable)
 panderOptions('table.emphasize.rownames', FALSE)
@@
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
-Full details of the text cleaning steps  and corresponding _R_ code are provided in Supplementary File X.
+_R_ code used for data extraction and cleaning is available from https://github.com/agbarnett/stats_section.
 
@@
 API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search criteria.
+In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings, including 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.
+Among DOIs that did not meet the partial matching criteria, initial search terms appeared in [TODO].
 
@@
 The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).
-[TODO: breakdown of number of records that had stats methods in supp material, other observations about those excluded at second stage filter]
 
@@
 
-```{r,fig.width=10,fig.align="center",fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}
+```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}
 
@@
 Figure \@ref(fig:plot-10-topics) displays topic clouds based on ten clusters.
-General themes reflected the use of statistical software, between-group hypothesis testing, definitions of statistical significance and regression modelling.
+Frequently occurring words reflected the use of statistical software (Topics 3 and 5), descriptive statistics (Topic 6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics 1 and 9).
+Statistical methods sections associated with regression (Topic 2) and meta-analysis (Topic 7) were also identified.
 
-Two topics reflected the use of statistical software: Prism GraphPad (Topic 3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
-Manual review of top matching sections assigned to Topic 3 showed strong evidence of boilerplate text.
-Out of the top 10 matches, 9 did not specify which statistical methods were used, but did give the software.
+Topics related to the use of statistical software differentiated between Prism GraphPad (Topic 3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
+Manual review of top matching sections assigned showed strong evidence of boilerplate text.
+Nine out ten top matches for Topic 3 stated the use of Prism GraphPad but did not specify which statistical methods were used; six out of ten top matches returned the same cluster score.
+Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
 
@@
 
-Boilerplate text was identified among topics that focused on two-group hypothesis testing as the only statistical method used (Box 2).
-Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance (n = 3,775).
-Similar methods sections were identified under Topic 9 (n = 6,104) but were differentiated by multiple thresholds for declaring statistical significance: \*p<0.05, \*\*p<0.01 and \*\*\*p<0.001.
+Definitions of statistical significance featured strongly in Topic 1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
+Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
+Topic 9 referenced similar methods combined with multiple thresholds for declaring statistical significance: \*p<0.05, \*\*p<0.01 and \*\*\*p<0.001.
+Examples of boilerplate text are give in Box 2.
 
@@
 
-Expanded descriptions of between-group analysis were found in three topics (4, 6 and 10), varying based on method(s) used and study design (Box 3 - include?).
-Sections assigned to these topics followed the general template of reporting descriptive statistics followed by methods of hypothesis testing depending on the number of groups being compared.
-Sections strongly associated with Topic 6 (n = 4,746) combined descriptive statistics with follow-up testing of group differences by Student's t-test, Mann Whitney U or one-way analysis of variance.
-Descriptive statistics were commonly reported as means with standard deviations or standard errors.
-Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates obtained under experimental designs.
-Variations to descriptions of one-way analysis of variance were identified under Topic 4 (n = 10,163), with an emphasis on post-hoc multiple comparisons.
+Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used (Box 3).
+For example, frequently occurring words in Topic 6 (n = 4,746) including 'mean', 'sd' and 'sem' reflected the use of descriptive statistics for summarising continuous variables.
+Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance.
+One-way analysis of variance also featured strongly in Topic 4 (n = 10,163), with text citing common methods for performing post-hoc multiple comparisons.
+Top matching sections across these topics did not include definitions of groups being compared.
+
+<!-- Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates collected under experimental designs.
+Top matching sections across these three topics did not define the groups being compared.
+[nw: excluded for now as evidence for boilerplate not as strong]-->
 
 Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183).
-Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling.
+
+<!-- Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling.
 Examples of multivariable models were (stepwise) logistic regression and Cox proportional hazards regression.
+-->
+
+___
+**Box 3: Examples of boilerplate text from _PLOS ONE_ (Between-group hypothesis testing, two or more groups)**
+
+Topic 4
+
+* *significant differences were determined using analysis-of-variance anova followed by tukey post-hoc tests for multiple comparisons.*
+* *statistical comparisons between study groups were performed using one-way anova test followed by post-hoc multiple comparison with dunnett test.
+p-value of less-than 0.05 were considered to be statistically significant.*
+* *one-way analysis-of-variance anova was used to compare multiple groups followed by bonferroni post-hoc test. statistical significance was set at p less-than 0.05.*
+
+Topic 6
+
+* *all results are expressed as means plus-or-minus standard deviation plus-or-minus sd.*
+* *data are presented as mean plus-or-minus sd or mean plus-or-minus sem. comparison of means was performed by anova test.*
+* *data are expressed as means plus-or-minus standard error of the mean sem. statistical analysis was performed using the student t-test. p less-than 0.05 was considered significant.*
+
+<!-- Topic 10 -->
+
+<!-- * *all statistical analyses were performed using microsoft excel.
+significance was assessed by the student t-test using data from at least three independent experiments.
+all images are representative of at least three independent experiments.* -->
+
+<!-- * *data are presented as the mean plus-or-minus sd of at least three experiments or are representative of experiments repeated at least three times.
+student t-test two-tailed paired was used to evaluate the statistical significance of results from proliferation assay experiments.* -->
+
+<!-- * *data are presented as the mean plus-or-minus sd. comparisons were made using a one-way anova. all experiments were repeated at least three times and representative experiments are shown.* -->
+
+___
 
@@
 Target sample size (per doubling) & 0.90 & 0.88, 0.92')
+
 ftab = flextable(table) %>%
-theme_box()
+theme_box() %>%
 autofit()
 ftab
+
 ```
@@
 Two sided p values of less than 0.05 will be considered statistically significant."
+
 ___

### 2021-03-22 code/6_process_meta.R

@@ -107,3 +107,56 @@ subject_patterns = distinct(doi_subject_matrix %>% select(-doi))
 
 hc = hclust(dist(subject_patterns))
 plot(hc)
+
+#excluded dois
+excluded_doi = setdiff(meta_dat_allrecords[['doi']],doi_list[['doi']]) %>% unique()
+#sample 100
+excluded_doi_sample100 = sample(excluded_doi,100)
+included_doi_sample100 = sample(doi_list[['doi']],100)
+
+prefix = c('data','statistical')
+suffix = c('analysis','analyses','method','methodology','modelling')
+search_terms = apply(expand.grid(prefix,suffix),1,paste,collapse=' ')
+q = str_c(search_terms,collapse='|')
+heading_search_result = list()
+library(rplos)
+library(XML)
+for (i in 1:length(excluded_doi)){
+  skip_to_next <- FALSE
+  tryCatch(plos_fulltext(excluded_doi[i]), error = function(e) { skip_to_next <<- TRUE})
+  if(skip_to_next) { next } 
+  
+  full_text <- plos_fulltext(excluded_doi[i])
+  full_text_xml <- xmlParse(full_text,asText=T,useInternalNodes = T,encoding='UTF-8')
+  methods_text = list()
+  node_info = getNodeSet(full_text_xml,"//sec")
+  
+  methods_text[['text_heading']] = sapply(node_info,xpathSApply,"./title",xmlValue)
+  methods_text[['text_heading']] =  unlist(as.character(methods_text[['text_heading']]))
+  
+  text_out = sapply(node_info,xpathSApply,"./p",xmlValue)
+  methods_text[['text_data']] = sapply(text_out,function(s) tolower(paste(s,collapse='')))
+  
+  if (!is.null(methods_text[['text_heading']]) & !is.null(methods_text[['text_data']])){
+    dat = do.call(cbind.data.frame,methods_text)
+    rownames(dat) <- NULL
+    dat = as_tibble(dat) %>% mutate_if(is.factor,as.character)
+    
+    #remove empty text fields
+    #dat = dat %>% filter(text_data!="")
+    
+    #filter to where original search terms appeared
+    heading_search_result[[i]] = dat %>% filter(grepl(q,ignore.case = T,text_data)|grepl(q,ignore.case = T,text_heading)) %>%
+      add_column(doi=excluded_doi[i],.before='text_heading')
+
+  }
+ 
+}
+
+
[…]

### 2021-03-29 manuscript/draft_manuscript.Rmd
plos results

@@
 library(flextable)
+library(tidytext)
 panderOptions('table.emphasize.rownames', FALSE)
@@
 
+```{r}
+load('plos.results.10topics.rda')
+
+ftab.plos = matches %>%
+  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')
+
+flextable(ftab.plos)
+
+```
+
 Figure \@ref(fig:plot-10-topics) displays topic clouds based on ten clusters.

### 2021-03-29 code/7_summarise_stats_sections.R

@@ -3,16 +3,46 @@ library(tidyverse)
 library(tidytext)
 library(openxlsx)
 dat = readRDS('data/stats_section_cleaned.rds')
-#ten topics
-matches = read.xlsx('manuscript/top.matches.10topics.xlsx')
-matches = left_join(matches,dat,by='doi')
-write.xlsx(matches,file='manuscript/top.matches.10topics.text.xlsx',colNames=T)
+#ten topics, PLOS ONE
+matches = read.csv('manuscript/plos_one_10topics.csv',header=T)
+matches = left_join(matches,dat,by=c('DOI'='doi')) %>% rename('doi'=DOI)
 
-#five topics
-matches = read.xlsx('manuscript/top.matches.5topics.xlsx')
-matches = left_join(matches,dat,by='doi')
-write.xlsx(matches,file='manuscript/top.matches.5topics.text.xlsx',colNames=T)
+#examples of boilerplate text: topic 3
+ tidy_matches = matches %>%
+   unnest_tokens(word,text_data_clean)
+ wordcounts <- tidy_matches %>%
+   group_by(topic_id,doi) %>%
+   summarise(words = n(),.groups='drop')
+ 
+matches = left_join(matches,wordcounts,by=c('doi','topic_id')) %>%
+  arrange(topic_id,-value) %>%
+  group_by(topic_id) %>% mutate(rank=row_number()) %>% ungroup() %>%
+  mutate(topic_id = paste('Topic',topic_id))
 
+
+save(matches,file='manuscript/plos.results.10topics.rda')
+
+
+#example - topics 3 and 5
+matches %>% filter(topic_id %in% c('Topic 1', 'Topic 3','Topic 5')) %>% ggplot(.,aes(x=rank,y=words)) + 
+  geom_point(alpha=0.1)+ geom_smooth(method='loess',se=T,colour='blue')+ facet_wrap(~topic_id,scales = 'free') + 
+  scale_x_continuous('Rank (1 = strongest topic match)')+scale_y_continuous('Word count (cleaned text)',breaks=seq(0,1500,100))+theme_minimal()
+
+
+#example - topics 1 and 9
+matches %>% filter(topic_id %in% c('Topic 1','Topic 9')) %>% ggplot(.,aes(x=rank,y=words)) + 
+  geom_point(alpha=0.1)+ geom_smooth(method='loess',se=T,colour='blue')+ facet_wrap(~topic_id,scales='free') + 
[…]

### 2021-03-29 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 ---
-title: "How many researchers use 'boilerplate' statistical analysis sections?"
+title: "On the use of 'boilerplate' statistical analysis sections?"
 author: "Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett"
@@
 Both of these guidelines were led by Doug Altman, who spoke often and for many years about the need for better statistical reporting.
-The awareness of these guidelines could likely be improved.
-There were 256 Google Scholar citations to the SAMPL paper (as at 15~March~2021) which is a good citation statistic for most papers, but feels low considering the millions of papers that use statistical analysis.
+The awareness and use of these guidelines could be improved.
+There were 256 Google Scholar citations to the SAMPL paper (as at 15 March 2021) which is a good citation statistic for most papers, but is low considering the millions of papers that use statistical analysis.
 <!--- Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked. --->
@@
 The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
-Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [wikipedia].
+Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [@Wikipedia].
 Use of these methods sections indicates that little thought has gone into the statistical analysis.
@@
 Articles must be in English.
-Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise.
+Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated areas of expertise.
 Submissions do not undergo formal statistical review.
-Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”.
+Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail” [@PLOS].
 All reviewers are asked the question: "Has the statistical analysis been performed appropriately and rigorously?", with the possible responses of "Yes", "No" and "I don't know".
@@
 Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2).
-By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_.
+By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_.
 Search results were indexed by a unique Digital Object Identifier (DOI).
-Attribute data collected per DOI included journal volume, subject classification(s) and total article views since publication.
+Attribute data collected per DOI included journal volume and subject classification(s).
+<!--- and total article views since publication.--->
 
@@
 Details required for registration follow a standardised template [@ANZCTR], which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes.
-Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR].
 The information provided must be in English.
-Studies are reviewed by ANZCTR staff for completeness of key information, which does not include the completeness of the statistical methods sections.
 Studies are not peer reviewed.
 <!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
+
+For the statistical methods section, researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR].
+Studies are reviewed by ANZCTR staff for completeness of key information, which does not include the completeness of the statistical methods sections.
 
@@
 
+We applied the same text cleaning to both data sources. 
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
@@
 For example, the Unicode characters corresponding to $\theta$ (<U+03B8>) were replaced with ‘theta’.
-Similarly, common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were converted into plain text, using functions available in the ‘textclean’ package [@textclean].
+Similarly, common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were converted into plain text using the ‘textclean’ package [@textclean].
 General formatting was removed, this included carriage returns, punctuation marks, in-text references (e.g.
 "[42]") centred equations, and other non-ascii characters.
-Text contained inside brackets was retained in the dataset to maximise content for analysis, with brackets removed.
+Text contained inside brackets was retained to maximise content for analysis, with brackets removed.
 

### 2021-03-29 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 General formatting was removed, this included carriage returns, punctuation marks, in-text references (e.g.
-"[42]") centred equations, and other non-ascii characters.
+"[42]") centred equations, and other non-ASCII characters.
 Text contained inside brackets was retained to maximise content for analysis, with brackets removed.
@@
 An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times.
-Further terms were sourced from index searches of three statistics textbooks [@Dobson2018,Diggle2013,Bland2015].
+Further terms were sourced from index searches of three statistics textbooks [@Dobson2018,@Diggle2013,@Bland2015].
 The final list is provided as Supplementary Material.
@@
 As a final step, common stop words including pronouns, contractions and selected prepositions were removed.
-We retained selected stop words that, if excluded, may have changed the context of statistical method(s) being described, for example ‘between’ and ‘against’.
+We retained selected stop words that, if excluded, may have changed the context of statistical methods being described, for example ‘between’ and ‘against’.
 
@@
 ## Clustering algorithm
+
+_Details to come_
+
+Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
 
@@
 to_plot = total.records[[2]] %>% select(-percentage) %>% gather(variable,value,-volume) %>%
-  mutate(variable=factor(variable,levels=c('API','Included in analysis')))
+  mutate(variable=factor(variable,levels=c('Retrieved from search','Included in analysis')))
 
@@
 
-g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
-  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count (log10 transformed)',breaks=seq(0,4,0.5))+g.theme
+g2 = ggplot(wordcounts,aes(y=words,x=volume,group=volume))+geom_boxplot()+
+  scale_x_continuous('Volume',breaks=1:14)+scale_y_log10('Word count (log-10 transformed)',breaks=c(0,10,100,200))+g.theme
 
@@
 
-g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = 'top',
+g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = c(0.8,0.2),
 legend.direction = 'horizontal',axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
@@
 The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`), corresponding to years 2013 and 2014.
-This trend aligned with the total number of papers published in _PLOS ONE_ over the same timeframe.
+This trend aligned with the total number of papers published in _PLOS ONE_ over the same period.
 The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).
@@
 
-Figure \@ref(fig:plot-10-topics) displays topic clouds based on ten clusters.
-Frequently occurring words reflected the use of statistical software (Topics 3 and 5), descriptive statistics (Topic 6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics 1 and 9).
-Statistical methods sections associated with regression (Topic 2) and meta-analysis (Topic 7) were also identified.
+The topic clouds based on ten clusters are in Figure~\@ref(fig:plot-10-topics).
+Frequently occurring words reflected the use of statistical software (Topics~3 and 5), descriptive statistics (Topic~6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9).
+There are also statistical methods sections associated with regression (Topic~2) and meta-analysis (Topic~7).
 
-Topics related to the use of statistical software differentiated between Prism GraphPad (Topic 3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
-Manual review of top matching sections assigned showed strong evidence of boilerplate text.
-Nine out ten top matches for Topic 3 stated the use of Prism GraphPad but did not specify which statistical methods were used; six out of ten top matches returned the same cluster score.
+Topics related to the use of statistical software differentiated between Prism GraphPad (Topic~3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
+A manual review of the top matching sections in these topics showed strong evidence of boilerplate text.
+Nine out ten top matches for Topic 3 stated the use of Prism GraphPad but did not specify which statistical methods were used; six out of ten top matches returned the same cluster score indicating near identical text.
 Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
@@
 
-* *statistical significance was determined by students t-tests.* \**p<0.05,* \*\* *p = <0.01,* \*\*\* *p<0.001.*
+* *statistical significance was determined by students t-tests. \* p<0.05, \*\* p = <0.01, \*\*\* p<0.001.*
 * *statistical analyses were performed using the nonparametric mann-whitney test. differences were considered significant for* \**p<0.05,* \*\* *p<0.01, \*\*\* *p<0.001.*

### 2021-03-29 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

@@
 An ideal statistical analysis will use appropriate methods to create insights from the data and inform the research questions.
-Unfortunately many current statistical analyses are far from ideal, with many researchers using the wrong methods, misinterpreting the results, or failing to adequately check their assumptions [@Goodman2008;@Leek2017].
+Unfortunately many current statistical analyses are far from ideal, with many researchers using the wrong methods, misinterpreting the results, or failing to adequately check their assumptions [@2008;@Leek2017].
 Some researchers take a "mechanistic" approach to statistics, copying the few methods they know regardless of their appropriateness, and then going through the motions of the analysis [@Stark2018].
@@
 
-Definitions of statistical significance featured strongly in Topic 1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
-Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
-Topic 9 referenced similar methods combined with multiple thresholds for declaring statistical significance: \*p<0.05, \*\*p<0.01 and \*\*\*p<0.001.
-Examples of boilerplate text are give in Box 2.
+Definitions of statistical significance featured strongly in Topic~1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
+Topic~1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
+Topic~9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: "\*p<0.05, \*\*p<0.01 and \*\*\*p<0.001", a practice that has been criticised [@Wasserstein2019].
+Examples of boilerplate text are in Box~2.
 
@@
 Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used (Box 3).
-For example, frequently occurring words in Topic 6 (n = 4,746) including 'mean', 'sd' and 'sem' reflected the use of descriptive statistics for summarising continuous variables.
+Frequently occurring words in Topic 6 (n = 4,746) including 'mean', 'sd' and 'sem' reflected the use of descriptive statistics for summarising continuous variables.
 Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance.
@@
 We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
-Of the 12,700 included studies, 9,523 had a statistical methods section, 75% (95% CI 74% to 76%).
+Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
+
+The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+Some methods sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".
+
+The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
+There were cases where the exact same method section had been re-used in a different study.
+
+We also found evidence of 'boilerplate' sections clustered as topic 3, example text
+
+___
+**Box x: Examples of boilerplate text from ANZCTR**
+
+Topic 3
+
+* "Shapiro Wilk test was used as normality test.
+Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
+Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test.
+Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed."
+* "Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant."
+___
 
@@
 ftab
-
 ```
-
-Some of the non-missing statistics sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".
-The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
-
-The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
-There were cases where the exact same method section had been ...
-
-We also found evidence of 'boilerplate' sections clustered as topic 3, example text
-
-___
-**Box x: Examples of boilerplate text from ANZCTR**
-
-Topic 3
-
-* "Shapiro Wilk test was used as normality test.
-Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
-Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test.
-Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed"
-* "Comparisons between categorical variables will be made either using chi square or Fisher exact test.
-Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
-Two sided p values of less than 0.05 will be considered statistically significant."
-
-___
 
@@
 # Discussion
-
-Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects.
-This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018].
-This is concerning because using the wrong statistical methods can reduce the potential of study, or worse, invalidate the entire study.
-Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].
 
@@
 A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5% level.
-This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [mcshane] and the common misinterpretations of p-values [goodman].
+This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [@McShane2019] and the common misinterpretations of p-values [@Goodman2008].
 
@@
 
+Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects.
+This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018].
+This is concerning because using the wrong statistical methods can reduce the value of study, or worse, invalidate the entire study.
+These mistakes are avoidable and are wasting of thousands of hours of researchers' time and the time of patients and volunteers.
+Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].
+
 ## Limitations
 
-We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods section would be fine.
+We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods might be adequate.
 
@@
 
-We only examined one journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.
+We only examined one large journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.
 

### 2021-04-20 manuscript/draft_manuscript.tex
tex files for draft_manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={How many researchers use `boilerplate' statistical analysis sections?},
  pdfauthor={Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{How many researchers use `boilerplate' statistical analysis sections?}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{An observational study of papers published in \emph{PLOS ONE}.}
\author{Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett}
\date{21/05/2020}

\begin{document}
\maketitle

An ideal statistical analysis will use appropriate methods to create
insights from the data and inform the research questions. Unfortunately
many current statistical analyses are far from ideal, with many
researchers using the wrong methods, misinterpreting the results, or
failing to adequately check their assumptions (Goodman
\protect\hyperlink{ref-Goodman2008}{2008}; Leek et al.
\protect\hyperlink{ref-Leek2017}{2017}). Some researchers take a
``mechanistic'' approach to statistics, copying the few methods they
know regardless of their appropriateness, and then going through the
motions of the analysis (Stark and Saltelli
\protect\hyperlink{ref-Stark2018}{2018}).

Many researchers lack adequate training in research methods, and
statistics is something they do with trepidation and even ignorance
(Altman \protect\hyperlink{ref-Altman1994}{1994}; King et al.
\protect\hyperlink{ref-King2019}{2019}). However, using the wrong
statistical methods can cause real harm (Altman
\protect\hyperlink{ref-Altman1994}{1994}) and bad statistical practices
are being to used abet weak science (Stark and Saltelli
\protect\hyperlink{ref-Stark2018}{2018}). Statistical mistakes are a key
source of waste in research and partly explain the current
reproducibility crisis in science (Allison et al.
\protect\hyperlink{ref-Allison2016}{2016}). Even when the correct
methods are used, many researchers fail to describe them adequately,
making it difficult to reproduce the results (Ernst and Albers
\protect\hyperlink{ref-Ernst2017}{2017}; Zhou and Skidmore
\protect\hyperlink{ref-Zhou2018}{2018}).

The International Committee of Medical Journal Editors recommend that
researchers should: ``Describe statistical methods with enough detail to
enable a knowledgeable reader with access to the original data to judge
its appropriateness for the study and to verify the reported results''
(ICJME \protect\hyperlink{ref-ICMJE2019}{2019}). Although the general
lack of statistical understanding from both authors and reviewers means
this recommendation may not be checked. A recent survey of editors found
that only 23\% of health and medical journals used expert statistical
review for all articles (Hardwicke and Goodman
\protect\hyperlink{ref-Hardwicke2020}{2020}), which was little different
from a survey from 22 years ago (Goodman, Altman, and George
\protect\hyperlink{ref-Goodman1998}{1998}).

Two statisticians on this paper (AB and NW) have heard researchers admit
that they sometimes copy-and-paste their statistical methods sections
from other papers, regardless of whether they are appropriate. The aim
of this paper is to use text-mining methods to estimate the extent that
researchers are using cut-and-paste or `boilerplate' statistical methods
sections. Use of these methods sections indicates that little thought
has gone into the statistical analysis.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-sources}{%
\subsection{Data sources}\label{data-sources}}

We used two openly available data sources to find statistical methods
sections.

\hypertarget{public-library-of-science-plos-one}{%
\subsubsection{Public Library of Science (PLOS
ONE)}\label{public-library-of-science-plos-one}}

\emph{PLOS ONE} is a large open access journal that publishes original
research across a wide range of scientific fields. Articles must be in
English. Article submissions are handled by an academic editor who
selects peer reviewers based on their self-nominated area(s) of
expertise. Submissions do not undergo formal statistical review.
Instead, reviewers are required to assess submissions against several
publication criteria, including whether: ``Experiments, statistics, and
other analyses are performed to a high technical standard and are
described in sufficient detail''. Authors are encouraged to follow
published reporting guidelines to ensure that chosen statistical methods
are appropriate for the study design, and adequate details are provided
to enable independent replication of results. Reporting guidelines are
available from the EQUATOR network which has developed guidelines for
study designs commonly used in health research, and these guidelines
were developed to tackle poor statistical application and reporting in
health and medical journals (Altman and Simera
\protect\hyperlink{ref-Altman2016}{2016}).

All \emph{PLOS ONE} articles are freely accessible via the PLOS
Application Programming Interface (API). This functionality enabled us
to conduct semi-automated searches of full-text articles and analyse
data on individual records, including text content and general
attributes such as publication date and field(s) of research. To find
papers with a statistical methods section we used targeted API searches
followed by article filtering based on section headings. The data were
downloaded on xx-xxxx-xxx.

\emph{Step 1}: Targeted API searches. API searches were completed using
the R package `rplos' (Chamberlain, Boettiger, and Ram
\protect\hyperlink{ref-rplos}{2020}). Search queries targeted the
presence of analysis-related terms anywhere in the article. Search terms
combined the words ``data'' or ``statistical'' with one of:
``analysis'', ``analyses'', ``method'', ``methodology'' or
``model(l)ing''. Search terms were intended to be broad whilst keeping
search results to a manageable number for full-text review (see Step 2).
By allowing terms to appear anywhere within the article, we accounted
for the possibility of relevant text being placed in different sections,
for example, in the \emph{Material and Methods} section versus
\emph{Results}. Search results were indexed by a unique Digital Object
Identifier (DOI). Attribute data collected per DOI included journal
volume, subject classification(s) and total article views since
publication.

\emph{Step 2}: Partial matching on section headings. Full text XML data
for all search results were downloaded and combined into a single
dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
ONE} does not prescribe standardised headings to preface statistical
methods sections, we performed partial matching on available headings
against frequently used terms in initial search results: `Statistical
analysis', `Statistical analyses', `Statistical method', `Statistics',
`Data analysis' and `Data analyses'. To determine the reliability of our
chosen filters, we manually reviewed full text data extracted for a
random sample of XXX articles that were not matched (File
S1).{[}TODO\ldots finish this thought\ldots{]}

\hypertarget{australia-and-new-zealand-clinical-trials-registry-anzctr}{%
\subsubsection{Australia and New Zealand Clinical Trials Registry
(ANZCTR)}\label{australia-and-new-zealand-clinical-trials-registry-anzctr}}

The ANZCTR was established in 2005 as part of a coordinated global
effort to improve research quality and transparency in clinical trials
reporting; observational studies can also be registered. All studies
registered on ANZCTR are publicly available and can be searched via an
online portal (\url{https://www.anzctr.org.au/}). Details required for
registration follow a standardised template (reference or supp file),
which covers participant eligibility, the intervention(s) being
evaluated, study design and outcomes. Researchers are asked to provide a
``brief description'' of the sample size calculations, statistical
methods and planned analyses, although this section is not compulsory
(anzctr \protect\hyperlink{ref-ANZCTR}{2019}). The information provided
must be in English. Studies are reviewed by ANZCTR staff for
completeness of key information, which does not include the completeness
of the statistical methods sections. Studies are not peer reviewed.

All studies available on ANZCTR were downloaded on 1 February 2020 in
XML format. We used all the text available in the ``Statistical
methods'' section. We also collated basic information about the study
including the study type (interventional or observational), submission
date, number of funders and target sample size. These variables were
chosen as we believed they might influence the completeness of the
statistical methods section, because we expected larger studies and
those with funding to be more complete, and we also were interested in
changes over time. Studies prior to 2013 were excluded as the
statistical methods section appeared to be introduced in 2013. Some
studies were first registered on the alternative trial database
\emph{clinicaltrials.gov} and then also posted to ANZCTR. We excluded
these studies because they almost all had no completed statistical
methods section as this requested by \emph{clinicaltrials.gov}.

\hypertarget{full-text-processing}{%
\subsection{Full-text processing}\label{full-text-processing}}

Text cleaning aimed to standardise notation and statistical terminology,
whilst minimising changes to article style and formatting. Full details
of text cleaning steps undertaken and corresponding R code are provided
in Supplementary File X.

Mathematical notation including Greek letters was converted from Unicode
characters to plain text. For example, the Unicode characters
corresponding to \(\theta\) (\textless U+03B8\textgreater) was replaced
with `theta'. Similarly, common symbols outside of Unicode blocks
including `\%' (percent) and `\textless{}' (`less-than') were converted
into plain text, using functions available in the `textclean' package
(Rinker \protect\hyperlink{ref-textclean}{2018}). General formatting was
removed, this included carriage returns, punctuation marks, in-text
references (e.g.~``{[}42{]}'') centred equations, and other non-ascii
characters. Text contained inside brackets was retained in the dataset
to maximise content for analysis, with brackets removed.

We compiled an extensive list of statistical terms to standardise
descriptions of statistical methods reported across both datasets. An
initial list was compiled by calculating individual word frequencies and
identifying relevant terms that appeared at least 100 times. Further
terms were sourced from index searches of statistics reference textbooks
{[}ref{]}. The final list is provided as Supplementary Material.
Possible variants including plurals (e.g.~`chi-squares') unhyphenated
(e.g `chi square') and combined (e.g.~`chisquare') terms were
transformed to singular, hyphenated form (e.g `chi-square'). Common
statistical tests were also hyphenated (e.g.~`hosmer lemeshow' to
`hosmer-lemeshow').

As a final step, common stop words including pronouns, contractions and
selected prepositions were removed. We retained selected stop words
that, if excluded, may have changed the context of statistical method(s)
being described, for example `between' and `against'.

\hypertarget{clustering-algorithm}{%
\subsection{Clustering algorithm}\label{clustering-algorithm}}

\hypertarget{missing-statistical-methods-sections}{%
\subsection{Missing statistical methods
sections}\label{missing-statistical-methods-sections}}

The statistical methods section for the ANZCTR data was somtimes missing
and so we examined if there were particular studies where this section
was more likely to be missing. We used a logistic regression model
fitted using a Bayesian paradigm. A small number of sections were
labelled as ``Not applicable'', ``Nil'' or ``None'' and we changed these
to missing.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{plos-one}{%
\subsection{PLOS ONE}\label{plos-one}}

\begin{figure}
\centering
\includegraphics{draft_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:plos-n}Search results by PLOS ONE volume (1st row);
word count per statistical methods section included in analysis (n =
111,731; 2nd row); subject classifications assigned to full-text records
included in analysis (3rd row)}
\end{figure}

API searches returned 131847 unique records, of which 111731 (85\%)
included a statistical methods section based on our search criteria.
Search results varied by journal volume (Figure 1A). The total number of
API search results peaked at volumes 8 (n = 19045) and 9 (n = 19045),
corresponding to years 2013 and 2014. This trend aligned with the total
number of papers published in PLOS ONE over the same timeframe. The
percentage of records that included a statistical methods section by
volume based on our proposed matching criteria varied between 64\%
(volume 2) and 86\% (volume 9).

The median length of statistical methods sections after text cleaning
was 127 per DOI (IQR: 61 to 254) (Figure 1B).

Across all records included in analysis, all included Biology and life
sciences (n = 107584), Earth sciences (n = 7605) and/or Computer and
information sciences (n = 5190) within their top 3 subject
classifications (Figure 1C).

{[}Figure here or as supplementary - 5 topic solution{]}

Results of topic modelling assuming five clusters showed clear
separation between statistical software, hypothesis testing and linear
modelling (Figure). The largest cluster, representing 48\% of all
included studies (Topic 2, n = 53,491).

Two topics (26\% of study sample) were dominated by terms related
statistical software, notably Prism/GraphPad (Topic 3, n = 10,940) and
SPSS (Topic 5, n = 18,029). Remaining topics focussed on hypothesis
testing (Topic, n = 12,531) and analysis of variance/multiple
comparisons (Topic 4, n = 16,740)

Top matches (Topic 3)

\begin{itemize}
\tightlist
\item
  10.1371/journal.pone.0043519: All statistical analyses were performed
  using GraphPad Prism (GraphPad Software Inc., San Diego, CA).
\item
  10.1371/journal.pone.0045453: GraphPad Prism (Graphpad Software, San
  Diego, CA) was used for all analyses.
\item
  10.1371/journal.pone.0170640: All statistical analysis of the data was
  performed using GraphPad Prism software (GraphPad Prism Software Inc.,
  La Jolla, Ca).
\item
  10.1371/journal.pone.0036750: All statistical analysis was performed
  using Graphpad Prism software.
\item
  10.1371/journal.pone.0058148: All statistical analysis was performed
  using the Graphpad Prism software.
\end{itemize}

Top matches (Topic 5) * 10.1371/journal.pone.0120046: Statistical
analysis was performed using the SPSS version 18.0 (SPSS Inc., Chicago,
IL, USA). Mann-Whitney test was used for the analysis of continuous
variables. Chi-square test, Fisher's exact test or one-way ANOVA test
was employed for the analysis of categorical variables. P \textless{}
0.05 was considered statistically significant.

\begin{itemize}
\item
  10.1371/journal.pone.0204950: Statistical analyses were performed
  using SPSS version 20.0 for Windows (SPSS Inc., Chicago, IL).
  Categorical and continuous variables were evaluated with the
  non-parametric chi-square test, the Kruskal-Wallis test, and the
  Mann-Whitney U test. All statistical tests were two-sided, and a p
  value \textless0.05 was considered statistically significant.
\item
  10.1371/journal.pone.0062685: SPSS version 21.0 for Windows (SPSS,
  Inc., Chicago, IL, USA) was used for statistical analysis. Categorical
  variables were analyzed by a chi-square test or Fisher's exact test.
  Continuous variables were analyzed by independent t-test or
  Mann-Whitney test. Logistic regression analysis was performed to
  evaluate the effects of independent variables on clinical outcomes. A
  p-value of \textless0.05 by the two-tailed test was considered
  statistically significant.
\item
  10.1371/journal.pone.0133783: Data are expressed as means ± SD or
  medians (with interquartile ranges). Statistical analyses were
  performed using SPSS software version 19.0 (SPSS, Inc., Chicago, IL,
  USA). Statistical significance for intergroup differences was assessed
  by the χ2 test or Fisher's exact test for categorical variables, and
  by the Student's t-test, or Mann--Whitney U test for continuous
  variables. Correlations between variables were tested with Spearman's
  correlation analysis. A P-value \textless0.05 was considered
  significant for all tests.
\item
  10.1371/journal.pone.0098797: Univariate analysis was performed using
  Mann--Whitney U-tests for continuous variables and Pearson's
  chi-square or Fisher's exact tests for categorical variables.
  Statistical significance was set at a P-value \textless0.05. SPSS
  statistical software version 19 (SPSS, Chicago, IL, USA) was used to
  perform the statistical analyses.
\end{itemize}

{[}Figure here or as supplementary - 10 topic solution{]}

\hypertarget{anzctr}{%
\subsection{ANZCTR}\label{anzctr}}

We downloaded 28,008 studies. The numbers of excluded studies are shown
in Figure\textasciitilde X. Of the 12,700 included studies, 9,523 had a
statistic section which is 75\% (95\% CI 74\% to 76\%).

We examined if four study characteristics were associated with a missing
statistics section. The odds ratios and 95\% credible intervals are in
Table\textasciitilde X. Observational studies were less likely to have a
missing methods section compared with interventional studies. Missing
sections became less likely over time. Studies with more funders and a
larger target sample size were less likely to have a missing methods
section.

Variable \& Odds ratio \& 95\% CI Study type = Observational \& 0.78 \&
0.69, 0.89 Date (per year) \& 0.90 \& 0.88, 0.91 Number of funders \&
0.80 \& 0.74, 0.86 Target sample size (per doubling) \& 0.90 \& 0.88,
0.92

Some of the non-missing statistics sections were only one word,
including ``ANOVA'', ``t-test'', ``SPSS'' and even ``SSPS''. The median
length of the section was 129 words with an inter-quartile range of 71
to 219 words.

\begin{itemize}
\tightlist
\item
  Final sample size
\end{itemize}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Many researchers are using lazy practice by copying a standard
``boilerplate'' statistical methods section, likely cut-and-pasting from
other researchers or projects. This is a strong sign of the ritualistic
practice of statistics where researchers go through the motions rather
than using conscientious practice (Stark and Saltelli
\protect\hyperlink{ref-Stark2018}{2018}). This is concerning because
using the wrong statistical methods can reduce the potential of study,
or worse, invalidate the entire study. Poor statistical practice is a
key driver of the ongoing reproducibility crisis in science (Ioannidis
et al. \protect\hyperlink{ref-Ioannidis2014}{2014}).

The first line in many statistical analysis sections in \emph{PLOS ONE}
was the software used and some entire sections in ANZCTR only stated the
software, implying that the software is the most important detail. As
Doug Altman said, ``Many people think that all you need to do statistics
is a computer and appropriate software'' (Altman
\protect\hyperlink{ref-Altman1994}{1994}). This is not the case, and
whilst it is important for researchers to mention the software and
version used for reproducibility purposes, it is a minor detail compared
with detailing what methods were used and why.

Despite the extensive array of statistical tests available, many authors
are reporting the same few methods.

One reason these inadequate sections get published is that most journals
do not use statistical reviewers, despite empirical evidence showing
they improve manuscript quality (Hardwicke and Goodman
\protect\hyperlink{ref-Hardwicke2020}{2020}).

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

We did not check whether papers used the correct methods, and for some
simple studies a `boilerplate' statistical methods section would be
fine.

We examined papers where there was a statistics section, and we missed
papers that used statistical analysis but did not include a statistical
analysis section. Reiterate outcomes of random sample checking here.

We only examined one journal and one trial registry and hence our
results may not be generalisable to all journals or registries,
especially those that consistently use a statistical reviewer.

We searched the full text of \emph{PLOS ONE} papers but not the
supporting information which may contain statistical methods sections
for some papers. The search terms we used to find statistical methods
appeared in the supporting information titles for xxx papers (x\%). We
did not include the supporting information because it is less structured
than the paper and could be in PDF or Word format.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allison2016}{}%
Allison, David B., Andrew W. Brown, Brandon J. George, and Kathryn A.
Kaiser. 2016. ``Reproducibility: A Tragedy of Errors.'' \emph{Nature}
530 (7588): 27--29. \url{https://doi.org/10.1038/530027a}.

\leavevmode\hypertarget{ref-Altman1994}{}%
Altman, D G. 1994. ``The Scandal of Poor Medical Research.'' \emph{BMJ}
308 (6924): 283--84. \url{https://doi.org/10.1136/bmj.308.6924.283}.

\leavevmode\hypertarget{ref-Altman2016}{}%
Altman, Douglas G, and Iveta Simera. 2016. ``A History of the Evolution
of Guidelines for Reporting Medical Research: The Long Road to the
EQUATOR Network.'' \emph{Journal of the Royal Society of Medicine} 109
(2): 67--77. \url{https://doi.org/10.1177/0141076815625599}.

\leavevmode\hypertarget{ref-ANZCTR}{}%
anzctr. 2019. ``ANZCTR Data Field Definitions V25.''
\url{https://www.anzctr.org.au/docs/ANZCTR\%20Data\%20field\%20explanation.pdf}.

\leavevmode\hypertarget{ref-rplos}{}%
Chamberlain, Scott, Carl Boettiger, and Karthik Ram. 2020. \emph{Rplos:
Interface to the Search API for 'PLoS' Journals}.
\url{https://CRAN.R-project.org/package=rplos}.

\leavevmode\hypertarget{ref-Ernst2017}{}%
Ernst, Anja F., and Casper J. Albers. 2017. ``Regression Assumptions in
Clinical Psychology Research Practicea Systematic Review of Common
Misconceptions.'' \emph{PeerJ} 5: e3323.
\url{https://doi.org/10.7717/peerj.3323}.

\leavevmode\hypertarget{ref-Goodman2008}{}%
Goodman, Steven. 2008. ``A Dirty Dozen: Twelve P-Value Misconceptions.''
\emph{Seminars in Hematology} 45 (3): 135--40.
\url{https://doi.org/10.1053/j.seminhematol.2008.04.003}.

\leavevmode\hypertarget{ref-Goodman1998}{}%
Goodman, Steven N., Douglas G. Altman, and Stephen L. George. 1998.
``Statistical Reviewing Policies of Medical Journals.'' \emph{Journal of
General Internal Medicine} 13 (11): 753--56.
\url{https://doi.org/10.1046/j.1525-1497.1998.00227.x}.

\leavevmode\hypertarget{ref-Hardwicke2020}{}%
Hardwicke, Tom E, and Steve Goodman. 2020. ``How Often Do Leading
Biomedical Journalsuse Statistical Experts to Evaluate Statistical
Methods? The Results of a Survey.'' MetaArXiv.
\url{https://doi.org/10.31222/osf.io/z27u4}.

\leavevmode\hypertarget{ref-ICMJE2019}{}%
ICJME. 2019. ``Recommendations for the Conduct, Reporting, Editing, and
Publication of Scholarly Work in Medical Journals.''
\url{http://www.icmje.org/icmje-recommendations.pdf}.

\leavevmode\hypertarget{ref-Ioannidis2014}{}%
Ioannidis, John P A, Sander Greenland, Mark A Hlatky, Muin J Khoury,
Malcolm R Macleod, David Moher, Kenneth F Schulz, and Robert Tibshirani.
2014. ``Increasing Value and Reducing Waste in Research Design, Conduct,
and Analysis.'' \emph{The Lancet} 383 (9912): 166--75.
\url{https://doi.org/10.1016/s0140-6736(13)62227-8}.

\leavevmode\hypertarget{ref-King2019}{}%
King, Kevin M., Michael D. Pullmann, Aaron R. Lyon, Shannon Dorsey, and
Cara C. Lewis. 2019. ``Using Implementation Science to Close the Gap
Between the Optimal and Typical Practice of Quantitative Methods in
Clinical Science.'' \emph{Journal of Abnormal Psychology} 128 (6):
547--62. \url{https://doi.org/10.1037/abn0000417}.

\leavevmode\hypertarget{ref-Leek2017}{}%
Leek, Jeff, Blakeley B. McShane, Andrew Gelman, David Colquhoun, Michèle
B. Nuijten, and Steven N. Goodman. 2017. ``Five Ways to Fix
Statistics.'' \emph{Nature} 551 (7682): 557--59.
\url{https://doi.org/10.1038/d41586-017-07522-z}.

\leavevmode\hypertarget{ref-textclean}{}%
Rinker, Tyler W. 2018. \emph{textclean: Text Cleaning Tools}. Buffalo,
New York. \url{https://github.com/trinker/textclean}.

\leavevmode\hypertarget{ref-Stark2018}{}%
Stark, Philip B., and Andrea Saltelli. 2018. ``Cargo-Cult Statistics and
Scientific Crisis.'' \emph{Significance} 15 (4): 40--43.
\url{https://doi.org/10.1111/j.1740-9713.2018.01174.x}.

\leavevmode\hypertarget{ref-Zhou2018}{}%
Zhou, Yuanyuan, and Susan Skidmore. 2018. ``A Reassessment of ANOVA
Reporting Practices: A Review of Three APA Journals.'' \emph{Journal of
Methods and Measurement in the Social Sciences} 8 (1): 3--19.
\url{https://doi.org/10.2458/v8i1.22019}.

\end{document}

### 2021-04-20 manuscript/asa_template/asa_template.Rmd
update to asa template

--- 2021-03-29 manuscript/draft_manuscript.Rmd
@@
 ---
-title: "On the use of 'boilerplate' statistical analysis sections?"
-author: "Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett"
-date: "21/05/2020"
-output: bookdown::pdf_document2
-toc: false
-link-citations: yes
+title: An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.
+
+# to produce blinded version set to 1
+blinded: 0
+
+authors: 
+- name: Nicole White
+  thanks: The authors gratefully acknowledge ...
+  affiliation: School of Public Health and Social Work, QUT
+
+- name: Thiru Balasubramaniam, Richi Nayak
+  affiliation: To add
+
+- name: Adrian Barnett
+  affiliation: School of Public Health and Social Work, QUT
+
+keywords:
+- 3 to 6 keywords
+- that do not appear in the title
+
+abstract: |
+  The text of your abstract.  200 or fewer words.
+
 bibliography: references.bib
-subtitle: An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.
-always_allow_html: yes
+output: rticles::asa_article
+fig_caption: yes
 ---
@@
 knitr::opts_chunk$set(echo = FALSE)
-library(dplyr)
-library(tidyr)
-library(pander)
-library(ggplot2)
+library(tidyverse)
 library(ggpubr)
@@
 library(tidytext)
-panderOptions('table.emphasize.rownames', FALSE)
-panderOptions('keep.trailing.zeros', TRUE)
-panderOptions('table.split.table', Inf)
-panderOptions('table.split.cells', Inf)
-panderOptions('big.mark', ',')
-panderOptions('table.emphasize.rownames', FALSE)
-panderOptions('keep.trailing.zeros', TRUE)
-panderOptions('table.split.table', Inf)
-panderOptions('table.split.cells', Inf)
-panderOptions('big.mark', ',')
-panderOptions('keep.line.breaks', TRUE)
-panderOptions('table.alignment.default','left')
 
-g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',
-                           axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
+g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal')
 ```
+
+# Introduction
 
@@
 
-There is guideance for researchers on how to write up their statistical methods and results.
+There is guidance for researchers on how to write up their statistical methods and results.
 The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019].
@@
 
-## Data sources
+\subsection{Data sources}
 
-We used two openly available data sources to find statistical methods sections.
+We used two openly available data sources to find statistical methods sections, covering peer-reviewed journal articles and registered clinical trial protocols.
 
-### Public Library of Science (PLOS ONE)
+\subsubsection{Public Library of Science (PLOS ONE)}
+\label{sec:methodsPLOS}
 
@@
 Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
-<!--- Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].
---->
 
@@
 Attribute data collected per DOI included journal volume and subject classification(s).
-<!--- and total article views since publication.--->
 
@@
 Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
-Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
-To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]
+Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings agai
 
-### Australia and New Zealand Clinical Trials Registry (ANZCTR)
+\subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}
+\label{sec:methodsANZCTR}
 
@@
 Studies are not peer reviewed.
-<!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
 
@@
 These variables were chosen as we believed they might influence the completeness of the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.
+
 Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
@@
 
-## Full-text processing
+\subsection{Full-text processing}
+\label{sec:methods-cleaning}
 
@@
 
-## Clustering algorithm
+nst frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
+To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]
+
+\subsection{Clustering algorithm}
 
@@
 
-## Missing statistical methods sections
+\subsection{Missing statistical methods sections}
 
@@
 ```{r}
-load('total.records.plos.rda')
+load('../total.records.plos.rda')
 n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
@@
 ```{r}
-load('wordcount.plos.rda')
+load('../wordcount.plos.rda')
 n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
@@
 
-```{r,fig.width=10,fig.height=12,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
-to_plot = total.records[[2]] %>% select(-percentage) %>% gather(variable,value,-volume) %>%
-  mutate(variable=factor(variable,levels=c('Retrieved from search','Included in analysis')))
+![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/figure1.jpg)
 
-g1 = ggplot(to_plot,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
-  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())
+```{r,fig.height=12,fig.width=6,fig.asp=0.6,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
 
-g2 = ggplot(wordcounts,aes(y=words,x=volume,group=volume))+geom_boxplot()+
-  scale_x_continuous('Volume',breaks=1:14)+scale_y_log10('Word count (log-10 transformed)',breaks=c(0,10,100,200))+g.theme
-
-load('stats_section_subject.classifications.rda')
+load('../stats_section_subject.classifications.rda')
 stats_section_subject = mutate(stats_section_subject,
                                top3 = ifelse(subject_level_number %in% 1:3,1,0))
+
 to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')
-
-to_plot.bar = to_plot %>% arrange(-n) %>% select(-n) %>% gather('Top 3 subject classification',value,-subject_level_name) %>%
-mutate(subject_level_name=factor(subject_level_name,levels=rev(unique(subject_level_name))),
-'Top 3 subject classification'=factor(`Top 3 subject classification`,levels = c('No','Yes')))
-
-g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = c(0.8,0.2),
-legend.direction = 'horizontal',axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
-
-ggarrange(g1,g2,g3,nrow=3,labels=LETTERS[1:3])
 
@@
 
-Search results varied by journal volume (Figure \@ref(fig:plos-n)A).
+Search results varied by journal volume (Figure \ref{fig:plos-n}A).
 The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`), corresponding to years 2013 and 2014.
@@
 
-The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \@ref(fig:plos-n)B).
+The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \ref{fig:plos-n}B).
 `r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
@@
 (n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications
-(Figure \@ref(fig:plos-n)C).
+(Figure \ref{fig:plos-n}C).
 
-```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}
-
-include_graphics("figures/plosone.10topics.jpg")
-
-```
+![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/plosone.10topics.jpg)
 
@@
 ```{r}
-load('plos.results.10topics.rda')
+load('../plos.results.10topics.rda')
 
@@
 Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
-
-___
-**Box 1: Examples of boilerplate text from _PLOS ONE_ (Statistical software)**
-
-Topic 3
-
-* *graphpad prism (graphpad software, san diego, ca) was used for all analyses.*
-* *all statistical analysis was performed using the graphpad prism software.*
-* *statistical analysis were done using graphpad prism software (graphpad software, inc., la jolla, ca).*
-
-Topic 5
-
-* *all statistical analyses were conducted in spss 19 (spss inc., chicago, il, usa).*
-* *all statistical analyses were performed using the spss version 15 for windows (spss inc., chicago, il, usa). a p-value less than 0.05 was considered a significant difference.*
-* *statistical analysis was performed using the spss 17.0 software package (spss, chicago, il, usa). all data were analyzed using t tests, values of p < 0.05 were considered significant.*
-
-___
-
-Definitions of statistical significance featured strongly in Topic~1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
-Topic~1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
-Topic~9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: "\*p<0.05, \*\*p<0.01 and \*\*\*p<0.001", a practice that has been criticised [@Wasserstein2019].
-Examples of boilerplate text are in Box~2.
-
-___
-**Box 2: Examples of boilerplate text from _PLOS ONE_ (Between-group hypothesis testing, two groups)**
-
-Topic 1
-
-* *student’s t-test was used for statistical analysis. a p value of <0.05 was considered statistically significant.*
-* *statistical analysis was performed using student`s t-test. a p-value <0.05 was considered to be statistically significant.*
-* *statistical analysis was performed using a two-tailed student’s t-test, and p-values <0.05 were considered statistically significant.*
-
-Topic 9
-
-* *statistical significance was determined by students t-tests. \* p<0.05, \*\* p = <0.01, \*\*\* p<0.001.*
-* *statistical analyses were performed using the nonparametric mann-whitney test. differences were considered significant for* \**p<0.05,* \*\* *p<0.01, \*\*\* *p<0.001.*
-* *comparisons between two groups were performed by student’s test. statistical significance was defined as \**p<0.05;* \*\* *p<0.01;* \*\*\* *p<0.001.*
-
-___
-
-Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used (Box 3).
-Frequently occurring words in Topic 6 (n = 4,746) including 'mean', 'sd' and 'sem' reflected the use of descriptive statistics for summarising continuous variables.
-Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance.
-One-way analysis of variance also featured strongly in Topic 4 (n = 10,163), with text citing common methods for performing post-hoc multiple comparisons.
-Top matching sections across these topics did not include definitions of groups being compared.
-
-<!-- Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates collected under experimental designs.
-Top matching sections across these three topics did not define the groups being compared.
-[nw: excluded for now as evidence for boilerplate not as strong]-->
-
-Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183).
-
-<!-- Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling.
-Examples of multivariable models were (stepwise) logistic regression and Cox proportional hazards regression.
--->
-
-___
-**Box 3: Examples of boilerplate text from _PLOS ONE_ (Between-group hypothesis testing, two or more groups)**
-
-Topic 4
-
-* *significant differences were determined using analysis-of-variance anova followed by tukey post-hoc tests for multiple comparisons.*
-* *statistical comparisons between study groups were performed using one-way anova test followed by post-hoc multiple comparison with dunnett test.
-p-value of less-than 0.05 were considered to be statistically significant.*
-* *one-way analysis-of-variance anova was used to compare multiple groups followed by bonferroni post-hoc test. statistical significance was set at p less-than 0.05.*
-
-Topic 6
-
-* *all results are expressed as means plus-or-minus standard deviation plus-or-minus sd.*
-* *data are presented as mean plus-or-minus sd or mean plus-or-minus sem. comparison of means was performed by anova test.*
-* *data are expressed as means plus-or-minus standard error of the mean sem. statistical analysis was performed using the student t-test. p less-than 0.05 was considered significant.*
-
-<!-- Topic 10 -->
-
-<!-- * *all statistical analyses were performed using microsoft excel.
-significance was assessed by the student t-test using data from at least three independent experiments.
-all images are representative of at least three independent experiments.* -->
-
-<!-- * *data are presented as the mean plus-or-minus sd of at least three experiments or are representative of experiments repeated at least three times.
-student t-test two-tailed paired was used to evaluate the statistical significance of results from proliferation assay experiments.* -->
-
-<!-- * *data are presented as the mean plus-or-minus sd. comparisons were made using a one-way anova. all experiments were repeated at least three times and representative experiments are shown.* -->
-
-___
-
-## ANZCTR
-
-We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
-Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
-
-The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
-Some methods sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".
-
-The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
-There were cases where the exact same method section had been re-used in a different study.
-
-We also found evidence of 'boilerplate' sections clustered as topic 3, example text
-
-___
-**Box x: Examples of boilerplate text from ANZCTR**
-
-Topic 3
-
-* "Shapiro Wilk test was used as normality test.
-Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
-Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test.
-Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed."
-* "Comparisons between categorical variables will be made either using chi square or Fisher exact test.
-Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
-Two sided p values of less than 0.05 will be considered statistically significant."
-___
-
-We examined if four study characteristics were associated with a missing statistics section.
-The odds ratios and 95% credible intervals are in Table~X.
-Observational studies were less likely to have a missing methods section compared with interventional studies.
-Missing sections became less likely over time.
-Studies with more funders and a larger target sample size were less likely to have a missing methods section.
-
-```{r}
-table =  read.table(header=TRUE, sep='&', text='
-Variable & Odds ratio & 95% CI
-Study type = Observational & 0.78 & 0.69, 0.89
-Date (per year) & 0.90 & 0.88, 0.91
-Number of funders & 0.80 & 0.74, 0.86
-Target sample size (per doubling) & 0.90 & 0.88, 0.92')
-
-ftab = flextable(table) %>%
-theme_box() %>%
-autofit()
-ftab
-```
-
-* Final sample size
-
-# Discussion
-
-The first line in many statistical analysis sections in _PLOS ONE_ was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail.
-As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994].
-This is far from the truth, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.
-
-A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5% level.
-This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [@McShane2019] and the common misinterpretations of p-values [@Goodman2008].
-
-Despite the extensive array of statistical tests available, many authors are reporting the same few methods.
-
-One reason these inadequate sections get published is that most journals do not use statistical reviewers, despite empirical evidence showing they improve manuscript quality [@Hardwicke2020].
-
-A related paper has criticised vague statistical methods sections because they deprive readers and reviewers for the opportunity to confirm that the appropriate methods were used [@Weissgerber2018].
-These authors checked hundreds of papers using ANOVA and found that 95\% did not contain the information needed to determine what type of ANOVA was performed.
-This lack of information could well be because the authors used a boilerplate statistical methods section that was missing key details.
-
-If authors shared their code then this would provide an alternative route for checking what statistical methods were used.
-This is not a perfect solution, as we still want authors to accurately report their methods, but it does increase transparency.
-However, a recent paper found that code sharing was very low in biomedical papers, with just 2\% of a sample of over 6,000 papers sharing code [@Serghiou2021].
-
-Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects.
-This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018].
-This is concerning because using the wrong statistical methods can reduce the value of study, or worse, invalidate the entire study.
-These mistakes are avoidable and are wasting of thousands of hours of researchers' time and the time of patients and volunteers.
-Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].
-
-## Limitations
-
-We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods might be adequate.
-
-We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section.
-Reiterate outcomes of random sample checking here.
-
-We only examined one large journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.
-
-We searched the full text of _PLOS ONE_ papers but not the supporting information which may contain statistical methods sections for some papers.
-The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x%).
-We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.
-
-# References

### 2021-04-20 manuscript/asa_template/asa_template.tex
update to asa template

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

% Pandoc citation processing

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf An observational study of papers published in \emph{PLOS ONE} and
studies posted to a trial registry.}

  \author{
        Nicole White \thanks{The authors gratefully acknowledge \ldots{}} \\
    School of Public Health and Social Work, QUT\\
     and \\     Thiru Balasubramaniam, Richi Nayak \\
    To add\\
     and \\     Adrian Barnett \\
    School of Public Health and Social Work, QUT\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf An observational study of papers published in \emph{PLOS ONE} and
studies posted to a trial registry.}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The text of your abstract. 200 or fewer words.
\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

An ideal statistical analysis will use appropriate methods to create
insights from the data and inform the research questions. Unfortunately
many current statistical analyses are far from ideal, with many
researchers using the wrong methods, misinterpreting the results, or
failing to adequately check their assumptions \citep{2008, Leek2017}.
Some researchers take a ``mechanistic'' approach to statistics, copying
the few methods they know regardless of their appropriateness, and then
going through the motions of the analysis \citep{Stark2018}.

Many researchers lack adequate training in research methods, and
statistics is something they do with trepidation and even ignorance
\citep{Altman1994, King2019}. However, using the wrong statistical
methods can cause real harm \citep{Altman1994, Brown2018} and bad
statistical practices are being to used abet weak science
\citep{Stark2018}. Statistical mistakes are a key source of waste in
research and partly explain the current reproducibility crisis in
science \citep{Allison2016}. Even when the correct methods are used,
many researchers fail to describe them adequately, making it difficult
to reproduce the results \citep{Ernst2017, Zhou2018}. Poor statistical
methods might not be caught by reviewers, as they may not be qualified
to judge the statistics. A recent survey of editors found that only 23\%
of health and medical journals used expert statistical review for all
articles \citep{Hardwicke2020}, which was little different from a survey
from 22 years ago \citep{Goodman1998}.

There is guidance for researchers on how to write up their statistical
methods and results. The International Committee of Medical Journal
Editors recommend that researchers should: ``Describe statistical
methods with enough detail to enable a knowledgeable reader with access
to the original data to judge its appropriateness for the study and to
verify the reported results'' \citep{ICMJE2019}. More detailed guideance
is given by the SAMPL and EQUATOR guidelines
\citep{Lang2013, Altman2016} with the latter covering all apsects of the
paper. Both of these guidelines were led by Doug Altman, who spoke often
and for many years about the need for better statistical reporting. The
awareness and use of these guidelines could be improved. There were 256
Google Scholar citations to the SAMPL paper (as at 15 March 2021) which
is a good citation statistic for most papers, but is low considering the
millions of papers that use statistical analysis.

Two statisticians on this paper (AB and NW) have heard researchers admit
that they have copied-and-pasted their statistical methods sections from
other papers, regardless of whether they are appropriate. The aim of
this paper is to use text-mining methods to estimate the extent that
researchers are using cut-and-paste or `boilerplate' statistical methods
sections. Boilerplate text is that ``which can be reused in new contexts
or applications without significant changes to the original''
\citep{Wikipedia}. Use of these methods sections indicates that little
thought has gone into the statistical analysis.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\subsection{Data sources}

We used two openly available data sources to find statistical methods
sections, covering peer-reviewed journal articles and registered
clinical trial protocols.

\subsubsection{Public Library of Science (PLOS ONE)}
\label{sec:methodsPLOS}

\emph{PLOS ONE} is a large open access journal that publishes original
research across a wide range of scientific fields. Articles must be in
English. Article submissions are handled by an academic editor who
selects peer reviewers based on their self-nominated areas of expertise.
Submissions do not undergo formal statistical review. Instead, reviewers
are required to assess submissions against several publication criteria,
including whether: ``Experiments, statistics, and other analyses are
performed to a high technical standard and are described in sufficient
detail'' \citep{PLOS}. All reviewers are asked the question: ``Has the
statistical analysis been performed appropriately and rigorously?'',
with the possible responses of ``Yes'', ``No'' and ``I don't know''.

Authors are encouraged to follow published reporting guidelines such as
EQUATOR, to ensure that chosen statistical methods are appropriate for
the study design, and adequate details are provided to enable
independent replication of results.

All \emph{PLOS ONE} articles are freely accessible via the PLOS
Application Programming Interface (API). This enabled us to conduct
semi-automated searches of full-text articles and analyse data on
individual records, including text content and general attributes such
as publication date and field(s) of research. To find papers with a
statistical methods section we used targeted API searches followed by
article filtering based on section headings. The data were downloaded on
3 July 2020.

\emph{Step 1}: Targeted API searches. API searches were completed using
the R package `rplos' \citep{rplos}. Search queries targeted the
presence of analysis-related terms anywhere in the article. Search terms
combined the words ``data'' or ``statistical'' with one of:
``analysis'', ``analyses'', ``method'', ``methodology'' or
``model(l)ing''. Search terms were intended to be broad whilst keeping
search results to a manageable number for full-text review (see Step 2).
By allowing terms to appear anywhere in the article, we accounted for
the possibility of relevant text being placed in different sections, for
example, in the \emph{Material and Methods} section versus
\emph{Results}. Search results were indexed by a unique Digital Object
Identifier (DOI). Attribute data collected per DOI included journal
volume and subject classification(s).

\emph{Step 2}: Partial matching on section headings. Full text XML data
for all search results were downloaded and combined into a single
dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
ONE} does not prescribe standardised headings to preface statistical
methods sections, we performed partial matching on available headings
agai

\subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}
\label{sec:methodsANZCTR}

The ANZCTR was established in 2005 as part of a coordinated global
effort to improve research quality and transparency in clinical trials
reporting; observational studies can also be registered. All studies
registered on ANZCTR are publicly available and can be searched via an
online portal (\url{https://www.anzctr.org.au}). Details required for
registration follow a standardised template \citep{ANZCTR}, which covers
participant eligibility, the intervention(s) being evaluated, study
design and outcomes. The information provided must be in English.
Studies are not peer reviewed.

For the statistical methods section, researchers are asked to provide a
``brief description'' of the sample size calculations, statistical
methods and planned analyses, although this section is not compulsory
\citep{ANZCTR}. Studies are reviewed by ANZCTR staff for completeness of
key information, which does not include the completeness of the
statistical methods sections.

All studies available on ANZCTR were downloaded on 1 February 2020 in
XML format. We used all the text available in the ``Statistical
methods'' section. We also collated basic information about the study
including the study type (interventional or observational), submission
date, number of funders and target sample size. These variables were
chosen as we believed they might influence the completeness of the
statistical methods section, because we expected larger studies and
those with funding to be more complete, and we also were interested in
changes over time.

Studies prior to 2013 were excluded as the statistical methods section
appeared to be introduced in 2013. Some studies were first registered on
the alternative trial database \emph{clinicaltrials.gov} and then also
posted to ANZCTR. We excluded these studies because they almost all had
no completed statistical methods section as this section is not included
in \emph{clinicaltrials.gov}.

\subsection{Full-text processing}
\label{sec:methods-cleaning}

We applied the same text cleaning to both data sources. Text cleaning
aimed to standardise notation and statistical terminology, whilst
minimising changes to article style and formatting. \emph{R} code used
for data extraction and cleaning is available from
\url{https://github.com/agbarnett/stats_section}.

Mathematical notation including Greek letters was converted from Unicode
characters to plain text. For example, the Unicode characters
corresponding to \(\theta\) (\textless U+03B8\textgreater) were replaced
with `theta'. Similarly, common symbols outside of Unicode blocks
including `\%' (percent) and `\textless{}' (`less-than') were converted
into plain text using the `textclean' package \citep{textclean}. General
formatting was removed, this included carriage returns, punctuation
marks, in-text references (e.g.~``{[}42{]}'') centred equations, and
other non-ASCII characters. Text contained inside brackets was retained
to maximise content for analysis, with brackets removed.

We compiled an extensive list of statistical terms to standardise
descriptions of statistical methods reported across both datasets. An
initial list was compiled by calculating individual word frequencies and
identifying relevant terms that appeared at least 100 times. Further
terms were sourced from index searches of three statistics textbooks
\citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. The final list
is provided as Supplementary Material. Plurals (e.g., `chi-squares')
unhyphenated (e.g., `chi square') and combined (e.g.~`chisquare') terms
were transformed to singular, hyphenated form (e.g., `chi-square').
Common statistical tests were also hyphenated (e.g., `hosmer lemeshow'
to `hosmer-lemeshow').

As a final step, common stop words including pronouns, contractions and
selected prepositions were removed. We retained selected stop words
that, if excluded, may have changed the context of statistical methods
being described, for example `between' and `against'.

nst frequently used terms in initial search results: `Statistical
analysis', `Statistical analyses', `Statistical method', `Statistics',
`Data analysis' and `Data analyses'. To determine the reliability of our
chosen filters, we manually reviewed full text data extracted for a
random sample of XXX articles that were not matched (File
S1).{[}TODO\ldots finish this thought\ldots{]}

\subsection{Clustering algorithm}

\emph{Details to come}

Results were transformed to lower case for the clustering, but examples
are given using the original capitalisation.

\subsection{Missing statistical methods sections}

The statistical methods section for the ANZCTR data was missing for some
studies and we examined if there were particular studies where this
section was more likely to be missing. We used four independent
variables of date, study type (observational or interventional), number
of funders and target sample size. We used a logistic regression model
fitted using a Bayesian paradigm. A small number of sections were
labelled as ``Not applicable'', ``Nil'' or ``None'' and we changed these
to missing.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{plos-one}{%
\subsection{\texorpdfstring{\emph{PLOS ONE}}{PLOS ONE}}\label{plos-one}}

\begin{figure}
\centering
\includegraphics{figures/figure1.jpg}
\caption{\label{fig:plos-n}A: Search results by \emph{PLOS ONE} volume;
B: word count per statistical methods section included in analysis (n =
111,73); C: subject classifications assigned to full-text records
included in analysis}
\end{figure}

API searches returned 131,847 unique records, of which 111,731 (85\%)
included a statistical methods section based on our search criteria. In
the final sample, 95,518 (85\%) DOIs returned an exact match against
common section headings, including 64,133 for `statistical analysis',
13,380 for `statistical analyses' and 13,627 for `data analysis'. Among
DOIs that did not meet the partial matching criteria, initial search
terms appeared in {[}TODO{]}.

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
total number of API search results peaked at volumes 8 (n = 19,045) and
9 (n = 19,045), corresponding to years 2013 and 2014. This trend aligned
with the total number of papers published in \emph{PLOS ONE} over the
same period. The percentage of records that included a statistical
methods section by volume based on our proposed matching criteria varied
between 64\% (volume 2) and 86\% (volume 9).

The median length of statistical methods sections was 127 words (IQR: 61
to 254 words) (Figure \ref{fig:plos-n}B). 7,450 articles (7\%) had a
statistical methods section of 500 words or more. 19,461 articles (17\%)
had a statistical methods section of 50 words or less, equal to the
length of this paragraph.

All papers included Biology and life sciences (n = 107,584), Earth
sciences (n = 7,605) and/or Computer and information sciences (n =
5,190) in their top 3 subject classifications (Figure
\ref{fig:plos-n}C).

\begin{figure}
\centering
\includegraphics{figures/plosone.10topics.jpg}
\caption{\label{fig:plot-10-topics}Topic clouds summarising common terms
in \emph{PLOS ONE} statistical methods sections assuming ten clusters}
\end{figure}

We applied the clustering algorithm to the cleaned dataset, varying the
number of clusters from 1 to 50. Increasing the number of clusters
decreased cluster quality based on global goodness-of-fit measures
(Supplementary Figure 1), with average silhouette score and
within-cluster dispersion leveling off around 20 clusters. This
indicated that the data comprised one large, heterogeneous cluster and
multiple smaller clusters.

\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-4-1.png}

The topic clouds based on ten clusters are in
Figure\textasciitilde@ref(fig:plot-10-topics). Frequently occurring
words reflected the use of statistical software (Topics\textasciitilde3
and 5), descriptive statistics (Topic\textasciitilde6), group based
hypothesis testing (Topics 1 and 4) and definitions of statistical
significance (Topics\textasciitilde1 and 9). There are also statistical
methods sections associated with regression (Topic\textasciitilde2) and
meta-analysis (Topic\textasciitilde7).

Topics related to the use of statistical software differentiated between
Prism GraphPad (Topic\textasciitilde3: n = 9,879; 8.8\%) and SPSS (Topic
5: n = 9,574; 8.6\%) (Box 1). A manual review of the top matching
sections in these topics showed strong evidence of boilerplate text.
Nine out ten top matches for Topic 3 stated the use of Prism GraphPad
but did not specify which statistical methods were used; six out of ten
top matches returned the same cluster score indicating near identical
text. Top matching sections for Topic 5 included information on SPSS
version numbers and definitions of statistical significance.

\bibliographystyle{agsm}
\bibliography{references.bib}

\end{document}

### 2021-04-20 manuscript/PLOSresults.R

@@ -1,0 +1,41 @@
+library(dplyr)
+library(tidyr)
+library(pander)
+library(ggplot2)
+library(ggpubr)
+library(flextable)
+library(tidytext)
+load('manuscript/total.records.plos.rda')
+load('manuscript/wordcount.plos.rda')
+
+to_plot = total.records[[2]] %>% select(-percentage) %>% 
+  rename('Retrieved from search'=API) %>%
+  gather(variable,value,-volume) %>%
+  mutate(variable=factor(variable,levels=c('Retrieved from search','Included in analysis')))
+
+g1 = ggplot(to_plot,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
+  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())
+
+g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
+  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count\n(log10 transformed)',breaks=seq(0,4,0.5)) + g.theme
+
+load('manuscript/stats_section_subject.classifications.rda')
+stats_section_subject = mutate(stats_section_subject,
+                               top3 = ifelse(subject_level_number %in% 1:3,1,0))
+
+to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')
+
+to_plot.bar = to_plot %>% arrange(-n) %>% select(-n) %>% gather('Top 3 subject classification',value,-subject_level_name) %>%
+  mutate(subject_level_name=factor(subject_level_name,levels=rev(unique(subject_level_name))),
+         'Top 3 subject classification'=factor(`Top 3 subject classification`,levels = c('No','Yes')))
+
+
+g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+
+  geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+
+  xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = c(0.55,0.2),
[…]

### 2021-04-20 manuscript/asa_template/asa_template.Rmd
update tex file

@@
 
-![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/plosone.10topics.jpg)
+![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/figure2.jpg)
 

### 2021-04-20 manuscript/asa_template/asa_template.Rmd
tidy up folder structure

@@
 There were 256 Google Scholar citations to the SAMPL paper (as at 15 March 2021) which is a good citation statistic for most papers, but is low considering the millions of papers that use statistical analysis.
-<!--- Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked. --->
 
@@
 All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au).
-<!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
+
 Details required for registration follow a standardised template [@ANZCTR], which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes.
@@
 
-![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/figure1.jpg)
+![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/plos.summary.jpg)
 
@@
 
-![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/figure2.jpg)
+![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/plos.wordclouds.jpg)
 
@@
 
-The topic clouds based on ten clusters are in Figure~\@ref(fig:plot-10-topics).
+The topic clouds based on ten clusters are in Figure~\ref{fig:plot-10-topics}.
 Frequently occurring words reflected the use of statistical software (Topics~3 and 5), descriptive statistics (Topic~6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9).

### 2021-04-20 manuscript/asa_template/asa_template.tex
update tex file
tidy up folder structure

@@
 registered on ANZCTR are publicly available and can be searched via an
-online portal (\url{https://www.anzctr.org.au}). Details required for
-registration follow a standardised template \citep{ANZCTR}, which covers
-participant eligibility, the intervention(s) being evaluated, study
-design and outcomes. The information provided must be in English.
-Studies are not peer reviewed.
+online portal (\url{https://www.anzctr.org.au}).
+
+Details required for registration follow a standardised template
+\citep{ANZCTR}, which covers participant eligibility, the
+intervention(s) being evaluated, study design and outcomes. The
+information provided must be in English. Studies are not peer reviewed.
 
@@
 \centering
-\includegraphics{figures/plosone.10topics.jpg}
+\includegraphics{figures/figure2.jpg}
 \caption{\label{fig:plot-10-topics}Topic clouds summarising common terms
@@
 The topic clouds based on ten clusters are in
-Figure\textasciitilde@ref(fig:plot-10-topics). Frequently occurring
+Figure\textasciitilde{}\ref{fig:plot-10-topics}. Frequently occurring
 words reflected the use of statistical software (Topics\textasciitilde3

### 2021-04-20 code/8_manuscript.figures.plos.R

@@ -1,0 +1,45 @@
+library(dplyr)
+library(tidyr)
+library(pander)
+library(ggplot2)
+library(ggpubr)
+library(flextable)
+library(tidytext)
+load('../results/total.records.plos.rda')
+load('../results/wordcount.plos.rda')
+load('../results/stats_section_subject.classifications.rda')
+
+g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',
+                           axis.text = element_text(size=12),axis.text.x = element_text(size=10),axis.text.y = element_text(size=10))
+
+plot.records = total.records[[2]] %>% select(-percentage) %>% 
+  rename('Retrieved from search'=API) %>%
+  gather(variable,value,-volume) %>%
+  mutate(variable=factor(variable,levels=c('Retrieved from search','Included in analysis')))
+
+g1 = ggplot(plot.records,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
+  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())
+
+g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
+  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count\n(log10 transformed)',breaks=seq(0,4,0.5)) + g.theme
+
+
+stats_section_subject = mutate(stats_section_subject,
+                               top3 = ifelse(subject_level_number %in% 1:3,1,0))
+
+plot.subjects = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')
+
+plot.subjects = plot.subjects %>% arrange(-n) %>% select(-n) %>% gather('Top 3 subject classification',value,-subject_level_name) %>%
+  mutate(subject_level_name=factor(subject_level_name,levels=rev(unique(subject_level_name))),
+         'Top 3 subject classification'=factor(`Top 3 subject classification`,levels = c('No','Yes')))
+
+
+g3 = ggplot(plot.subjects,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+
[…]

### 2021-04-20 code/8_manuscript.figures.plos.R
separate script for mansucript figures

@@ -40,6 +40,6 @@ g3 = ggplot(plot.subjects,aes(x=subject_level_name,y=value/10000,fill=`Top 3 sub
                                                                                                                                                                                                                                                                                                                       legend.direction = 'horizontal')
 
 
-jpeg('../manuscript/asa_template/figures/plots.summary.jpg',width=480,height=600,quality=100)
+jpeg('../manuscript/asa_template/figures/plos.summary.jpg',width=480,height=600,quality=100)
 ggarrange(g1,g2,g3,nrow=3,labels=LETTERS[1:3])
 dev.off()

### 2021-04-21 manuscript/asa_template/asa_template.Rmd
update figures and tex files

@@
 
+Results were examined for evidence of boilerplate text in two ways.
+We first reviewed the top ten results that represented the strongest matches to each cluster.
+To compare results within a cluster, we calculated cosine similarity score for all pairs of records; higher scores denoted a higher degree of similarity in text between a pair of records.
+
 \subsection{Missing statistical methods sections}
@@
 ```{r}
-load('../total.records.plos.rda')
+load('../../results/total.records.plos.rda')
 n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
@@
 ```{r}
-load('../wordcount.plos.rda')
+load('../../results/wordcount.plos.rda')
 n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
@@
 
-![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/plos.summary.jpg)
+```{r, echo=FALSE, fig.align='center', out.width="95%", fig.cap="PLOS ONE result summary."}
+knitr::include_graphics("figures/plos.summary.jpg")
+```
 
-```{r,fig.height=12,fig.width=6,fig.asp=0.6,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
+<!-- ![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/plos.summary.jpg) -->
 
-load('../stats_section_subject.classifications.rda')
+```{r,fig.height=12,fig.width=6,fig.asp=0.55,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
+
+load('../../results/stats_section_subject.classifications.rda')
 stats_section_subject = mutate(stats_section_subject,
@@
 
-![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/plos.wordclouds.jpg)
+```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="PLOS ONE topic clouds."}
+knitr::include_graphics("figures/plos.wordclouds.jpg")
+```
 
@@
 ```{r}
-load('../plos.results.10topics.rda')
+load('../../results/plos.results.10topics.rda')
 
@@
 Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
+
+\subsection{ANZCTR}
+
+```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR topic clouds."}
+knitr::include_graphics("figures/anzctr.wordclouds.jpg")
+```
+
+\section{Supplementary}
+
+\subsection{PLOS}
+
+```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE clustering metrics."}
+knitr::include_graphics("figures/supplementary/supp_plosone.50clusters.jpg")
+```
+
+```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE topic values."}
+knitr::include_graphics("figures/supplementary/plos.topicvalues.jpg")
+```
+
+\subsection{ANZCTR}
+
+```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR clustering metrics."}
+knitr::include_graphics("figures/supplementary/supp_anzctr.50clusters.jpg")
+```
+
+```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR topic values."}
+knitr::include_graphics("figures/supplementary/anzctr.topicvalues.jpg")
+```

### 2021-04-21 manuscript/asa_template/asa_template.tex
update figures and tex files

@@
 
+Results were examined for evidence of boilerplate text in two ways. We
+first reviewed the top ten results that represented the strongest
+matches to each cluster. To compare results within a cluster, we
+calculated cosine similarity score for all pairs of records; higher
+scores denoted a higher degree of similarity in text between a pair of
+records.
+
 \subsection{Missing statistical methods sections}
@@
 \begin{figure}
-\centering
-\includegraphics{figures/figure1.jpg}
-\caption{\label{fig:plos-n}A: Search results by \emph{PLOS ONE} volume;
-B: word count per statistical methods section included in analysis (n =
-111,73); C: subject classifications assigned to full-text records
-included in analysis}
+
+{\centering \includegraphics[width=0.95\linewidth]{figures/plos.summary}
+
+}
+
+\caption{PLOS ONE result summary.}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{figures/figure2.jpg}
-\caption{\label{fig:plot-10-topics}Topic clouds summarising common terms
-in \emph{PLOS ONE} statistical methods sections assuming ten clusters}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/plos.wordclouds}
+
+}
+
+\caption{PLOS ONE topic clouds.}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-4-1.png}
+\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}
 
@@
 
+\subsection{ANZCTR}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/anzctr.wordclouds}
+
+}
+
+\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+\section{Supplementary}
+
+\subsection{PLOS}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/supplementary/supp_plosone.50clusters}
+
+}
+
+\caption{PLOS ONE clustering metrics.}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/supplementary/plos.topicvalues}
+
+}
+
+\caption{PLOS ONE topic values.}\label{fig:unnamed-chunk-9}
+\end{figure}
+
+\subsection{ANZCTR}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/supplementary/supp_anzctr.50clusters}
+
+}
+
+\caption{ANZCTR clustering metrics.}\label{fig:unnamed-chunk-10}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/supplementary/anzctr.topicvalues}
+
+}
+
+\caption{ANZCTR topic values.}\label{fig:unnamed-chunk-11}
+\end{figure}
+
 \bibliographystyle{agsm}

### 2021-04-21 code/7_summarise_stats_sections.R

@@ -2,14 +2,14 @@
 library(tidyverse)
 library(tidytext)
 library(openxlsx)
-dat = readRDS('data/stats_section_cleaned.rds')
+
+dat = readRDS('../data/stats_section_cleaned.rds')
 #ten topics, PLOS ONE
-matches = read.csv('manuscript/plos_one_10topics.csv',header=T)
+matches = read.csv('../results/plos_one_10topics.csv',header=T)
 matches = left_join(dat,matches,by=c('doi'='DOI')) %>%
   mutate(value = as.numeric(value)) %>%
   distinct(doi,text_heading,.keep_all = T)
 
-#examples of boilerplate text: topic 3
  tidy_matches = matches %>%
    unnest_tokens(word,text_data_clean)
  

### 2021-04-21 code/8_manuscript.figures.plos.R

@@ -9,8 +9,11 @@ load('../results/total.records.plos.rda')
 load('../results/wordcount.plos.rda')
 load('../results/stats_section_subject.classifications.rda')
 
-g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',
-                           axis.text = element_text(size=12),axis.text.x = element_text(size=10),axis.text.y = element_text(size=10))
+#FIGURE: summary of search results: number found, word count, subject classifications
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',legend.text = element_text(size=10),
+                           axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
 
 plot.records = total.records[[2]] %>% select(-percentage) %>% 
   rename('Retrieved from search'=API) %>%
@@ -18,15 +21,13 @@ plot.records = total.records[[2]] %>% select(-percentage) %>%
   mutate(variable=factor(variable,levels=c('Retrieved from search','Included in analysis')))
 
 g1 = ggplot(plot.records,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
-  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())
+  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+theme(legend.title = element_blank(),legend.position = 'top',legend.direction = 'horizontal')
 
 g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
-  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count\n(log10 transformed)',breaks=seq(0,4,0.5)) + g.theme
-
+  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count\n(log10 transformed)',breaks=seq(0,4,0.5))
 
 stats_section_subject = mutate(stats_section_subject,
                                top3 = ifelse(subject_level_number %in% 1:3,1,0))
-
[…]

### 2021-04-21 manuscript/asa_template/asa_template.Rmd
add plos boilerplate text

@@
 
+* Results of cosine similaritites
+
+```{r}
+load('../../results/plos.cosinesim.10topics.rda')
+
+ftab = flextable(ftab.cosine.plos) %>% theme_box() %>% autofit()
+ftab <- set_caption(ftab, "Summary of cosine similarity scores by topic: PLOS ONE")
+ftab
+
+```
+
 \subsection{ANZCTR}
@@
 ```
+
+We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
+Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
+
+The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+Some methods sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".
+
+The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
+There were cases where the exact same method section had been re-used in a different study.
+
+We also found evidence of 'boilerplate' sections clustered as topic 3, example text
+
+**Box x: Examples of boilerplate text from ANZCTR**
+
+Topic 3
+
+* "Shapiro Wilk test was used as normality test.
+Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
+Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test.
+Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed."
+* "Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant."
+
+We examined if four study characteristics were associated with a missing statistics section.
+The odds ratios and 95% credible intervals are in Table~X.
+Observational studies were less likely to have a missing methods section compared with interventional studies.
+Missing sections became less likely over time.
+Studies with more funders and a larger target sample size were less likely to have a missing methods section.
+
+```{r}
+table =  read.table(header=TRUE, sep='&', text='
+Variable & Odds ratio & 95% CI
+Study type = Observational & 0.78 & 0.69, 0.89
+Date (per year) & 0.90 & 0.88, 0.91
+Number of funders & 0.80 & 0.74, 0.86
+Target sample size (per doubling) & 0.90 & 0.88, 0.92')
+
+ftab = flextable(table) %>%
+theme_box() %>%
+autofit()
+ftab
+```
+
+* Final sample size
+
+# Discussion
+
+The first line in many statistical analysis sections in _PLOS ONE_ was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail.
+As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994].
+This is far from the truth, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.
+
+A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5% level.
+This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [@McShane2019] and the common misinterpretations of p-values [@Goodman2008].
+
+Despite the extensive array of statistical tests available, many authors are reporting the same few methods.
+
+One reason these inadequate sections get published is that most journals do not use statistical reviewers, despite empirical evidence showing they improve manuscript quality [@Hardwicke2020].
+
+A related paper has criticised vague statistical methods sections because they deprive readers and reviewers for the opportunity to confirm that the appropriate methods were used [@Weissgerber2018].
+These authors checked hundreds of papers using ANOVA and found that 95\% did not contain the information needed to determine what type of ANOVA was performed.
+This lack of information could well be because the authors used a boilerplate statistical methods section that was missing key details.
+
+If authors shared their code then this would provide an alternative route for checking what statistical methods were used.
+This is not a perfect solution, as we still want authors to accurately report their methods, but it does increase transparency.
+However, a recent paper found that code sharing was very low in biomedical papers, with just 2\% of a sample of over 6,000 papers sharing code [@Serghiou2021].
+
+Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects.
+This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018].
+This is concerning because using the wrong statistical methods can reduce the value of study, or worse, invalidate the entire study.
+These mistakes are avoidable and are wasting of thousands of hours of researchers' time and the time of patients and volunteers.
+Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].
+
+## Limitations
+
+We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods might be adequate.
+
+We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section.
+Reiterate outcomes of random sample checking here.
+
+We only examined one large journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.
+
+We searched the full text of _PLOS ONE_ papers but not the supporting information which may contain statistical methods sections for some papers.
+The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x%).
+We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.
 

### 2021-04-21 manuscript/asa_template/asa_template.tex
add plos boilerplate text

@@
 
+\begin{itemize}
+\tightlist
+\item
+  Results of cosine similaritites
+\end{itemize}
+
+\includegraphics[width=6.13in,height=3.49in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-7-1.png}
+
 \subsection{ANZCTR}
@@
 
-\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-7}
+\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-8}
 \end{figure}
+
+We downloaded 28,008 studies. The numbers of excluded studies are shown
+in Figure\textasciitilde X. Of the 12,700 included studies, 9,523 (75\%)
+had a statistical methods section.
+
+The median length of the section was 129 words with an inter-quartile
+range of 71 to 219 words. Some methods sections were only one word,
+including ``ANOVA'', ``t-test'', ``SPSS'' and even ``SSPS''.
+
+The clustering algorithm found groups that were purely sample size
+calculations (topic 2), pilot studies (topic 5), safety/tolerability
+studies (topic 6) and repeated measures ANOVA (topic 10). There were
+cases where the exact same method section had been re-used in a
+different study.
+
+We also found evidence of `boilerplate' sections clustered as topic 3,
+example text
+
+\textbf{Box x: Examples of boilerplate text from ANZCTR}
+
+Topic 3
+
+\begin{itemize}
+\tightlist
+\item
+  ``Shapiro Wilk test was used as normality test. Continuous variables
+  were compared using Student t-test and Mann-Whitney U test when the
+  data were not normally distributed. Categorical variables were
+  compared using Pearson's chi-squared test and Fisher's exact test.
+  Paired data were analyzed using Paired t-test and Wilcoxon signed rank
+  test when data were not normally distributed.''
+\item
+  ``Comparisons between categorical variables will be made either using
+  chi square or Fisher exact test. Continuous data will be compared
+  using the Student's t-test or Mann-Whitney U test. Two sided p values
+  of less than 0.05 will be considered statistically significant.''
+\end{itemize}
+
+We examined if four study characteristics were associated with a missing
+statistics section. The odds ratios and 95\% credible intervals are in
+Table\textasciitilde X. Observational studies were less likely to have a
+missing methods section compared with interventional studies. Missing
+sections became less likely over time. Studies with more funders and a
+larger target sample size were less likely to have a missing methods
+section.
+
+\includegraphics[width=4.66in,height=1.55in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-9-1.png}
+
+\begin{itemize}
+\tightlist
+\item
+  Final sample size
+\end{itemize}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+The first line in many statistical analysis sections in \emph{PLOS ONE}
+was the software used and some entire sections in ANZCTR only stated the
+software, implying that the software is the most important detail. As
+Doug Altman said, ``Many people think that all you need to do statistics
+is a computer and appropriate software'' \citep{Altman1994}. This is far
+from the truth, and whilst it is important for researchers to mention
+the software and version used for reproducibility purposes, it is a
+minor detail compared with detailing what methods were used and why.
+
+A frequent theme in the boilerplate statistical methods is the
+definition of statistical significance, nearly always using a p-value at
+the 5\% level. This widespread use of statistical significance is
+troubling giving the bright-line thinking it engenders
+\citep{McShane2019} and the common misinterpretations of p-values
+\citep{Goodman2008}.
+
+Despite the extensive array of statistical tests available, many authors
+are reporting the same few methods.
+
+One reason these inadequate sections get published is that most journals
+do not use statistical reviewers, despite empirical evidence showing
+they improve manuscript quality \citep{Hardwicke2020}.
+
+A related paper has criticised vague statistical methods sections
+because they deprive readers and reviewers for the opportunity to
+confirm that the appropriate methods were used \citep{Weissgerber2018}.
+These authors checked hundreds of papers using ANOVA and found that 95\%
+did not contain the information needed to determine what type of ANOVA
+was performed. This lack of information could well be because the
+authors used a boilerplate statistical methods section that was missing
+key details.
+
+If authors shared their code then this would provide an alternative
+route for checking what statistical methods were used. This is not a
+perfect solution, as we still want authors to accurately report their
+methods, but it does increase transparency. However, a recent paper
+found that code sharing was very low in biomedical papers, with just 2\%
+of a sample of over 6,000 papers sharing code \citep{Serghiou2021}.
+
+Many researchers are using lazy practice by copying a standard
+``boilerplate'' statistical methods section, likely cut-and-pasting from
+other researchers or projects. This is a strong sign of the ritualistic
+practice of statistics where researchers go through the motions rather
+than using conscientious practice \citep{Stark2018}. This is concerning
+because using the wrong statistical methods can reduce the value of
+study, or worse, invalidate the entire study. These mistakes are
+avoidable and are wasting of thousands of hours of researchers' time and
+the time of patients and volunteers. Poor statistical practice is a key
+driver of the ongoing reproducibility crisis in science
+\citep{Ioannidis2014}.
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+We did not check whether papers used the correct methods, and for some
+simple studies a `boilerplate' statistical methods might be adequate.
+
+We examined papers where there was a statistics section, and we missed
+papers that used statistical analysis but did not include a statistical
+analysis section. Reiterate outcomes of random sample checking here.
+
+We only examined one large journal and one trial registry and hence our
+results may not be generalisable to all journals or registries,
+especially those that consistently use a statistical reviewer.
+
+We searched the full text of \emph{PLOS ONE} papers but not the
+supporting information which may contain statistical methods sections
+for some papers. The search terms we used to find statistical methods
+appeared in the supporting information titles for xxx papers (x\%). We
+did not include the supporting information because it is less structured
+than the paper and could be in PDF or Word format.
 
@@
 
-\caption{PLOS ONE clustering metrics.}\label{fig:unnamed-chunk-8}
+\caption{PLOS ONE clustering metrics.}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 
-\caption{PLOS ONE topic values.}\label{fig:unnamed-chunk-9}
+\caption{PLOS ONE topic values.}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 
-\caption{ANZCTR clustering metrics.}\label{fig:unnamed-chunk-10}
+\caption{ANZCTR clustering metrics.}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
-\caption{ANZCTR topic values.}\label{fig:unnamed-chunk-11}
+\caption{ANZCTR topic values.}\label{fig:unnamed-chunk-13}
 \end{figure}

### 2021-04-21 code/7_summarise_stats_sections.R

@@ -6,11 +6,12 @@ library(openxlsx)
 dat = readRDS('../data/stats_section_cleaned.rds')
 #ten topics, PLOS ONE
 matches = read.csv('../results/plos_one_10topics.csv',header=T)
+
 matches = left_join(dat,matches,by=c('doi'='DOI')) %>%
   mutate(value = as.numeric(value)) %>%
   distinct(doi,text_heading,.keep_all = T)
 
- tidy_matches = matches %>%
+tidy_matches = matches %>%
    unnest_tokens(word,text_data_clean)
  
 wordcounts <- tidy_matches %>%
@@ -23,21 +24,6 @@ matches = left_join(matches,wordcounts,by=c('doi','topic_id')) %>%
   mutate(topic_id = paste('Topic',topic_id))
 
 
-save(matches,file='manuscript/plos.results.10topics.rda')
-
-
-#example - topics 3 and 5
-matches %>% filter(topic_id %in% c('Topic 1', 'Topic 3','Topic 5')) %>% ggplot(.,aes(x=rank,y=words)) + 
-  geom_point(alpha=0.1)+ geom_smooth(method='loess',se=T,colour='blue')+ facet_wrap(~topic_id,scales = 'free') + 
-  scale_x_continuous('Rank (1 = strongest topic match)')+scale_y_continuous('Word count (cleaned text)',breaks=seq(0,1500,100))+theme_minimal()
-
-
-#example - topics 1 and 9
-matches %>% filter(topic_id %in% c('Topic 1','Topic 9')) %>% ggplot(.,aes(x=rank,y=words)) + 
-  geom_point(alpha=0.1)+ geom_smooth(method='loess',se=T,colour='blue')+ facet_wrap(~topic_id,scales='free') + 
-  scale_x_continuous('Rank (1 = strongest topic match)')+scale_y_continuous('Word count (cleaned text)')+theme_minimal()
-
-
 cos.sim <- function(ix,distances.mat) 
 {
   A = distances.mat[,ix[1]]
@@ -45,7 +31,7 @@ cos.sim <- function(ix,distances.mat)
   return( sum(A*B)/sqrt(sum(A^2)*sum(B^2)) )
 } 
 
-distances = tidy_matches %>% left_join(.,matches %>% select(doi,rank),by='doi') %>% filter(rank<=100)
[…]

### 2021-04-21 manuscript/asa_template/asa_template.Rmd
update r scripts for boilerplate text

@@
 ftab = flextable(ftab.cosine.plos) %>% theme_box() %>% autofit()
-ftab <- set_caption(ftab, "Summary of cosine similarity scores by topic: PLOS ONE")
+ftab <- set_caption(ftab, "Summary of cosine similarity scores among top 500 DOIs assigned to each topic: PLOS ONE")
 ftab

### 2021-04-21 code/7_summarise_stats_sections.R

@@ -80,3 +80,6 @@ boilerplate.text.plos = boilerplate.dois %>% left_join(matches %>% select(doi,te
 save(matches,file='../results/plos.results.10topics.rda')
 save(stats_section.sim,ftab.cosine.plos,boilerplate.text.plos,file='../results/plos.cosinesim.10topics.rda')
 
+#save boilerplate text as separate excel file
+write.xlsx(boilerplate.text.plos,file='../results/plos.boilerplate.xlsx')
+

### 2021-04-21 code/7_summarise_stats_sections_anzctr.R
add boilerplate examples for anzctr

@@ -1,0 +1,86 @@
+#7_summarise_stats_sections.R
+library(tidyverse)
+library(tidytext)
+library(openxlsx)
+
+dat = readRDS('../data/stats_section_anzctr_cleaned.rds')
+#ten topics, PLOS ONE
+matches = read.csv('../results/anzctr_10topics.csv',header=T)
+
+matches = left_join(dat,matches,by=c('number'='DOI')) %>%
+  mutate(value = as.numeric(value)) %>%
+  distinct(number,.keep_all = T)
+
+tidy_matches = matches %>%
+  unnest_tokens(word,text_data_clean)
+
+wordcounts <- tidy_matches %>%
+  group_by(topic_id,number) %>%
+  summarise(words = n(),.groups='drop')
+
+matches = left_join(matches,wordcounts,by=c('number','topic_id')) %>%
+  arrange(topic_id,-value) %>%
+  group_by(topic_id) %>% mutate(rank=row_number()) %>% ungroup() %>%
+  mutate(topic_id = paste('Topic',topic_id))
+
+
+cos.sim <- function(ix,distances.mat) 
+{
+  A = distances.mat[,ix[1]]
+  B = distances.mat[,ix[2]]
+  return( sum(A*B)/sqrt(sum(A^2)*sum(B^2)) )
+} 
+
+distances = tidy_matches %>% left_join(.,matches %>% select(number,rank),by='number') %>% filter(rank<=500) %>% arrange(rank)
+
+calc_dist_mat <- function(indata=distances,topicNumber){
+  to_stat = indata %>% filter(topic_id==topicNumber) %>%
+    distinct(number,word,.keep_all=F) %>%
+    group_by(number) %>% count(word) %>%
+    spread(number,n,fill=0) 
+  
+  distances.mat = as.matrix(to_stat[,-1])
+  distances.number = colnames(to_stat[,-1])
+  
+  #calc cos similarity
+  n <- ncol(distances.mat) 
+  cmb <- expand.grid(i=1:n, j=1:n) %>% filter(i<j)
+  
+  cos.sim.vec <- apply(cmb,1,function(x) cos.sim(x,distances.mat))
+  
+  to_plot = cmb %>% add_column(sim=cos.sim.vec)
+  
+  return(list(to_plot=to_plot,number.list=distances.number))
+}
+
+stats_section.sim = lapply(1:10, function(tt)calc_dist_mat(topicNumber=tt))
+#summarise for table
[…]

### 2021-04-22 manuscript/asa_template/asa_template.Rmd
update asa version of manuscript

@@
 - name: Nicole White
-  thanks: The authors gratefully acknowledge ...
+  thanks: The authors gratefully acknowledge computational resources and services used in this work provided by the eResearch Office, Queensland University of Technology, Brisbane, Australia.
   affiliation: School of Public Health and Social Work, QUT
@@
 
-We used two openly available data sources to find statistical methods sections, covering peer-reviewed journal articles and registered clinical trial protocols.
+We used two openly available data sources to find statistical methods sections: research articles published in _PLOS ONE_ and study protocols registered on the Australian and New Zealand Clinical Trials Registry (ANZCTR).
+Data sources were chosen as examples of common research outputs that include descriptions of statistical methods that were used, or are planned, for analysing study outcomes.
 
@@
 All _PLOS ONE_ articles are freely accessible via the PLOS Application Programming Interface (API).
-This enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
+This enabled us to conduct searches of full-text articles and analyse data on articles' text content and general attributes such as publication date and field(s) of research.
 To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings.
@@
 Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
-Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings agai
+Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
+For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed.
+
+<!-- For records that did not pass this second stage filter, we reviewed where keywords appeared in the full-text to ascertain the extent of missing statistical analysis sections.
+...[TODO: finish this whilst draft is with coauthors]-->
 
@@
 
-nst frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
-To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]
-
 \subsection{Clustering algorithm}
@@
 
+We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
+
 Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
 
-Results were examined for evidence of boilerplate text in two ways.
+For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways.
 We first reviewed the top ten results that represented the strongest matches to each cluster.
-To compare results within a cluster, we calculated cosine similarity score for all pairs of records; higher scores denoted a higher degree of similarity in text between a pair of records.
+Records assigned to the same cluster were also compared by calculating pairwise cosine similarities; higher scores denoted a higher degree of similarity in text between a pair of records.
 
@@
 
-```{r, echo=FALSE, fig.align='center', out.width="95%", fig.cap="PLOS ONE result summary."}
+```{r, echo=FALSE, fig.align='center', out.width="70%",fig.cap="\\label{fig:plos-n}PLOS summary"}
+
 knitr::include_graphics("figures/plos.summary.jpg")
@@
 
-```{r,fig.height=12,fig.width=6,fig.asp=0.55,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
+```{r}
 
@@
 
-API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search criteria.
+API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search strategy.
 In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings, including 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.
@@
 
-```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="PLOS ONE topic clouds."}
+```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="\\label{fig:plos-wordclouds}PLOS ONE topic clouds."}
 knitr::include_graphics("figures/plos.wordclouds.jpg")
@@
 
-We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
 Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters.
@@
 
-```{r}
-load('../../results/plos.results.10topics.rda')
+The topic clouds based on ten clusters are in Figure~\ref{fig:plos-wordclouds}.
+Frequently occurring words reflected the use of statistical software (Topics 3 and 5), descriptive statistics (Topic 6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9).
+Statistical methods sections associated with regression (Topic 2) and meta-analysis (Topic 7) were also identified.
 
-ftab.plos = matches %>%
-  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')
+<!-- ```{r} -->
+<!-- load('../../results/plos.results.10topics.rda') -->
 
-flextable(ftab.plos)
+<!-- ftab.plos = matches %>% -->
+<!--   group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')  -->
 
-```
+<!-- flextable(ftab.plos) -->
 
-The topic clouds based on ten clusters are in Figure~\ref{fig:plot-10-topics}.
-Frequently occurring words reflected the use of statistical software (Topics~3 and 5), descriptive statistics (Topic~6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9).
-There are also statistical methods sections associated with regression (Topic~2) and meta-analysis (Topic~7).
+<!-- ``` -->
 
@@
 
-We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
-Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
+We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~\ref{fig:anzctr-missing}. Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
+
+```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR search results."}
+
+knitr::include_graphics("figures/excluded_anzctr_missing.jpg")
+
+```
 

### 2021-04-22 manuscript/asa_template/asa_template.tex
update asa version of manuscript

@@
   \author{
-        Nicole White \thanks{The authors gratefully acknowledge \ldots{}} \\
+        Nicole White \thanks{The authors gratefully acknowledge computational resources and services
+used in this work provided by the eResearch Office, Queensland
+University of Technology, Brisbane, Australia.} \\
     School of Public Health and Social Work, QUT\\
@@
 We used two openly available data sources to find statistical methods
-sections, covering peer-reviewed journal articles and registered
-clinical trial protocols.
+sections: research articles published in \emph{PLOS ONE} and study
+protocols registered on the Australian and New Zealand Clinical Trials
+Registry (ANZCTR). Data sources were chosen as examples of common
+research outputs that include descriptions of statistical methods that
+were used, or are planned, for analysing study outcomes.
 
@@
 Application Programming Interface (API). This enabled us to conduct
-semi-automated searches of full-text articles and analyse data on
-individual records, including text content and general attributes such
-as publication date and field(s) of research. To find papers with a
-statistical methods section we used targeted API searches followed by
-article filtering based on section headings. The data were downloaded on
-3 July 2020.
+searches of full-text articles and analyse data on articles' text
+content and general attributes such as publication date and field(s) of
+research. To find papers with a statistical methods section we used
+targeted API searches followed by article filtering based on section
+headings. The data were downloaded on 3 July 2020.
 
@@
 methods sections, we performed partial matching on available headings
-agai
+against frequently used terms in initial search results: `Statistical
+analysis', `Statistical analyses', `Statistical method', `Statistics',
+`Data analysis' and `Data analyses'. For records that did not pass this
+second stage filter, we selected a random sample of XXX records and
+reviewed where initial search terms appeared in the full-text, to
+estimate the likely proportion of statistical methods sections that were
+missed.
 
@@
 
-nst frequently used terms in initial search results: `Statistical
-analysis', `Statistical analyses', `Statistical method', `Statistics',
-`Data analysis' and `Data analyses'. To determine the reliability of our
-chosen filters, we manually reviewed full text data extracted for a
-random sample of XXX articles that were not matched (File
-S1).{[}TODO\ldots finish this thought\ldots{]}
-
 \subsection{Clustering algorithm}
@@
 
+We applied the clustering algorithm to the cleaned dataset, varying the
+number of clusters from 1 to 50.
+
 Results were transformed to lower case for the clustering, but examples
@@
 
-Results were examined for evidence of boilerplate text in two ways. We
-first reviewed the top ten results that represented the strongest
-matches to each cluster. To compare results within a cluster, we
-calculated cosine similarity score for all pairs of records; higher
-scores denoted a higher degree of similarity in text between a pair of
-records.
+For each dataset, records assigned to invididual clusters were examined
+for evidence of boilerplate text in two ways. We first reviewed the top
+ten results that represented the strongest matches to each cluster.
+Records assigned to the same cluster were also compared by calculating
+pairwise cosine similarities; higher scores denoted a higher degree of
+similarity in text between a pair of records.
 
@@
 
-{\centering \includegraphics[width=0.95\linewidth]{figures/plos.summary}
+{\centering \includegraphics[width=0.7\linewidth]{figures/plos.summary}
 
@@
 
-\caption{PLOS ONE result summary.}\label{fig:unnamed-chunk-3}
+\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 API searches returned 131,847 unique records, of which 111,731 (85\%)
-included a statistical methods section based on our search criteria. In
+included a statistical methods section based on our search strategy. In
 the final sample, 95,518 (85\%) DOIs returned an exact match against
@@
 
-\caption{PLOS ONE topic clouds.}\label{fig:unnamed-chunk-5}
+\caption{\label{fig:plos-wordclouds}PLOS ONE topic clouds.}\label{fig:unnamed-chunk-5}
 \end{figure}
 
-We applied the clustering algorithm to the cleaned dataset, varying the
-number of clusters from 1 to 50. Increasing the number of clusters
-decreased cluster quality based on global goodness-of-fit measures
-(Supplementary Figure 1), with average silhouette score and
-within-cluster dispersion leveling off around 20 clusters. This
-indicated that the data comprised one large, heterogeneous cluster and
-multiple smaller clusters.
-
-\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}
+Increasing the number of clusters decreased cluster quality based on
+global goodness-of-fit measures (Supplementary Figure 1), with average
+silhouette score and within-cluster dispersion leveling off around 20
+clusters. This indicated that the data comprised one large,
+heterogeneous cluster and multiple smaller clusters.
 
 The topic clouds based on ten clusters are in
-Figure\textasciitilde{}\ref{fig:plot-10-topics}. Frequently occurring
-words reflected the use of statistical software (Topics\textasciitilde3
-and 5), descriptive statistics (Topic\textasciitilde6), group based
-hypothesis testing (Topics 1 and 4) and definitions of statistical
-significance (Topics\textasciitilde1 and 9). There are also statistical
-methods sections associated with regression (Topic\textasciitilde2) and
-meta-analysis (Topic\textasciitilde7).
+Figure\textasciitilde{}\ref{fig:plos-wordclouds}. Frequently occurring
+words reflected the use of statistical software (Topics 3 and 5),
+descriptive statistics (Topic 6), group based hypothesis testing (Topics
+1 and 4) and definitions of statistical significance
+(Topics\textasciitilde1 and 9). Statistical methods sections associated
+with regression (Topic 2) and meta-analysis (Topic 7) were also
+identified.
 
@@
 
-\includegraphics[width=6.13in,height=3.49in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-7-1.png}
+\includegraphics[width=6.13in,height=3.49in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}
 
@@
 
-\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-8}
+\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 We downloaded 28,008 studies. The numbers of excluded studies are shown
-in Figure\textasciitilde X. Of the 12,700 included studies, 9,523 (75\%)
-had a statistical methods section.
+in Figure\textasciitilde{}\ref{fig:anzctr-missing}. Of the 12,700
+included studies, 9,523 (75\%) had a statistical methods section.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/excluded_anzctr_missing}
+
+}
+
+\caption{ANZCTR search results.}\label{fig:unnamed-chunk-8}
+\end{figure}
 

### 2021-04-22 manuscript/asa_template/asa_template.Rmd
update format of .xlsx boilerplate files

@@
 output: rticles::asa_article
+header-includes:
+      - \usepackage{pdflscape}
 fig_caption: yes
@@
 
-```{r, echo=FALSE, fig.align='center', out.width="70%",fig.cap="\\label{fig:plos-n}PLOS summary"}
+```{r, echo=FALSE, fig.align='center', out.width="60%",fig.cap="\\label{fig:plos-n}PLOS summary"}
 
@@
 
-API searches returned `r format(n.api,big.mark=',')` unique records, of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search strategy.
-In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings, including 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.
+API searches returned `r format(n.api,big.mark=',')` results (DOIs), of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search strategy.
+In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings: 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.
+
 Among DOIs that did not meet the partial matching criteria, initial search terms appeared in [TODO].
@@
 
-The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \ref{fig:plos-n}B).
+Statistical methods sections had a median length of `r n.words[['med']]` words and inter-quartile range of `r n.words[['q1']]` to `r n.words[['q3']]` words (Figure \ref{fig:plos-n}B).
 `r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
-`r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.
+`r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had sections with 50 words or less, equal to the length of this paragraph.
 
-All papers included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
+All DOIs included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
 (n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications
@@
 
-```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="\\label{fig:plos-wordclouds}PLOS ONE topic clouds."}
+<!-- Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters.
+This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.
+-->
+
+```{r, echo=FALSE,fig.align='center', out.width="70%",fig.cap="\\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE"}
+
 knitr::include_graphics("figures/plos.wordclouds.jpg")
+
 ```
 
-Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters.
-This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.
-
-The topic clouds based on ten clusters are in Figure~\ref{fig:plos-wordclouds}.
+The topic clouds based on ten clusters are in Figure~\ref{fig:plos-clouds}.
 Frequently occurring words reflected the use of statistical software (Topics 3 and 5), descriptive statistics (Topic 6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9).
@@
 
-<!-- ```{r} -->
-<!-- load('../../results/plos.results.10topics.rda') -->
+```{r}
+load('../../results/plos.results.10topics.rda')
 
-<!-- ftab.plos = matches %>% -->
-<!--   group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')  -->
+ftab.plos = matches %>%
+  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')
 
-<!-- flextable(ftab.plos) -->
+flextable(ftab.plos)
 
-<!-- ``` -->
+```
 
-Topics related to the use of statistical software differentiated between Prism GraphPad (Topic~3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) (Box 1).
-A manual review of the top matching sections in these topics showed strong evidence of boilerplate text.
-Nine out ten top matches for Topic 3 stated the use of Prism GraphPad but did not specify which statistical methods were used; six out of ten top matches returned the same cluster score indicating near identical text.
+Topics related to the use of statistical software differentiated between Prism GraphPad (Topic~3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) .
+Manual review of results showed that nine out of the ten top matching DOIs for Topic 3 stated the use of Prism GraphPad, but did not specify which statistical methods were used.
 Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
-
-* Results of cosine similaritites
+Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.
+Examples correspond to top ranking DOIs, and DOIs with the higher number of matches based on cosine similarity score greater than 0.8.
+Further results from cosine similarity scores are provided in Supplementary File 2.
 
 ```{r}
-load('../../results/plos.cosinesim.10topics.rda')
+FitFlextableToPage <- function(ft, pgwidth = 8){
 
-ftab = flextable(ftab.cosine.plos) %>% theme_box() %>% autofit()
-ftab <- set_caption(ftab, "Summary of cosine similarity scores among top 500 DOIs assigned to each topic: PLOS ONE")
+  ft_out <- ft %>% autofit()
+
+  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
+  return(ft_out)
+}
+boilerplate.plos =  read.table(header=TRUE, sep='&', text='
+Theme & Topic & Examples
+Statistical software & 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.
+Statistical software &3 & All statistical analysis was performed using Graphpad Prism software.
+Statistical software & 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.
+Statistical software & 5  & The results are presented as the mean ± SEM.
+Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
+The difference was considered statistically significant at P<0.05.
+Between-group hypothesis testing, two groups & 1 & Student’s t-test was used for statistical analysis. A p value of <0.05 was considered statistically significant.,
+Between-group hypothesis testing, two groups& 1 & Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant
+Between-group hypothesis testing, two groups& 9 & Statistical significance was determined by Students t-tests. * P<0.05, ** P = <0.01, *** P<0.001.
+Between-group hypothesis testing, two groups& 9 & Data are presented as the mean ± SD.
+Statistical analysis was performed using the Students unpaired t-test.
+Differences were considered statistically significant at *p<0.05, **p<0.01 and ***p<0.001
+Between-group hypothesis testing, two or more groups & 4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.
+Between-group hypothesis testing, two or more groups& 4 & Data are expressed as the mean ± SEM.
+Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
+P values less than 0.05 were considered significant.
+Descriptive statistics & 6 & All results are expressed as means ± standard deviation (± SD).
+Descriptive statistics & 6 & Data are expressed as mean ± Standard error of the mean (SEM).
+Statistical analysis of the data was performed by the Student’s t-test.
+A value of P<0.05 was considered statistically significant.
+')
+boilerplate.plos = boilerplate.plos %>% mutate(Topic=replace_na(Topic,''))
+ftab = flextable(boilerplate.plos) %>% empty_blanks %>% 
+  merge_v(j=c('Theme','Topic')) %>% 
+  theme_booktabs %>%
+FitFlextableToPage
 ftab
@@
 
+Definitions of statistical significance featured strongly in Topic 1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
+Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
+Topic 9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: ``$^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$''., a practice that has been criticised [add wasserstein2019].
+
+Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used.
+One-way analysis of variance also featured strongly in Topic 4 (n = 10,163), with text citing common methods for performing post-hoc multiple comparisons.
+Top matching sections across these topics did not include definitions of groups being compared.
+
+Frequently occurring words in Topic 6 (n = 4,746) including 'Mean', 'SD' and 'SEM' reflected the use of descriptive statistics for summarising continuous variables.
+Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance.
+
+<!-- Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183).
+-->
+
 \subsection{ANZCTR}
-
-```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR topic clouds."}
-knitr::include_graphics("figures/anzctr.wordclouds.jpg")
-```
 
@@
 knitr::include_graphics("figures/excluded_anzctr_missing.jpg")
+
+```
+
+```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR topic clouds."}
+
+knitr::include_graphics("figures/anzctr.wordclouds.jpg")
 

### 2021-04-22 manuscript/asa_template/asa_template.tex
update format of .xlsx boilerplate files

@@
 % Pandoc citation processing
+
+\usepackage{pdflscape}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/plos.summary}
+{\centering \includegraphics[width=0.6\linewidth]{figures/plos.summary}
 
@@
 
-API searches returned 131,847 unique records, of which 111,731 (85\%)
+API searches returned 131,847 results (DOIs), of which 111,731 (85\%)
 included a statistical methods section based on our search strategy. In
 the final sample, 95,518 (85\%) DOIs returned an exact match against
-common section headings, including 64,133 for `statistical analysis',
-13,380 for `statistical analyses' and 13,627 for `data analysis'. Among
-DOIs that did not meet the partial matching criteria, initial search
-terms appeared in {[}TODO{]}.
+common section headings: 64,133 for `statistical analysis', 13,380 for
+`statistical analyses' and 13,627 for `data analysis'.
+
+Among DOIs that did not meet the partial matching criteria, initial
+search terms appeared in {[}TODO{]}.
 
@@
 
-The median length of statistical methods sections was 127 words (IQR: 61
-to 254 words) (Figure \ref{fig:plos-n}B). 7,450 articles (7\%) had a
-statistical methods section of 500 words or more. 19,461 articles (17\%)
-had a statistical methods section of 50 words or less, equal to the
-length of this paragraph.
+Statistical methods sections had a median length of 127 words and
+inter-quartile range of 61 to 254 words (Figure \ref{fig:plos-n}B).
+7,450 articles (7\%) had a statistical methods section of 500 words or
+more. 19,461 articles (17\%) had sections with 50 words or less, equal
+to the length of this paragraph.
 
-All papers included Biology and life sciences (n = 107,584), Earth
+All DOIs included Biology and life sciences (n = 107,584), Earth
 sciences (n = 7,605) and/or Computer and information sciences (n =
@@
 
-\caption{\label{fig:plos-wordclouds}PLOS ONE topic clouds.}\label{fig:unnamed-chunk-5}
+\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE (left) and ANZCTR (right)}\label{fig:unnamed-chunk-5}
 \end{figure}
 
-Increasing the number of clusters decreased cluster quality based on
-global goodness-of-fit measures (Supplementary Figure 1), with average
-silhouette score and within-cluster dispersion leveling off around 20
-clusters. This indicated that the data comprised one large,
-heterogeneous cluster and multiple smaller clusters.
+The topic clouds based on ten clusters are in
+Figure\textasciitilde{}\ref{fig:plos-clouds}. Frequently occurring words
+reflected the use of statistical software (Topics 3 and 5), descriptive
+statistics (Topic 6), group based hypothesis testing (Topics 1 and 4)
+and definitions of statistical significance (Topics\textasciitilde1 and
+9). Statistical methods sections associated with regression (Topic 2)
+and meta-analysis (Topic 7) were also identified.
 
-The topic clouds based on ten clusters are in
-Figure\textasciitilde{}\ref{fig:plos-wordclouds}. Frequently occurring
-words reflected the use of statistical software (Topics 3 and 5),
-descriptive statistics (Topic 6), group based hypothesis testing (Topics
-1 and 4) and definitions of statistical significance
-(Topics\textasciitilde1 and 9). Statistical methods sections associated
-with regression (Topic 2) and meta-analysis (Topic 7) were also
-identified.
+\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}
 
@@
 Prism GraphPad (Topic\textasciitilde3: n = 9,879; 8.8\%) and SPSS (Topic
-5: n = 9,574; 8.6\%) (Box 1). A manual review of the top matching
-sections in these topics showed strong evidence of boilerplate text.
-Nine out ten top matches for Topic 3 stated the use of Prism GraphPad
-but did not specify which statistical methods were used; six out of ten
-top matches returned the same cluster score indicating near identical
-text. Top matching sections for Topic 5 included information on SPSS
-version numbers and definitions of statistical significance.
+5: n = 9,574; 8.6\%) . Manual review of results showed that nine out of
+the ten top matching DOIs for Topic 3 stated the use of Prism GraphPad,
+but did not specify which statistical methods were used. Top matching
+sections for Topic 5 included information on SPSS version numbers and
+definitions of statistical significance. Examples of boilerplate text
+for selected topics are presented in Table
+\ref{tab:plos-example-boilerplate}. Examples correspond to top ranking
+DOIs, and DOIs with the higher number of matches based on cosine
+similarity score greater than 0.8. Further results from cosine
+similarity scores are provided in Supplementary File 2.
 
-\begin{itemize}
-\tightlist
-\item
-  Results of cosine similaritites
-\end{itemize}
+\includegraphics[width=8.00in,height=4.12in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-7-1.png}
 
-\includegraphics[width=6.13in,height=3.49in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}
+Definitions of statistical significance featured strongly in Topic 1 (n
+= 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for
+comparing differences between two groups. Topic 1 reflected applications
+of Student's t-test assuming a 5\% level of statistical significance.
+Topic 9 referenced similar methods combined with multiple thresholds for
+declaring statistical significance by asterisk: ``\(^{*}p<0.05\),
+\(^{**}p<0.01\) and \(^{***}p<0.001\)''., a practice that has been
+criticised {[}add wasserstein2019{]}.
+
+Group-based hypothesis testing was a recurring theme across topics, with
+text descriptions varying based on method(s) used. One-way analysis of
+variance also featured strongly in Topic 4 (n = 10,163), with text
+citing common methods for performing post-hoc multiple comparisons. Top
+matching sections across these topics did not include definitions of
+groups being compared.
+
+Frequently occurring words in Topic 6 (n = 4,746) including `Mean', `SD'
+and `SEM' reflected the use of descriptive statistics for summarising
+continuous variables. Sections associated with this topic appeared to be
+expanded versions of Topics 1 and 9, showing evidence of boilerplate
+text in the form of descriptive statistics followed by hypothesis
+testing using Student's t-test, Mann Whitney U or one-way analysis of
+variance.
 
 \subsection{ANZCTR}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/anzctr.wordclouds}
-
-}
-
-\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-7}
-\end{figure}
 
@@
 \caption{ANZCTR search results.}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/anzctr.wordclouds}
+
+}
+
+\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 
-\includegraphics[width=4.66in,height=1.55in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-9-1.png}
+\includegraphics[width=4.66in,height=1.55in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-10-1.png}
 
@@
 
-\caption{PLOS ONE clustering metrics.}\label{fig:unnamed-chunk-10}
+\caption{PLOS ONE clustering metrics.}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 
-\caption{PLOS ONE topic values.}\label{fig:unnamed-chunk-11}
+\caption{PLOS ONE topic values.}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
-\caption{ANZCTR clustering metrics.}\label{fig:unnamed-chunk-12}
+\caption{ANZCTR clustering metrics.}\label{fig:unnamed-chunk-13}
 \end{figure}
@@
 
-\caption{ANZCTR topic values.}\label{fig:unnamed-chunk-13}
+\caption{ANZCTR topic values.}\label{fig:unnamed-chunk-14}
 \end{figure}

### 2021-04-22 manuscript/draft_manuscript.Rmd
update to asa template
update format of .xlsx boilerplate files

@@
 
-Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183).
-
 <!-- Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling.

### 2021-04-22 code/7_summarise_stats_sections.R

@@ -66,16 +66,20 @@ ftab.cosine.plos = mutate(ftab.cosine.plos,'Median (IQR)' = paste0(Median,' (',Q
   select(topic,'Median (IQR)',pgt80,pgt90,peq1) %>%
   rename('Similarity > 0.8'=pgt80,'Similarity > 0.9' = pgt90, 'Similarity = 1' = peq1) 
 
-#exact matches, topic 1
-boilerplate = lapply(1:10,function(x) stats_section.sim.top500 %>% filter(topic==x,sim>0.8) %>% mutate(pair = row_number(),
-                                                                               doi.i = stats_section.doi.top500[[x]][i],
-                                                                               doi.j = stats_section.doi.top500[[x]][j]) %>% select(topic,pair,sim,doi.i,doi.j))
+#boilerplate text, pairwise
+boilerplate = lapply(1:10,function(x) stats_section.sim.top500 %>% filter(topic==x,sim>0.5) %>% mutate(pair = row_number(),
+                                                                               doi_1 = stats_section.doi.top500[[x]][i],
+                                                                               doi_2 = stats_section.doi.top500[[x]][j]) %>% select(topic,pair,sim,doi_1,doi_2))
   
 
-boilerplate.dois = bind_rows(boilerplate) %>% gather(variable,doi,-topic,-pair,-sim) %>% arrange(topic,-sim,pair) %>% select(-variable)
+boilerplate.dois = bind_rows(boilerplate) ##%>% gather(variable,doi,-topic,-pair,-sim) %>% arrange(topic,-sim,pair)
 
 #join to text
-boilerplate.text.plos = boilerplate.dois %>% left_join(matches %>% select(doi,text_heading,text_data,text_data_clean,rank,words),by='doi')
+boilerplate.text.plos = boilerplate.dois %>% left_join(matches %>% select(doi,text_data,rank,words),by=c('doi_1'='doi')) %>%
+  rename('stats_section_1'=text_data,'rank_1'=rank,'words_1'=words) %>%
+  left_join(matches %>% select(doi,text_data,rank,words),by=c('doi_2'='doi')) %>%
+  rename('stats_section_2'=text_data,'rank_2'=rank,'words_2'=words) %>%
+  select(topic,pair,sim,doi_1,rank_1,words_1,doi_2,rank_2,words_2,stats_section_1,stats_section_2) %>% arrange(topic,-sim)
 
[…]

### 2021-04-22 code/7_summarise_stats_sections_anzctr.R

@@ -62,21 +62,27 @@ ftab.cosine.anzctr = stats_section.sim.top500 %>% group_by(topic) %>% summarise(
                                                                               pgt80=sum(sim>0.8),pgt90=sum(sim>0.9),peq1=sum(sim==1)) %>% arrange(as.numeric(topic),.groups='drop') %>%
   mutate_if(is.numeric,function(x) round(x,2))
 
-ftab.cosine.plos = mutate(ftab.cosine.plos,'Median (IQR)' = paste0(Median,' (',Q1,' to ',Q3,')')) %>%
+ftab.cosine.anzctr = mutate(ftab.cosine.anzctr,'Median (IQR)' = paste0(Median,' (',Q1,' to ',Q3,')')) %>%
   select(topic,'Median (IQR)',pgt80,pgt90,peq1) %>%
   rename('Similarity > 0.8'=pgt80,'Similarity > 0.9' = pgt90, 'Similarity = 1' = peq1) 
 
 #exact matches, topic 1
 boilerplate = lapply(1:10,function(x) stats_section.sim.top500 %>% filter(topic==x,sim>0.8) %>% mutate(pair = row_number(),
-                                                                                                       num.i = stats_section.number.top500[[x]][i],
-                                                                                                       num.j = stats_section.number.top500[[x]][j]) %>% 
-                       select(topic,pair,sim,num.i,num.j))
+                                                                                                       number_1 = stats_section.number.top500[[x]][i],
+                                                                                                       number_2 = stats_section.number.top500[[x]][j]) %>% 
+                       select(topic,pair,sim,number_1,number_2))
 
 
-boilerplate.numbers = bind_rows(boilerplate) %>% gather(variable,number,-topic,-pair,-sim) %>% arrange(topic,-sim,pair) %>% select(-variable)
+boilerplate.numbers = bind_rows(boilerplate)
 
 #join to text
-boilerplate.text.anzctr = boilerplate.numbers %>% left_join(matches %>% select(number,text_data,text_data_clean,rank,words),by='number')
+boilerplate.text.anzctr = boilerplate.numbers %>% left_join(matches %>% select(number,text_data,text_data_clean,rank,words),by=c('number_1'='number')) %>%
+  rename('stats_section_1'=text_data,'rank_1'=rank,'words_1'=words) %>%
[…]

### 2021-04-23 manuscript/asa_template/asa_template.Rmd
revise manuscript anzctr plos

@@
 
-We applied the same text cleaning to both data sources. 
+We applied the same text cleaning steps to both data sources using functions from the `textclean package' [@textclean]. 
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
@@
 Mathematical notation including Greek letters was converted from Unicode characters to plain text.
-For example, the Unicode characters corresponding to $\theta$ (<U+03B8>) were replaced with ‘theta’.
-Similarly, common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were converted into plain text using the ‘textclean’ package [@textclean].
+For example, the Unicode character corresponding to $\theta$ (<U+03B8>) was replaced with ‘theta’.
+Common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were similarly converted into plain text.
 General formatting was removed, this included carriage returns, punctuation marks, in-text references (e.g.
@@
 Further terms were sourced from index searches of three statistics textbooks [@Dobson2018,@Diggle2013,@Bland2015].
-The final list is provided as Supplementary Material.
+The final list is provided in Supplementary File 1.
 Plurals (e.g., ‘chi-squares’) unhyphenated (e.g., ‘chi square’) and combined (e.g.
@@
 We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
-
-Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
+Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster dispersion.
 
 For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways.
-We first reviewed the top ten results that represented the strongest matches to each cluster.
-Records assigned to the same cluster were also compared by calculating pairwise cosine similarities; higher scores denoted a higher degree of similarity in text between a pair of records.
+We first reviewed the strongest matches to each cluster based on topic scores[may not be the right word].
+We also compared similarities in text between pairs of records assigned to each cluster.
+Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster.
+Higher similarity score represent greater overlap in text content between statistical methods sections.
+Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
 
@@
 
-API searches returned `r format(n.api,big.mark=',')` results (DOIs), of which `r format(n.include,big.mark=',')` (`r perc.include`%) included a statistical methods section based on our search strategy.
+API searches returned `r format(n.api,big.mark=',')` results (DOIs) based on chosen search terms.
+After partial matching, `r format(n.include,big.mark=',')` (`r perc.include`%) statistical methods sections were identified.
 In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings: 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.
-
-Among DOIs that did not meet the partial matching criteria, initial search terms appeared in [TODO].
+For our random sample of DOIs that did not meet the partial matching criteria, initial search terms appeared in Introduction (number), Results (number) only.
+xxx appeared in the Methods section without a section heading and xxx had a non-standard section heading.
+[TODO].
 
 Search results varied by journal volume (Figure \ref{fig:plos-n}A).
-The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`), corresponding to years 2013 and 2014.
+The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) in years 2013 and 2014.
 This trend aligned with the total number of papers published in _PLOS ONE_ over the same period.
@@
 All DOIs included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
-(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications
-(Figure \ref{fig:plos-n}C).
+(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications (Figure \ref{fig:plos-n}C).
 
@@
 
-```{r, echo=FALSE,fig.align='center', out.width="70%",fig.cap="\\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE"}
+```{r, echo=FALSE,fig.align='center', out.width="80%",fig.cap="\\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE"}
 
@@
 
-The topic clouds based on ten clusters are in Figure~\ref{fig:plos-clouds}.
-Frequently occurring words reflected the use of statistical software (Topics 3 and 5), descriptive statistics (Topic 6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics~1 and 9).
-Statistical methods sections associated with regression (Topic 2) and meta-analysis (Topic 7) were also identified.
+The topic clouds based on ten clusters are in Figure \ref{fig:plos-clouds}.
+Frequently occurring words reflected the use of statistical software (topics 3 and 5), descriptive statistics (topic 6), group based hypothesis testing (topics 1 and 4) and definitions of statistical significance (topics 1 and 9).
+Statistical methods sections related to regression (topic 2) and meta-analysis (topic 7) were also identified.
 
@@
 ftab.plos = matches %>%
-  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop')
+  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop') %>%
+  mutate(perc = round(100*DOIs/sum(DOIs),1))
 
-flextable(ftab.plos)
+#flextable(ftab.plos)
 
@@
 
-Topics related to the use of statistical software differentiated between Prism GraphPad (Topic~3: n = 9,879; 8.8%) and SPSS (Topic 5: n = 9,574; 8.6%) .
-Manual review of results showed that nine out of the ten top matching DOIs for Topic 3 stated the use of Prism GraphPad, but did not specify which statistical methods were used.
-Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
-Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.
-Examples correspond to top ranking DOIs, and DOIs with the higher number of matches based on cosine similarity score greater than 0.8.
-Further results from cosine similarity scores are provided in Supplementary File 2.
+Topics related to statistical software differentiated between Prism GraphPad (topic 3: n = `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(perc)`%) and SPSS (topic 5: n = `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(perc)`%) .
+Manual review of results showed that nine out of the ten top matching DOIs for topic 3 stated the use of Prism GraphPad, but did not specify which statistical methods were used.
+Top matching sections for topic 5 included information on SPSS version numbers and definitions of statistical significance.
 
-```{r}
-FitFlextableToPage <- function(ft, pgwidth = 8){
+\begin{landscape}
+\begin{table}[]
+\centering
+\caption{Example boilerplate text from PLOS ONE dataset}
+\label{tab:plos-example-boilerplate}
+\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\hline
+\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
+1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
+& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
+\hline
+3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
+ & All statistical analysis was performed using Graphpad Prism software.\\
+ \hline
+4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
+& Data are expressed as the mean ± SEM.
+Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
+P values less than 0.05 were considered significant.\\
+\hline
+5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
+  & The results are presented as the mean ± SEM.
+Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
+The difference was considered statistically significant at P$<$0.05.\\
+  \hline
+6 & All results are expressed as means ± standard deviation (± SD).\\
+& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
+\hline
+9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
+ & Data are presented as the mean ± SD.
+Statistical analysis was performed using the Students unpaired t-test.
+Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
+ \hline
+\end{tabular}
+\end{table}
+\end{landscape}
 
-  ft_out <- ft %>% autofit()
-
-  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
-  return(ft_out)
-}
-boilerplate.plos =  read.table(header=TRUE, sep='&', text='
-Theme & Topic & Examples
-Statistical software & 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.
-Statistical software &3 & All statistical analysis was performed using Graphpad Prism software.
-Statistical software & 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.
-Statistical software & 5  & The results are presented as the mean ± SEM.
-Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
-The difference was considered statistically significant at P<0.05.
-Between-group hypothesis testing, two groups & 1 & Student’s t-test was used for statistical analysis. A p value of <0.05 was considered statistically significant.,
-Between-group hypothesis testing, two groups& 1 & Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant
-Between-group hypothesis testing, two groups& 9 & Statistical significance was determined by Students t-tests. * P<0.05, ** P = <0.01, *** P<0.001.
-Between-group hypothesis testing, two groups& 9 & Data are presented as the mean ± SD.
-Statistical analysis was performed using the Students unpaired t-test.
-Differences were considered statistically significant at *p<0.05, **p<0.01 and ***p<0.001
-Between-group hypothesis testing, two or more groups & 4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.
-Between-group hypothesis testing, two or more groups& 4 & Data are expressed as the mean ± SEM.
-Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
-P values less than 0.05 were considered significant.
-Descriptive statistics & 6 & All results are expressed as means ± standard deviation (± SD).
-Descriptive statistics & 6 & Data are expressed as mean ± Standard error of the mean (SEM).
-Statistical analysis of the data was performed by the Student’s t-test.
-A value of P<0.05 was considered statistically significant.
-')
-boilerplate.plos = boilerplate.plos %>% mutate(Topic=replace_na(Topic,''))
-ftab = flextable(boilerplate.plos) %>% empty_blanks %>% 
-  merge_v(j=c('Theme','Topic')) %>% 
-  theme_booktabs %>%
-FitFlextableToPage
-ftab
-
-```
-
-Definitions of statistical significance featured strongly in Topic 1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
+Definitions of statistical significance featured strongly in topic 1 (n = `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(perc)`%) and topic 9 (n = `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(perc)`%), combined with text describing hypothesis testing for comparing differences between two groups.
 Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
-Topic 9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: ``$^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$''., a practice that has been criticised [add wasserstein2019].
+Topic 9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: ``$^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$'', a practice that has been criticised [@Wasserstein2019].
 
 Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used.
-One-way analysis of variance also featured strongly in Topic 4 (n = 10,163), with text citing common methods for performing post-hoc multiple comparisons.
-Top matching sections across these topics did not include definitions of groups being compared.
-
-Frequently occurring words in Topic 6 (n = 4,746) including 'Mean', 'SD' and 'SEM' reflected the use of descriptive statistics for summarising continuous variables.
-Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance.
+One-way analysis of variance also featured strongly in topic 4 (n = `r filter(ftab.plos,topic_id=='Topic 4') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 4') %>% pull(perc)`%), combined with common methods for performing post-hoc multiple comparisons.
+Frequently occurring words in topic 6 (n = `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(perc)`%) reflected mentions of descriptive statistics for summarising continuous variables, for example: ``Data are expressed as means ± standard error of the mean (SEM).'' Sections assigned to this topic appeared to be expanded versions of topics 1 and 9.
+Examples of boilerplate text were in the form of descriptive statistics followed by hypothesis testing, for example, Student's t-test, Mann Whitney U or one-way analysis of variance.
 
@@
 
+Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}. Detailed findings based on cosine similarity scores are provided in Supplementary File 2.
+
 \subsection{ANZCTR}
 
-We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~\ref{fig:anzctr-missing}. Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
+We downloaded 28,008 studies.
+The numbers of excluded studies are shown in Figure \ref{fig:anzctr-missing}.
+Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
+The median length of sections was 129 words with an inter-quartile range of 71 to 219 words.
 
-```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR search results."}
+```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="\\label{fig:anzctr-missing}ANZCTR search results."}
 
@@
 
-The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+The clustering algorithm found groups that were purely sample size calculations (topic 2, n = 1834; topic 4, n = 909), pilot studies (topic 5, n = 834), safety/tolerability studies (topic 6, n = 524), intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852).
+Examples of boilerplate text are provided in Table \ref{tab:anzctr-example-boilerplate} and Supplementary File 3.
+
 Some methods sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".
+There were cases where the exact same method section had been re-used in a different study.
+For example, in topic 7 (n = 333), 232 sections stated 'descriptive statistics' or 'descriptive statistics used' with no additional details provided.
+Other instances with expanded descriptions of methods included topic 3 (n = 1277) outlining descriptive analyses only, and topic 6 (n = 524), for hypothesis testing methods, statistical significance and software.
 
-The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
-There were cases where the exact same method section had been re-used in a different study.
+In other cases, text had been slightly modified to account for changes in primary and secondary outcomes.
+Examples of these text changes were found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances related to sample size calculations for patient recruitment to different studies (see Supplementary File 3).
 
-We also found evidence of 'boilerplate' sections clustered as topic 3, example text
+\begin{landscape}
+\begin{table}[]
+\centering
+\caption{Example boilerplate text from ANZCTR dataset}
+\label{tab:anzctr-example-boilerplate}
+\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\hline
+\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
+3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant.\\
+ &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
+The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate.
+\\ \hline
+5 & Pilot study \\
+ &  No formal sample size calculation was performed \\ \hline
+7 & Descriptive statistics \\
+ &  Descriptive statistics used \\ \hline
+9 & Linear mixed models will be used to analyse the data. \\ \hline
+10 & Repeated measures of ANOVA \\
+ &  Pre-, during, post- and follow-up variables will be subjected to mixed methods and repeated measures analyses to determine significant changes over (group and) time. \\ \hline
+\end{tabular}
+\end{table}
+\end{landscape}
 
-**Box x: Examples of boilerplate text from ANZCTR**
-
-Topic 3
-
-* "Shapiro Wilk test was used as normality test.
+<!-- # **Box x: Examples of boilerplate text from ANZCTR (moved to table)** -->
+<!-- #  -->
+<!-- # Topic 3 -->
+<!-- #  -->
+<!-- # "Shapiro Wilk test was used as normality test.
 Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
 Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test.
-Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed."
-* "Comparisons between categorical variables will be made either using chi square or Fisher exact test.
-Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
-Two sided p values of less than 0.05 will be considered statistically significant."
+Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed." -->
 
 We examined if four study characteristics were associated with a missing statistics section.
-The odds ratios and 95% credible intervals are in Table~X.
+The odds ratios and 95% credible intervals are in Table \ref{tab:anzctr-missing-odds}.
 Observational studies were less likely to have a missing methods section compared with interventional studies.
@@
 
-```{r}
-table =  read.table(header=TRUE, sep='&', text='
-Variable & Odds ratio & 95% CI
-Study type = Observational & 0.78 & 0.69, 0.89
-Date (per year) & 0.90 & 0.88, 0.91
-Number of funders & 0.80 & 0.74, 0.86
-Target sample size (per doubling) & 0.90 & 0.88, 0.92')
-
-ftab = flextable(table) %>%
-theme_box() %>%
-autofit()
-ftab
-```
-
-* Final sample size
+\begin{table}[]
+\centering
+\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
+\label{tab:anzctr-missing-odds}
+\begin{tabular}{lcc}
+\hline
+Variable & Odds ratio & 95$\%$ CI \\
+\hline
+Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
+Date (per year) & 0.90 & (0.88, 0.91) \\
+Number of funders & 0.80 & (0.74, 0.86) \\
+Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\section{Supplementary}
+<!-- \section{Supplementary} -->
 
-\subsection{PLOS}
+<!-- \subsection{PLOS} -->
 
-```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE clustering metrics."}
-knitr::include_graphics("figures/supplementary/supp_plosone.50clusters.jpg")
-```
+<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE clustering metrics."} -->
+<!-- knitr::include_graphics("figures/supplementary/supp_plosone.50clusters.jpg") -->
+<!-- ``` -->
 
-```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE topic values."}
-knitr::include_graphics("figures/supplementary/plos.topicvalues.jpg")
-```
+<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE topic values."} -->
+<!-- knitr::include_graphics("figures/supplementary/plos.topicvalues.jpg") -->
+<!-- ``` -->
 
-\subsection{ANZCTR}
+<!-- \subsection{ANZCTR} -->
 
-```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR clustering metrics."}
-knitr::include_graphics("figures/supplementary/supp_anzctr.50clusters.jpg")
-```
+<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR clustering metrics."} -->
+<!-- knitr::include_graphics("figures/supplementary/supp_anzctr.50clusters.jpg") -->
+<!-- ``` -->
 
-```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR topic values."}
-knitr::include_graphics("figures/supplementary/anzctr.topicvalues.jpg")
-```
+<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR topic values."} -->
+<!-- knitr::include_graphics("figures/supplementary/anzctr.topicvalues.jpg") -->
+<!-- ``` -->

### 2021-04-23 manuscript/asa_template/asa_template.tex
revise manuscript anzctr plos

@@
 
-We applied the same text cleaning to both data sources. Text cleaning
+We applied the same text cleaning steps to both data sources using
+functions from the `textclean package' \citep{textclean}. Text cleaning
 aimed to standardise notation and statistical terminology, whilst
@@
 Mathematical notation including Greek letters was converted from Unicode
-characters to plain text. For example, the Unicode characters
-corresponding to \(\theta\) (\textless U+03B8\textgreater) were replaced
-with `theta'. Similarly, common symbols outside of Unicode blocks
-including `\%' (percent) and `\textless{}' (`less-than') were converted
-into plain text using the `textclean' package \citep{textclean}. General
-formatting was removed, this included carriage returns, punctuation
-marks, in-text references (e.g.~``{[}42{]}'') centred equations, and
-other non-ASCII characters. Text contained inside brackets was retained
-to maximise content for analysis, with brackets removed.
+characters to plain text. For example, the Unicode character
+corresponding to \(\theta\) (\textless U+03B8\textgreater) was replaced
+with `theta'. Common symbols outside of Unicode blocks including `\%'
+(percent) and `\textless{}' (`less-than') were similarly converted into
+plain text. General formatting was removed, this included carriage
+returns, punctuation marks, in-text references (e.g.~``{[}42{]}'')
+centred equations, and other non-ASCII characters. Text contained inside
+brackets was retained to maximise content for analysis, with brackets
+removed.
 
@@
 \citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. The final list
-is provided as Supplementary Material. Plurals (e.g., `chi-squares')
+is provided in Supplementary File 1. Plurals (e.g., `chi-squares')
 unhyphenated (e.g., `chi square') and combined (e.g.~`chisquare') terms
@@
 We applied the clustering algorithm to the cleaned dataset, varying the
-number of clusters from 1 to 50.
-
-Results were transformed to lower case for the clustering, but examples
-are given using the original capitalisation.
+number of clusters from 1 to 50. Clustering solutions were assessed
+using silhouette score, within-cluster dispersion and between-cluster
+dispersion.
 
 For each dataset, records assigned to invididual clusters were examined
-for evidence of boilerplate text in two ways. We first reviewed the top
-ten results that represented the strongest matches to each cluster.
-Records assigned to the same cluster were also compared by calculating
-pairwise cosine similarities; higher scores denoted a higher degree of
-similarity in text between a pair of records.
+for evidence of boilerplate text in two ways. We first reviewed the
+strongest matches to each cluster based on topic scores{[}may not be the
+right word{]}. We also compared similarities in text between pairs of
+records assigned to each cluster. Similarities were determined from
+calculating cosine similarity scores, with comparisons limited to the
+top 500 records per cluster. Higher similarity score represent greater
+overlap in text content between statistical methods sections. Results
+were transformed to lower case for the clustering, but examples are
+given using the original capitalisation.
 
@@
 
-API searches returned 131,847 results (DOIs), of which 111,731 (85\%)
-included a statistical methods section based on our search strategy. In
-the final sample, 95,518 (85\%) DOIs returned an exact match against
-common section headings: 64,133 for `statistical analysis', 13,380 for
-`statistical analyses' and 13,627 for `data analysis'.
-
-Among DOIs that did not meet the partial matching criteria, initial
-search terms appeared in {[}TODO{]}.
+API searches returned 131,847 results (DOIs) based on chosen search
+terms. After partial matching, 111,731 (85\%) statistical methods
+sections were identified. In the final sample, 95,518 (85\%) DOIs
+returned an exact match against common section headings: 64,133 for
+`statistical analysis', 13,380 for `statistical analyses' and 13,627 for
+`data analysis'. For our random sample of DOIs that did not meet the
+partial matching criteria, initial search terms appeared in Introduction
+(number), Results (number) only. xxx appeared in the Methods section
+without a section heading and xxx had a non-standard section heading.
+{[}TODO{]}.
 
@@
 total number of API search results peaked at volumes 8 (n = 19,045) and
-9 (n = 19,045), corresponding to years 2013 and 2014. This trend aligned
-with the total number of papers published in \emph{PLOS ONE} over the
-same period. The percentage of records that included a statistical
-methods section by volume based on our proposed matching criteria varied
-between 64\% (volume 2) and 86\% (volume 9).
+9 (n = 19,045) in years 2013 and 2014. This trend aligned with the total
+number of papers published in \emph{PLOS ONE} over the same period. The
+percentage of records that included a statistical methods section by
+volume based on our proposed matching criteria varied between 64\%
+(volume 2) and 86\% (volume 9).
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/plos.wordclouds}
+{\centering \includegraphics[width=0.8\linewidth]{figures/plos.wordclouds}
 
@@
 
-\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE (left) and ANZCTR (right)}\label{fig:unnamed-chunk-5}
+\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE}\label{fig:unnamed-chunk-5}
 \end{figure}
 
-The topic clouds based on ten clusters are in
-Figure\textasciitilde{}\ref{fig:plos-clouds}. Frequently occurring words
-reflected the use of statistical software (Topics 3 and 5), descriptive
-statistics (Topic 6), group based hypothesis testing (Topics 1 and 4)
-and definitions of statistical significance (Topics\textasciitilde1 and
-9). Statistical methods sections associated with regression (Topic 2)
-and meta-analysis (Topic 7) were also identified.
+The topic clouds based on ten clusters are in Figure
+\ref{fig:plos-clouds}. Frequently occurring words reflected the use of
+statistical software (topics 3 and 5), descriptive statistics (topic 6),
+group based hypothesis testing (topics 1 and 4) and definitions of
+statistical significance (topics 1 and 9). Statistical methods sections
+related to regression (topic 2) and meta-analysis (topic 7) were also
+identified.
 
-\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-6-1.png}
+Topics related to statistical software differentiated between Prism
+GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%)
+. Manual review of results showed that nine out of the ten top matching
+DOIs for topic 3 stated the use of Prism GraphPad, but did not specify
+which statistical methods were used. Top matching sections for topic 5
+included information on SPSS version numbers and definitions of
+statistical significance.
 
-Topics related to the use of statistical software differentiated between
-Prism GraphPad (Topic\textasciitilde3: n = 9,879; 8.8\%) and SPSS (Topic
-5: n = 9,574; 8.6\%) . Manual review of results showed that nine out of
-the ten top matching DOIs for Topic 3 stated the use of Prism GraphPad,
-but did not specify which statistical methods were used. Top matching
-sections for Topic 5 included information on SPSS version numbers and
-definitions of statistical significance. Examples of boilerplate text
-for selected topics are presented in Table
-\ref{tab:plos-example-boilerplate}. Examples correspond to top ranking
-DOIs, and DOIs with the higher number of matches based on cosine
-similarity score greater than 0.8. Further results from cosine
-similarity scores are provided in Supplementary File 2.
+\begin{landscape}
+\begin{table}[]
+\centering
+\caption{Example boilerplate text from PLOS ONE dataset}
+\label{tab:plos-example-boilerplate}
+\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\hline
+\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
+1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
+& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
+\hline
+3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
+ & All statistical analysis was performed using Graphpad Prism software.\\
+ \hline
+4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
+& Data are expressed as the mean ± SEM.
+Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
+P values less than 0.05 were considered significant.\\
+\hline
+5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
+  & The results are presented as the mean ± SEM.
+Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
+The difference was considered statistically significant at P$<$0.05.\\
+  \hline
+6 & All results are expressed as means ± standard deviation (± SD).\\
+& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
+\hline
+9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
+ & Data are presented as the mean ± SD.
+Statistical analysis was performed using the Students unpaired t-test.
+Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
+ \hline
+\end{tabular}
+\end{table}
+\end{landscape}
 
-\includegraphics[width=8.00in,height=4.12in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-7-1.png}
-
-Definitions of statistical significance featured strongly in Topic 1 (n
-= 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for
-comparing differences between two groups. Topic 1 reflected applications
-of Student's t-test assuming a 5\% level of statistical significance.
-Topic 9 referenced similar methods combined with multiple thresholds for
-declaring statistical significance by asterisk: ``\(^{*}p<0.05\),
-\(^{**}p<0.01\) and \(^{***}p<0.001\)''., a practice that has been
-criticised {[}add wasserstein2019{]}.
+Definitions of statistical significance featured strongly in topic 1 (n
+= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%), combined with text
+describing hypothesis testing for comparing differences between two
+groups. Topic 1 reflected applications of Student's t-test assuming a
+5\% level of statistical significance. Topic 9 referenced similar
+methods combined with multiple thresholds for declaring statistical
+significance by asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and
+\(^{***}p<0.001\)'', a practice that has been criticised
+\citep{Wasserstein2019}.
 
@@
 text descriptions varying based on method(s) used. One-way analysis of
-variance also featured strongly in Topic 4 (n = 10,163), with text
-citing common methods for performing post-hoc multiple comparisons. Top
-matching sections across these topics did not include definitions of
-groups being compared.
+variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
+with common methods for performing post-hoc multiple comparisons.
+Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
+mentions of descriptive statistics for summarising continuous variables,
+for example: ``Data are expressed as means ± standard error of the mean
+(SEM).'' Sections assigned to this topic appeared to be expanded
+versions of topics 1 and 9. Examples of boilerplate text were in the
+form of descriptive statistics followed by hypothesis testing, for
+example, Student's t-test, Mann Whitney U or one-way analysis of
+variance.
 
-Frequently occurring words in Topic 6 (n = 4,746) including `Mean', `SD'
-and `SEM' reflected the use of descriptive statistics for summarising
-continuous variables. Sections associated with this topic appeared to be
-expanded versions of Topics 1 and 9, showing evidence of boilerplate
-text in the form of descriptive statistics followed by hypothesis
-testing using Student's t-test, Mann Whitney U or one-way analysis of
-variance.
+Examples of boilerplate text for selected topics are presented in Table
+\ref{tab:plos-example-boilerplate}. Detailed findings based on cosine
+similarity scores are provided in Supplementary File 2.
 
@@
 We downloaded 28,008 studies. The numbers of excluded studies are shown
-in Figure\textasciitilde{}\ref{fig:anzctr-missing}. Of the 12,700
-included studies, 9,523 (75\%) had a statistical methods section.
+in Figure \ref{fig:anzctr-missing}. Of the 12,700 included studies,
+9,523 (75\%) had a statistical methods section. The median length of
+sections was 129 words with an inter-quartile range of 71 to 219 words.
 
@@
 
-\caption{ANZCTR search results.}\label{fig:unnamed-chunk-8}
+\caption{\label{fig:anzctr-missing}ANZCTR search results.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-9}
+\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-8}
 \end{figure}
 
-The median length of the section was 129 words with an inter-quartile
-range of 71 to 219 words. Some methods sections were only one word,
-including ``ANOVA'', ``t-test'', ``SPSS'' and even ``SSPS''.
+The clustering algorithm found groups that were purely sample size
+calculations (topic 2, n = 1834; topic 4, n = 909), pilot studies (topic
+5, n = 834), safety/tolerability studies (topic 6, n = 524),
+intervention studies (topic 8, n = 1020) and repeated measures ANOVA
+(topic 10 = 852). Examples of boilerplate text are provided in Table
+\ref{tab:anzctr-example-boilerplate} and Supplementary File 3.
 
-The clustering algorithm found groups that were purely sample size
-calculations (topic 2), pilot studies (topic 5), safety/tolerability
-studies (topic 6) and repeated measures ANOVA (topic 10). There were
-cases where the exact same method section had been re-used in a
-different study.
+Some methods sections were only one word, including ``ANOVA'',
+``t-test'', ``SPSS'' and even ``SSPS''. There were cases where the exact
+same method section had been re-used in a different study. For example,
+in topic 7 (n = 333), 232 sections stated `descriptive statistics' or
+`descriptive statistics used' with no additional details provided. Other
+instances with expanded descriptions of methods included topic 3 (n =
+1277) outlining descriptive analyses only, and topic 6 (n = 524), for
+hypothesis testing methods, statistical significance and software.
 
-We also found evidence of `boilerplate' sections clustered as topic 3,
-example text
+In other cases, text had been slightly modified to account for changes
+in primary and secondary outcomes. Examples of these text changes were
+found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances
+related to sample size calculations for patient recruitment to different
+studies (see Supplementary File 3).
 
-\textbf{Box x: Examples of boilerplate text from ANZCTR}
-
-Topic 3
-
-\begin{itemize}
-\tightlist
-\item
-  ``Shapiro Wilk test was used as normality test. Continuous variables
-  were compared using Student t-test and Mann-Whitney U test when the
-  data were not normally distributed. Categorical variables were
-  compared using Pearson's chi-squared test and Fisher's exact test.
-  Paired data were analyzed using Paired t-test and Wilcoxon signed rank
-  test when data were not normally distributed.''
-\item
-  ``Comparisons between categorical variables will be made either using
-  chi square or Fisher exact test. Continuous data will be compared
-  using the Student's t-test or Mann-Whitney U test. Two sided p values
-  of less than 0.05 will be considered statistically significant.''
-\end{itemize}
+\begin{landscape}
+\begin{table}[]
+\centering
+\caption{Example boilerplate text from ANZCTR dataset}
+\label{tab:anzctr-example-boilerplate}
+\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\hline
+\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
+3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant.\\
+ &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
+The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate.
+\\ \hline
+5 & Pilot study \\
+ &  No formal sample size calculation was performed \\ \hline
+7 & Descriptive statistics \\
+ &  Descriptive statistics used \\ \hline
+9 & Linear mixed models will be used to analyse the data. \\ \hline
+10 & Repeated measures of ANOVA \\
+ &  Pre-, during, post- and follow-up variables will be subjected to mixed methods and repeated measures analyses to determine significant changes over (group and) time. \\ \hline
+\end{tabular}
+\end{table}
+\end{landscape}
 
@@
 statistics section. The odds ratios and 95\% credible intervals are in
-Table\textasciitilde X. Observational studies were less likely to have a
-missing methods section compared with interventional studies. Missing
-sections became less likely over time. Studies with more funders and a
-larger target sample size were less likely to have a missing methods
-section.
+Table \ref{tab:anzctr-missing-odds}. Observational studies were less
+likely to have a missing methods section compared with interventional
+studies. Missing sections became less likely over time. Studies with
+more funders and a larger target sample size were less likely to have a
+missing methods section.
 
-\includegraphics[width=4.66in,height=1.55in,keepaspectratio]{asa_template_files/figure-latex/unnamed-chunk-10-1.png}
-
-\begin{itemize}
-\tightlist
-\item
-  Final sample size
-\end{itemize}
+\begin{table}[]
+\centering
+\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
+\label{tab:anzctr-missing-odds}
+\begin{tabular}{lcc}
+\hline
+Variable & Odds ratio & 95$\%$ CI \\
+\hline
+Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
+Date (per year) & 0.90 & (0.88, 0.91) \\
+Number of funders & 0.80 & (0.74, 0.86) \\
+Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\section{Supplementary}
-
-\subsection{PLOS}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/supplementary/supp_plosone.50clusters}
-
-}
-
-\caption{PLOS ONE clustering metrics.}\label{fig:unnamed-chunk-11}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/supplementary/plos.topicvalues}
-
-}
-
-\caption{PLOS ONE topic values.}\label{fig:unnamed-chunk-12}
-\end{figure}
-
-\subsection{ANZCTR}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/supplementary/supp_anzctr.50clusters}
-
-}
-
-\caption{ANZCTR clustering metrics.}\label{fig:unnamed-chunk-13}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/supplementary/anzctr.topicvalues}
-
-}
-
-\caption{ANZCTR topic values.}\label{fig:unnamed-chunk-14}
-\end{figure}
-
 \bibliographystyle{agsm}

### 2021-04-23 code/8_manuscript.figures.plos.R

@@ -23,8 +23,8 @@ plot.records = total.records[[2]] %>% select(-percentage) %>%
 g1 = ggplot(plot.records,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
   scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+theme(legend.title = element_blank(),legend.position = 'top',legend.direction = 'horizontal')
 
-g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
-  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count\n(log10 transformed)',breaks=seq(0,4,0.5))
+g2 = ggplot(wordcounts,aes(y=words,x=volume,group=volume))+geom_boxplot()+
+  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count',breaks=seq(0,3500,250))
 
 stats_section_subject = mutate(stats_section_subject,
                                top3 = ifelse(subject_level_number %in% 1:3,1,0))

### 2021-04-24 code/3_extract_title_abstract_xml.R
add script to extract plos titles and abstracts

@@ -1,0 +1,44 @@
+#3_extract_titles_abstracts
+
+library(rplos)
+library(XML)
+library(tidyverse)
+
+#plos full text - e,g, batch 1
+batch.number = 1
+meta_dat = readRDS(paste0('./data/plos_meta_',batch.number,'.RDS'))
+
+doi_list = meta_dat %>% pull(doi)
+dat = list()
+
+for(i in 1:100){
+  #extract xml of full text. if no xml available, skip to the next doi
+  full_text <- tryCatch(
+    plos_fulltext(doi_list[i]),
+    error=function(e) e
+  )
+  
+  if(inherits(full_text, "error")) next
+
+  full_text_xml <- xmlParse(full_text,asText=T,useInternalNodes = T,encoding='UTF-8')
+  methods_text = list()
+  title_abstract = list()
+  node_info.title = getNodeSet(full_text_xml,"//front//title-group//article-title")
+  node_info.abstract = getNodeSet(full_text_xml,"//front//abstract")
+  
+  title_abstract[['title']] = xmlValue(node_info.title)
+  #first try extracting abstract minus section headings (eg aim, methods)
+  title_abstract[['abstract']] = paste(unlist(as.character(sapply(node_info.abstract,xpathSApply,"./sec/p",xmlValue))),collapse=' ')
+  #if nothing returned due to no section headings available, extract abstract as single block of text
+  if (title_abstract[['abstract']]=='list()'){
+    title_abstract[['abstract']] = xmlValue(node_info.abstract)
+  }
+  
+  
+  if (!is.null(title_abstract[['title']]) & !is.null(title_abstract[['abstract']])){
+    dat[[i]] = bind_cols(title_abstract) %>% add_column(doi=doi_list[i]) %>% mutate_if(is.factor,as.character)
+
+  }
+}
+dat = bind_rows(dat)
+save(dat,file=paste0('./data/title_abstract_batch_',batch.number,'.rda'))

### 2021-04-24 code/4_combine_batch_info.R
add code to store excluded plos sections

@@ -8,6 +8,7 @@ analysis_keywords = c('statistical analysis','statistical analyses',
 #loop over batches
 batch = 1:5
 stats_section = list()
+stats_section_excluded = list()
 for (b in batch){
   folder_name = paste0('data/full text records/batch_',b)
   results_by_doi = list.files(folder_name)
@@ -20,6 +21,9 @@ for (b in batch){
   text_results = bind_rows(text_results,.id='doi')
   stats_section[[b]] = text_results %>% filter(grepl(paste(analysis_keywords,collapse='|'),text_heading,
                                        ignore.case=T,fixed=F))
+  stats_section_excluded[[b]] = text_results %>% filter(!grepl(paste(analysis_keywords,collapse='|'),text_heading,
+                                                              ignore.case=T,fixed=F))
 }
 
 save(stats_section,file='data/stats_section_info.rda')
+save(stats_section_excluded,file='data/stats_section_excluded_plos.rda')

### 2021-04-25 code/3_extract_title_abstract_xml.R
amend xml script to extract subject classifications

@@ -4,14 +4,15 @@ library(rplos)
 library(XML)
 library(tidyverse)
 
-#plos full text - e,g, batch 1
-batch.number = 1
-meta_dat = readRDS(paste0('./data/plos_meta_',batch.number,'.RDS'))
-
-doi_list = meta_dat %>% pull(doi)
+#plos full text - loop over batches
 dat = list()
+for (batch.number in 1:5){
+meta_dat = readRDS(paste0('./data/plos_meta_',batch.number,'.RDS'))
+doi_list = meta_dat %>% pull(doi)
 
-for(i in 1:100){
+dat.batch = list()
+
+for(i in 1:10){ #length(doi_list)
   #extract xml of full text. if no xml available, skip to the next doi
   full_text <- tryCatch(
     plos_fulltext(doi_list[i]),
@@ -21,11 +22,17 @@ for(i in 1:100){
   if(inherits(full_text, "error")) next
 
   full_text_xml <- xmlParse(full_text,asText=T,useInternalNodes = T,encoding='UTF-8')
-  methods_text = list()
+  
+  #title and abstract
   title_abstract = list()
   node_info.title = getNodeSet(full_text_xml,"//front//title-group//article-title")
   node_info.abstract = getNodeSet(full_text_xml,"//front//abstract")
   
+  #subject classifications
+  subjects = list()
+  node_info.subjects = getNodeSet(full_text_xml,"//front//article-categories/subj-group")
+    
+    
   title_abstract[['title']] = xmlValue(node_info.title)
   #first try extracting abstract minus section headings (eg aim, methods)
   title_abstract[['abstract']] = paste(unlist(as.character(sapply(node_info.abstract,xpathSApply,"./sec/p",xmlValue))),collapse=' ')
@@ -34,11 +41,28 @@ for(i in 1:100){
     title_abstract[['abstract']] = xmlValue(node_info.abstract)
   }
   
-  
+  #level 1 classification (e..g biology and life sciences)
+  #use as.characters() to force blank cells
+  subjects[['level1']] = unlist(as.character(sapply(node_info.subjects,xpathSApply,"./subject",xmlValue)))
+  subjects[['level2']] = unlist(as.character(sapply(node_info.subjects,xpathSApply,"./subj-group/subject",xmlValue)))
[…]

### 2021-04-26 code/3_extract_title_abstract_xml.R
update xml code for subject classifications

@@ -5,14 +5,15 @@ library(XML)
 library(tidyverse)
 
 #plos full text - loop over batches
-dat = list()
+
 for (batch.number in 1:5){
 meta_dat = readRDS(paste0('./data/plos_meta_',batch.number,'.RDS'))
 doi_list = meta_dat %>% pull(doi)
 
-dat.batch = list()
+dat.batch_titleabs = list()
+dat.batch_subjects = list()
 
-for(i in 1:10){ #length(doi_list)
+for(i in 1:length(doi_list)){
   #extract xml of full text. if no xml available, skip to the next doi
   full_text <- tryCatch(
     plos_fulltext(doi_list[i]),
@@ -48,7 +49,7 @@ for(i in 1:10){ #length(doi_list)
   subjects[['level3']] = unlist(as.character(sapply(node_info.subjects,xpathSApply,"./subj-group/subj-group/subject",xmlValue)))
     
   if (!is.null(title_abstract[['title']]) & !is.null(title_abstract[['abstract']])){
-    dat.batch[[i]] = bind_cols(title_abstract) %>% add_column(doi=doi_list[i],.before = 1) %>% mutate_if(is.factor,as.character)
+    dat.batch_titleabs[[i]] = bind_cols(title_abstract) %>% add_column(doi=doi_list[i],.before = 1) %>% mutate_if(is.factor,as.character)
   }
   
   if (!is.null(subjects[['level1']])){ #must have level 1 subject classification at a minimum
@@ -56,13 +57,22 @@ for(i in 1:10){ #length(doi_list)
       arrange(level1,level2,level3) %>% 
       mutate_at(c("level1","level2","level3"),function(z) str_remove_all(z,'list\\(\\)')) %>% 
       add_column(doi=doi_list[i],.before = 1) %>% mutate_if(is.factor,as.character)
+    #collapse level 3 by ';', then create one line for levels 1-3
+    dat.batch_subjects[[i]] = dat.batch_subjects[[i]] %>% group_by(doi,level1,level2) %>% 
+      summarise(level3=str_c(level3,collapse=';'),.groups='drop') %>%
+      mutate(classifications = paste(level1,level2,level3,sep='/')) %>% select(doi,classifications) 
+    
   }
   
 }
 
-dat.title.abstract[[batch.number]] = bind_rows(dat.batch)
-dat.subject.class[[batch.number]] = bind_rows(dat.batch_subjects)
[…]

### 2021-05-10 code/7_summarise_stats_sections.R
add sentence level code for PLOS datsaat

@@ -2,120 +2,40 @@
 library(tidyverse)
 library(tidytext)
 library(openxlsx)
-
-dat = readRDS('../data/stats_section_cleaned.rds')
+dat = readRDS('data/stats_section_cleaned.rds')
 #ten topics, PLOS ONE
-matches = read.csv('../results/plos_one_10topics.csv',header=T)
-
+matches = read.csv('manuscript/plos_one_10topics.csv',header=T)
 matches = left_join(dat,matches,by=c('doi'='DOI')) %>%
   mutate(value = as.numeric(value)) %>%
   distinct(doi,text_heading,.keep_all = T)
 
-tidy_matches = matches %>%
+ tidy_matches = matches %>%
    unnest_tokens(word,text_data_clean)
  
 wordcounts <- tidy_matches %>%
    group_by(topic_id,doi) %>%
-   summarise(words = n(),.groups='drop')
+   summarise(words = n(),.groups='drop') 
  
 matches = left_join(matches,wordcounts,by=c('doi','topic_id')) %>%
   arrange(topic_id,-value) %>%
   group_by(topic_id) %>% mutate(rank=row_number()) %>% ungroup() %>%
   mutate(topic_id = paste('Topic',topic_id))
 
-# Attached information of subejct classifications
-#take subject classifications from meta-data afind top keywords/classifications per topic
-load('../data/plos_meta_data.rda')
 
-#remove level1 subject classifications common to all records (e.g biology and life sciences)
-meta_dat_allrecords = meta_dat_allrecords %>% mutate(level_1 = paste0('/',
-                                                                      str_replace_all(subject_level_1,pattern=',',replacement = '/|/'),
-                                                                      '/'))
-meta_dat_allrecords = meta_dat_allrecords %>% mutate(subjects = str_remove_all(subject,level_1), 
-                                                     subjects = str_replace_all(subjects,pattern='/',replacement = ','),
-                                                     subjects = str_split(subjects,',')) 
+save(matches,file='manuscript/plos.results.10topics.rda')
 
 
-
-#join with matches
[…]

### 2021-05-13 manuscript/asa_template/asa_template.Rmd
update manuscript methods results

@@
 
-<!-- For records that did not pass this second stage filter, we reviewed where keywords appeared in the full-text to ascertain the extent of missing statistical analysis sections.
-...[TODO: finish this whilst draft is with coauthors]-->
-
 \subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}
@@
 
-_Details to come_
+Let $P \in R^{M \times N}$ denotes the paper content matrix where the  statistical methods sections of $N$ papers consist of $M$ distinct terms.
+Text clustering for identifying common writing themes in these papers requires this matrix to represent with a vector space model.
+Let Matrix $P$ be modeled with the unique terms in the processed sections represented with the tf\*idf (term frequent * inverse document frequency) weighting schema to consider both common and rare terms.
 
-We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
-Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster dispersion.
+Text clustering is known to face the curse of dimensionality due to the high number of terms in doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine].
+Therefore text-based methods based on distance, density or probability face difficulties [@park2018examining,@mohotti2018corpus,@mohotti2019concept].
+Specifically, the distance difference between near and far points becomes negligible in high-dimensional data [@aggarwal2012mining].
+This directly affects the distance-based methods such as \textit{k}-means \cite{jain} in accurately identifying the common subgroups.
+In addition, the sparseness of this high dimensional matrix representation does not allow to differentiate the user groups based on density differences [@mohotti2018corpus, @mohotti2018efficient].
 
-For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways.
-We first reviewed the strongest matches to each cluster based on topic scores[may not be the right word].
+Non-negative Matrix factorization (NMF) which maps the high-dimensional data to a lower-dimensional space has been found to provide an effective solution by allowing to form clusters in the lower-dimensional space [@kim2014algorithms,@luong2019clustering].
+NMF approximates topic vectors in linear manner considering the context of terms.
+
+In traditional NMF [@aggarwal2012mining], the high-dimensional matrix \( P \in R^{M \times N} \) is approximated by learning two factor matrices \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \)  where $g$ is the number of cluster groups or common topics in the data collection, as follows.
+
+\begin{equation}
+\label{eq:2}
+ P \approx WH^{T}
+\end{equation}
+
+The matrix factorization process approximates the lower dimensional non-negative factor matrices $W$ and $H$ such that they can represent high dimensional $P$ with the least error.
+The objective function in Eq \ref{eq:Eq2} iteratively attempts to update the entries in $W$ and $H$ to find the optimum values which possess minimum sum of square error for all the elements in both of those matrices.
+NMF uses Frobenius norm as its objective function and will find the optimum value of \textit{W} and \textit{H} iteratively.
+
+\begin{equation}
+\label{eq:Eq2}
+\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
+\end{equation}
+
+Matrix \textit{H} contains the information regarding topic membership of each document.
+Topic membership of each method section in $P$ is obtained considering the maximum coefficient value in $H$ for a method section.
+
+We applied the NMF clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
+Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster
+dispersion.
+
+\subsection{Within topic analysis}
+
+Results from the clustering algorithm were visualised by word clouds, which summarised frequently occurring terms associated with topic membership.
+For both _PLOS ONE_ and ANZCTR datasets, we reviewed the strongest matches to each cluster based on topic coherence.
+
+Within the _PLOS ONE_ dataset, we used the strongest matches to investigate evidence of boilerplate text, by comparing text with other studies assigned to the same topic.
+Comparisons were carried out by calculating Jaccard similarity scores between studies at the sentence level.
+We chose the Jaccard score as an intuitive measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either the strongest match or the sentence being compared.
+Boilerplate text was defined based on word count (plus or minus 3 words) and a Jaccard score of 0.9 or higher.
+
 We also compared similarities in text between pairs of records assigned to each cluster.

### 2021-05-13 manuscript/asa_template/asa_template.tex
update manuscript methods results

@@
 
-\emph{Details to come}
+Let \(P \in R^{M \times N}\) denotes the paper content matrix where the
+statistical methods sections of \(N\) papers consist of \(M\) distinct
+terms. Text clustering for identifying common writing themes in these
+papers requires this matrix to represent with a vector space model. Let
+Matrix \(P\) be modeled with the unique terms in the processed sections
+represented with the tf*idf (term frequent * inverse document frequency)
+weighting schema to consider both common and rare terms.
 
-We applied the clustering algorithm to the cleaned dataset, varying the
-number of clusters from 1 to 50. Clustering solutions were assessed
+Text clustering is known to face the curse of dimensionality due to the
+high number of terms in doc\(\times\)term matrix representation
+\citep[\citet{sutanto2018fine}]{aggarwal2012mining}. Therefore
+text-based methods based on distance, density or probability face
+difficulties
+\citep[\citet{mohotti2018corpus},\citet{mohotti2019concept}]{park2018examining}.
+Specifically, the distance difference between near and far points
+becomes negligible in high-dimensional data \citep{aggarwal2012mining}.
+This directly affects the distance-based methods such as
+\textit{k}-means \cite{jain} in accurately identifying the common
+subgroups. In addition, the sparseness of this high dimensional matrix
+representation does not allow to differentiate the user groups based on
+density differences
+\citep[\citet{mohotti2018efficient}]{mohotti2018corpus}.
+
+Non-negative Matrix factorization (NMF) which maps the high-dimensional
+data to a lower-dimensional space has been found to provide an effective
+solution by allowing to form clusters in the lower-dimensional space
+\citep[\citet{luong2019clustering}]{kim2014algorithms}. NMF approximates
+topic vectors in linear manner considering the context of terms.
+
+In traditional NMF \citep{aggarwal2012mining}, the high-dimensional
+matrix \(P \in R^{M \times N}\) is approximated by learning two factor
+matrices \(W \in R^{M \times g}\) and \(H \in R ^{N \times g}\) where
+\(g\) is the number of cluster groups or common topics in the data
+collection, as follows.
+
+\begin{equation}
+\label{eq:2}
+ P \approx WH^{T}
+\end{equation}
+
+The matrix factorization process approximates the lower dimensional
+non-negative factor matrices \(W\) and \(H\) such that they can
+represent high dimensional \(P\) with the least error. The objective
+function in Eq \ref{eq:Eq2} iteratively attempts to update the entries
+in \(W\) and \(H\) to find the optimum values which possess minimum sum
+of square error for all the elements in both of those matrices. NMF uses
+Frobenius norm as its objective function and will find the optimum value
+of \textit{W} and \textit{H} iteratively.
+
+\begin{equation}
+\label{eq:Eq2}
+\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
+\end{equation}
+
+Matrix \textit{H} contains the information regarding topic membership of
+each document. Topic membership of each method section in \(P\) is
+obtained considering the maximum coefficient value in \(H\) for a method
+section.
+
+We applied the NMF clustering algorithm to the cleaned dataset, varying
+the number of clusters from 1 to 50. Clustering solutions were assessed
 using silhouette score, within-cluster dispersion and between-cluster
@@
 
-For each dataset, records assigned to invididual clusters were examined
-for evidence of boilerplate text in two ways. We first reviewed the
-strongest matches to each cluster based on topic scores{[}may not be the
-right word{]}. We also compared similarities in text between pairs of
-records assigned to each cluster. Similarities were determined from
-calculating cosine similarity scores, with comparisons limited to the
-top 500 records per cluster. Higher similarity score represent greater
-overlap in text content between statistical methods sections. Results
-were transformed to lower case for the clustering, but examples are
-given using the original capitalisation.
+\subsection{Within topic analysis}
+
+Results from the clustering algorithm were visualised by word clouds,
+which summarised frequently occurring terms associated with topic
+membership. For both \emph{PLOS ONE} and ANZCTR datasets, we reviewed
+the strongest matches to each cluster based on topic coherence.
+
+Within the \emph{PLOS ONE} dataset, we used the strongest matches to
+investigate evidence of boilerplate text, by comparing text with other
+studies assigned to the same topic. Comparisons were carried out by
+calculating Jaccard similarity scores between studies at the sentence
+level. We chose the Jaccard score as an intuitive measure, which
+summarises the similarity between two sentences by the number of words
+common to each sentence (intersection), divided by the number of words
+that appeared in either the strongest match or the sentence being
+compared. Boilerplate text was defined based on word count (plus or
+minus 3 words) and a Jaccard score of 0.9 or higher.
+
+We also compared similarities in text between pairs of records assigned
+to each cluster. Similarities were determined from calculating cosine
+similarity scores, with comparisons limited to the top 500 records per
+cluster. Higher similarity score represent greater overlap in text
+content between statistical methods sections. Results were transformed
+to lower case for the clustering, but examples are given using the
+original capitalisation.
 

### 2021-05-13 manuscript/draft_manuscript.Rmd
update manuscript methods results

@@
 
-_Details to come_
+\subsection{Clustering algorithm}
 
+Let $P \in R^{M \times N}$ denotes the paper content matrix where the  statistical methods sections of $N$ papers consist of $M$ distinct terms.
+Text clustering for identifying common writing themes in these papers requires this matrix to represent with a vector space model.
+Let Matrix $P$ be modeled with the unique terms in the processed sections represented with the tf\*idf (term frequent * inverse document frequency) weighting schema to consider both common and rare terms.
+
+Text clustering is known to face the curse of dimensionality due to the high number of terms in doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine].
+Therefore text-based methods based on distance, density or probability face difficulties [@park2018examining,@mohotti2018corpus,@mohotti2019concept].
+Specifically, the distance difference between near and far points becomes negligible in high-dimensional data [@aggarwal2012mining].
+This directly affects the distance-based methods such as \textit{k}-means \cite{jain} in accurately identifying the common subgroups.
+In addition, the sparseness of this high dimensional matrix representation does not allow to differentiate the user groups based on density differences [@mohotti2018corpus, @mohotti2018efficient].
+
+Non-negative Matrix factorization (NMF) which maps the high-dimensional data to a lower-dimensional space has been found to provide an effective solution by allowing to form clusters in the lower-dimensional space [@kim2014algorithms,@luong2019clustering].
+NMF approximates topic vectors in linear manner considering the context of terms.
+
+In traditional NMF [@aggarwal2012mining], the high-dimensional matrix \( P \in R^{M \times N} \) is approximated by learning two factor matrices \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \)  where $g$ is the number of cluster groups or common topics in the data collection, as follows.
+
+\begin{equation}
+\label{eq:2}
+ P \approx WH^{T} 
+\end{equation}
+
+The matrix factorization process approximates the lower dimensional non-negative factor matrices $W$ and $H$ such that they can represent high dimensional $P$ with the least error.
+The objective function in Eq \ref{eq:Eq2} iteratively attempts to update the entries in $W$ and $H$ to find the optimum values which possess minimum sum of square error for all the elements in both of those matrices.
+NMF uses Frobenius norm as its objective function and will find the optimum value of \textit{W} and \textit{H} iteratively.
+
+\begin{equation}
+\label{eq:Eq2}
+\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
+\end{equation}
+
+Matrix \textit{H} contains the information regarding topic membership of each document.
+Topic membership of each method section in $P$ is obtained considering the maximum coefficient value in $H$ for a method section.
+
+We applied the NMF clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
+Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster
+dispersion.
+
+For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways.
+We first reviewed the strongest matches to each cluster based on topic coherence.
+We also compared similarities in text between pairs of records assigned to each cluster.
+Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster.
+Higher similarity score represent greater overlap in text content between statistical methods sections.
 Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
@@
 
-```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}
+<!-- ```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"} -->
 
-include_graphics("figures/plosone.10topics.jpg")
+<!-- include_graphics("figures/plosone.10topics.jpg") -->
 
-```
+<!-- ``` -->
 

### 2021-05-13 manuscript/draft_manuscript.tex
update manuscript methods results

@@
 \hypersetup{
-  pdftitle={How many researchers use `boilerplate' statistical analysis sections?},
+  pdftitle={On the use of `boilerplate' statistical analysis sections?},
   pdfauthor={Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett},
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx,grffile}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 
-\title{How many researchers use `boilerplate' statistical analysis sections?}
+\title{On the use of `boilerplate' statistical analysis sections?}
 \usepackage{etoolbox}
@@
 \makeatother
-\subtitle{An observational study of papers published in \emph{PLOS ONE}.}
+\subtitle{An observational study of papers published in \emph{PLOS ONE} and studies posted to a trial registry.}
 \author{Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett}
@@
 
-An ideal statistical analysis will use appropriate methods to create
-insights from the data and inform the research questions. Unfortunately
-many current statistical analyses are far from ideal, with many
-researchers using the wrong methods, misinterpreting the results, or
-failing to adequately check their assumptions (Goodman
-\protect\hyperlink{ref-Goodman2008}{2008}; Leek et al.
-\protect\hyperlink{ref-Leek2017}{2017}). Some researchers take a
-``mechanistic'' approach to statistics, copying the few methods they
-know regardless of their appropriateness, and then going through the
-motions of the analysis (Stark and Saltelli
-\protect\hyperlink{ref-Stark2018}{2018}).
+An ideal statistical analysis will use appropriate methods to create insights from the data and inform the research questions.
+Unfortunately many current statistical analyses are far from ideal, with many researchers using the wrong methods, misinterpreting the results, or failing to adequately check their assumptions ({\textbf{???}}; Leek et al.
+\protect\hyperlink{ref-Leek2017}{2017}).
+Some researchers take a ``mechanistic'' approach to statistics, copying the few methods they know regardless of their appropriateness, and then going through the motions of the analysis (Stark and Saltelli \protect\hyperlink{ref-Stark2018}{2018}).
 
-Many researchers lack adequate training in research methods, and
-statistics is something they do with trepidation and even ignorance
-(Altman \protect\hyperlink{ref-Altman1994}{1994}; King et al.
-\protect\hyperlink{ref-King2019}{2019}). However, using the wrong
-statistical methods can cause real harm (Altman
-\protect\hyperlink{ref-Altman1994}{1994}) and bad statistical practices
-are being to used abet weak science (Stark and Saltelli
-\protect\hyperlink{ref-Stark2018}{2018}). Statistical mistakes are a key
-source of waste in research and partly explain the current
-reproducibility crisis in science (Allison et al.
-\protect\hyperlink{ref-Allison2016}{2016}). Even when the correct
-methods are used, many researchers fail to describe them adequately,
-making it difficult to reproduce the results (Ernst and Albers
-\protect\hyperlink{ref-Ernst2017}{2017}; Zhou and Skidmore
-\protect\hyperlink{ref-Zhou2018}{2018}).
+Many researchers lack adequate training in research methods, and statistics is something they do with trepidation and even ignorance (Altman \protect\hyperlink{ref-Altman1994}{1994}; King et al.
+\protect\hyperlink{ref-King2019}{2019}).
+However, using the wrong statistical methods can cause real harm (Altman \protect\hyperlink{ref-Altman1994}{1994}; Brown, Kaiser, and Allison \protect\hyperlink{ref-Brown2018}{2018}) and bad statistical practices are being to used abet weak science (Stark and Saltelli \protect\hyperlink{ref-Stark2018}{2018}).
+Statistical mistakes are a key source of waste in research and partly explain the current reproducibility crisis in science (Allison et al.
+\protect\hyperlink{ref-Allison2016}{2016}).
+Even when the correct methods are used, many researchers fail to describe them adequately, making it difficult to reproduce the results (Ernst and Albers \protect\hyperlink{ref-Ernst2017}{2017}; Zhou and Skidmore \protect\hyperlink{ref-Zhou2018}{2018}).
+Poor statistical methods might not be caught by reviewers, as they may not be qualified to judge the statistics.
+A recent survey of editors found that only 23\% of health and medical journals used expert statistical review for all articles (Hardwicke and Goodman \protect\hyperlink{ref-Hardwicke2020}{2020}), which was little different from a survey from 22 years ago (Goodman, Altman, and George \protect\hyperlink{ref-Goodman1998}{1998}).
 
-The International Committee of Medical Journal Editors recommend that
-researchers should: ``Describe statistical methods with enough detail to
-enable a knowledgeable reader with access to the original data to judge
-its appropriateness for the study and to verify the reported results''
-(ICJME \protect\hyperlink{ref-ICMJE2019}{2019}). Although the general
-lack of statistical understanding from both authors and reviewers means
-this recommendation may not be checked. A recent survey of editors found
-that only 23\% of health and medical journals used expert statistical
-review for all articles (Hardwicke and Goodman
-\protect\hyperlink{ref-Hardwicke2020}{2020}), which was little different
-from a survey from 22 years ago (Goodman, Altman, and George
-\protect\hyperlink{ref-Goodman1998}{1998}).
+There is guidance for researchers on how to write up their statistical methods and results.
+The International Committee of Medical Journal Editors recommend that researchers should: ``Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results'' (ICJME \protect\hyperlink{ref-ICMJE2019}{2019}).
+More detailed guideance is given by the SAMPL and EQUATOR guidelines (Lang and Altman \protect\hyperlink{ref-Lang2013}{2013}; Altman and Simera \protect\hyperlink{ref-Altman2016}{2016}) with the latter covering all apsects of the paper.
+Both of these guidelines were led by Doug Altman, who spoke often and for many years about the need for better statistical reporting.
+The awareness and use of these guidelines could be improved.
+There were 256 Google Scholar citations to the SAMPL paper (as at 15 March 2021) which is a good citation statistic for most papers, but is low considering the millions of papers that use statistical analysis.
 
-Two statisticians on this paper (AB and NW) have heard researchers admit
-that they sometimes copy-and-paste their statistical methods sections
-from other papers, regardless of whether they are appropriate. The aim
-of this paper is to use text-mining methods to estimate the extent that
-researchers are using cut-and-paste or `boilerplate' statistical methods
-sections. Use of these methods sections indicates that little thought
-has gone into the statistical analysis.
+Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate.
+The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or `boilerplate' statistical methods sections.
+Boilerplate text is that ``which can be reused in new contexts or applications without significant changes to the original'' (Wikipedia \protect\hyperlink{ref-Wikipedia}{2021}).
+Use of these methods sections indicates that little thought has gone into the statistical analysis.
 
@@
 
-We used two openly available data sources to find statistical methods
-sections.
+We used two openly available data sources to find statistical methods sections.
 
 \hypertarget{public-library-of-science-plos-one}{%
-\subsubsection{Public Library of Science (PLOS
-ONE)}\label{public-library-of-science-plos-one}}
+\subsubsection{Public Library of Science (PLOS ONE)}\label{public-library-of-science-plos-one}}
 
-\emph{PLOS ONE} is a large open access journal that publishes original
-research across a wide range of scientific fields. Articles must be in
-English. Article submissions are handled by an academic editor who
-selects peer reviewers based on their self-nominated area(s) of
-expertise. Submissions do not undergo formal statistical review.
-Instead, reviewers are required to assess submissions against several
-publication criteria, including whether: ``Experiments, statistics, and
-other analyses are performed to a high technical standard and are
-described in sufficient detail''. Authors are encouraged to follow
-published reporting guidelines to ensure that chosen statistical methods
-are appropriate for the study design, and adequate details are provided
-to enable independent replication of results. Reporting guidelines are
-available from the EQUATOR network which has developed guidelines for
-study designs commonly used in health research, and these guidelines
-were developed to tackle poor statistical application and reporting in
-health and medical journals (Altman and Simera
-\protect\hyperlink{ref-Altman2016}{2016}).
+\emph{PLOS ONE} is a large open access journal that publishes original research across a wide range of scientific fields.
+Articles must be in English.
+Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated areas of expertise.
+Submissions do not undergo formal statistical review.
+Instead, reviewers are required to assess submissions against several publication criteria, including whether: ``Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail'' (PLOS \protect\hyperlink{ref-PLOS}{2021}).
+All reviewers are asked the question: ``Has the statistical analysis been performed appropriately and rigorously?'', with the possible responses of ``Yes'', ``No'' and ``I don't know''.
 
-All \emph{PLOS ONE} articles are freely accessible via the PLOS
-Application Programming Interface (API). This functionality enabled us
-to conduct semi-automated searches of full-text articles and analyse
-data on individual records, including text content and general
-attributes such as publication date and field(s) of research. To find
-papers with a statistical methods section we used targeted API searches
-followed by article filtering based on section headings. The data were
-downloaded on xx-xxxx-xxx.
+Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.
 
-\emph{Step 1}: Targeted API searches. API searches were completed using
-the R package `rplos' (Chamberlain, Boettiger, and Ram
-\protect\hyperlink{ref-rplos}{2020}). Search queries targeted the
-presence of analysis-related terms anywhere in the article. Search terms
-combined the words ``data'' or ``statistical'' with one of:
-``analysis'', ``analyses'', ``method'', ``methodology'' or
-``model(l)ing''. Search terms were intended to be broad whilst keeping
-search results to a manageable number for full-text review (see Step 2).
-By allowing terms to appear anywhere within the article, we accounted
-for the possibility of relevant text being placed in different sections,
-for example, in the \emph{Material and Methods} section versus
-\emph{Results}. Search results were indexed by a unique Digital Object
-Identifier (DOI). Attribute data collected per DOI included journal
-volume, subject classification(s) and total article views since
-publication.
+All \emph{PLOS ONE} articles are freely accessible via the PLOS Application Programming Interface (API).
+This enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research.
+To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings.
+The data were downloaded on 3 July 2020.
 
-\emph{Step 2}: Partial matching on section headings. Full text XML data
-for all search results were downloaded and combined into a single
-dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
-ONE} does not prescribe standardised headings to preface statistical
-methods sections, we performed partial matching on available headings
-against frequently used terms in initial search results: `Statistical
-analysis', `Statistical analyses', `Statistical method', `Statistics',
-`Data analysis' and `Data analyses'. To determine the reliability of our
-chosen filters, we manually reviewed full text data extracted for a
-random sample of XXX articles that were not matched (File
-S1).{[}TODO\ldots finish this thought\ldots{]}
+\emph{Step 1}: Targeted API searches.
+API searches were completed using the R package `rplos' (Chamberlain, Boettiger, and Ram \protect\hyperlink{ref-rplos}{2020}).
+Search queries targeted the presence of analysis-related terms anywhere in the article.
+Search terms combined the words ``data'' or ``statistical'' with one of: ``analysis'', ``analyses'', ``method'', ``methodology'' or ``model(l)ing''.
+Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2).
+By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the \emph{Material and Methods} section versus \emph{Results}.
+Search results were indexed by a unique Digital Object Identifier (DOI).
+Attribute data collected per DOI included journal volume and subject classification(s).
+
+\emph{Step 2}: Partial matching on section headings.
+Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
+Since \emph{PLOS ONE} does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: `Statistical analysis', `Statistical analyses', `Statistical method', `Statistics', `Data analysis' and `Data analyses'.
+To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).{[}TODO\ldots finish this thought\ldots{]}
 
 \hypertarget{australia-and-new-zealand-clinical-trials-registry-anzctr}{%
-\subsubsection{Australia and New Zealand Clinical Trials Registry
-(ANZCTR)}\label{australia-and-new-zealand-clinical-trials-registry-anzctr}}
+\subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}\label{australia-and-new-zealand-clinical-trials-registry-anzctr}}
 
-The ANZCTR was established in 2005 as part of a coordinated global
-effort to improve research quality and transparency in clinical trials
-reporting; observational studies can also be registered. All studies
-registered on ANZCTR are publicly available and can be searched via an
-online portal (\url{https://www.anzctr.org.au/}). Details required for
-registration follow a standardised template (reference or supp file),
-which covers participant eligibility, the intervention(s) being
-evaluated, study design and outcomes. Researchers are asked to provide a
-``brief description'' of the sample size calculations, statistical
-methods and planned analyses, although this section is not compulsory
-(anzctr \protect\hyperlink{ref-ANZCTR}{2019}). The information provided
-must be in English. Studies are reviewed by ANZCTR staff for
-completeness of key information, which does not include the completeness
-of the statistical methods sections. Studies are not peer reviewed.
+The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered.
+All studies registered on ANZCTR are publicly available and can be searched via an online portal (\url{https://www.anzctr.org.au}).
+Details required for registration follow a standardised template (ANZCTR \protect\hyperlink{ref-ANZCTR}{2019}), which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes.
+The information provided must be in English.
+Studies are not peer reviewed.
 
-All studies available on ANZCTR were downloaded on 1 February 2020 in
-XML format. We used all the text available in the ``Statistical
-methods'' section. We also collated basic information about the study
-including the study type (interventional or observational), submission
-date, number of funders and target sample size. These variables were
-chosen as we believed they might influence the completeness of the
-statistical methods section, because we expected larger studies and
-those with funding to be more complete, and we also were interested in
-changes over time. Studies prior to 2013 were excluded as the
-statistical methods section appeared to be introduced in 2013. Some
-studies were first registered on the alternative trial database
-\emph{clinicaltrials.gov} and then also posted to ANZCTR. We excluded
-these studies because they almost all had no completed statistical
-methods section as this requested by \emph{clinicaltrials.gov}.
+For the statistical methods section, researchers are asked to provide a ``brief description'' of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory (ANZCTR \protect\hyperlink{ref-ANZCTR}{2019}).
+Studies are reviewed by ANZCTR staff for completeness of key information, which does not include the completeness of the statistical methods sections.
+
+All studies available on ANZCTR were downloaded on 1 February 2020 in XML format.
+We used all the text available in the ``Statistical methods'' section.
+We also collated basic information about the study including the study type (interventional or observational), submission date, number of funders and target sample size.
+These variables were chosen as we believed they might influence the completeness of the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.
+Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
+Some studies were first registered on the alternative trial database \emph{clinicaltrials.gov} and then also posted to ANZCTR.
+We excluded these studies because they almost all had no completed statistical methods section as this section is not included in \emph{clinicaltrials.gov}.
 
@@
 
-Text cleaning aimed to standardise notation and statistical terminology,
-whilst minimising changes to article style and formatting. Full details
-of text cleaning steps undertaken and corresponding R code are provided
-in Supplementary File X.
+We applied the same text cleaning to both data sources.
+Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
+\emph{R} code used for data extraction and cleaning is available from \url{https://github.com/agbarnett/stats_section}.
 
-Mathematical notation including Greek letters was converted from Unicode
-characters to plain text. For example, the Unicode characters
-corresponding to \(\theta\) (\textless U+03B8\textgreater) was replaced
-with `theta'. Similarly, common symbols outside of Unicode blocks
-including `\%' (percent) and `\textless{}' (`less-than') were converted
-into plain text, using functions available in the `textclean' package
-(Rinker \protect\hyperlink{ref-textclean}{2018}). General formatting was
-removed, this included carriage returns, punctuation marks, in-text
-references (e.g.~``{[}42{]}'') centred equations, and other non-ascii
-characters. Text contained inside brackets was retained in the dataset
-to maximise content for analysis, with brackets removed.
+Mathematical notation including Greek letters was converted from Unicode characters to plain text.
+For example, the Unicode characters corresponding to \(\theta\) (\textless U+03B8\textgreater) were replaced with `theta'.
+Similarly, common symbols outside of Unicode blocks including `\%' (percent) and `\textless{}' (`less-than') were converted into plain text using the `textclean' package (Rinker \protect\hyperlink{ref-textclean}{2018}).
+General formatting was removed, this included carriage returns, punctuation marks, in-text references (e.g.~``{[}42{]}'') centred equations, and other non-ASCII characters.
+Text contained inside brackets was retained to maximise content for analysis, with brackets removed.
 
-We compiled an extensive list of statistical terms to standardise
-descriptions of statistical methods reported across both datasets. An
-initial list was compiled by calculating individual word frequencies and
-identifying relevant terms that appeared at least 100 times. Further
-terms were sourced from index searches of statistics reference textbooks
-{[}ref{]}. The final list is provided as Supplementary Material.
-Possible variants including plurals (e.g.~`chi-squares') unhyphenated
-(e.g `chi square') and combined (e.g.~`chisquare') terms were
-transformed to singular, hyphenated form (e.g `chi-square'). Common
-statistical tests were also hyphenated (e.g.~`hosmer lemeshow' to
-`hosmer-lemeshow').
+We compiled an extensive list of statistical terms to standardise descriptions of statistical methods reported across both datasets.
+An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times.
+Further terms were sourced from index searches of three statistics textbooks (Dobson and Barnett \protect\hyperlink{ref-Dobson2018}{2018}, @Diggle2013, @Bland2015).
+The final list is provided as Supplementary Material.
+Plurals (e.g., `chi-squares') unhyphenated (e.g., `chi square') and combined (e.g.~`chisquare') terms were transformed to singular, hyphenated form (e.g., `chi-square').
+Common statistical tests were also hyphenated (e.g., `hosmer lemeshow' to `hosmer-lemeshow').
 
-As a final step, common stop words including pronouns, contractions and
-selected prepositions were removed. We retained selected stop words
-that, if excluded, may have changed the context of statistical method(s)
-being described, for example `between' and `against'.
+As a final step, common stop words including pronouns, contractions and selected prepositions were removed.
+We retained selected stop words that, if excluded, may have changed the context of statistical methods being described, for example `between' and `against'.
 
@@
 
+\subsection{Clustering algorithm}
+
+Let \(P \in R^{M \times N}\) denotes the paper content matrix where the statistical methods sections of \(N\) papers consist of \(M\) distinct terms.
+Text clustering for identifying common writing themes in these papers requires this matrix to represent with a vector space model.
+Let Matrix \(P\) be modeled with the unique terms in the processed sections represented with the tf*idf (term frequent * inverse document frequency) weighting schema to consider both common and rare terms.
+
+Text clustering is known to face the curse of dimensionality due to the high number of terms in doc\(\times\)term matrix representation ({\textbf{???}}).
+Therefore text-based methods based on distance, density or probability face difficulties ({\textbf{???}}).
+Specifically, the distance difference between near and far points becomes negligible in high-dimensional data ({\textbf{???}}).
+This directly affects the distance-based methods such as \textit{k}-means \cite{jain} in accurately identifying the common subgroups.
+In addition, the sparseness of this high dimensional matrix representation does not allow to differentiate the user groups based on density differences ({\textbf{???}}).
+
+Non-negative Matrix factorization (NMF) which maps the high-dimensional data to a lower-dimensional space has been found to provide an effective solution by allowing to form clusters in the lower-dimensional space ({\textbf{???}}).
+NMF approximates topic vectors in linear manner considering the context of terms.
+
+In traditional NMF ({\textbf{???}}), the high-dimensional matrix \(P \in R^{M \times N}\) is approximated by learning two factor matrices \(W \in R^{M \times g}\) and \(H \in R ^{N \times g}\) where \(g\) is the number of cluster groups or common topics in the data collection, as follows.
+
+\begin{equation}
+\label{eq:2}
+ P \approx WH^{T} 
+\end{equation}
+
+The matrix factorization process approximates the lower dimensional non-negative factor matrices \(W\) and \(H\) such that they can represent high dimensional \(P\) with the least error.
+The objective function in Eq \ref{eq:Eq2} iteratively attempts to update the entries in \(W\) and \(H\) to find the optimum values which possess minimum sum of square error for all the elements in both of those matrices.
+NMF uses Frobenius norm as its objective function and will find the optimum value of \textit{W} and \textit{H} iteratively.
+
+\begin{equation}
+\label{eq:Eq2}
+\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
+\end{equation}
+
+Matrix \textit{H} contains the information regarding topic membership of each document.
+Topic membership of each method section in \(P\) is obtained considering the maximum coefficient value in \(H\) for a method section.
+
+We applied the NMF clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
+Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster
+dispersion.
+
+For each dataset, records assigned to invididual clusters were examined for evidence of boilerplate text in two ways.
+We first reviewed the strongest matches to each cluster based on topic coherence.
+We also compared similarities in text between pairs of records assigned to each cluster.
+Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster.
+Higher similarity score represent greater overlap in text content between statistical methods sections.
+Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
+
 \hypertarget{missing-statistical-methods-sections}{%
-\subsection{Missing statistical methods
-sections}\label{missing-statistical-methods-sections}}
+\subsection{Missing statistical methods sections}\label{missing-statistical-methods-sections}}
 
-The statistical methods section for the ANZCTR data was somtimes missing
-and so we examined if there were particular studies where this section
-was more likely to be missing. We used a logistic regression model
-fitted using a Bayesian paradigm. A small number of sections were
-labelled as ``Not applicable'', ``Nil'' or ``None'' and we changed these
-to missing.
+The statistical methods section for the ANZCTR data was missing for some studies and we examined if there were particular studies where this section was more likely to be missing.
+We used four independent variables of date, study type (observational or interventional), number of funders and target sample size.
+We used a logistic regression model fitted using a Bayesian paradigm. A small number of sections were labelled as ``Not applicable'', ``Nil'' or ``None'' and we changed these to missing.
 
@@
 \hypertarget{plos-one}{%
-\subsection{PLOS ONE}\label{plos-one}}
+\subsection{\texorpdfstring{\emph{PLOS ONE}}{PLOS ONE}}\label{plos-one}}
 
@@
 \includegraphics{draft_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{\label{fig:plos-n}Search results by PLOS ONE volume (1st row);
-word count per statistical methods section included in analysis (n =
-111,731; 2nd row); subject classifications assigned to full-text records
-included in analysis (3rd row)}
+\caption{\label{fig:unnamed-chunk-3}\label{fig:plos-n}Search results by \emph{PLOS ONE} volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)}
 \end{figure}
 
-API searches returned 131847 unique records, of which 111731 (85\%)
-included a statistical methods section based on our search criteria.
-Search results varied by journal volume (Figure 1A). The total number of
-API search results peaked at volumes 8 (n = 19045) and 9 (n = 19045),
-corresponding to years 2013 and 2014. This trend aligned with the total
-number of papers published in PLOS ONE over the same timeframe. The
-percentage of records that included a statistical methods section by
-volume based on our proposed matching criteria varied between 64\%
-(volume 2) and 86\% (volume 9).
+API searches returned 131,847 unique records, of which 111,731 (85\%) included a statistical methods section based on our search criteria.
+In the final sample, 95,518 (85\%) DOIs returned an exact match against common section headings, including 64,133 for `statistical analysis', 13,380 for `statistical analyses' and 13,627 for `data analysis'.
+Among DOIs that did not meet the partial matching criteria, initial search terms appeared in {[}TODO{]}.
 
-The median length of statistical methods sections after text cleaning
-was 127 per DOI (IQR: 61 to 254) (Figure 1B).
+Search results varied by journal volume (Figure \ref{fig:plos-n}A).
+The total number of API search results peaked at volumes 8 (n = 19,045) and 9 (n = 19,045), corresponding to years 2013 and 2014.
+This trend aligned with the total number of papers published in \emph{PLOS ONE} over the same period.
+The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between 64\% (volume 2) and 86\% (volume 9).
 
-Across all records included in analysis, all included Biology and life
-sciences (n = 107584), Earth sciences (n = 7605) and/or Computer and
-information sciences (n = 5190) within their top 3 subject
-classifications (Figure 1C).
+The median length of statistical methods sections was 127 words (IQR: 61 to 253 words) (Figure \ref{fig:plos-n}B).
+7,348 articles (7\%) had a statistical methods section of 500 words or more.
+19,479 articles (17\%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.
 
-{[}Figure here or as supplementary - 5 topic solution{]}
+All papers included Biology and life sciences (n = 107,584), Earth sciences
+(n = 7,605) and/or Computer and information sciences (n = 5,190) in their top 3 subject classifications
+(Figure \ref{fig:plos-n}C).
 
-Results of topic modelling assuming five clusters showed clear
-separation between statistical software, hypothesis testing and linear
-modelling (Figure). The largest cluster, representing 48\% of all
-included studies (Topic 2, n = 53,491).
+We applied the clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
+Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters.
+This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.
 
-Two topics (26\% of study sample) were dominated by terms related
-statistical software, notably Prism/GraphPad (Topic 3, n = 10,940) and
-SPSS (Topic 5, n = 18,029). Remaining topics focussed on hypothesis
-testing (Topic, n = 12,531) and analysis of variance/multiple
-comparisons (Topic 4, n = 16,740)
+\includegraphics[width=3.75in,height=2.75in,keepaspectratio]{draft_manuscript_files/figure-latex/unnamed-chunk-4-1.png}
 
-Top matches (Topic 3)
+The topic clouds based on ten clusters are in Figure\textasciitilde\ref{fig:plot-10-topics}.
+Frequently occurring words reflected the use of statistical software (Topics\textasciitilde3 and 5), descriptive statistics (Topic\textasciitilde6), group based hypothesis testing (Topics 1 and 4) and definitions of statistical significance (Topics\textasciitilde1 and 9).
+There are also statistical methods sections associated with regression (Topic\textasciitilde2) and meta-analysis (Topic\textasciitilde7).
+
+Topics related to the use of statistical software differentiated between Prism GraphPad (Topic\textasciitilde3: n = 9,879; 8.8\%) and SPSS (Topic 5: n = 9,574; 8.6\%) (Box 1).
+A manual review of the top matching sections in these topics showed strong evidence of boilerplate text.
+Nine out ten top matches for Topic 3 stated the use of Prism GraphPad but did not specify which statistical methods were used; six out of ten top matches returned the same cluster score indicating near identical text.
+Top matching sections for Topic 5 included information on SPSS version numbers and definitions of statistical significance.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Box 1: Examples of boilerplate text from \emph{PLOS ONE} (Statistical software)}
+
+Topic 3
 
@@
 \item
-  10.1371/journal.pone.0043519: All statistical analyses were performed
-  using GraphPad Prism (GraphPad Software Inc., San Diego, CA).
+  \emph{graphpad prism (graphpad software, san diego, ca) was used for all analyses.}
 \item
-  10.1371/journal.pone.0045453: GraphPad Prism (Graphpad Software, San
-  Diego, CA) was used for all analyses.
+  \emph{all statistical analysis was performed using the graphpad prism software.}
 \item
-  10.1371/journal.pone.0170640: All statistical analysis of the data was
-  performed using GraphPad Prism software (GraphPad Prism Software Inc.,
-  La Jolla, Ca).
-\item
-  10.1371/journal.pone.0036750: All statistical analysis was performed
-  using Graphpad Prism software.
-\item
-  10.1371/journal.pone.0058148: All statistical analysis was performed
-  using the Graphpad Prism software.
+  \emph{statistical analysis were done using graphpad prism software (graphpad software, inc., la jolla, ca).}
 \end{itemize}
 
-Top matches (Topic 5) * 10.1371/journal.pone.0120046: Statistical
-analysis was performed using the SPSS version 18.0 (SPSS Inc., Chicago,
-IL, USA). Mann-Whitney test was used for the analysis of continuous
-variables. Chi-square test, Fisher's exact test or one-way ANOVA test
-was employed for the analysis of categorical variables. P \textless{}
-0.05 was considered statistically significant.
+Topic 5
 
 \begin{itemize}
+\tightlist
 \item
-  10.1371/journal.pone.0204950: Statistical analyses were performed
-  using SPSS version 20.0 for Windows (SPSS Inc., Chicago, IL).
-  Categorical and continuous variables were evaluated with the
-  non-parametric chi-square test, the Kruskal-Wallis test, and the
-  Mann-Whitney U test. All statistical tests were two-sided, and a p
-  value \textless0.05 was considered statistically significant.
+  \emph{all statistical analyses were conducted in spss 19 (spss inc., chicago, il, usa).}
 \item
-  10.1371/journal.pone.0062685: SPSS version 21.0 for Windows (SPSS,
-  Inc., Chicago, IL, USA) was used for statistical analysis. Categorical
-  variables were analyzed by a chi-square test or Fisher's exact test.
-  Continuous variables were analyzed by independent t-test or
-  Mann-Whitney test. Logistic regression analysis was performed to
-  evaluate the effects of independent variables on clinical outcomes. A
-  p-value of \textless0.05 by the two-tailed test was considered
-  statistically significant.
+  \emph{all statistical analyses were performed using the spss version 15 for windows (spss inc., chicago, il, usa). a p-value less than 0.05 was considered a significant difference.}
 \item
-  10.1371/journal.pone.0133783: Data are expressed as means ± SD or
-  medians (with interquartile ranges). Statistical analyses were
-  performed using SPSS software version 19.0 (SPSS, Inc., Chicago, IL,
-  USA). Statistical significance for intergroup differences was assessed
-  by the χ2 test or Fisher's exact test for categorical variables, and
-  by the Student's t-test, or Mann--Whitney U test for continuous
-  variables. Correlations between variables were tested with Spearman's
-  correlation analysis. A P-value \textless0.05 was considered
-  significant for all tests.
-\item
-  10.1371/journal.pone.0098797: Univariate analysis was performed using
-  Mann--Whitney U-tests for continuous variables and Pearson's
-  chi-square or Fisher's exact tests for categorical variables.
-  Statistical significance was set at a P-value \textless0.05. SPSS
-  statistical software version 19 (SPSS, Chicago, IL, USA) was used to
-  perform the statistical analyses.
+  \emph{statistical analysis was performed using the spss 17.0 software package (spss, chicago, il, usa).
+all data were analyzed using t tests, values of p \textless{} 0.05 were considered significant.}
 \end{itemize}
 
-{[}Figure here or as supplementary - 10 topic solution{]}
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+Definitions of statistical significance featured strongly in Topic\textasciitilde1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups.
+Topic\textasciitilde1 reflected applications of Student's t-test assuming a 5\% level of statistical significance.
+Topic\textasciitilde9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: ``*p\textless0.05, **p\textless0.01 and ***p\textless0.001'', a practice that has been criticised (Wasserstein, Schirm, and Lazar \protect\hyperlink{ref-Wasserstein2019}{2019}).
+Examples of boilerplate text are in Box\textasciitilde2.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Box 2: Examples of boilerplate text from \emph{PLOS ONE} (Between-group hypothesis testing, two groups)}
+
+Topic 1
+
+\begin{itemize}
+\tightlist
+\item
+  \emph{student's t-test was used for statistical analysis. a p value of \textless0.05 was considered statistically significant.}
+\item
+  \emph{statistical analysis was performed using student`s t-test. a p-value \textless0.05 was considered to be statistically significant.}
+\item
+  \emph{statistical analysis was performed using a two-tailed student's t-test, and p-values \textless0.05 were considered statistically significant.}
+\end{itemize}
+
+Topic 9
+
+\begin{itemize}
+\tightlist
+\item
+  \emph{statistical significance was determined by students t-tests. * p\textless0.05, ** p = \textless0.01, *** p\textless0.001.}
+\item
+  \emph{statistical analyses were performed using the nonparametric mann-whitney test.
+differences were considered significant for} *\emph{p\textless0.05,} ** \emph{p\textless0.01, *** }p\textless0.001.*
+\item
+  \emph{comparisons between two groups were performed by student's test. statistical significance was defined as *}p\textless0.05;* ** \emph{p\textless0.01;} *** \emph{p\textless0.001.}
+\end{itemize}
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used (Box 3).
+Frequently occurring words in Topic 6 (n = 4,746) including `mean', `sd' and `sem' reflected the use of descriptive statistics for summarising continuous variables.
+Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance.
+One-way analysis of variance also featured strongly in Topic 4 (n = 10,163), with text citing common methods for performing post-hoc multiple comparisons.
+Top matching sections across these topics did not include definitions of groups being compared.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Box 3: Examples of boilerplate text from \emph{PLOS ONE} (Between-group hypothesis testing, two or more groups)}
+
+Topic 4
+
+\begin{itemize}
+\tightlist
+\item
+  \emph{significant differences were determined using analysis-of-variance anova followed by tukey post-hoc tests for multiple comparisons.}
+\item
+  \emph{statistical comparisons between study groups were performed using one-way anova test followed by post-hoc multiple comparison with dunnett test.
+p-value of less-than 0.05 were considered to be statistically significant.}
+\item
+  \emph{one-way analysis-of-variance anova was used to compare multiple groups followed by bonferroni post-hoc test. statistical significance was set at p less-than 0.05.}
+\end{itemize}
+
+Topic 6
+
+\begin{itemize}
+\tightlist
+\item
+  \emph{all results are expressed as means plus-or-minus standard deviation plus-or-minus sd.}
+\item
+  \emph{data are presented as mean plus-or-minus sd or mean plus-or-minus sem. comparison of means was performed by anova test.}
+\item
+  \emph{data are expressed as means plus-or-minus standard error of the mean sem. statistical analysis was performed using the student t-test. p less-than 0.05 was considered significant.}
+\end{itemize}
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
 
@@
 
-We downloaded 28,008 studies. The numbers of excluded studies are shown
-in Figure\textasciitilde X. Of the 12,700 included studies, 9,523 had a
-statistic section which is 75\% (95\% CI 74\% to 76\%).
+We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure\textasciitilde X.
+Of the 12,700 included studies, 9,523 (75\%) had a statistical methods section.
 
-We examined if four study characteristics were associated with a missing
-statistics section. The odds ratios and 95\% credible intervals are in
-Table\textasciitilde X. Observational studies were less likely to have a
-missing methods section compared with interventional studies. Missing
-sections became less likely over time. Studies with more funders and a
-larger target sample size were less likely to have a missing methods
-section.
+The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
+Some methods sections were only one word, including ``ANOVA'', ``t-test'', ``SPSS'' and even ``SSPS''.
 
-Variable \& Odds ratio \& 95\% CI Study type = Observational \& 0.78 \&
-0.69, 0.89 Date (per year) \& 0.90 \& 0.88, 0.91 Number of funders \&
-0.80 \& 0.74, 0.86 Target sample size (per doubling) \& 0.90 \& 0.88,
-0.92
+The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10).
+There were cases where the exact same method section had been re-used in a different study.
 
-Some of the non-missing statistics sections were only one word,
-including ``ANOVA'', ``t-test'', ``SPSS'' and even ``SSPS''. The median
-length of the section was 129 words with an inter-quartile range of 71
-to 219 words.
+We also found evidence of `boilerplate' sections clustered as topic 3, example text
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Box x: Examples of boilerplate text from ANZCTR}
+
+Topic 3
+
+\begin{itemize}
+\tightlist
+\item
+  ``Shapiro Wilk test was used as normality test.
+Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed.
+Categorical variables were compared using Pearson's chi-squared test and Fisher's exact test.
+Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed.''
+\item
+  ``Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student's t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant.''
+  \_\_\_
+\end{itemize}
+
+We examined if four study characteristics were associated with a missing statistics section.
+The odds ratios and 95\% credible intervals are in Table\textasciitilde X.
+Observational studies were less likely to have a missing methods section compared with interventional studies.
+Missing sections became less likely over time.
+Studies with more funders and a larger target sample size were less likely to have a missing methods section.
+
+\includegraphics[width=4.66in,height=1.55in,keepaspectratio]{draft_manuscript_files/figure-latex/unnamed-chunk-5-1.png}
 
@@
 
-Many researchers are using lazy practice by copying a standard
-``boilerplate'' statistical methods section, likely cut-and-pasting from
-other researchers or projects. This is a strong sign of the ritualistic
-practice of statistics where researchers go through the motions rather
-than using conscientious practice (Stark and Saltelli
-\protect\hyperlink{ref-Stark2018}{2018}). This is concerning because
-using the wrong statistical methods can reduce the potential of study,
-or worse, invalidate the entire study. Poor statistical practice is a
-key driver of the ongoing reproducibility crisis in science (Ioannidis
-et al. \protect\hyperlink{ref-Ioannidis2014}{2014}).
+The first line in many statistical analysis sections in \emph{PLOS ONE} was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail.
+As Doug Altman said, ``Many people think that all you need to do statistics is a computer and appropriate software'' (Altman \protect\hyperlink{ref-Altman1994}{1994}).
+This is far from the truth, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.
 
-The first line in many statistical analysis sections in \emph{PLOS ONE}
-was the software used and some entire sections in ANZCTR only stated the
-software, implying that the software is the most important detail. As
-Doug Altman said, ``Many people think that all you need to do statistics
-is a computer and appropriate software'' (Altman
-\protect\hyperlink{ref-Altman1994}{1994}). This is not the case, and
-whilst it is important for researchers to mention the software and
-version used for reproducibility purposes, it is a minor detail compared
-with detailing what methods were used and why.
+A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5\% level.
+This widespread use of statistical significance is troubling giving the bright-line thinking it engenders (McShane et al.
+\protect\hyperlink{ref-McShane2019}{2019}) and the common misinterpretations of p-values (Goodman \protect\hyperlink{ref-Goodman2008}{2008}).
 
-Despite the extensive array of statistical tests available, many authors
-are reporting the same few methods.
+Despite the extensive array of statistical tests available, many authors are reporting the same few methods.
 
-One reason these inadequate sections get published is that most journals
-do not use statistical reviewers, despite empirical evidence showing
-they improve manuscript quality (Hardwicke and Goodman
-\protect\hyperlink{ref-Hardwicke2020}{2020}).
+One reason these inadequate sections get published is that most journals do not use statistical reviewers, despite empirical evidence showing they improve manuscript quality (Hardwicke and Goodman \protect\hyperlink{ref-Hardwicke2020}{2020}).
+
+A related paper has criticised vague statistical methods sections because they deprive readers and reviewers for the opportunity to confirm that the appropriate methods were used (Weissgerber Tracey et al.
+\protect\hyperlink{ref-Weissgerber2018}{2018}).
+These authors checked hundreds of papers using ANOVA and found that 95\% did not contain the information needed to determine what type of ANOVA was performed.
+This lack of information could well be because the authors used a boilerplate statistical methods section that was missing key details.
+
+If authors shared their code then this would provide an alternative route for checking what statistical methods were used.
+This is not a perfect solution, as we still want authors to accurately report their methods, but it does increase transparency.
+However, a recent paper found that code sharing was very low in biomedical papers, with just 2\% of a sample of over 6,000 papers sharing code (Serghiou \protect\hyperlink{ref-Serghiou2021}{2021}).
+
+Many researchers are using lazy practice by copying a standard ``boilerplate'' statistical methods section, likely cut-and-pasting from other researchers or projects.
+This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice (Stark and Saltelli \protect\hyperlink{ref-Stark2018}{2018}).
+This is concerning because using the wrong statistical methods can reduce the value of study, or worse, invalidate the entire study.
+These mistakes are avoidable and are wasting of thousands of hours of researchers' time and the time of patients and volunteers.
+Poor statistical practice is a key driver of the ongoing reproducibility crisis in science (Ioannidis et al.
+\protect\hyperlink{ref-Ioannidis2014}{2014}).
 
@@
 
-We did not check whether papers used the correct methods, and for some
-simple studies a `boilerplate' statistical methods section would be
-fine.
+We did not check whether papers used the correct methods, and for some simple studies a `boilerplate' statistical methods might be adequate.
 
-We examined papers where there was a statistics section, and we missed
-papers that used statistical analysis but did not include a statistical
-analysis section. Reiterate outcomes of random sample checking here.
+We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section.
+Reiterate outcomes of random sample checking here.
 
-We only examined one journal and one trial registry and hence our
-results may not be generalisable to all journals or registries,
-especially those that consistently use a statistical reviewer.
+We only examined one large journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.
 
-We searched the full text of \emph{PLOS ONE} papers but not the
-supporting information which may contain statistical methods sections
-for some papers. The search terms we used to find statistical methods
-appeared in the supporting information titles for xxx papers (x\%). We
-did not include the supporting information because it is less structured
-than the paper and could be in PDF or Word format.
+We searched the full text of \emph{PLOS ONE} papers but not the supporting information which may contain statistical methods sections for some papers.
+The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x\%).
+We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.
 
@@
 \leavevmode\hypertarget{ref-Allison2016}{}%
-Allison, David B., Andrew W. Brown, Brandon J. George, and Kathryn A.
-Kaiser. 2016. ``Reproducibility: A Tragedy of Errors.'' \emph{Nature}
-530 (7588): 27--29. \url{https://doi.org/10.1038/530027a}.
+Allison, David B., Andrew W. Brown, Brandon J. George, and Kathryn A. Kaiser. 2016. ``Reproducibility: A Tragedy of Errors.'' \emph{Nature} 530 (7588): 27--29. \url{https://doi.org/10.1038/530027a}.
 
 \leavevmode\hypertarget{ref-Altman1994}{}%
-Altman, D G. 1994. ``The Scandal of Poor Medical Research.'' \emph{BMJ}
-308 (6924): 283--84. \url{https://doi.org/10.1136/bmj.308.6924.283}.
+Altman, D G. 1994. ``The Scandal of Poor Medical Research.'' \emph{BMJ} 308 (6924): 283--84. \url{https://doi.org/10.1136/bmj.308.6924.283}.
 
 \leavevmode\hypertarget{ref-Altman2016}{}%
-Altman, Douglas G, and Iveta Simera. 2016. ``A History of the Evolution
-of Guidelines for Reporting Medical Research: The Long Road to the
-EQUATOR Network.'' \emph{Journal of the Royal Society of Medicine} 109
-(2): 67--77. \url{https://doi.org/10.1177/0141076815625599}.
+Altman, Douglas G, and Iveta Simera.
+2016.
+``A History of the Evolution of Guidelines for Reporting Medical Research: The Long Road to the EQUATOR Network.'' \emph{Journal of the Royal Society of Medicine} 109 (2): 67--77.
+\url{https://doi.org/10.1177/0141076815625599}.
 
 \leavevmode\hypertarget{ref-ANZCTR}{}%
-anzctr. 2019. ``ANZCTR Data Field Definitions V25.''
-\url{https://www.anzctr.org.au/docs/ANZCTR\%20Data\%20field\%20explanation.pdf}.
+ANZCTR. 2019. ``ANZCTR Data Field Definitions V25.'' \url{https://www.anzctr.org.au/docs/ANZCTR\%20Data\%20field\%20explanation.pdf}.
+
+\leavevmode\hypertarget{ref-Bland2015}{}%
+Bland, M. 2015. \emph{An Introduction to Medical Statistics}. Oxford Medical Publications. Oxford University Press.
+
+\leavevmode\hypertarget{ref-Brown2018}{}%
+Brown, Andrew W., Kathryn A.
+Kaiser, and David B.
+Allison.
+2018.
+``Issues with Data and Analyses: Errors, Underlying Themes, and Potential Solutions.'' \emph{Proceedings of the National Academy of Sciences} 115 (11): 2563--70.
+\url{https://doi.org/10.1073/pnas.1708279115}.
 
 \leavevmode\hypertarget{ref-rplos}{}%
-Chamberlain, Scott, Carl Boettiger, and Karthik Ram. 2020. \emph{Rplos:
-Interface to the Search API for 'PLoS' Journals}.
-\url{https://CRAN.R-project.org/package=rplos}.
+Chamberlain, Scott, Carl Boettiger, and Karthik Ram. 2020. \emph{Rplos: Interface to the Search API for 'PLoS' Journals}. \url{https://CRAN.R-project.org/package=rplos}.
+
+\leavevmode\hypertarget{ref-Diggle2013}{}%
+Diggle, P., P. Heagerty, K. Y. Liang, and S. Zeger. 2013. \emph{Analysis of Longitudinal Data}. Oxford Statistical Science Series. OUP Oxford.
+
+\leavevmode\hypertarget{ref-Dobson2018}{}%
+Dobson, A. J., and A. G. Barnett. 2018. \emph{An Introduction to Generalized Linear Models}. Chapman \& Hall/Crc Texts in Statistical Science. CRC Press.
 
 \leavevmode\hypertarget{ref-Ernst2017}{}%
-Ernst, Anja F., and Casper J. Albers. 2017. ``Regression Assumptions in
-Clinical Psychology Research Practicea Systematic Review of Common
-Misconceptions.'' \emph{PeerJ} 5: e3323.
+Ernst, Anja F., and Casper J.
+Albers.
+2017.
+``Regression Assumptions in Clinical Psychology Research Practicea Systematic Review of Common Misconceptions.'' \emph{PeerJ} 5: e3323.
 \url{https://doi.org/10.7717/peerj.3323}.
@@
 \leavevmode\hypertarget{ref-Goodman2008}{}%
-Goodman, Steven. 2008. ``A Dirty Dozen: Twelve P-Value Misconceptions.''
-\emph{Seminars in Hematology} 45 (3): 135--40.
-\url{https://doi.org/10.1053/j.seminhematol.2008.04.003}.
+Goodman, Steven. 2008. ``A Dirty Dozen: Twelve P-Value Misconceptions.'' \emph{Seminars in Hematology} 45 (3): 135--40. \url{https://doi.org/10.1053/j.seminhematol.2008.04.003}.
 
 \leavevmode\hypertarget{ref-Goodman1998}{}%
-Goodman, Steven N., Douglas G. Altman, and Stephen L. George. 1998.
-``Statistical Reviewing Policies of Medical Journals.'' \emph{Journal of
-General Internal Medicine} 13 (11): 753--56.
+Goodman, Steven N., Douglas G.
+Altman, and Stephen L.
+George.
+1998.
+``Statistical Reviewing Policies of Medical Journals.'' \emph{Journal of General Internal Medicine} 13 (11): 753--56.
 \url{https://doi.org/10.1046/j.1525-1497.1998.00227.x}.
@@
 \leavevmode\hypertarget{ref-Hardwicke2020}{}%
-Hardwicke, Tom E, and Steve Goodman. 2020. ``How Often Do Leading
-Biomedical Journalsuse Statistical Experts to Evaluate Statistical
-Methods? The Results of a Survey.'' MetaArXiv.
+Hardwicke, Tom E, and Steve Goodman.
+2020.
+``How Often Do Leading Biomedical Journalsuse Statistical Experts to Evaluate Statistical Methods?
+The Results of a Survey.'' MetaArXiv.
 \url{https://doi.org/10.31222/osf.io/z27u4}.
@@
 \leavevmode\hypertarget{ref-ICMJE2019}{}%
-ICJME. 2019. ``Recommendations for the Conduct, Reporting, Editing, and
-Publication of Scholarly Work in Medical Journals.''
-\url{http://www.icmje.org/icmje-recommendations.pdf}.
+ICJME. 2019. ``Recommendations for the Conduct, Reporting, Editing, and Publication of Scholarly Work in Medical Journals.'' \url{http://www.icmje.org/icmje-recommendations.pdf}.
 
 \leavevmode\hypertarget{ref-Ioannidis2014}{}%
-Ioannidis, John P A, Sander Greenland, Mark A Hlatky, Muin J Khoury,
-Malcolm R Macleod, David Moher, Kenneth F Schulz, and Robert Tibshirani.
-2014. ``Increasing Value and Reducing Waste in Research Design, Conduct,
-and Analysis.'' \emph{The Lancet} 383 (9912): 166--75.
+Ioannidis, John P A, Sander Greenland, Mark A Hlatky, Muin J Khoury, Malcolm R Macleod, David Moher, Kenneth F Schulz, and Robert Tibshirani.
+2014.
+``Increasing Value and Reducing Waste in Research Design, Conduct, and Analysis.'' \emph{The Lancet} 383 (9912): 166--75.
 \url{https://doi.org/10.1016/s0140-6736(13)62227-8}.
@@
 \leavevmode\hypertarget{ref-King2019}{}%
-King, Kevin M., Michael D. Pullmann, Aaron R. Lyon, Shannon Dorsey, and
-Cara C. Lewis. 2019. ``Using Implementation Science to Close the Gap
-Between the Optimal and Typical Practice of Quantitative Methods in
-Clinical Science.'' \emph{Journal of Abnormal Psychology} 128 (6):
-547--62. \url{https://doi.org/10.1037/abn0000417}.
+King, Kevin M., Michael D.
+Pullmann, Aaron R.
+Lyon, Shannon Dorsey, and Cara C.
+Lewis.
+2019.
+``Using Implementation Science to Close the Gap Between the Optimal and Typical Practice of Quantitative Methods in Clinical Science.'' \emph{Journal of Abnormal Psychology} 128 (6): 547--62.
+\url{https://doi.org/10.1037/abn0000417}.
+
+\leavevmode\hypertarget{ref-Lang2013}{}%
+Lang, T, and D Altman.
+2013.
+``Basic Statistical Reporting for Articles Published in Clinical Medical Journals: The SAMPL Guidelines.'' In \emph{Science Editors' Handbook}, edited by P Smart, H Maisonneuve, and A Polderman.
+European Association of Science Editors.
 
 \leavevmode\hypertarget{ref-Leek2017}{}%
-Leek, Jeff, Blakeley B. McShane, Andrew Gelman, David Colquhoun, Michèle
-B. Nuijten, and Steven N. Goodman. 2017. ``Five Ways to Fix
-Statistics.'' \emph{Nature} 551 (7682): 557--59.
+Leek, Jeff, Blakeley B.
+McShane, Andrew Gelman, David Colquhoun, Michèle B.
+Nuijten, and Steven N.
+Goodman.
+2017.
+``Five Ways to Fix Statistics.'' \emph{Nature} 551 (7682): 557--59.
 \url{https://doi.org/10.1038/d41586-017-07522-z}.
 
+\leavevmode\hypertarget{ref-McShane2019}{}%
+McShane, Blakeley B., David Gal, Andrew Gelman, Christian Robert, and Jennifer L.
+Tackett.
+2019.
+``Abandon Statistical Significance.'' \emph{The American Statistician} 73 (sup1): 235--45.
+\url{https://doi.org/10.1080/00031305.2018.1527253}.
+
+\leavevmode\hypertarget{ref-PLOS}{}%
+PLOS. 2021. ``PLOS One: Accelerating the Publication of Peer-Reviewed Science.'' \url{https://journals.plos.org/plosone/s/criteria-for-publication}.
+
 \leavevmode\hypertarget{ref-textclean}{}%
-Rinker, Tyler W. 2018. \emph{textclean: Text Cleaning Tools}. Buffalo,
-New York. \url{https://github.com/trinker/textclean}.
+Rinker, Tyler W. 2018. \emph{textclean: Text Cleaning Tools}. Buffalo, New York. \url{https://github.com/trinker/textclean}.
+
+\leavevmode\hypertarget{ref-Serghiou2021}{}%
+Serghiou, Despina G.
+AND Boyack, Stylianos AND Contopoulos-Ioannidis.
+2021.
+``Assessment of Transparency Indicators Across the Biomedical Literature: How Open Is Open?'' \emph{PLOS Biology} 19 (3): 1--26.
+\url{https://doi.org/10.1371/journal.pbio.3001107}.
 
 \leavevmode\hypertarget{ref-Stark2018}{}%
-Stark, Philip B., and Andrea Saltelli. 2018. ``Cargo-Cult Statistics and
-Scientific Crisis.'' \emph{Significance} 15 (4): 40--43.
-\url{https://doi.org/10.1111/j.1740-9713.2018.01174.x}.
+Stark, Philip B., and Andrea Saltelli. 2018. ``Cargo-Cult Statistics and Scientific Crisis.'' \emph{Significance} 15 (4): 40--43. \url{https://doi.org/10.1111/j.1740-9713.2018.01174.x}.
+
+\leavevmode\hypertarget{ref-Wasserstein2019}{}%
+Wasserstein, Ronald L., Allen L.
+Schirm, and Nicole A.
+Lazar.
+2019.
+``Moving to a World Beyond ``P0.167em\(\less\)0.167em0.05''.'' \emph{The American Statistician} 73 (sup1): 1--19.
+\url{https://doi.org/10.1080/00031305.2019.1583913}.
+
+\leavevmode\hypertarget{ref-Weissgerber2018}{}%
+Weissgerber Tracey, L., Oscar Garcia-Valencia, Vesna D.
+Garovic, Natasa M.
+Milic, and Stacey J.
+Winham.
+2018.
+``Why We Need to Report More Than 'Data Were Analyzed by T-Tests or ANOVA'.'' \emph{eLife} 7.
+\url{https://gateway.library.qut.edu.au/login?url=https://search.proquest.com/docview/2174217344?accountid=13380}.
+
+\leavevmode\hypertarget{ref-Wikipedia}{}%
+Wikipedia. 2021. ``Boilerplate Text.'' \url{https://en.wikipedia.org/wiki/Boilerplate_text}.
 
 \leavevmode\hypertarget{ref-Zhou2018}{}%
-Zhou, Yuanyuan, and Susan Skidmore. 2018. ``A Reassessment of ANOVA
-Reporting Practices: A Review of Three APA Journals.'' \emph{Journal of
-Methods and Measurement in the Social Sciences} 8 (1): 3--19.
+Zhou, Yuanyuan, and Susan Skidmore.
+2018.
+``A Reassessment of ANOVA Reporting Practices: A Review of Three APA Journals.'' \emph{Journal of Methods and Measurement in the Social Sciences} 8 (1): 3--19.
 \url{https://doi.org/10.2458/v8i1.22019}.

### 2021-05-13 code/7_summarise_stats_sections.R

@@ -3,8 +3,9 @@ library(tidyverse)
 library(tidytext)
 library(openxlsx)
 dat = readRDS('data/stats_section_cleaned.rds')
+load('data/plos_meta_data.rda')
 #ten topics, PLOS ONE
-matches = read.csv('manuscript/plos_one_10topics.csv',header=T)
+matches = read.csv('results/plos_one_10topics.csv',header=T)
 matches = left_join(dat,matches,by=c('doi'='DOI')) %>%
   mutate(value = as.numeric(value)) %>%
   distinct(doi,text_heading,.keep_all = T)
@@ -21,9 +22,11 @@ matches = left_join(matches,wordcounts,by=c('doi','topic_id')) %>%
   group_by(topic_id) %>% mutate(rank=row_number()) %>% ungroup() %>%
   mutate(topic_id = paste('Topic',topic_id))
 
+#add volume to wordcounts
+wordcounts = wordcounts %>% left_join(select(meta_dat_allrecords,doi,volume),by='doi')
 
 save(matches,file='manuscript/plos.results.10topics.rda')
-
+save(wordcounts,file='manuscript/wordcount.plos.rda')
 
 #example - topics 3 and 5
 matches %>% filter(topic_id %in% c('Topic 1', 'Topic 3','Topic 5')) %>% ggplot(.,aes(x=rank,y=words)) + 

### 2021-05-13 manuscript/asa_template/asa_template.Rmd
update manuscript .tex file

@@
 library(tidytext)
+library(openxlsx)
 
@@
 Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate.
-The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
+The aim of this paper is to use text-mining methods to analyse the content of statistical methods sections included in common research outputs.
+Results are then evaluated to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
 Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [@Wikipedia].
@@
 _PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields.
-Articles must be in English.
 Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated areas of expertise.
@@
 Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
-Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2).
 By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_.
@@
 _Step 2_: Partial matching on section headings.
-Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
+Full text XML data were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
 Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
-For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed.
+
+<!-- For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed.
+-->
 
@@
 
-We applied the same text cleaning steps to both data sources using functions from the `textclean package' [@textclean]. 
+Both datasets were cleaned using functions from the `textclean package' [@textclean]. 
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
@@
 
-Results from the clustering algorithm were visualised by word clouds, which summarised frequently occurring terms associated with topic membership.
-For both _PLOS ONE_ and ANZCTR datasets, we reviewed the strongest matches to each cluster based on topic coherence.
+<!-- Clustering results were visualised by word clouds that summarised frequently occurring terms associated with membership to different topics.  -->
 
-Within the _PLOS ONE_ dataset, we used the strongest matches to investigate evidence of boilerplate text, by comparing text with other studies assigned to the same topic.
+Within the _PLOS ONE_ dataset, statistical methods sections with high topic coherence scores were used to investigate evidence of boilerplate text, compared with other studies assigned to the same topic.
 Comparisons were carried out by calculating Jaccard similarity scores between studies at the sentence level.
@@
 
-We also compared similarities in text between pairs of records assigned to each cluster.
-Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster.
-Higher similarity score represent greater overlap in text content between statistical methods sections.
 Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
@@
 
-\begin{landscape}
-\begin{table}[]
-\centering
-\caption{Example boilerplate text from PLOS ONE dataset}
-\label{tab:plos-example-boilerplate}
-\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
-\hline
-\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
-1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
-& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
-\hline
-3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
- & All statistical analysis was performed using Graphpad Prism software.\\
- \hline
-4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
-& Data are expressed as the mean ± SEM.
+<!-- \begin{landscape} -->
+<!-- \begin{table}[] -->
+<!-- \centering -->
+<!-- \caption{Example boilerplate text from PLOS ONE dataset} -->
+<!-- \label{tab:plos-example-boilerplate} -->
+<!-- \begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}} -->
+<!-- \hline -->
+<!-- \textbf{Topic} & \textbf{Statistical methods section} \\ \hline -->
+<!-- 1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\ -->
+<!-- & Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\ -->
+<!-- \hline -->
+<!-- 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\ -->
+<!--  & All statistical analysis was performed using Graphpad Prism software.\\ -->
+<!--  \hline -->
+<!-- 4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\ -->
+<!-- & Data are expressed as the mean ± SEM.
 Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
-P values less than 0.05 were considered significant.\\
-\hline
-5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
-  & The results are presented as the mean ± SEM.
+P values less than 0.05 were considered significant.\\ -->
+<!-- \hline -->
+<!-- 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\ -->
+<!--   & The results are presented as the mean ± SEM.
 Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
-The difference was considered statistically significant at P$<$0.05.\\
-  \hline
-6 & All results are expressed as means ± standard deviation (± SD).\\
-& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
-\hline
-9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
- & Data are presented as the mean ± SD.
+The difference was considered statistically significant at P$<$0.05.\\ -->
+<!--   \hline -->
+<!-- 6 & All results are expressed as means ± standard deviation (± SD).\\ -->
+<!-- & Data are expressed as mean ± Standard error of the mean (SEM).
+Statistical analysis of the data was performed by the Student’s t-test.
+A value of P$<$0.05 was considered statistically significant.\\ -->
+<!-- \hline -->
+<!-- 9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\ -->
+<!--  & Data are presented as the mean ± SD.
 Statistical analysis was performed using the Students unpaired t-test.
-Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
- \hline
-\end{tabular}
-\end{table}
-\end{landscape}
+Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\ -->
+<!--  \hline -->
+<!-- \end{tabular} -->
+<!-- \end{table} -->
+<!-- \end{landscape} -->
+
+```{r}
+jaccard.matches = read.xlsx('../../results/plos.jaccard_matches.xlsx')
+ftab = jaccard.matches %>% mutate(Topic = paste('Topic',topic)) %>%
+  select(Topic,sentences,matches) %>%  arrange(Topic,-matches) %>%
+  rename('Statistical methods text' = sentences,
+         'Number of matching studies' = matches)
+knitr::kable(ftab,format = "latex", caption='Example boilerplate text from PLOS ONE dataset')
+```
 
@@
 
-Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}. Detailed findings based on cosine similarity scores are provided in Supplementary File 2.
+Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.
 
@@
 The median length of sections was 129 words with an inter-quartile range of 71 to 219 words.
+
+We examined if four study characteristics were associated with a missing statistics section.
+The odds ratios and 95% credible intervals are in Table \ref{tab:anzctr-missing-odds}.
+Observational studies were less likely to have a missing methods section compared with interventional studies.
+Missing sections became less likely over time.
+Studies with more funders and a larger target sample size were less likely to have a missing methods section.
+
+\begin{table}[]
+\centering
+\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
+\label{tab:anzctr-missing-odds}
+\begin{tabular}{lcc}
+\hline
+Variable & Odds ratio & 95$\%$ CI \\
+\hline
+Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
+Date (per year) & 0.90 & (0.88, 0.91) \\
+Number of funders & 0.80 & (0.74, 0.86) \\
+Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 Examples of these text changes were found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances related to sample size calculations for patient recruitment to different studies (see Supplementary File 3).
+
+Finally, targeted searches for the term 'statistician' revealed instances across all topics where no methods were specified.
+Instead, all analyses were the responsibility of an appointed statistician, for example, "A statistician employed by hospital was used" and ``Pilot study at this point will use a statistician professionally to determine sample size calculations as required''.
 
@@
 Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed." -->
-
-We examined if four study characteristics were associated with a missing statistics section.
-The odds ratios and 95% credible intervals are in Table \ref{tab:anzctr-missing-odds}.
-Observational studies were less likely to have a missing methods section compared with interventional studies.
-Missing sections became less likely over time.
-Studies with more funders and a larger target sample size were less likely to have a missing methods section.
-
-\begin{table}[]
-\centering
-\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
-\label{tab:anzctr-missing-odds}
-\begin{tabular}{lcc}
-\hline
-Variable & Odds ratio & 95$\%$ CI \\
-\hline
-Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
-Date (per year) & 0.90 & (0.88, 0.91) \\
-Number of funders & 0.80 & (0.74, 0.86) \\
-Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
-\hline
-\end{tabular}
-\end{table}
 

### 2021-05-13 manuscript/asa_template/asa_template.tex
update manuscript .tex file

@@
 other papers, regardless of whether they are appropriate. The aim of
-this paper is to use text-mining methods to estimate the extent that
-researchers are using cut-and-paste or `boilerplate' statistical methods
-sections. Boilerplate text is that ``which can be reused in new contexts
-or applications without significant changes to the original''
+this paper is to use text-mining methods to analyse the content of
+statistical methods sections included in common research outputs.
+Results are then evaluated to estimate the extent that researchers are
+using cut-and-paste or `boilerplate' statistical methods sections.
+Boilerplate text is that ``which can be reused in new contexts or
+applications without significant changes to the original''
 \citep{Wikipedia}. Use of these methods sections indicates that little
@@
 Registry (ANZCTR). Data sources were chosen as examples of common
-research outputs that include descriptions of statistical methods that
+research outputs that should include descriptions of statistical methods that
 were used, or are planned, for analysing study outcomes.
@@
 \emph{PLOS ONE} is a large open access journal that publishes original
-research across a wide range of scientific fields. Articles must be in
-English. Article submissions are handled by an academic editor who
-selects peer reviewers based on their self-nominated areas of expertise.
-Submissions do not undergo formal statistical review. Instead, reviewers
-are required to assess submissions against several publication criteria,
-including whether: ``Experiments, statistics, and other analyses are
-performed to a high technical standard and are described in sufficient
-detail'' \citep{PLOS}. All reviewers are asked the question: ``Has the
+research across a wide range of scientific fields. Article submissions
+are handled by an academic editor who selects peer reviewers based on
+their self-nominated areas of expertise. Submissions do not undergo
+formal statistical review. Instead, reviewers are required to assess
+submissions against several publication criteria, including whether:
+``Experiments, statistics, and other analyses are performed to a high
+technical standard and are described in sufficient detail''
+\citep{PLOS}. All reviewers are asked the question: ``Has the
 statistical analysis been performed appropriately and rigorously?'',
@@
 
-\emph{Step 1}: Targeted API searches. API searches were completed using
-the R package `rplos' \citep{rplos}. Search queries targeted the
-presence of analysis-related terms anywhere in the article. Search terms
-combined the words ``data'' or ``statistical'' with one of:
+\emph{Step 1: Targeted API searches}. API searches were completed using
+the `rplos' package \citep{rplos}. Search queries targeted analysis-related terms, combining 
+the words ``data'' or ``statistical'' with one of:
 ``analysis'', ``analyses'', ``method'', ``methodology'' or
-``model(l)ing''. Search terms were intended to be broad whilst keeping
-search results to a manageable number for full-text review (see Step 2).
-By allowing terms to appear anywhere in the article, we accounted for
-the possibility of relevant text being placed in different sections, for
-example, in the \emph{Material and Methods} section versus
+``model(l)ing''. We allowed terms to appear anywhere in the article, to allow for the possibility of relevant text being placed in different
+sections, for example, in the \emph{Material and Methods} section versus
 \emph{Results}. Search results were indexed by a unique Digital Object
@@
 
-\emph{Step 2}: Partial matching on section headings. Full text XML data
-for all search results were downloaded and combined into a single
-dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
-ONE} does not prescribe standardised headings to preface statistical
-methods sections, we performed partial matching on available headings
-against frequently used terms in initial search results: `Statistical
-analysis', `Statistical analyses', `Statistical method', `Statistics',
-`Data analysis' and `Data analyses'. For records that did not pass this
-second stage filter, we selected a random sample of XXX records and
-reviewed where initial search terms appeared in the full-text, to
-estimate the likely proportion of statistical methods sections that were
-missed.
+\emph{Step 2: Partial matching on section headings}. Full text XML data
+were downloaded and combined into a single dataset, organised by DOI and
+subsection heading(s). Since \emph{PLOS ONE} does not prescribe
+standardised headings to preface statistical methods sections, we
+performed partial matching on available headings against frequently used
+terms in initial search results: `Statistical analysis', `Statistical
+analyses', `Statistical method', `Statistics', `Data analysis' and `Data
+analyses'.
 
@@
 
-We applied the same text cleaning steps to both data sources using
-functions from the `textclean package' \citep{textclean}. Text cleaning
-aimed to standardise notation and statistical terminology, whilst
-minimising changes to article style and formatting. \emph{R} code used
-for data extraction and cleaning is available from
-\url{https://github.com/agbarnett/stats_section}.
+ Text cleaning aimed to standardise notation and
+statistical terminology, whilst minimising changes to article style and
+formatting. Datasets were cleaned using the `textclean package'
+\citep{textclean}. \emph{R} code used for data extraction and cleaning is
+available from \url{https://github.com/agbarnett/stats_section}.
 
@@
 terms were sourced from index searches of three statistics textbooks
-\citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. The final list
+\citep{Diggle2013,Bland2015,Dobson2018}. The final list
 is provided in Supplementary File 1. Plurals (e.g., `chi-squares')
@@
 high number of terms in doc\(\times\)term matrix representation
-\citep[\citet{sutanto2018fine}]{aggarwal2012mining}. Therefore
+\citep{sutanto2018fine,aggarwal2012mining}. Therefore
 text-based methods based on distance, density or probability face
 difficulties
-\citep[\citet{mohotti2018corpus},\citet{mohotti2019concept}]{park2018examining}.
+\citep{mohotti2018corpus,mohotti2019concept,park2018examining}.
 Specifically, the distance difference between near and far points
@@
 This directly affects the distance-based methods such as
-\textit{k}-means \cite{jain} in accurately identifying the common
+\textit{k}-means \citep{jain} in accurately identifying the common
 subgroups. In addition, the sparseness of this high dimensional matrix
@@
 density differences
-\citep[\citet{mohotti2018efficient}]{mohotti2018corpus}.
+\citep{mohotti2018efficient,mohotti2018corpus}.
 
@@
 solution by allowing to form clusters in the lower-dimensional space
-\citep[\citet{luong2019clustering}]{kim2014algorithms}. NMF approximates
+\citep{luong2019clustering,kim2014algorithms}. NMF approximates
 topic vectors in linear manner considering the context of terms.
@@
 
-Results from the clustering algorithm were visualised by word clouds,
-which summarised frequently occurring terms associated with topic
-membership. For both \emph{PLOS ONE} and ANZCTR datasets, we reviewed
-the strongest matches to each cluster based on topic coherence.
-
-Within the \emph{PLOS ONE} dataset, we used the strongest matches to
-investigate evidence of boilerplate text, by comparing text with other
-studies assigned to the same topic. Comparisons were carried out by
-calculating Jaccard similarity scores between studies at the sentence
-level. We chose the Jaccard score as an intuitive measure, which
-summarises the similarity between two sentences by the number of words
-common to each sentence (intersection), divided by the number of words
-that appeared in either the strongest match or the sentence being
-compared. Boilerplate text was defined based on word count (plus or
-minus 3 words) and a Jaccard score of 0.9 or higher.
-
-We also compared similarities in text between pairs of records assigned
-to each cluster. Similarities were determined from calculating cosine
-similarity scores, with comparisons limited to the top 500 records per
-cluster. Higher similarity score represent greater overlap in text
-content between statistical methods sections. Results were transformed
-to lower case for the clustering, but examples are given using the
-original capitalisation.
+Within the \emph{PLOS ONE} dataset, statistical methods sections with
+high topic coherence scores were used to investigate evidence of
+boilerplate text, compared with other studies assigned to the same
+topic. Comparisons were carried out by calculating Jaccard similarity
+scores between studies at the sentence level. We chose the Jaccard score
+as an intuitive measure, which summarises the similarity between two
+sentences as the number of words common to each sentence (intersection),
+divided by the number of words that appeared in either the strongest
+match or the sentence being compared. Boilerplate text was defined based
+on word count (plus or minus 3 words) and a Jaccard score of 0.9 or
+higher. Results were transformed to lower case for the clustering, but examples
+are given using the original capitalisation.
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{figures/plos.summary}
+{\centering \includegraphics[width=0.8\linewidth]{figures/plossummary.jpg}
 
@@
 
-\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-3}
+\caption{\label{fig:plos-n} A: Search results by PLOS ONE volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{figures/plos.wordclouds}
+{\centering \includegraphics[width=0.8\linewidth]{figures/ploswordclouds.jpg}
 
@@
 \begin{landscape}
-\begin{table}[]
+\begin{table}
+
+\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE dataset.
+Sentences marked with an asterisk $^{\dagger}$ correspond to the statistical methods section with the strongest match based on topic coherence score.}
 \centering
-\caption{Example boilerplate text from PLOS ONE dataset}
-\label{tab:plos-example-boilerplate}
-\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\resizebox{1.15\textwidth}{!}{\begin{tabular}[t]{p{0.05\linewidth} p{0.80\linewidth} p{0.15\linewidth}}
 \hline
-\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
-1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
-& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
+Topic & Statistical methods text & Matching DOIs\\
 \hline
-3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
- & All statistical analysis was performed using Graphpad Prism software.\\
+1  & $^{\dagger}$Students t-test was used for statistical analysis & 29\\
+ & $^{\dagger}$A p value of less-than 0.05 was considered statistically significant. & 1883\\
+\cline{2-3}
+ & Statistical analysis was performed using the students t-test & 380\\
+  & p values of less-than 0.05 were considered significant. & 1526\\
 \hline
-4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
-& Data are expressed as the mean ± SEM.
-Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
-P values less than 0.05 were considered significant.\\
+3 & $^{\dagger}$GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
+\cline{2-3}
+ & All statistical analysis was performed using Graphpad Prism software. & 1312\\
 \hline
-5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
-  & The results are presented as the mean ± SEM.
-Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
-The difference was considered statistically significant at P$<$0.05.\\
+4 & $^{\dagger}$Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons. & 2216\\
+\cline{2-3}
+ & Data are expressed as the mean ± SEM & 860\\
+ & Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test & 4516\\
+ & P values less than 0.05 were considered significant & 2414\\
 \hline
-6 & All results are expressed as means ± standard deviation (± SD).\\
-& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
+5 & $^{\dagger}$SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
+\cline{2-3}
+ & The results are presented as the mean ± SEM. & 420\\
+ & Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA). & 19\\
+ & The difference was considered statistically significant at P$<$0.05. & 2555\\
 \hline
-9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
- & Data are presented as the mean ± SD.
-Statistical analysis was performed using the Students unpaired t-test.
-Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
+6 & $^{\dagger}$All results are expressed as means ± standard deviation (± SD). & 102\\
+\cline{2-3}
+ & Data are expressed as mean ± Standard error of the mean (SEM) & 604\\
+ & Statistical analysis of the data was performed by the Student’s t-test & 444\\
+ & A value of P$<$0.05 was considered statistically significant. & 1786\\
 \hline
-\end{tabular}
+9 & $^{\dagger}$Statistical significance was determined by Students t-tests. & 49\\
+ & $^{\dagger}$$^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$. & 6\\
+\cline{2-3}
+ & Data are presented as the mean ± SD. & 378\\
+ & Statistical analysis was performed using the Students unpaired t-test. & 752\\
+ & Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$. & 413\\
+
+\hline
+\end{tabular}}
 \end{table}
@@
 Group-based hypothesis testing was a recurring theme across topics, with
-text descriptions varying based on method(s) used. One-way analysis of
-variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
-with common methods for performing post-hoc multiple comparisons.
+text varying based on method(s) used. One-way analysis of
+variance featured strongly in topic 4 (n = 10212; 8.8\%), combined
+with common methods for performing post-hoc multiple comparisons, for example, Tukey's test and Dunnett's test.
 Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
@@
 Examples of boilerplate text for selected topics are presented in Table
-\ref{tab:plos-example-boilerplate}. Detailed findings based on cosine
-similarity scores are provided in Supplementary File 2.
+\ref{tab:plos-example-boilerplate}.
 
@@
 
+We examined if four study characteristics were associated with a missing
+statistics section. The odds ratios and 95\% credible intervals are in
+Table \ref{tab:anzctr-missing-odds}. Observational studies were less
+likely to have a missing methods section compared with interventional
+studies. Missing sections became less likely over time. Studies with
+more funders and a larger target sample size were less likely to have a
+missing methods section.
+
+\begin{table}[]
+\centering
+\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
+\label{tab:anzctr-missing-odds}
+\begin{tabular}{lcc}
+\hline
+Variable & Odds ratio & 95$\%$ CI \\
+\hline
+Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
+Date (per year) & 0.90 & (0.88, 0.91) \\
+Number of funders & 0.80 & (0.74, 0.86) \\
+Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
+\hline
+\end{tabular}
+\end{table}
+
 \begin{figure}
@@
 
-\caption{\label{fig:anzctr-missing}ANZCTR search results.}\label{fig:unnamed-chunk-7}
+\caption{\label{fig:anzctr-missing}ANZCTR search results.}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/anzctr.wordclouds}
+{\centering \includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
 
@@
 
-\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-8}
+\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 studies (see Supplementary File 3).
+
+Targeted searches for the term `statistician' revealed
+instances across all topics where no methods were specified. Instead,
+all analyses were the responsibility of an appointed statistician, for
+example, ``A statistician employed by hospital was used'' and ``Pilot
+study at this point will use a statistician professionally to determine
+sample size calculations as required''.
 
@@
 \end{landscape}
-
-We examined if four study characteristics were associated with a missing
-statistics section. The odds ratios and 95\% credible intervals are in
-Table \ref{tab:anzctr-missing-odds}. Observational studies were less
-likely to have a missing methods section compared with interventional
-studies. Missing sections became less likely over time. Studies with
-more funders and a larger target sample size were less likely to have a
-missing methods section.
-
-\begin{table}[]
-\centering
-\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
-\label{tab:anzctr-missing-odds}
-\begin{tabular}{lcc}
-\hline
-Variable & Odds ratio & 95$\%$ CI \\
-\hline
-Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
-Date (per year) & 0.90 & (0.88, 0.91) \\
-Number of funders & 0.80 & (0.74, 0.86) \\
-Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
-\hline
-\end{tabular}
-\end{table}
 

### 2021-05-24 manuscript/asa_template/asa_template.Rmd
update paper add plos flow diagram

@@
 This enabled us to conduct searches of full-text articles and analyse data on articles' text content and general attributes such as publication date and field(s) of research.
-To find papers with a statistical methods section we used targeted API searches followed by article filtering based on section headings.
-The data were downloaded on 3 July 2020.
+Statistical methods sections were extracted using a two-step approach:
 
-_Step 1_: Targeted API searches.
-API searches were completed using the R package ‘rplos’ [@rplos].
-Search queries targeted the presence of analysis-related terms anywhere in the article.
-Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
-By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_.
+_Step 1_: Targeted API searches were completed using the R package ‘rplos’ [@rplos].
+Search queries targeted analysis-related terms, combining the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”.
+Terms could appear anywhere within the main body of the article, to account for the placement of relevant text in different sections, for example, in the _Material and Methods_ section versus _Results_.
 Search results were indexed by a unique Digital Object Identifier (DOI).
@@
 
-_Step 2_: Partial matching on section headings.
-Full text XML data were downloaded and combined into a single dataset, organised by DOI and subsection heading(s).
-Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
+_Step 2_: _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections.
+To address this, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.
 
-<!-- For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed.
--->
+<!-- Full-text XML data were combined into a single dataset organised by DOI and subsection heading(s).  -->
+
+Data included in our analysis were downloaded on 3 July 2020.
+For records that did not pass Step 2, we reviewed where initial search terms appeared in the full-text, to evaluate the likely proportion of statistical methods sections that were missed.
 
@@
 
+The statistical methods section for the ANZCTR data was missing for some studies and we examined if there were particular studies where this section was more likely to be missing. 
+We used four independent variables of date, study type (observational or interventional), number of funders and target sample size. 
+We used a logistic regression model fitted using a Bayesian paradigm. A small number of sections were labelled as "Not applicable", "Nil" or "None" and we changed these to missing.
+
 \subsection{Full-text processing}
@@
 
-Both datasets were cleaned using functions from the `textclean package' [@textclean]. 
+<!-- Datasets were cleaned using functions from the `textclean package' [@textclean].  -->
+
 Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting.
@@
 
-Mathematical notation including Greek letters was converted from Unicode characters to plain text.
+Mathematical notation was converted from Unicode characters to plain text.
 For example, the Unicode character corresponding to $\theta$ (<U+03B8>) was replaced with ‘theta’.
 Common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were similarly converted into plain text.
-General formatting was removed, this included carriage returns, punctuation marks, in-text references (e.g.
+General formatting was removed, including carriage returns, punctuation marks, in-text references (e.g.
 "[42]") centred equations, and other non-ASCII characters.
 Text contained inside brackets was retained to maximise content for analysis, with brackets removed.
+Common stop words including pronouns, contractions and selected prepositions were removed.
+We retained selected stop words that, if excluded, may have changed the context of statistical methods being described, for example ‘between’ and ‘against’.
 
-We compiled an extensive list of statistical terms to standardise descriptions of statistical methods reported across both datasets.
+We compiled an extensive list of statistical terms to standardise reported descriptions of statistical methods.
 An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times.
 Further terms were sourced from index searches of three statistics textbooks [@Dobson2018,@Diggle2013,@Bland2015].
-The final list is provided in Supplementary File 1.
 Plurals (e.g., ‘chi-squares’) unhyphenated (e.g., ‘chi square’) and combined (e.g.
@@
 Common statistical tests were also hyphenated (e.g., ‘hosmer lemeshow’ to ‘hosmer-lemeshow’).
-
-As a final step, common stop words including pronouns, contractions and selected prepositions were removed.
-We retained selected stop words that, if excluded, may have changed the context of statistical methods being described, for example ‘between’ and ‘against’.
-
-<!-- Questions: -->
-
-<!-- * Reference updates to stats reporting guidelines; e.g see: -->
-
-<!-- - https://everyone.plos.org/2019/09/26/new-plos-one-statistical-reporting-guidelines/. -->
-<!-- - https://web.archive.org/web/20190607174803/https://journals.plos.org/plosone/s/submission-guidelines -->
-<!-- - https://web.archive.org/web/20150507175314/https://journals.plos.org/plosone/s/submission-guidelines -->
+The final list is provided in Supplementary File 1.
 
@@
 
-Let $P \in R^{M \times N}$ denotes the paper content matrix where the  statistical methods sections of $N$ papers consist of $M$ distinct terms.
-Text clustering for identifying common writing themes in these papers requires this matrix to represent with a vector space model.
-Let Matrix $P$ be modeled with the unique terms in the processed sections represented with the tf\*idf (term frequent * inverse document frequency) weighting schema to consider both common and rare terms.
+Text from statistical methods sections was analysed using Non-Negative Matrix Factorization (NMF).
+NMF is an established approach for topic modelling, and provides an effective solution for text-based clustering when dealing with high-dimensional data [@kim2014algorithms,@luong2019clustering].
+In this section, we outline key details of NMF and its strengths compared with other text-based clustering methods.
 
-Text clustering is known to face the curse of dimensionality due to the high number of terms in doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine].
-Therefore text-based methods based on distance, density or probability face difficulties [@park2018examining,@mohotti2018corpus,@mohotti2019concept].
-Specifically, the distance difference between near and far points becomes negligible in high-dimensional data [@aggarwal2012mining].
-This directly affects the distance-based methods such as \textit{k}-means \cite{jain} in accurately identifying the common subgroups.
-In addition, the sparseness of this high dimensional matrix representation does not allow to differentiate the user groups based on density differences [@mohotti2018corpus, @mohotti2018efficient].
+For $N$ studies, let $P \in R^{M \times N}$ denote a content matrix, representing text from statistical methods sections as $M$ unique terms.
+Text clustering algorithms for identifying common topics across studies requires $P$ to be represented with a vector space model.
+In our case, unique terms in $P$ are modelled using the tf\*idf (term frequency $\times$ inverse document frequency) weighting schema, to account for the relative importance of common and rare terms.
 
-Non-negative Matrix factorization (NMF) which maps the high-dimensional data to a lower-dimensional space has been found to provide an effective solution by allowing to form clusters in the lower-dimensional space [@kim2014algorithms,@luong2019clustering].
-NMF approximates topic vectors in linear manner considering the context of terms.
+A common problem facing text clustering algorithms is the curse of dimensionality due to the high number of terms in the doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine].
+Applying text-based methods based on distance, density or probability therefore face difficulties in high-dimensional settings [@park2018examining,@mohotti2018corpus,@mohotti2019concept].
+Specifically, distances between near and far points becomes negligible [@aggarwal2012mining].
+This behaviour directly affects the performance of distance-based clustering methods such as \textit{k}-means [@jain] in accurately identifying subgroups (topics) present in the data.
+Furthermore, sparseness associated with high-dimensional matrix representations does not allow for differentiation between topics based on density differences [@mohotti2018corpus, @mohotti2018efficient].
 
-In traditional NMF [@aggarwal2012mining], the high-dimensional matrix \( P \in R^{M \times N} \) is approximated by learning two factor matrices \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \)  where $g$ is the number of cluster groups or common topics in the data collection, as follows.
-
+To address these limitations, NMF deals with high-dimensional data by mapping it to a lower-dimensional space.
+This mapping this achieved by approximating $P$ with two factor matrices: \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \) [@aggarwal2012mining].
+The number of subgroups of common topic inferred from the data is represented by $g$, leading to the approximation:
 \begin{equation}
-\label{eq:2}
- P \approx WH^{T}
+\label{eq:approxPNMF}
+ P \approx WH^{T}.
 \end{equation}
@@
 The matrix factorization process approximates the lower dimensional non-negative factor matrices $W$ and $H$ such that they can represent high dimensional $P$ with the least error.
-The objective function in Eq \ref{eq:Eq2} iteratively attempts to update the entries in $W$ and $H$ to find the optimum values which possess minimum sum of square error for all the elements in both of those matrices.
-NMF uses Frobenius norm as its objective function and will find the optimum value of \textit{W} and \textit{H} iteratively.
+Estimation of $W$ and $H$ is achieved by optimising an objective function; for NMB, the Fronbeius norm is used, equivalent to minimising the sum of squares for all elements of $P$:
 
 \begin{equation}
-\label{eq:Eq2}
+\label{eq:NMFobjectivefn}
 \min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
@@
 
-Matrix \textit{H} contains the information regarding topic membership of each document.
-Topic membership of each method section in $P$ is obtained considering the maximum coefficient value in $H$ for a method section.
+Following estimation, \textit{H} contains the information regarding topic membership for all studies.
+In our case, topic membership $(1,\ldots,g)$ for a statistical methods section is inferred from the maximum coefficient value in the corresponding row of $H$.
 
-We applied the NMF clustering algorithm to the cleaned dataset, varying the number of clusters from 1 to 50.
-Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster
-dispersion.
+For our two datasets, we applied NMF varying $g$ from 1 to 50. Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster dispersion.
 
@@
 
-\subsection{Missing statistical methods sections}
-
-The statistical methods section for the ANZCTR data was missing for some studies and we examined if there were particular studies where this section was more likely to be missing. 
-We used four independent variables of date, study type (observational or interventional), number of funders and target sample size. 
-We used a logistic regression model fitted using a Bayesian paradigm. A small number of sections were labelled as "Not applicable", "Nil" or "None" and we changed these to missing.
+<!-- \subsection{Missing statistical methods sections} -->
 
@@
 
-## _PLOS ONE_
+## Included studies
+
+### PLOS ONE
 
@@
 
-knitr::include_graphics("figures/plos.summary.jpg")
+knitr::include_graphics("figures/plos_summary.jpg")
 ```
@@
 
-API searches returned `r format(n.api,big.mark=',')` results (DOIs) based on chosen search terms.
+Searching with the _PLOS ONE_ API returned `r format(n.api,big.mark=',')` results (DOIs) based on chosen search terms (\autoref{fig:plos-flowchart}).
 After partial matching, `r format(n.include,big.mark=',')` (`r perc.include`%) statistical methods sections were identified.
@@
 
-Statistical methods sections had a median length of `r n.words[['med']]` words and inter-quartile range of `r n.words[['q1']]` to `r n.words[['q3']]` words (Figure \ref{fig:plos-n}B).
+All DOIs included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
+(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications (Figure \ref{fig:plos-n}B).
+
+Statistical methods sections had a median length of `r n.words[['med']]` words and inter-quartile range of `r n.words[['q1']]` to `r n.words[['q3']]` words.
 `r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more.
-`r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had sections with 50 words or less, equal to the length of this paragraph.
-
-All DOIs included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
-(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications (Figure \ref{fig:plos-n}C).
+`r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had statistical methods sections with 50 words or less, equal to the length of this paragraph.
 
@@
 
-knitr::include_graphics("figures/plos.wordclouds.jpg")
+knitr::include_graphics("figures/ploswordclouds.jpg")
 
@@
 
-knitr::include_graphics("figures/anzctr.wordclouds.jpg")
+knitr::include_graphics("figures/anzctrwordclouds.jpg")
 

### 2021-05-24 manuscript/asa_template/asa_template.tex
update paper add plos flow diagram

--- 2021-04-23 manuscript/asa_template/asa_template.tex
@@
 content and general attributes such as publication date and field(s) of
-research. To find papers with a statistical methods section we used
-targeted API searches followed by article filtering based on section
-headings. The data were downloaded on 3 July 2020.
+research. Statistical methods sections were extracted using a two-step
+approach:
 
-\emph{Step 1}: Targeted API searches. API searches were completed using
-the R package `rplos' \citep{rplos}. Search queries targeted the
-presence of analysis-related terms anywhere in the article. Search terms
-combined the words ``data'' or ``statistical'' with one of:
+\emph{Step 1}: Targeted API searches were completed using the R package
+`rplos' \citep{rplos}. Search queries targeted analysis-related terms,
+combining the words ``data'' or ``statistical'' with one of:
 ``analysis'', ``analyses'', ``method'', ``methodology'' or
-``model(l)ing''. Search terms were intended to be broad whilst keeping
-search results to a manageable number for full-text review (see Step 2).
-By allowing terms to appear anywhere in the article, we accounted for
-the possibility of relevant text being placed in different sections, for
-example, in the \emph{Material and Methods} section versus
+``model(l)ing''. Terms could appear anywhere within the main body of the
+article, to account for the placement of relevant text in different
+sections, for example, in the \emph{Material and Methods} section versus
 \emph{Results}. Search results were indexed by a unique Digital Object
@@
 
-\emph{Step 2}: Partial matching on section headings. Full text XML data
-for all search results were downloaded and combined into a single
-dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
-ONE} does not prescribe standardised headings to preface statistical
-methods sections, we performed partial matching on available headings
-against frequently used terms in initial search results: `Statistical
-analysis', `Statistical analyses', `Statistical method', `Statistics',
-`Data analysis' and `Data analyses'. For records that did not pass this
-second stage filter, we selected a random sample of XXX records and
-reviewed where initial search terms appeared in the full-text, to
-estimate the likely proportion of statistical methods sections that were
-missed.
+\emph{Step 2}: \emph{PLOS ONE} does not prescribe standardised headings
+to preface statistical methods sections. To address this, we performed
+partial matching on available headings against frequently used terms in
+initial search results: `Statistical analysis', `Statistical analyses',
+`Statistical method', `Statistics', `Data analysis' and `Data analyses'.
+
+Data included in our analysis were downloaded on 3 July 2020. For
+records that did not pass Step 2, we reviewed where initial search terms
+appeared in the full-text, to evaluate the likely proportion of
+statistical methods sections that were missed.
 
@@
 
-\subsection{Full-text processing}
-\label{sec:methods-cleaning}
-
-We applied the same text cleaning steps to both data sources using
-functions from the `textclean package' \citep{textclean}. Text cleaning
-aimed to standardise notation and statistical terminology, whilst
-minimising changes to article style and formatting. \emph{R} code used
-for data extraction and cleaning is available from
-\url{https://github.com/agbarnett/stats_section}.
-
-Mathematical notation including Greek letters was converted from Unicode
-characters to plain text. For example, the Unicode character
-corresponding to \(\theta\) (\textless U+03B8\textgreater) was replaced
-with `theta'. Common symbols outside of Unicode blocks including `\%'
-(percent) and `\textless{}' (`less-than') were similarly converted into
-plain text. General formatting was removed, this included carriage
-returns, punctuation marks, in-text references (e.g.~``{[}42{]}'')
-centred equations, and other non-ASCII characters. Text contained inside
-brackets was retained to maximise content for analysis, with brackets
-removed.
-
-We compiled an extensive list of statistical terms to standardise
-descriptions of statistical methods reported across both datasets. An
-initial list was compiled by calculating individual word frequencies and
-identifying relevant terms that appeared at least 100 times. Further
-terms were sourced from index searches of three statistics textbooks
-\citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. The final list
-is provided in Supplementary File 1. Plurals (e.g., `chi-squares')
-unhyphenated (e.g., `chi square') and combined (e.g.~`chisquare') terms
-were transformed to singular, hyphenated form (e.g., `chi-square').
-Common statistical tests were also hyphenated (e.g., `hosmer lemeshow'
-to `hosmer-lemeshow').
-
-As a final step, common stop words including pronouns, contractions and
-selected prepositions were removed. We retained selected stop words
-that, if excluded, may have changed the context of statistical methods
-being described, for example `between' and `against'.
-
-\subsection{Clustering algorithm}
-
-\emph{Details to come}
-
-We applied the clustering algorithm to the cleaned dataset, varying the
-number of clusters from 1 to 50. Clustering solutions were assessed
-using silhouette score, within-cluster dispersion and between-cluster
-dispersion.
-
-For each dataset, records assigned to invididual clusters were examined
-for evidence of boilerplate text in two ways. We first reviewed the
-strongest matches to each cluster based on topic scores{[}may not be the
-right word{]}. We also compared similarities in text between pairs of
-records assigned to each cluster. Similarities were determined from
-calculating cosine similarity scores, with comparisons limited to the
-top 500 records per cluster. Higher similarity score represent greater
-overlap in text content between statistical methods sections. Results
-were transformed to lower case for the clustering, but examples are
-given using the original capitalisation.
-
-\subsection{Missing statistical methods sections}
-
 The statistical methods section for the ANZCTR data was missing for some
@@
 
+\subsection{Full-text processing}
+\label{sec:methods-cleaning}
+
+Text cleaning aimed to standardise notation and statistical terminology,
+whilst minimising changes to article style and formatting. \emph{R} code
+used for data extraction and cleaning is available from
+\url{https://github.com/agbarnett/stats_section}.
+
+Mathematical notation was converted from Unicode characters to plain
+text. For example, the Unicode character corresponding to \(\theta\)
+(\textless U+03B8\textgreater) was replaced with `theta'. Common symbols
+outside of Unicode blocks including `\%' (percent) and `\textless{}'
+(`less-than') were similarly converted into plain text. General
+formatting was removed, including carriage returns, punctuation marks,
+in-text references (e.g.~``{[}42{]}'') centred equations, and other
+non-ASCII characters. Text contained inside brackets was retained to
+maximise content for analysis, with brackets removed. Common stop words
+including pronouns, contractions and selected prepositions were removed.
+We retained selected stop words that, if excluded, may have changed the
+context of statistical methods being described, for example `between'
+and `against'.
+
+We compiled an extensive list of statistical terms to standardise
+reported descriptions of statistical methods. An initial list was
+compiled by calculating individual word frequencies and identifying
+relevant terms that appeared at least 100 times. Further terms were
+sourced from index searches of three statistics textbooks
+\citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. Plurals (e.g.,
+`chi-squares') unhyphenated (e.g., `chi square') and combined
+(e.g.~`chisquare') terms were transformed to singular, hyphenated form
+(e.g., `chi-square'). Common statistical tests were also hyphenated
+(e.g., `hosmer lemeshow' to `hosmer-lemeshow'). The final list is
+provided in Supplementary File 1.
+
+\subsection{Clustering algorithm}
+
+Text from statistical methods sections was analysed using Non-Negative
+Matrix Factorization (NMF). NMF is an established approach for topic
+modelling, and provides an effective solution for text-based clustering
+when dealing with high-dimensional data
+\citep[\citet{luong2019clustering}]{kim2014algorithms}. In this section,
+we outline key details of NMF and its strengths compared with other
+text-based clustering methods.
+
+For \(N\) studies, let \(P \in R^{M \times N}\) denote a content matrix,
+representing text from statistical methods sections as \(M\) unique
+terms. Text clustering algorithms for identifying common topics across
+studies requires \(P\) to be represented with a vector space model. In
+our case, unique terms in \(P\) are modelled using the tf*idf (term
+frequency \(\times\) inverse document frequency) weighting schema, to
+account for the relative importance of common and rare terms.
+
+A common problem facing text clustering algorithms is the curse of
+dimensionality due to the high number of terms in the doc\(\times\)term
+matrix representation
+\citep[\citet{sutanto2018fine}]{aggarwal2012mining}. Applying text-based
+methods based on distance, density or probability therefore face
+difficulties in high-dimensional settings
+\citep[\citet{mohotti2018corpus},\citet{mohotti2019concept}]{park2018examining}.
+Specifically, distances between near and far points becomes negligible
+\citep{aggarwal2012mining}. This behaviour directly affects the
+performance of distance-based clustering methods such as
+\textit{k}-means \citep{jain} in accurately identifying subgroups
+(topics) present in the data. Furthermore, sparseness associated with
+high-dimensional matrix representations does not allow for
+differentiation between topics based on density differences
+\citep[\citet{mohotti2018efficient}]{mohotti2018corpus}.
+
+To address these limitations, NMF deals with high-dimensional data by
+mapping it to a lower-dimensional space. This mapping this achieved by
+approximating \(P\) with two factor matrices: \(W \in R^{M \times g}\)
+and \(H \in R ^{N \times g}\) \citep{aggarwal2012mining}. The number of
+subgroups of common topic inferred from the data is represented by
+\(g\), leading to the approximation: \begin{equation}
+\label{eq:approxPNMF}
+ P \approx WH^{T}.
+\end{equation}
+
+The matrix factorization process approximates the lower dimensional
+non-negative factor matrices \(W\) and \(H\) such that they can
+represent high dimensional \(P\) with the least error. Estimation of
+\(W\) and \(H\) is achieved by optimising an objective function; for
+NMB, the Fronbeius norm is used, equivalent to minimising the sum of
+squares for all elements of \(P\):
+
+\begin{equation}
+\label{eq:NMFobjectivefn}
+\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
+\end{equation}
+
+Following estimation, \textit{H} contains the information regarding
+topic membership for all studies. In our case, topic membership
+\((1,\ldots,g)\) for a statistical methods section is inferred from the
+maximum coefficient value in the corresponding row of \(H\).
+
+For our two datasets, we applied NMF varying \(g\) from 1 to 50.
+Clustering solutions were assessed using silhouette score,
+within-cluster dispersion and between-cluster dispersion.
+
+\subsection{Within topic analysis}
+
+Within the \emph{PLOS ONE} dataset, statistical methods sections with
+high topic coherence scores were used to investigate evidence of
+boilerplate text, compared with other studies assigned to the same
+topic. Comparisons were carried out by calculating Jaccard similarity
+scores between studies at the sentence level. We chose the Jaccard score
+as an intuitive measure, which summarises the similarity between two
+sentences by the number of words common to each sentence (intersection),
+divided by the number of words that appeared in either the strongest
+match or the sentence being compared. Boilerplate text was defined based
+on word count (plus or minus 3 words) and a Jaccard score of 0.9 or
+higher.
+
+Results were transformed to lower case for the clustering, but examples
+are given using the original capitalisation.
+
 \hypertarget{results}{%
@@
 
+\hypertarget{included-studies}{%
+\subsection{Included studies}\label{included-studies}}
+
 \hypertarget{plos-one}{%
-\subsection{\texorpdfstring{\emph{PLOS ONE}}{PLOS ONE}}\label{plos-one}}
+\subsubsection{PLOS ONE}\label{plos-one}}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{figures/plos.summary}
+{\centering \includegraphics[width=0.6\linewidth]{figures/plos_summary}
 
@@
 
-API searches returned 131,847 results (DOIs) based on chosen search
-terms. After partial matching, 111,731 (85\%) statistical methods
-sections were identified. In the final sample, 95,518 (85\%) DOIs
-returned an exact match against common section headings: 64,133 for
-`statistical analysis', 13,380 for `statistical analyses' and 13,627 for
-`data analysis'. For our random sample of DOIs that did not meet the
-partial matching criteria, initial search terms appeared in Introduction
+Searching with the \emph{PLOS ONE} API returned 131,847 results (DOIs)
+based on chosen search terms (\autoref{fig:plos-flowchart}). After
+partial matching, 111,731 (85\%) statistical methods sections were
+identified. In the final sample, 95,518 (85\%) DOIs returned an exact
+match against common section headings: 64,133 for `statistical
+analysis', 13,380 for `statistical analyses' and 13,627 for `data
+analysis'. For our random sample of DOIs that did not meet the partial
+matching criteria, initial search terms appeared in Introduction
 (number), Results (number) only. xxx appeared in the Methods section
@@
 
-Statistical methods sections had a median length of 127 words and
-inter-quartile range of 61 to 254 words (Figure \ref{fig:plos-n}B).
-7,450 articles (7\%) had a statistical methods section of 500 words or
-more. 19,461 articles (17\%) had sections with 50 words or less, equal
-to the length of this paragraph.
-
 All DOIs included Biology and life sciences (n = 107,584), Earth
@@
 5,190) in their top 3 subject classifications (Figure
-\ref{fig:plos-n}C).
+\ref{fig:plos-n}B).
+
+Statistical methods sections had a median length of 127 words and
+inter-quartile range of 61 to 254 words. 7,450 articles (7\%) had a
+statistical methods section of 500 words or more. 19,461 articles (17\%)
+had statistical methods sections with 50 words or less, equal to the
+length of this paragraph.
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{figures/plos.wordclouds}
+{\centering \includegraphics[width=0.8\linewidth]{figures/ploswordclouds}
 
@@
 
-\begin{landscape}
-\begin{table}[]
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-7}Example boilerplate text from PLOS ONE dataset}
 \centering
-\caption{Example boilerplate text from PLOS ONE dataset}
-\label{tab:plos-example-boilerplate}
-\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\begin{tabular}[t]{l|l|r}
 \hline
-\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
-1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
-& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
+Topic & Statistical methods text & Number of matching studies\\
 \hline
-3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
- & All statistical analysis was performed using Graphpad Prism software.\\
+Topic 1 & a p value of less-than 0.05 was considered statistically significant. & 1883\\
 \hline
-4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
-& Data are expressed as the mean ± SEM.
-Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test.
-P values less than 0.05 were considered significant.\\
+Topic 1 & p values of less-than 0.05 were considered significant. & 1526\\
 \hline
-5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
-  & The results are presented as the mean ± SEM.
-Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA).
-The difference was considered statistically significant at P$<$0.05.\\
+Topic 1 & statistical analysis was performed using the students t-test & 380\\
 \hline
-6 & All results are expressed as means ± standard deviation (± SD).\\
-& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
+Topic 1 & students t-test was used for statistical analysis & 29\\
 \hline
-9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
- & Data are presented as the mean ± SD.
-Statistical analysis was performed using the Students unpaired t-test.
-Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
+Topic 3 & graphpad prism graphpad software san diego ca was used for all analyses. & 1941\\
+\hline
+Topic 3 & all statistical analysis was performed using graphpad prism software. & 1312\\
+\hline
+Topic 4 & statistical analysis was performed using one-way analysis of variance anova test followed by dunnetts multiple comparison post hoc test & 4516\\
+\hline
+Topic 4 & p values less-than 0.05 were considered significant. & 2414\\
+\hline
+Topic 4 & significant differences were determined using analysis of variance anova followed by tukey post-hoc tests for multiple comparisons. & 2216\\
+\hline
+Topic 4 & data are expressed as the mean plus-or-minus sem & 860\\
+\hline
+Topic 5 & the difference was considered statistically significant at p less-than 0.05. & 2555\\
+\hline
+Topic 5 & spss software version 17.0 spss chicago il usa was used for statistical analysis. & 441\\
+\hline
+Topic 5 & the results are presented as the mean plus-or-minus sem & 420\\
+\hline
+Topic 5 & data were analyzed by one-way anova and lsd tests using the spss 13.0 software spss inc chicago il usa & 19\\
+\hline
+Topic 6 & a value of p less-than 0.05 was considered statistically significant. & 1786\\
+\hline
+Topic 6 & data are expressed as mean plus-or-minus standard error of the mean sem & 604\\
+\hline
+Topic 6 & statistical analysis of the data was performed by the students t-test & 444\\
+\hline
+Topic 6 & all results are expressed as means plus-or-minus standard deviation plus-or-minus sd. & 102\\
+\hline
+Topic 9 & statistical analysis was performed using the students unpaired t-test & 752\\
+\hline
+Topic 9 & differences were considered statistically significant at p less-than 0.05 p less-than 0.01 and p less-than 0.001. & 413\\
+\hline
+Topic 9 & data are presented as the mean plus-or-minus sd & 378\\
+\hline
+Topic 9 & statistical significance was determined by students t-tests & 49\\
+\hline
+Topic 9 & p less-than 0.05 p equal-to less-than 0.01 p less-than 0.001. & 6\\
 \hline
@@
 \end{table}
-\end{landscape}
 
@@
 studies (see Supplementary File 3).
+
+Finally, targeted searches for the term `statistician' revealed
+instances across all topics where no methods were specified. Instead,
+all analyses were the responsibility of an appointed statistician, for
+example, ``A statistician employed by hospital was used'' and ``Pilot
+study at this point will use a statistician professionally to determine
+sample size calculations as required''.
 

### 2021-05-24 code/3_plot_excluded.plos.R

@@ -1,0 +1,41 @@
+#3_plot_excluded
+# CONSORT style diagram of included/excluded studies. PLOS One case study
+# May 2021
+library(dplyr)
+library(ggplot2)
+library(diagram)
+
+#load metadata
+load('results/total.records.plos.rda')
+#number downloaded through target search queries
+n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
+#number of stats sections included following partial matching
+n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
+
+count_exclude = tibble(reason=c('Did not pass second\nstage filter'),n=c(n.api -  n.include))
+
+
+## a) function to make diagram ANZCTR
+make_diagram_plos = function(){
+  par(mai=c(0,0,0,0))
+  labels = c(paste('Downloaded\n n = ', format(n.api, big.mark = ','), sep=''),
+             paste('Excluded\n- ', count_exclude$reason[1], ', n = ', format(count_exclude$n[1],big.mark=',')),
+             paste('Analysed\n n =', format(n.include, big.mark = ',')))
+  n_labels = length(labels)
+  M = matrix(nrow=n_labels, ncol=n_labels)
+  M[3,1] = "' '" 
+  pos = matrix(data=c(0.29,0.8,
+                      0.7,0.55,
+                      0.29,0.3), ncol=2, byrow=TRUE)
+  sizes=c(1.5,2.3,1.5) / 10
+  props = c(0.5,0.5,0.5) # narrower for first and last
+  plotmat(M, name=labels, pos=pos, box.type = 'rect', box.size=sizes, box.prop = props, curve = 0, arr.pos=0.85)
+  shape::Arrows(x0=0.29, x1=0.45, y0=0.55, y1=0.55, arr.width=0.2, arr.length=0.22, arr.type='triangle')
+  # heading
+  text(0.5, 0.95, "PLOS ONE", font=2)
+}
+
+# export
+jpeg('figures/excluded_plosone.jpg', width=5, height=4, units='in', res=300)
+make_diagram_plos()
+dev.off()

### 2021-05-24 manuscript/asa_template/asa_template_almostdone.tex
add edited tex file

--- 2021-05-24 manuscript/asa_template/asa_template.tex
@@
 
-The statistical methods section for the ANZCTR data was missing for some
-studies and we examined if there were particular studies where this
-section was more likely to be missing. We used four independent
-variables of date, study type (observational or interventional), number
-of funders and target sample size. We used a logistic regression model
-fitted using a Bayesian paradigm. A small number of sections were
-labelled as ``Not applicable'', ``Nil'' or ``None'' and we changed these
-to missing.
+Statistical methods sections were missing for some studies downloaded
+from ANZCTR, including a small number of sections labelled as ``Not
+applicable'', ``Nil'' or ``None''. We therefore examined if there were
+particular studies where this section was more likely to be missing.
+Analysis considered a logistic regression model estimated in the
+Bayesian framework {[}\citet{INLA}; www.r-inla.org{]}, with missing
+statistical methods section (yes/no) as the dependent variable. Date,
+study type (observational or interventional), number of funders and
+target sample size were used as independent variables. Results were
+reported as odds ratios with 95\% credible intervals (CI).
 
@@
 approximating \(P\) with two factor matrices: \(W \in R^{M \times g}\)
-and \(H \in R ^{N \times g}\) \citep{aggarwal2012mining}. The number of
-subgroups of common topic inferred from the data is represented by
-\(g\), leading to the approximation: \begin{equation}
-\label{eq:approxPNMF}
- P \approx WH^{T}.
-\end{equation}
+and \(H \in R ^{N \times g}\) \citep{aggarwal2012mining}, such that
+\(P \approx WH^{T}\). The number of subgroups of common topic inferred
+from the data is given by \(g\).
 
@@
 \((1,\ldots,g)\) for a statistical methods section is inferred from the
-maximum coefficient value in the corresponding row of \(H\).
+maximum coefficient value in the corresponding row of \(H\), also known
+as the topic coherence score. For our two datasets, we applied NMF with
+\(g=10\).
 
-For our two datasets, we applied NMF varying \(g\) from 1 to 50.
-Clustering solutions were assessed using silhouette score,
-within-cluster dispersion and between-cluster dispersion.
+\subsection{Content analysis}
 
-\subsection{Within topic analysis}
+Clustering results were visualised by word clouds to summarise
+frequently occurring terms associated with topic membership. Whilst
+these results were useful for inferring common themes across statistical
+methods sections, they did not indicate evidence of boilerplate text
+beyond the use of common terms. Follow up analysis therefore considered
+similarities in text between sections assigned to the same topic.
 
-Within the \emph{PLOS ONE} dataset, statistical methods sections with
-high topic coherence scores were used to investigate evidence of
-boilerplate text, compared with other studies assigned to the same
-topic. Comparisons were carried out by calculating Jaccard similarity
-scores between studies at the sentence level. We chose the Jaccard score
-as an intuitive measure, which summarises the similarity between two
-sentences by the number of words common to each sentence (intersection),
-divided by the number of words that appeared in either the strongest
-match or the sentence being compared. Boilerplate text was defined based
-on word count (plus or minus 3 words) and a Jaccard score of 0.9 or
-higher.
-
-Results were transformed to lower case for the clustering, but examples
-are given using the original capitalisation.
+We used the Jaccard similarity index to compare the statistical methods
+section with the highest topic coherence score with all other sections
+assigned to the same topic. Indices were calculated at the sentence
+level, to account for instances of boilerplate text within larger
+methods sections. We chose the Jaccard index as an easy to interpret
+measure, which summarises the similarity between two sentences by the
+number of words common to each sentence (intersection), divided by the
+number of words that appeared in either sentence (union). Instances of
+boilerplate text were defined by a Jaccard index of 0.9 or higher and a
+difference in word count of plus or minus three words.
 
@@
 
-\hypertarget{included-studies}{%
-\subsection{Included studies}\label{included-studies}}
+\hypertarget{plos-one}{%
+\subsection{PLOS ONE}\label{plos-one}}
 
-\hypertarget{plos-one}{%
-\subsubsection{PLOS ONE}\label{plos-one}}
+\begin{figure}
+
+{\centering \includegraphics[width=0.49\linewidth]{figures/excluded_plosone} \includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}
+
+}
+
+\caption{\label{fig:consort-diagrams}Search results for PLOS ONE (left) and ANZCTR (right).}
+\end{figure}
 
@@
 
-\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-3}
+\caption{\label{fig:plos-n} A: Search results by PLOS ONE volume; B: Subject classifications assigned to full-text recorded included in the analysis}
 \end{figure}
 
-Searching with the \emph{PLOS ONE} API returned 131,847 results (DOIs)
-based on chosen search terms (\autoref{fig:plos-flowchart}). After
-partial matching, 111,731 (85\%) statistical methods sections were
-identified. In the final sample, 95,518 (85\%) DOIs returned an exact
-match against common section headings: 64,133 for `statistical
-analysis', 13,380 for `statistical analyses' and 13,627 for `data
-analysis'. For our random sample of DOIs that did not meet the partial
-matching criteria, initial search terms appeared in Introduction
-(number), Results (number) only. xxx appeared in the Methods section
-without a section heading and xxx had a non-standard section heading.
-{[}TODO{]}.
+API searches returned 131,847 results (DOIs)
+(\autoref{fig:consort-diagrams}). After partial matching, 111,731 (85\%)
+statistical methods sections were identified. In the final sample,
+95,518 (85\%) DOIs returned an exact match against common section
+headings: 64,133 for `statistical analysis', 13,380 for `statistical
+analyses' and 13,627 for `data analysis'.
 
@@
 volume based on our proposed matching criteria varied between 64\%
-(volume 2) and 86\% (volume 9).
-
-All DOIs included Biology and life sciences (n = 107,584), Earth
-sciences (n = 7,605) and/or Computer and information sciences (n =
-5,190) in their top 3 subject classifications (Figure
-\ref{fig:plos-n}B).
+(volume 2) and 86\% (volume 9). All DOIs included Biology and life
+sciences (n = 107,584), Earth sciences (n = 7,605) and/or Computer and
+information sciences (n = 5,190) in their top 3 subject classifications
+(Figure \ref{fig:plos-n}B).
 
@@
 
+For studies excluded based on section headings, initial search terms
+appeared in Introduction (number), Results (number) only. xxx appeared
+in the Methods section without a section heading and xxx had a
+non-standard section heading. {[}TODO{]}.
+
 \begin{figure}
@@
 
-\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE}\label{fig:unnamed-chunk-5}
+\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE}\label{fig:unnamed-chunk-6}
 \end{figure}
 
-The topic clouds based on ten clusters are in Figure
-\ref{fig:plos-clouds}. Frequently occurring words reflected the use of
-statistical software (topics 3 and 5), descriptive statistics (topic 6),
-group based hypothesis testing (topics 1 and 4) and definitions of
-statistical significance (topics 1 and 9). Statistical methods sections
-related to regression (topic 2) and meta-analysis (topic 7) were also
-identified.
+The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}.
+Frequently occurring words reflected the use of statistical software
+(topics 3 and 5), descriptive statistics (topic 6), group based
+hypothesis testing (topics 1 and 4) and definitions of statistical
+significance (topics 1 and 9). Statistical methods sections related to
+regression (topic 2) and meta-analysis (topic 7) were also identified.
+Examples of boilerplate text for selected topics are presented in Table
+\ref{tab:plos-example-boilerplate}.
 
 Topics related to statistical software differentiated between Prism
-GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%)
-. Manual review of results showed that nine out of the ten top matching
-DOIs for topic 3 stated the use of Prism GraphPad, but did not specify
-which statistical methods were used. Top matching sections for topic 5
-included information on SPSS version numbers and definitions of
-statistical significance.
+GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%).
+Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
+Diego, CA) was used for all analyses'' was found in 1,941 studies.
+Similar phrasing was identified for topic 5: ``SPSS software version
+17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'', with
+similar text found in 441 studies.
 
+\begin{landscape}
 \begin{table}
 
-\caption{\label{tab:unnamed-chunk-7}Example boilerplate text from PLOS ONE dataset}
+\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE datase (sentence level).
+Statistical methods text corresponds to the DOI with the highest topic coherence score within a topic}
 \centering
-\begin{tabular}[t]{l|l|r}
+\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
 \hline
-Topic & Statistical methods text & Number of matching studies\\
+Topic & Statistical methods text & Matching DOIs\\
 \hline
-Topic 1 & a p value of less-than 0.05 was considered statistically significant. & 1883\\
+Topic 1 & Students t-test was used for statistical analysis. & 29\\
+\cline{2-3}
+ & A p value of $<$ 0.05 was considered statistically significant. & 1883\\
 \hline
-Topic 1 & p values of less-than 0.05 were considered significant. & 1526\\
+Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
 \hline
-Topic 1 & statistical analysis was performed using the students t-test & 380\\
+Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
 \hline
-Topic 1 & students t-test was used for statistical analysis & 29\\
+Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
 \hline
-Topic 3 & graphpad prism graphpad software san diego ca was used for all analyses. & 1941\\
+Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
 \hline
-Topic 3 & all statistical analysis was performed using graphpad prism software. & 1312\\
-\hline
-Topic 4 & statistical analysis was performed using one-way analysis of variance anova test followed by dunnetts multiple comparison post hoc test & 4516\\
-\hline
-Topic 4 & p values less-than 0.05 were considered significant. & 2414\\
-\hline
-Topic 4 & significant differences were determined using analysis of variance anova followed by tukey post-hoc tests for multiple comparisons. & 2216\\
-\hline
-Topic 4 & data are expressed as the mean plus-or-minus sem & 860\\
-\hline
-Topic 5 & the difference was considered statistically significant at p less-than 0.05. & 2555\\
-\hline
-Topic 5 & spss software version 17.0 spss chicago il usa was used for statistical analysis. & 441\\
-\hline
-Topic 5 & the results are presented as the mean plus-or-minus sem & 420\\
-\hline
-Topic 5 & data were analyzed by one-way anova and lsd tests using the spss 13.0 software spss inc chicago il usa & 19\\
-\hline
-Topic 6 & a value of p less-than 0.05 was considered statistically significant. & 1786\\
-\hline
-Topic 6 & data are expressed as mean plus-or-minus standard error of the mean sem & 604\\
-\hline
-Topic 6 & statistical analysis of the data was performed by the students t-test & 444\\
-\hline
-Topic 6 & all results are expressed as means plus-or-minus standard deviation plus-or-minus sd. & 102\\
-\hline
-Topic 9 & statistical analysis was performed using the students unpaired t-test & 752\\
-\hline
-Topic 9 & differences were considered statistically significant at p less-than 0.05 p less-than 0.01 and p less-than 0.001. & 413\\
-\hline
-Topic 9 & data are presented as the mean plus-or-minus sd & 378\\
-\hline
-Topic 9 & statistical significance was determined by students t-tests & 49\\
-\hline
-Topic 9 & p less-than 0.05 p equal-to less-than 0.01 p less-than 0.001. & 6\\
+Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
+\cline{2-3}
+& $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
 \hline
@@
 \end{table}
+\end{landscape}
 
 Definitions of statistical significance featured strongly in topic 1 (n
-= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%), combined with text
-describing hypothesis testing for comparing differences between two
-groups. Topic 1 reflected applications of Student's t-test assuming a
-5\% level of statistical significance. Topic 9 referenced similar
-methods combined with multiple thresholds for declaring statistical
-significance by asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and
-\(^{***}p<0.001\)'', a practice that has been criticised
-\citep{Wasserstein2019}.
+= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%). Topic 1 reflected
+applications of two-tailed and unpaired Student's t-tests, however,
+instances of boilerplate text emphasised the 5\% level of statistical
+significance (\ref{tab:plos-example-boilerplate}). In contrast, Topic 9
+focused on multiple thresholds for declaring statistical significance by
+asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
+practice that has been criticised \citep{Wasserstein2019}.
 
@@
 variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
-with common methods for performing post-hoc multiple comparisons.
+with common methods for performing post-hoc multiple comparisons. Based
+on the Jaccard index, 1 in 5 studies were matched to the sentence
+``Significant differences were determined using analysis of variance
+(ANOVA) follwo by Tukey post-hoc tests for multiple comparisons''.
 Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
 mentions of descriptive statistics for summarising continuous variables,
-for example: ``Data are expressed as means ± standard error of the mean
-(SEM).'' Sections assigned to this topic appeared to be expanded
-versions of topics 1 and 9. Examples of boilerplate text were in the
-form of descriptive statistics followed by hypothesis testing, for
-example, Student's t-test, Mann Whitney U or one-way analysis of
-variance.
-
-Examples of boilerplate text for selected topics are presented in Table
-\ref{tab:plos-example-boilerplate}.
+namely means with standard deviations or standard errors.
 
@@
 We downloaded 28,008 studies. The numbers of excluded studies are shown
-in Figure \ref{fig:anzctr-missing}. Of the 12,700 included studies,
+in Figure \ref{fig:consort-diagrams}. Of the 12,700 included studies,
 9,523 (75\%) had a statistical methods section. The median length of
@@
 
-We examined if four study characteristics were associated with a missing
-statistics section. The odds ratios and 95\% credible intervals are in
-Table \ref{tab:anzctr-missing-odds}. Observational studies were less
-likely to have a missing methods section compared with interventional
+Odds ratios and 95\% credible intervals for study characteristics
+included in the logistic regression model are presented in Table
+\ref{tab:anzctr-missing-odds}. Observational studies were less likely to
+have a missing statistical methods section compared with interventional
 studies. Missing sections became less likely over time. Studies with
 more funders and a larger target sample size were less likely to have a
-missing methods section.
+missing statistical methods section.
 
@@
 \centering
-\caption{Regression results for factors associated with missing statistical methods sections in ANZCTR}
+\caption{Logistic regression results for study characteristics associated with missing statistical methods sections in ANZCTR}
 \label{tab:anzctr-missing-odds}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/excluded_anzctr_missing}
-
-}
-
-\caption{\label{fig:anzctr-missing}ANZCTR search results.}\label{fig:unnamed-chunk-8}
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
@@
 The clustering algorithm found groups that were purely sample size
-calculations (topic 2, n = 1834; topic 4, n = 909), pilot studies (topic
-5, n = 834), safety/tolerability studies (topic 6, n = 524),
-intervention studies (topic 8, n = 1020) and repeated measures ANOVA
-(topic 10 = 852). Examples of boilerplate text are provided in Table
-\ref{tab:anzctr-example-boilerplate} and Supplementary File 3.
+calculations (topic 2, n = 1834), pilot studies (topic 5, n = 834),
+safety/tolerability studies (topic 6, n = 524), intervention studies
+(topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852).
+Examples of boilerplate text are provided in Table
+\ref{tab:anzctr-example-boilerplate} .
 
@@
 
-Finally, targeted searches for the term `statistician' revealed
-instances across all topics where no methods were specified. Instead,
-all analyses were the responsibility of an appointed statistician, for
-example, ``A statistician employed by hospital was used'' and ``Pilot
-study at this point will use a statistician professionally to determine
-sample size calculations as required''.
+Since studies included in this dataset outlined planned analyses, we
+suspected that some studies did not specify statistical methods because
+they had yet to consult with a statistician. Targeted searches for the
+term ``statistician'' across all topics returned xx matching records,
+with examples including ``A statistician employed by hospital was used''
+and ``Pilot study at this point will use a statistician professionally
+to determine sample size calculations as required''.
 

### 2021-05-24 code/3_plot_excluded.plos.R

@@ -32,10 +32,10 @@ make_diagram_plos = function(){
   plotmat(M, name=labels, pos=pos, box.type = 'rect', box.size=sizes, box.prop = props, curve = 0, arr.pos=0.85)
   shape::Arrows(x0=0.29, x1=0.45, y0=0.55, y1=0.55, arr.width=0.2, arr.length=0.22, arr.type='triangle')
   # heading
-  text(0.5, 0.95, "PLOS ONE", font=2)
+  #text(0.5, 0.95, "PLOS ONE", font=2)
 }
 
 # export
-jpeg('figures/excluded_plosone.jpg', width=5, height=4, units='in', res=300)
+jpeg('manuscript/asa_template/figures/excluded_plosone.jpg', width=5, height=4, units='in', res=300)
 make_diagram_plos()
 dev.off()

### 2021-05-24 code/3_extract_title_abstract_xml.R
update title abstract code plos

@@ -5,14 +5,17 @@ library(XML)
 library(tidyverse)
 
 #plos full text - loop over batches
-dat = list()
-for (batch.number in 1:5){
+dat.title.abstract = list()
+dat.subject.class = list()
+
+for (batch.number in 2:5){
 meta_dat = readRDS(paste0('./data/plos_meta_',batch.number,'.RDS'))
 doi_list = meta_dat %>% pull(doi)
 
-dat.batch = list()
+dat.batch_titleabs = list()
+dat.batch_subjects = list()
 
-for(i in 1:10){ #length(doi_list)
+for(i in 1:length(doi_list)){ 
   #extract xml of full text. if no xml available, skip to the next doi
   full_text <- tryCatch(
     plos_fulltext(doi_list[i]),
@@ -48,7 +51,7 @@ for(i in 1:10){ #length(doi_list)
   subjects[['level3']] = unlist(as.character(sapply(node_info.subjects,xpathSApply,"./subj-group/subj-group/subject",xmlValue)))
     
   if (!is.null(title_abstract[['title']]) & !is.null(title_abstract[['abstract']])){
-    dat.batch[[i]] = bind_cols(title_abstract) %>% add_column(doi=doi_list[i],.before = 1) %>% mutate_if(is.factor,as.character)
+    dat.batch_titleabs[[i]] = bind_cols(title_abstract) %>% add_column(doi=doi_list[i],.before = 1) %>% mutate_if(is.factor,as.character)
   }
   
   if (!is.null(subjects[['level1']])){ #must have level 1 subject classification at a minimum
@@ -60,9 +63,10 @@ for(i in 1:10){ #length(doi_list)
   
 }
 
-dat.title.abstract[[batch.number]] = bind_rows(dat.batch)
+dat.title.abstract[[batch.number]] = bind_rows(dat.batch_titleabs)
 dat.subject.class[[batch.number]] = bind_rows(dat.batch_subjects)
 }
 
-dat = bind_rows(dat)
-save(dat,file='./data/title_abstract_plos.rda')
+dat.title.abstract = bind_rows(dat.title.abstract)
+dat.subject.class = bind_rows(dat.subject.class)
+save(list(title_abstract = dat.title.abstract,subjects = dat.subject.class),file='./data/title_abstract_subjects_plos.rda')

### 2021-05-24 manuscript/asa_template/asa_template.Rmd

@@
 
-The statistical methods section for the ANZCTR data was missing for some studies and we examined if there were particular studies where this section was more likely to be missing. 
-We used four independent variables of date, study type (observational or interventional), number of funders and target sample size. 
-We used a logistic regression model fitted using a Bayesian paradigm. A small number of sections were labelled as "Not applicable", "Nil" or "None" and we changed these to missing.
+Statistical methods sections were missing for some studies downloaded from ANZCTR, including a small number of sections labelled as "Not applicable", "Nil" or "None".
+We therefore examined if there were particular studies where this section was more likely to be missing.
+Analysis considered a logistic regression model estimated in the Bayesian framework [@INLA; www.r-inla.org], with missing statistical methods section (yes/no) as the dependent variable.
+Date, study type (observational or interventional), number of funders and target sample size were used as independent variables.
+Results were reported as odds ratios with 95% credible intervals (CI).
 
@@
 To address these limitations, NMF deals with high-dimensional data by mapping it to a lower-dimensional space.
-This mapping this achieved by approximating $P$ with two factor matrices: \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \) [@aggarwal2012mining].
-The number of subgroups of common topic inferred from the data is represented by $g$, leading to the approximation:
-\begin{equation}
-\label{eq:approxPNMF}
- P \approx WH^{T}.
-\end{equation}
+This mapping this achieved by approximating $P$ with two factor matrices: \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \) [@aggarwal2012mining], such that $P \approx WH^{T}$.
+The number of subgroups of common topic inferred from the data is given by $g$.
 
@@
 Following estimation, \textit{H} contains the information regarding topic membership for all studies.
-In our case, topic membership $(1,\ldots,g)$ for a statistical methods section is inferred from the maximum coefficient value in the corresponding row of $H$.
+In our case, topic membership $(1,\ldots,g)$ for a statistical methods section is inferred from the maximum coefficient value in the corresponding row of $H$, also known as the topic coherence score.
+For our two datasets, we applied NMF with $g=10$.
 
-For our two datasets, we applied NMF varying $g$ from 1 to 50. Clustering solutions were assessed using silhouette score, within-cluster dispersion and between-cluster dispersion.
+\subsection{Content analysis}
 
-\subsection{Within topic analysis}
+Clustering results were visualised by word clouds to summarise frequently occurring terms associated with topic membership.
+Whilst these results were useful for inferring common themes across statistical methods sections, they did not indicate evidence of boilerplate text beyond the use of common terms.
+Follow up analysis therefore considered similarities in text between sections assigned to the same topic.
 
-<!-- Clustering results were visualised by word clouds that summarised frequently occurring terms associated with membership to different topics.  -->
-
-Within the _PLOS ONE_ dataset, statistical methods sections with high topic coherence scores were used to investigate evidence of boilerplate text, compared with other studies assigned to the same topic.
-Comparisons were carried out by calculating Jaccard similarity scores between studies at the sentence level.
-We chose the Jaccard score as an intuitive measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either the strongest match or the sentence being compared.
-Boilerplate text was defined based on word count (plus or minus 3 words) and a Jaccard score of 0.9 or higher.
-
-Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.
-
-<!-- \subsection{Missing statistical methods sections} -->
+We used the Jaccard similarity index to compare the statistical methods section with the highest topic coherence score with all other sections assigned to the same topic.
+Indices were calculated at the sentence level, to account for instances of boilerplate text within larger methods sections.
+We chose the Jaccard index as an easy to interpret measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either sentence (union).
+Instances of boilerplate text were defined by a Jaccard index of 0.9 or higher and a difference in word count of plus or minus three words.
 
@@
 
-## Included studies
-
-### PLOS ONE
+## PLOS ONE
 
@@
 n.extremes = wordcounts %>% summarise(morethan500 = sum(words>500),lessthan50=sum(words<50))
+```
+
+```{r, echo=FALSE, fig.show='hold',fig.align='center', out.width="49%", fig.cap="\\label{fig:consort-diagrams}Search results for PLOS ONE (left) and ANZCTR (right)."}
+
+knitr::include_graphics(c("figures/excluded_plosone.jpg","figures/excluded_anzctr_missing.jpg"))
+
 ```
@@
 
-Searching with the _PLOS ONE_ API returned `r format(n.api,big.mark=',')` results (DOIs) based on chosen search terms (\autoref{fig:plos-flowchart}).
+API searches returned `r format(n.api,big.mark=',')` results (DOIs) (\autoref{fig:consort-diagrams}).
 After partial matching, `r format(n.include,big.mark=',')` (`r perc.include`%) statistical methods sections were identified.
 In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings: 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.
-For our random sample of DOIs that did not meet the partial matching criteria, initial search terms appeared in Introduction (number), Results (number) only.
-xxx appeared in the Methods section without a section heading and xxx had a non-standard section heading.
-[TODO].
 
@@
 The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).
-
 All DOIs included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
@@
 
-<!-- Increasing the number of clusters decreased cluster quality based on global goodness-of-fit measures (Supplementary Figure 1), with average silhouette score and within-cluster dispersion leveling off around 20 clusters.
-This indicated that the data comprised one large, heterogeneous cluster and multiple smaller clusters.
--->
+For studies excluded based on section headings, initial search terms appeared in Introduction (number), Results (number) only.
+xxx appeared in the Methods section without a section heading and xxx had a non-standard section heading.
+[TODO].
 
@@
 
-The topic clouds based on ten clusters are in Figure \ref{fig:plos-clouds}.
+The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}.
 Frequently occurring words reflected the use of statistical software (topics 3 and 5), descriptive statistics (topic 6), group based hypothesis testing (topics 1 and 4) and definitions of statistical significance (topics 1 and 9).
 Statistical methods sections related to regression (topic 2) and meta-analysis (topic 7) were also identified.
+Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.
 
@@
 
-Topics related to statistical software differentiated between Prism GraphPad (topic 3: n = `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(perc)`%) and SPSS (topic 5: n = `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(perc)`%) .
-Manual review of results showed that nine out of the ten top matching DOIs for topic 3 stated the use of Prism GraphPad, but did not specify which statistical methods were used.
-Top matching sections for topic 5 included information on SPSS version numbers and definitions of statistical significance.
+Topics related to statistical software differentiated between Prism GraphPad (topic 3: n = `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(perc)`%) and SPSS (topic 5: n = `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(perc)`%).
+Within topic 3, the sentence "GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses" was found in 1,941 studies.
+Similar phrasing was identified for topic 5: "SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis", with similar text found in 441 studies.
 
@@
 ftab = jaccard.matches %>% mutate(Topic = paste('Topic',topic)) %>%
-  select(Topic,sentences,matches) %>%  arrange(Topic,-matches) %>%
+  select(Topic,sentences,matches) %>%  arrange(Topic) %>%
   rename('Statistical methods text' = sentences,
-         'Number of matching studies' = matches)
+         'Matching DOIs' = matches)
 knitr::kable(ftab,format = "latex", caption='Example boilerplate text from PLOS ONE dataset')
@@
 
-Definitions of statistical significance featured strongly in topic 1 (n = `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(perc)`%) and topic 9 (n = `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(perc)`%), combined with text describing hypothesis testing for comparing differences between two groups.
-Topic 1 reflected applications of Student's t-test assuming a 5% level of statistical significance.
-Topic 9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: ``$^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$'', a practice that has been criticised [@Wasserstein2019].
+Definitions of statistical significance featured strongly in topic 1 (n = `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(perc)`%) and topic 9 (n = `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(perc)`%).
+Topic 1 reflected applications of two-tailed and unpaired Student's t-tests, however, instances of boilerplate text emphasised the 5% level of statistical significance (\ref{tab:plos-example-boilerplate}).
+In contrast, Topic 9 focused on multiple thresholds for declaring statistical significance by asterisk: ``$^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$'', a practice that has been criticised [@Wasserstein2019].
 
@@
 One-way analysis of variance also featured strongly in topic 4 (n = `r filter(ftab.plos,topic_id=='Topic 4') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 4') %>% pull(perc)`%), combined with common methods for performing post-hoc multiple comparisons.
-Frequently occurring words in topic 6 (n = `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(perc)`%) reflected mentions of descriptive statistics for summarising continuous variables, for example: ``Data are expressed as means ± standard error of the mean (SEM).'' Sections assigned to this topic appeared to be expanded versions of topics 1 and 9.
-Examples of boilerplate text were in the form of descriptive statistics followed by hypothesis testing, for example, Student's t-test, Mann Whitney U or one-way analysis of variance.
+Based on the Jaccard index, 1 in 5 studies were matched to the sentence ``Significant differences were determined using analysis of variance (ANOVA) follwo by Tukey post-hoc tests for multiple comparisons''.
+Frequently occurring words in topic 6 (n = `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(perc)`%) reflected mentions of descriptive statistics for summarising continuous variables, namely means with standard deviations or standard errors.
 
@@
 
-Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.
-
 \subsection{ANZCTR}
@@
 We downloaded 28,008 studies.
-The numbers of excluded studies are shown in Figure \ref{fig:anzctr-missing}.
+The numbers of excluded studies are shown in Figure \ref{fig:consort-diagrams}.
 Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.
@@
 
-We examined if four study characteristics were associated with a missing statistics section.
-The odds ratios and 95% credible intervals are in Table \ref{tab:anzctr-missing-odds}.
-Observational studies were less likely to have a missing methods section compared with interventional studies.
+Odds ratios and 95% credible intervals for study characteristics included in the logistic regression model are presented in Table \ref{tab:anzctr-missing-odds}.
+Observational studies were less likely to have a missing statistical methods section compared with interventional studies.
 Missing sections became less likely over time.
-Studies with more funders and a larger target sample size were less likely to have a missing methods section.
+Studies with more funders and a larger target sample size were less likely to have a missing statistical methods section.
 
@@
 
-```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="\\label{fig:anzctr-missing}ANZCTR search results."}
-
-knitr::include_graphics("figures/excluded_anzctr_missing.jpg")
-
-```
-
 ```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR topic clouds."}
@@
 
-The clustering algorithm found groups that were purely sample size calculations (topic 2, n = 1834; topic 4, n = 909), pilot studies (topic 5, n = 834), safety/tolerability studies (topic 6, n = 524), intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852).
-Examples of boilerplate text are provided in Table \ref{tab:anzctr-example-boilerplate} and Supplementary File 3.
+The clustering algorithm found groups that were purely sample size calculations (topic 2, n = 1834), pilot studies (topic 5, n = 834), safety/tolerability studies (topic 6, n = 524), intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852).
+Examples of boilerplate text are provided in Table \ref{tab:anzctr-example-boilerplate} .
 
@@
 
-Finally, targeted searches for the term 'statistician' revealed instances across all topics where no methods were specified.
-Instead, all analyses were the responsibility of an appointed statistician, for example, "A statistician employed by hospital was used" and ``Pilot study at this point will use a statistician professionally to determine sample size calculations as required''.
+Since studies included in this dataset outlined planned analyses, we suspected that some studies did not specify statistical methods because they had yet to consult with a statistician.
+Targeted searches for the term "statistician" across all topics returned xx matching records, with examples including "A statistician employed by hospital was used" and ``Pilot study at this point will use a statistician professionally to determine sample size calculations as required''.
 

### 2021-05-24 manuscript/asa_template/asa_template_almostdone.tex
update manuscript file and bib file
update tex files

@@
-% !TeX program = pdfLaTeX
 \documentclass[12pt]{article}
@@
 
-Two statisticians on this paper (AB and NW) have heard researchers admit
+A potential contributor to poor statistical reporting is the temptation for researchers to replicate
+descriptions of statistical methods, in an effort to make their papers resemble those of their peers 
+and increase perceived chances of publication  \citep{Diong2018}. Two statisticians on this paper 
+(AB and NW) have heard researchers admit
 that they have copied-and-pasted their statistical methods sections from
-other papers, regardless of whether they are appropriate. The aim of
-this paper is to use text-mining methods to analyse the content of
-statistical methods sections included in common research outputs.
-Results are then evaluated to estimate the extent that researchers are
+other papers, regardless of whether they are appropriate. For this paper, we applied a text-based 
+clustering method to analyse the content of statistical methods sections included in common scientific research outputs.
+Clustering results are then evaluated to estimate the extent that researchers are
 using cut-and-paste or `boilerplate' statistical methods sections.
@@
 records that did not pass Step 2, we reviewed where initial search terms
-appeared in the full-text, to evaluate the likely proportion of
-statistical methods sections that were missed.
+appeared in the full-text, to determine the proportion of
+statistical methods sections that were missed under our chosen approach.
 
@@
 number of papers published in \emph{PLOS ONE} over the same period. The
-percentage of records that included a statistical methods section by
-volume based on our proposed matching criteria varied between 64\%
+percentage of records that included a statistical methods section varied between 64\%
 (volume 2) and 86\% (volume 9). All DOIs included Biology and life
@@
 hypothesis testing (topics 1 and 4) and definitions of statistical
-significance (topics 1 and 9). Statistical methods sections related to
+significance (topics 1 and 9). Topics related to
 regression (topic 2) and meta-analysis (topic 7) were also identified.
@@
 Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
-Diego, CA) was used for all analyses'' was found in 1,941 studies.
-Similar phrasing was identified for topic 5: ``SPSS software version
-17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'', with
-similar text found in 441 studies.
+Diego, CA) was used for all analyses'' was matched in 1,941 studies based on our definition of boilerplate text.
+Similarly, for topic 5, ``SPSS software version
+17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 studies.
 
@@
 \hline
-Topic 1 & Students t-test was used for statistical analysis. & 29\\
+Topic 1 & $^{\dagger}$Students t-test was used for statistical analysis. & 29\\
 \cline{2-3}
- & A p value of $<$ 0.05 was considered statistically significant. & 1883\\
+ & $^{\dagger}$A p value of $<$ 0.05 was considered statistically significant. & 1883\\
 \hline
-Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
+Topic 3 & $^{\dagger}$GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
 \hline
-Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
+Topic 4 & $^{\dagger}$Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
 \hline
-Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
+Topic 5 & $^{\dagger}$SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
 \hline
-Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
+Topic 6 & $^{\dagger}$All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
 \hline
-Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
+Topic 9 & $^{\dagger}$Statistical significance was determined by Students t-tests. & 49\\
 \cline{2-3}
-& $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
+& $^{\dagger}$ $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
 \hline
@@
 instances of boilerplate text emphasised the 5\% level of statistical
-significance (\ref{tab:plos-example-boilerplate}). In contrast, Topic 9
+significance (Table \ref{tab:plos-example-boilerplate}). In contrast, Topic 9
 focused on multiple thresholds for declaring statistical significance by
@@
 ``Significant differences were determined using analysis of variance
-(ANOVA) follwo by Tukey post-hoc tests for multiple comparisons''.
+(ANOVA) followed by Tukey post-hoc tests for multiple comparisons''.
 Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
@@
 related to sample size calculations for patient recruitment to different
-studies (see Supplementary File 3).
+studies.
 
 Since studies included in this dataset outlined planned analyses, we
-suspected that some studies did not specify statistical methods because
+hypothesised that some studies did not specify statistical methods because
 they had yet to consult with a statistician. Targeted searches for the
@@
 \bibliographystyle{agsm}
-\bibliography{references.bib}
+\bibliography{references}
 

### 2021-05-24 code/4_combine_batch_info_excluded.R
add script for excluded plos articles

@@ -1,0 +1,36 @@
+#combine_batch_info_excluded
+library(tidyverse)
+
+#filter: identify dois with statitical analyses subsection based on text_heading (partial string matching)
+analysis_keywords = c('statistical analysis','statistical analyses',
+                      'statistical method','statistics',
+                      'data analysis', 'data analyses')
+
+
+load('../data/plos_meta_data.rda')
+stats_sections = readRDS('../data/stats_section_cleaned.rds')
+
+excluded_dois = anti_join(meta_dat_allrecords,stats_sections,by='doi') %>% pull(doi)
+
+#loop over batches
+batch = 1:5
+stats_section = list()
+stats_section_excluded = list()
+for (b in batch){
+  folder_name = paste0('data/full text records/batch_',b)
+  results_by_doi = list.files(folder_name)
+  doi_list  = gsub('.rda','',results_by_doi)
+  text_results = lapply(results_by_doi,
+                        function(x) {load(paste(folder_name,x,sep='/'))
+                          return(dat %>% mutate_if(is.factor,as.character))}
+  )
+  names(text_results) <- doi_list
+  text_results = bind_rows(text_results,.id='doi')
+  stats_section[[b]] = text_results %>% filter(grepl(paste(analysis_keywords,collapse='|'),text_heading,
+                                                     ignore.case=T,fixed=F))
+  stats_section_excluded[[b]] = text_results %>% filter(!grepl(paste(analysis_keywords,collapse='|'),text_heading,
+                                                               ignore.case=T,fixed=F))
+}
+
+save(stats_section,file='data/stats_section_info.rda')
+save(stats_section_excluded,file='data/stats_section_excluded_plos.rda')

### 2021-05-24 code/4_combine_batch_info_excluded.R
update script for excluded plos studies

@@ -1,36 +1,36 @@
 #combine_batch_info_excluded
 library(tidyverse)
 
-#filter: identify dois with statitical analyses subsection based on text_heading (partial string matching)
-analysis_keywords = c('statistical analysis','statistical analyses',
-                      'statistical method','statistics',
-                      'data analysis', 'data analyses')
-
+#original search terms
+prefix = c('data','statistical')
+suffix = c('analysis','analyses','method','methodology','modelling')
+search_terms = apply(expand.grid(prefix,suffix),1,paste,collapse=' ')
+search_terms = str_c(search_terms,collapse='|')
 
 load('../data/plos_meta_data.rda')
 stats_sections = readRDS('../data/stats_section_cleaned.rds')
-
 excluded_dois = anti_join(meta_dat_allrecords,stats_sections,by='doi') %>% pull(doi)
 
 #loop over batches
 batch = 1:5
-stats_section = list()
+#stats_section = list()
 stats_section_excluded = list()
 for (b in batch){
-  folder_name = paste0('data/full text records/batch_',b)
-  results_by_doi = list.files(folder_name)
-  doi_list  = gsub('.rda','',results_by_doi)
+  folder_name = paste0('Z:/rplos/data/full text records/batch_',b) #full text on hpc drive
+  results_by_doi_orig = list.files(folder_name) %>% str_remove_all(.,'doi_|.rda')
+  results_by_doi = intersect(results_by_doi_orig,excluded_dois  %>% str_replace_all(.,pattern='/',replacement='.')) #filter to excluded dois
   text_results = lapply(results_by_doi,
-                        function(x) {load(paste(folder_name,x,sep='/'))
+                        function(x) {load(paste0(folder_name,'/doi_',x,'.rda'))
                           return(dat %>% mutate_if(is.factor,as.character))}
   )
-  names(text_results) <- doi_list
+  names(text_results) <- results_by_doi
   text_results = bind_rows(text_results,.id='doi')
-  stats_section[[b]] = text_results %>% filter(grepl(paste(analysis_keywords,collapse='|'),text_heading,
-                                                     ignore.case=T,fixed=F))
-  stats_section_excluded[[b]] = text_results %>% filter(!grepl(paste(analysis_keywords,collapse='|'),text_heading,
-                                                               ignore.case=T,fixed=F))
[…]

### 2021-05-24 code/4_combine_batch_info_excluded.R
upload excluded plos results

@@ -5,7 +5,6 @@ library(tidyverse)
 prefix = c('data','statistical')
 suffix = c('analysis','analyses','method','methodology','modelling')
 search_terms = apply(expand.grid(prefix,suffix),1,paste,collapse=' ')
-search_terms = str_c(search_terms,collapse='|')
 
 load('../data/plos_meta_data.rda')
 stats_sections = readRDS('../data/stats_section_cleaned.rds')

### 2021-05-24 code/4_combine_batch_info_excluded.R
upload excluded plos script

@@ -32,4 +32,4 @@ for (b in batch){
 }
 
 #save(stats_section,file='data/stats_section_info.rda')
-save(stats_section_excluded,file='data/stats_section_excluded_plos.rda')
+save(stats_section_excluded,file='../data/stats_section_excluded_plos.rda')

### 2021-05-25 manuscript/asa_template/asa_template.tex

@@
 {
-  \title{\bf An observational study of papers published in \emph{PLOS ONE} and
-studies posted to a trial registry.}
+  \title{\bf An observational study of papers published in \emph{PLOS
+ONE} and studies posted to a trial registry.}
 
   \author{
-        Nicole White \thanks{The authors gratefully acknowledge computational resources and services
-used in this work provided by the eResearch Office, Queensland
-University of Technology, Brisbane, Australia.} \\
+        Nicole White \thanks{The authors gratefully acknowledge
+computational resources and services used in this work provided by the
+eResearch Office, Queensland University of Technology, Brisbane,
+Australia.} \\
     School of Public Health and Social Work, QUT\\
@@
   \begin{center}
-    {\LARGE\bf An observational study of papers published in \emph{PLOS ONE} and
-studies posted to a trial registry.}
+    {\LARGE\bf An observational study of papers published in \emph{PLOS
+ONE} and studies posted to a trial registry.}
   \end{center}
@@
 
-\hypertarget{included-studies}{%
-\subsection{Included studies}\label{included-studies}}
+\hypertarget{plos-one}{%
+\subsection{PLOS ONE}\label{plos-one}}
 
-\hypertarget{plos-one}{%
-\subsubsection{PLOS ONE}\label{plos-one}}
+\begin{figure}
+
+{\centering \includegraphics[width=0.49\linewidth]{figures/excluded_plosone} \includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}
+
+}
+
+\caption{\label{fig:consort-diagrams}Search results for PLOS ONE (left) and ANZCTR (right).}\label{fig:unnamed-chunk-3}
+\end{figure}
 
@@
 
-\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-3}
+\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-4}
 \end{figure}
 
-Searching with the \emph{PLOS ONE} API returned 131,847 results (DOIs)
-based on chosen search terms (\autoref{fig:plos-flowchart}). After
-partial matching, 111,731 (85\%) statistical methods sections were
-identified. In the final sample, 95,518 (85\%) DOIs returned an exact
-match against common section headings: 64,133 for `statistical
-analysis', 13,380 for `statistical analyses' and 13,627 for `data
-analysis'. For our random sample of DOIs that did not meet the partial
-matching criteria, initial search terms appeared in Introduction
-(number), Results (number) only. xxx appeared in the Methods section
-without a section heading and xxx had a non-standard section heading.
-{[}TODO{]}.
+API searches returned 131,847 results (DOIs)
+(\autoref{fig:consort-diagrams}). After partial matching, 111,731 (85\%)
+statistical methods sections were identified. In the final sample,
+95,518 (85\%) DOIs returned an exact match against common section
+headings: 64,133 for `statistical analysis', 13,380 for `statistical
+analyses' and 13,627 for `data analysis'.
 
@@
 
-\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE}\label{fig:unnamed-chunk-5}
+\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE}\label{fig:unnamed-chunk-6}
 \end{figure}
 
-The topic clouds based on ten clusters are in Figure
-\ref{fig:plos-clouds}. Frequently occurring words reflected the use of
-statistical software (topics 3 and 5), descriptive statistics (topic 6),
-group based hypothesis testing (topics 1 and 4) and definitions of
-statistical significance (topics 1 and 9). Statistical methods sections
-related to regression (topic 2) and meta-analysis (topic 7) were also
-identified.
+The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}.
+Frequently occurring words reflected the use of statistical software
+(topics 3 and 5), descriptive statistics (topic 6), group based
+hypothesis testing (topics 1 and 4) and definitions of statistical
+significance (topics 1 and 9). Statistical methods sections related to
+regression (topic 2) and meta-analysis (topic 7) were also identified.
+Examples of boilerplate text for selected topics are presented in Table
+\ref{tab:plos-example-boilerplate}.
 
 Topics related to statistical software differentiated between Prism
-GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%)
-. Manual review of results showed that nine out of the ten top matching
-DOIs for topic 3 stated the use of Prism GraphPad, but did not specify
-which statistical methods were used. Top matching sections for topic 5
-included information on SPSS version numbers and definitions of
-statistical significance.
+GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%).
+Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
+Diego, CA) was used for all analyses'' was found in 1,941 studies.
+Similar phrasing was identified for topic 5: ``SPSS software version
+17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'', with
+similar text found in 441 studies.
 
@@
 
-\caption{\label{tab:unnamed-chunk-7}Example boilerplate text from PLOS ONE dataset}
+\caption{\label{tab:unnamed-chunk-8}Example boilerplate text from PLOS ONE dataset}
 \centering
@@
 \hline
-Topic & Statistical methods text & Number of matching studies\\
+Topic & Statistical methods text & Matching DOIs\\
+\hline
+Topic 1 & students t-test was used for statistical analysis & 29\\
 \hline
@@
 \hline
-Topic 1 & p values of less-than 0.05 were considered significant. & 1526\\
-\hline
 Topic 1 & statistical analysis was performed using the students t-test & 380\\
 \hline
-Topic 1 & students t-test was used for statistical analysis & 29\\
+Topic 1 & p values of less-than 0.05 were considered significant. & 1526\\
 \hline
@@
 \hline
-Topic 4 & statistical analysis was performed using one-way analysis of variance anova test followed by dunnetts multiple comparison post hoc test & 4516\\
-\hline
-Topic 4 & p values less-than 0.05 were considered significant. & 2414\\
-\hline
 Topic 4 & significant differences were determined using analysis of variance anova followed by tukey post-hoc tests for multiple comparisons. & 2216\\
@@
 \hline
-Topic 5 & the difference was considered statistically significant at p less-than 0.05. & 2555\\
+Topic 4 & statistical analysis was performed using one-way analysis of variance anova test followed by dunnetts multiple comparison post hoc test & 4516\\
+\hline
+Topic 4 & p values less-than 0.05 were considered significant. & 2414\\
 \hline
@@
 \hline
-Topic 6 & a value of p less-than 0.05 was considered statistically significant. & 1786\\
+Topic 5 & the difference was considered statistically significant at p less-than 0.05. & 2555\\
+\hline
+Topic 6 & all results are expressed as means plus-or-minus standard deviation plus-or-minus sd. & 102\\
 \hline
@@
 \hline
-Topic 6 & all results are expressed as means plus-or-minus standard deviation plus-or-minus sd. & 102\\
+Topic 6 & a value of p less-than 0.05 was considered statistically significant. & 1786\\
+\hline
+Topic 9 & statistical significance was determined by students t-tests & 49\\
+\hline
+Topic 9 & p less-than 0.05 p equal-to less-than 0.01 p less-than 0.001. & 6\\
+\hline
+Topic 9 & data are presented as the mean plus-or-minus sd & 378\\
 \hline
@@
 \hline
-Topic 9 & data are presented as the mean plus-or-minus sd & 378\\
-\hline
-Topic 9 & statistical significance was determined by students t-tests & 49\\
-\hline
-Topic 9 & p less-than 0.05 p equal-to less-than 0.01 p less-than 0.001. & 6\\
-\hline
 \end{tabular}
@@
 Definitions of statistical significance featured strongly in topic 1 (n
-= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%), combined with text
-describing hypothesis testing for comparing differences between two
-groups. Topic 1 reflected applications of Student's t-test assuming a
-5\% level of statistical significance. Topic 9 referenced similar
-methods combined with multiple thresholds for declaring statistical
-significance by asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and
-\(^{***}p<0.001\)'', a practice that has been criticised
-\citep{Wasserstein2019}.
+= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%). Topic 1 reflected
+applications of two-tailed and unpaired Student's t-tests, however,
+instances of boilerplate text emphasised the 5\% level of statistical
+significance (\ref{tab:plos-example-boilerplate}). In contrast, Topic 9
+focused on multiple thresholds for declaring statistical significance by
+asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'\,', a
+practice that has been criticised \citep{Wasserstein2019}.
 
@@
 variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
-with common methods for performing post-hoc multiple comparisons.
+with common methods for performing post-hoc multiple comparisons. Based
+on the Jaccard index, 1 in 5 studies were matched to the sentence
+``Significant differences were determined using analysis of variance
+(ANOVA) follwo by Tukey post-hoc tests for multiple comparisons'\,'.
 Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
 mentions of descriptive statistics for summarising continuous variables,
-for example: ``Data are expressed as means ± standard error of the mean
-(SEM).'' Sections assigned to this topic appeared to be expanded
-versions of topics 1 and 9. Examples of boilerplate text were in the
-form of descriptive statistics followed by hypothesis testing, for
-example, Student's t-test, Mann Whitney U or one-way analysis of
-variance.
-
-Examples of boilerplate text for selected topics are presented in Table
-\ref{tab:plos-example-boilerplate}.
+namely means with standard deviations or standard errors.
 
@@
 
-Finally, targeted searches for the term `statistician' revealed
-instances across all topics where no methods were specified. Instead,
-all analyses were the responsibility of an appointed statistician, for
-example, ``A statistician employed by hospital was used'' and ``Pilot
-study at this point will use a statistician professionally to determine
-sample size calculations as required''.
+Since studies included in this dataset outlined planned analyses, we
+suspected that some studies did not specify statistical methods because
+they had yet to consult with a statistician. Targeted searches for the
+term ``statistician'' across all topics returned xx matching records,
+with examples including ``A statistician employed by hospital was used''
+and ``Pilot study at this point will use a statistician professionally
+to determine sample size calculations as required'\,'.
 

### 2021-05-25 manuscript/asa_template/asa_template_almostdone.tex
update manuscript plos results

@@
 
-For studies excluded based on section headings, initial search terms
-appeared in Introduction (number), Results (number) only. xxx appeared
-in the Methods section without a section heading and xxx had a
-non-standard section heading. {[}TODO{]}.
+For studies excluded based on section headings (n = 20,116), 2,136 DOIs included search terms as part of the Materials and Methods section.
+For studies without a matching Materials and Methods section,
+ search terms oftern featured in the Introdction (n = 728) and Results and/or Discussion (n = 981) sections.
+4,010 DOIs included at least 1 heading that included the terms ``analysis'' or ``analyses'' without direct reference to statistical method(s) (e.g.
+Microarray analysis).
 

### 2021-05-25 code/3_plot_excluded.plos.R

@@ -1,8 +1,7 @@
 #3_plot_excluded
 # CONSORT style diagram of included/excluded studies. PLOS One case study
 # May 2021
-library(dplyr)
-library(ggplot2)
+library(tidyverse)
 library(diagram)
 
 #load metadata
@@ -12,14 +11,45 @@ n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
 #number of stats sections included following partial matching
 n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
 
-count_exclude = tibble(reason=c('Did not pass second\nstage filter'),n=c(n.api -  n.include))
 
+#original search terms
+prefix = c('data','statistical')
+suffix = c('analysis','analyses','method','methodology','modelling')
+search_terms = apply(expand.grid(prefix,suffix),1,paste,collapse=' ') %>% str_c(.,collapse='|')
+
+excluded_text_heading = stats_section_excluded %>% filter(grepl(search_terms,text_heading))
+excluded_text_data = stats_section_excluded %>% filter(grepl(search_terms,text_data))
+
+#view most common sections
+excluded_headings_n = excluded_text_data %>% count(text_heading) %>% arrange(-n)
+
+generic_methods_heading = excluded_headings_n %>% filter(grepl('^materials$|^methods$|material(.*)method|method(.*)material',text_heading,fixed=F,ignore.case = T))
+analysis_heading = excluded_headings_n %>% filter(grepl('Analysis',text_heading,ignore.case = T)) %>% pull(n) %>% sum()
+results_discussion_headings = excluded_headings_n %>% filter(grepl('results|discussion',text_heading,ignore.case=T)) %>% pull(n) %>% sum()
+intro_headings = excluded_headings_n %>% filter(grepl('introduction',text_heading,ignore.case=T)) %>% pull(n) %>% sum()
+
+#materials and methods section
+materials_methods_n = generic_methods_heading %>% pull(n) %>% sum()
+materials_methods_doi = excluded_text_data %>% filter(grepl('^materials$|^methods$|material(.*)method|method(.*)material',
+                                    text_heading,fixed=F,ignore.case = T)) %>% pull(doi) %>% unique()
+
+#section headings excluded materials and methods dois
[…]

### 2021-05-25 code/7_summarise_stats_sections.R
update r script for jaccard index

@@ -116,6 +116,36 @@ dat.sentences = lapply(1:nrow(matches), function(i) str_split(matches[i,]$text_d
 
 matches.bysentence = dat.sentences %>% bind_rows() %>% left_join(.,matches %>% select(doi,topic_id,rank),by='doi')
 
+#starting at rank=1 within each topic (e.g. topic 1 as starting point)
+target_s = matches.bysentence %>% filter(topic_id=='Topic 3',rank==1) %>% pull(text_data_clean_s)
+compare_s_all = matches.bysentence %>% filter(topic_id=='Topic 3',rank!=1) %>% 
+  mutate(rowid = row_number())
+
+compare_s = filter(compare_s_all,rank>1) 
+
+
+calculate_jaccard_index = function(target_sentence,compare_sentences){
+words_target = target_sentence %>% str_split(.,pattern=' ') %>% unlist()#test sentence
+
+jaccard.index = matrix(0,nrow(compare_sentences),2)
+for (x in 1:nrow(compare_sentences)){
+words_compare =  compare_sentences[x,'text_data_clean_s'] %>% str_split(.,pattern=' ') %>% unlist()
+intersection_ab = length(intersect(words_target,words_compare))
+union_ab = length(words_target)+length(words_compare) - intersection_ab
+jaccard.index[x,1] = intersection_ab/union_ab
+words_abs.diff = abs(length(words_target)-length(words_compare))
+jaccard.index[x,2] = words_abs.diff
+}
+colnames(jaccard.index) = c('jaccard.index','diff.words')
+jaccard.index = data.frame(jaccard.index) %>% add_column(doi = compare_sentences[['doi']],rowid = compare_sentences[['rowid']],.before = 'jaccard.index')
+return(jaccard.index)
+}
+
+test_jaccard = lapply(target_s,function(s) calculate_jaccard_index(target_sentence = s,compare_sentences = compare_s)) 
+names(test_jaccard) = target_s
+
+test_result = bind_rows(test_jaccard,.id='target_sentence') %>% filter(jaccard.index>=0.9,diff.words<=3)
+test_result
 
 #example sentence
 test_str = 'student t-test was used for statistical analysis'

### 2021-05-25 manuscript/asa_template/asa_template_almostdone.tex
update almostdone pdf and bib files

--- 2021-05-24 manuscript/asa_template/asa_template_almostdone.tex
@@
 {
-  \title{\bf An observational study of papers published in \emph{PLOS ONE} and
-studies posted to a trial registry.}
+  \title{\bf ``A p-value of $<$ 0.05 was considered statistically significant'': a content analysis of published statistical methods sections}
 
   \author{
-        Nicole White \thanks{The authors gratefully acknowledge computational resources and services
+        Nicole M White \thanks{The authors gratefully acknowledge computational resources and services
 used in this work provided by the eResearch Office, Queensland
 University of Technology, Brisbane, Australia.} \\
-    School of Public Health and Social Work, QUT\\
+    \small{Australian Centre for Health Services Innovation and Centre for Healthcare Transformation} \\
+\small{School of Public Health and Social Work, Queensland University of Technology} \\
      and \\     Thiru Balasubramaniam, Richi Nayak \\
-    To add\\
-     and \\     Adrian Barnett \\
-    School of Public Health and Social Work, QUT\\
+    \small{Centre for Data Science and School of Computer Science} \\
+\small{Queensland University of Technology} \\
+     and \\     Adrian G Barnett \\
+    \small{Australian Centre for Health Services Innovation and Centre for Healthcare Transformation} \\
+\small{School of Public Health and Social Work, Queensland University of Technology} \\
       }
@@
   \begin{center}
-    {\LARGE\bf An observational study of papers published in \emph{PLOS ONE} and
-studies posted to a trial registry.}
+    {\LARGE\bf ``A p-value of $<$ 0.05 was considered statistically significant'': a textual analysis of published statistical methods sections}
   \end{center}
@@
 
-A potential contributor to poor statistical reporting is the temptation for researchers to replicate
+A potential contributor to poor statistical reporting is the temptation for researchers to re-use
 descriptions of statistical methods, in an effort to make their papers resemble those of their peers 
@@
 other papers, regardless of whether they are appropriate. For this paper, we applied a text-based 
-clustering method to analyse the content of statistical methods sections included in common scientific research outputs.
+clustering method to analyse the content of published statistical methods sections.
 Clustering results are then evaluated to estimate the extent that researchers are
@@
 
-\emph{PLOS ONE} is a large open access journal that publishes original
+\emph{PLOS ONE} is a open access mega-journal that publishes original
 research across a wide range of scientific fields. Article submissions
 are handled by an academic editor who selects peer reviewers based on
-their self-nominated areas of expertise. Submissions do not undergo
+their self-nominated areas of expertise. Currently there are 324 academic editors out of 9,648 (3\%) 
+with the keywords of "statistics (mathematics)" or "statistical methods" in their expertise list (web search on 25-May-2021,
+ https://journals.plos.org/plosone/static/editorial-board). Submissions do not undergo
 formal statistical review. Instead, reviewers are required to assess
@@
 
-Data included in our analysis were downloaded on 3 July 2020. For
+Data were downloaded on 3 July 2020. For
 records that did not pass Step 2, we reviewed where initial search terms
@@
 Statistical methods sections were missing for some studies downloaded
-from ANZCTR, including a small number of sections labelled as ``Not
-applicable'', ``Nil'' or ``None''. We therefore examined if there were
-particular studies where this section was more likely to be missing.
+from ANZCTR, including sections labelled as ``Not
+applicable'', ``Nil'' or ``None''. We examined if there were
+particular studies where the statistical methods section was more likely to be missing.
 Analysis considered a logistic regression model estimated in the
 Bayesian framework {[}\citet{INLA}; www.r-inla.org{]}, with missing
-statistical methods section (yes/no) as the dependent variable. Date,
+statistical methods section (yes/no) as the dependent variable. The independent variables were date,
 study type (observational or interventional), number of funders and
-target sample size were used as independent variables. Results were
+target sample size. Results were
 reported as odds ratios with 95\% credible intervals (CI).
@@
 sourced from index searches of three statistics textbooks
-\citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. Plurals (e.g.,
+\citep{Diggle2013,Bland2015,Dobson2018}. Plurals (e.g.,
 `chi-squares') unhyphenated (e.g., `chi square') and combined
@@
 
-For \(N\) studies, let \(P \in R^{M \times N}\) denote a content matrix,
-representing text from statistical methods sections as \(M\) unique
+For \(N\) studies, let \(P \in R^{M \times N}\) denote a content matrix of text from statistical methods sections as \(M\) unique
 terms. Text clustering algorithms for identifying common topics across
 studies requires \(P\) to be represented with a vector space model. In
-our case, unique terms in \(P\) are modelled using the tf*idf (term
+our case, unique terms in \(P\) are modelled using the tf-idf (term
 frequency \(\times\) inverse document frequency) weighting schema, to
@@
 performance of distance-based clustering methods such as
-\textit{k}-means \citep{jain} in accurately identifying subgroups
+\textit{k}-means \citep{jain2010} in accurately identifying subgroups
 (topics) present in the data. Furthermore, sparseness associated with
@@
 as the topic coherence score. For our two datasets, we applied NMF with
-\(g=10\).
+\(g=10\) topics.
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\clearpage
 
-\hypertarget{plos-one}{%
-\subsection{PLOS ONE}\label{plos-one}}
+\section{Results}\label{results}
 
-\begin{figure}
+\subsection{PLOS ONE}\label{plos-one}
+
+API searches returned 131,847 papers (DOIs)
+(\autoref{fig:consort-diagrams}). After partial matching, 111,731 (85\%)
+statistical methods sections were identified. In the final sample,
+95,518 (85\%) DOIs returned an exact match against common section
+headings: 64,133 for `statistical analysis', 13,380 for `statistical
+analyses' and 13,627 for `data analysis'.
+
+\begin{figure}[htbp]
 
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 
@@
 
-API searches returned 131,847 results (DOIs)
-(\autoref{fig:consort-diagrams}). After partial matching, 111,731 (85\%)
-statistical methods sections were identified. In the final sample,
-95,518 (85\%) DOIs returned an exact match against common section
-headings: 64,133 for `statistical analysis', 13,380 for `statistical
-analyses' and 13,627 for `data analysis'.
-
 Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
@@
 percentage of records that included a statistical methods section varied between 64\%
-(volume 2) and 86\% (volume 9). All DOIs included Biology and life
-sciences (n = 107,584), Earth sciences (n = 7,605) and/or Computer and
-information sciences (n = 5,190) in their top 3 subject classifications
+(volume 2) and 86\% (volume 9). All DOIs included ``Biology and life
+sciences'' (n = 107,584), ``Earth sciences'' (n = 7,605) and/or ``Computer and
+information sciences'' (n = 5,190) in their top 3 subject classifications
 (Figure \ref{fig:plos-n}B).
@@
 
-For studies excluded based on section headings, initial search terms
-appeared in Introduction (number), Results (number) only. xxx appeared
-in the Methods section without a section heading and xxx had a
-non-standard section heading. {[}TODO{]}.
+For studies excluded based on section headings (n = 20,116), analysis-related search terms for 2,136 DOIs appeared in the general text of the Materials and Methods section.
+For studies without a matching Materials and Methods section,
+ search terms oftern appeared in the Introduction (n = 728) and Results and/or Discussion (n = 981) sections.
+4,010 DOIs included at least 1 section heading that contained the terms ``analysis'' or ``analyses'' without direct reference to statistical method(s) (e.g.
+Microarray analysis).
 
@@
 
-\caption{\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE}\label{fig:unnamed-chunk-6}
+\caption{\label{fig:plos-clouds}Word clouds for ten topics for statistical methods sections published in PLOS ONE}\label{fig:unnamed-chunk-6}
 \end{figure}
 
-The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}.
+The word clouds for ten topics are in Figure \ref{fig:plos-clouds}.
 Frequently occurring words reflected the use of statistical software
@@
 regression (topic 2) and meta-analysis (topic 7) were also identified.
-Examples of boilerplate text for selected topics are presented in Table
+Examples of boilerplate text for selected topics are in Table
 \ref{tab:plos-example-boilerplate}.
@@
 Topics related to statistical software differentiated between Prism
-GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%).
+GraphPad (topic 3: n = 9974) and SPSS (topic 5: n = 9648).
 Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
-Diego, CA) was used for all analyses'' was matched in 1,941 studies based on our definition of boilerplate text.
+Diego, CA) was used for all analyses'' was matched in 1,941 papers based on our definition of boilerplate text.
 Similarly, for topic 5, ``SPSS software version
-17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 studies.
+17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 papers.
 
@@
 
-\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE datase (sentence level).
-Statistical methods text corresponds to the DOI with the highest topic coherence score within a topic}
+\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE papers (sentence level).
+The number of matching to each sentence was based on a Jaccard score of 0.9 or higher and a difference in word count of $\pm$ 3 words}
 \centering
@@
 \hline
-Topic & Statistical methods text & Matching DOIs\\
+Topic & Statistical methods text & Matching papers\\
 \hline
-Topic 1 & $^{\dagger}$Students t-test was used for statistical analysis. & 29\\
+Topic 1 & Students t-test was used for statistical analysis. & 29\\
 \cline{2-3}
- & $^{\dagger}$A p value of $<$ 0.05 was considered statistically significant. & 1883\\
+ & A p value of $<$ 0.05 was considered statistically significant. & 1883\\
 \hline
-Topic 3 & $^{\dagger}$GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
+Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
 \hline
-Topic 4 & $^{\dagger}$Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
+Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
 \hline
-Topic 5 & $^{\dagger}$SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
+Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
 \hline
-Topic 6 & $^{\dagger}$All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
+Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
 \hline
-Topic 9 & $^{\dagger}$Statistical significance was determined by Students t-tests. & 49\\
+Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
 \cline{2-3}
-& $^{\dagger}$ $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
+&  $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
 \hline
@@
 Definitions of statistical significance featured strongly in topic 1 (n
-= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%). Topic 1 reflected
+= 3784) and topic 9 (n = 6195). Topic 1 reflected
 applications of two-tailed and unpaired Student's t-tests, however,
@@
 text descriptions varying based on method(s) used. One-way analysis of
-variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
+variance also featured strongly in topic 4 (n = 10212), combined
 with common methods for performing post-hoc multiple comparisons. Based
@@
 (ANOVA) followed by Tukey post-hoc tests for multiple comparisons''.
-Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
+Frequently occurring words in topic 6 (n = 4764) reflected
 mentions of descriptive statistics for summarising continuous variables,
@@
 Odds ratios and 95\% credible intervals for study characteristics
-included in the logistic regression model are presented in Table
+associated with missing statistical methods sections are in Table
 \ref{tab:anzctr-missing-odds}. Observational studies were less likely to
@@
 
-\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-9}
+\caption{Word clouds for ten topics for statistical methods sections published in ANZCTR}\label{fig:unnamed-chunk-9}
 \end{figure}
 
-The clustering algorithm found groups that were purely sample size
+The text clustering algorithm found topics that were purely sample size
 calculations (topic 2, n = 1834), pilot studies (topic 5, n = 834),
@@
 hypothesis testing methods, statistical significance and software. 
-
 In other cases, text had been slightly modified to account for changes
 in primary and secondary outcomes. Examples of these text changes were
-found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances
+found in topics 2 and 4; identified instances
 related to sample size calculations for patient recruitment to different
@@
 
-Since studies included in this dataset outlined planned analyses, we
+Since studies in this dataset described planned analyses, we
 hypothesised that some studies did not specify statistical methods because
-they had yet to consult with a statistician. Targeted searches for the
-term ``statistician'' across all topics returned xx matching records,
+they had yet to consult with a statistician. Targeted searches for ``statistician'' across all topics returned 1,277 matching studies,
 with examples including ``A statistician employed by hospital was used''

### 2021-05-25 code/7_summarise_stats_sections_anzctr.R

@@ -86,7 +86,7 @@ boilerplate.text.anzctr = boilerplate.numbers %>% left_join(matches %>% select(n
 
 save(matches,file='../results/anzctr.results.10topics.rda')
 save(stats_section.sim,ftab.cosine.anzctr,boilerplate.text.anzctr,file='../results/anzctr.cosinesim.10topics.rda')
-
+matches %>% filter(grepl('stati(.*)ti(.*)ian',text_data_clean))
 #save boilerplate text as separate excel file
 write.xlsx(boilerplate.text.anzctr,file='../results/anzctr.boilerplate.xlsx')
 

### 2021-05-25 code/7_summarise_stats_sections_anzctr.R
add top studies for anzctr for jaccard index

@@ -88,5 +88,10 @@ save(matches,file='../results/anzctr.results.10topics.rda')
 save(stats_section.sim,ftab.cosine.anzctr,boilerplate.text.anzctr,file='../results/anzctr.cosinesim.10topics.rda')
 matches %>% filter(grepl('stati(.*)ti(.*)ian',text_data_clean))
 #save boilerplate text as separate excel file
-write.xlsx(boilerplate.text.anzctr,file='../results/anzctr.boilerplate.xlsx')
+#save top ranking numbers as separate exclude
+top_matches = matches %>% group_by(topic_id) %>% filter(rank==1)
 
+#which value is the most frequently occuring?
+frequent_matches = matches %>% group_by(topic_id) %>% count(value) %>% filter(n==max(n))
+
+write.xlsx(top_matches,file='results/anzctr.boilerplate.xlsx')

### 2021-05-25 code/7_summarise_stats_sections_anzctr.R
move old files

@@ -23,7 +23,7 @@ matches = left_join(matches,wordcounts,by=c('number','topic_id')) %>%
   group_by(topic_id) %>% mutate(rank=row_number()) %>% ungroup() %>%
   mutate(topic_id = paste('Topic',topic_id))
 
-
+save(matches,file='../results/anzctr.results.10topics.rda')
 cos.sim <- function(ix,distances.mat) 
 {
   A = distances.mat[,ix[1]]
@@ -84,7 +84,6 @@ boilerplate.text.anzctr = boilerplate.numbers %>% left_join(matches %>% select(n
   mutate(topic=as.numeric(topic)) %>%
   arrange(topic,-sim)
 
-save(matches,file='../results/anzctr.results.10topics.rda')
 save(stats_section.sim,ftab.cosine.anzctr,boilerplate.text.anzctr,file='../results/anzctr.cosinesim.10topics.rda')
 matches %>% filter(grepl('stati(.*)ti(.*)ian',text_data_clean))
 #save boilerplate text as separate excel file
@@ -93,5 +92,4 @@ top_matches = matches %>% group_by(topic_id) %>% filter(rank==1)
 
 #which value is the most frequently occuring?
 frequent_matches = matches %>% group_by(topic_id) %>% count(value) %>% filter(n==max(n))
-
 write.xlsx(top_matches,file='results/anzctr.boilerplate.xlsx')

### 2021-05-26 code/5_process_stats_section_actrn.R
upload correct  anzctr.results.10topics.rda

@@ -143,8 +143,6 @@ stats_section = stats_section %>% mutate(text_data_clean = replace_curly_quote(t
 stats_section$text_data_clean = strip(stats_section$text_data_clean,char.keep = c("~~",".","-"),apostrophe.remove=T,digit.remove=F)
 
 
-
-
 #4. make common statistical terms and methods consistent
 #for each hyphenated term, create combined and unique plural terms
 #plurals includes entries from stat_terms_model eg regressions to regression

### 2021-05-26 code/7_summarise_stats_sections_anzctr.R

@@ -4,6 +4,12 @@ library(tidytext)
 library(openxlsx)
 
 dat = readRDS('../data/stats_section_anzctr_cleaned.rds')
+#UPDATE: replace text_data_clean with version sent to Thiru and Richi. Did not save updated .rds at the time //TODO save as combined .rds file
+ad = read.table('../stats_section_cleaned_anzctr_check.txt',sep='\t',header=T)
+
+dat = left_join(dat %>% select(-text_data_clean),ad,by='number')
+
+
 #ten topics, PLOS ONE
 matches = read.csv('../results/anzctr_10topics.csv',header=T)
 

### 2021-06-28 manuscript/asa_template/asa_template_almostdone.tex
revise manuscript

@@
 
-An ideal statistical analysis will use appropriate methods to create
+An ideal statistical analysis will use appropriate methods to draw
 insights from the data and inform the research questions. Unfortunately
-many current statistical analyses are far from ideal, with many
-researchers using the wrong methods, misinterpreting the results, or
+many current statistical analyses are far from ideal, with 
+researchers often using the wrong methods, misinterpreting the results, or
 failing to adequately check their assumptions \citep{2008, Leek2017}.
@@
 the few methods they know regardless of their appropriateness, and then
-going through the motions of the analysis \citep{Stark2018}.
+going through the motions of the analysis \citep{Stark2018}. Acceptance of methodological illiteracy in favour of 
+research quantitiy and perceived novelty is at odds
+with principles of scientific inquiry, yet continues to pervade published scientific research \citep{VanCalster2021}. 
+This paradox has been exemplified during the COVID-19 pandemic, which has led to unprecendented levels of published research of largely 
+poor quality due to biases in conduct and reporting \citep{Glasziou2020,Wynants2020}.
 
@@
 statistical practices are being to used abet weak science
-\citep{Stark2018}. Statistical mistakes are a key source of waste in
-research and partly explain the current reproducibility crisis in
+\citep{Stark2018}. Statistical mistakes are a key source of research waste and are contributing to 
+the current reproducibility crisis in
 science \citep{Allison2016}. Even when the correct methods are used,
@@
 \citep{Lang2013, Altman2016} with the latter covering all apsects of the
-paper. Both of these guidelines were led by Doug Altman, who spoke often
+paper tailored to different study designs. Both of these guidelines were led by Doug Altman, who spoke often
 and for many years about the need for better statistical reporting. The
@@
 descriptions of statistical methods, in an effort to make their papers resemble those of their peers 
-and increase perceived chances of publication  \citep{Diong2018}. Two statisticians on this paper 
+and increase perceived chances of publication  \citep{Diong2018}. As descriptions become more common, rebuke by 
+reviewers and journal editors can be difficult, as their past use may be argued by researchers as offering precedent for the 
+conduct of analysis within their discipline \citep{Altman2002}. Two statisticians on this paper 
 (AB and NW) have heard researchers admit
 that they have copied-and-pasted their statistical methods sections from
-other papers, regardless of whether they are appropriate. For this paper, we applied a text-based 
+other papers, regardless of whether they are appropriate. To investigate the extent of this practice, we applied a text-based 
 clustering method to analyse the content of published statistical methods sections.
@@
 We used two openly available data sources to find statistical methods
-sections: research articles published in \emph{PLOS ONE} and study
+sections: study
 protocols registered on the Australian and New Zealand Clinical Trials
-Registry (ANZCTR). Data sources were chosen as examples of common
+Registry (ANZCTR) and research articles published in \emph{PLOS ONE}. Data sources were chosen as examples of common
 research outputs that include descriptions of statistical methods that
-were used, or are planned, for analysing study outcomes.
-
-\subsubsection{Public Library of Science (PLOS ONE)}
-\label{sec:methodsPLOS}
-
-\emph{PLOS ONE} is a open access mega-journal that publishes original
-research across a wide range of scientific fields. Article submissions
-are handled by an academic editor who selects peer reviewers based on
-their self-nominated areas of expertise. Currently there are 324 academic editors out of 9,648 (3\%) 
-with the keywords of "statistics (mathematics)" or "statistical methods" in their expertise list (web search on 25-May-2021,
- https://journals.plos.org/plosone/static/editorial-board). Submissions do not undergo
-formal statistical review. Instead, reviewers are required to assess
-submissions against several publication criteria, including whether:
-``Experiments, statistics, and other analyses are performed to a high
-technical standard and are described in sufficient detail''
-\citep{PLOS}. All reviewers are asked the question: ``Has the
-statistical analysis been performed appropriately and rigorously?'',
-with the possible responses of ``Yes'', ``No'' and ``I don't know''.
-
-Authors are encouraged to follow published reporting guidelines such as
-EQUATOR, to ensure that chosen statistical methods are appropriate for
-the study design, and adequate details are provided to enable
-independent replication of results.
-
-All \emph{PLOS ONE} articles are freely accessible via the PLOS
-Application Programming Interface (API). This enabled us to conduct
-searches of full-text articles and analyse data on articles' text
-content and general attributes such as publication date and field(s) of
-research. Statistical methods sections were extracted using a two-step
-approach:
-
-\emph{Step 1}: Targeted API searches were completed using the R package
-`rplos' \citep{rplos}. Search queries targeted analysis-related terms,
-combining the words ``data'' or ``statistical'' with one of:
-``analysis'', ``analyses'', ``method'', ``methodology'' or
-``model(l)ing''. Terms could appear anywhere within the main body of the
-article, to account for the placement of relevant text in different
-sections, for example, in the \emph{Material and Methods} section versus
-\emph{Results}. Search results were indexed by a unique Digital Object
-Identifier (DOI). Attribute data collected per DOI included journal
-volume and subject classification(s).
-
-\emph{Step 2}: \emph{PLOS ONE} does not prescribe standardised headings
-to preface statistical methods sections. To address this, we performed
-partial matching on available headings against frequently used terms in
-initial search results: `Statistical analysis', `Statistical analyses',
-`Statistical method', `Statistics', `Data analysis' and `Data analyses'.
-
-Data were downloaded on 3 July 2020. For
-records that did not pass Step 2, we reviewed where initial search terms
-appeared in the full-text, to determine the proportion of
-statistical methods sections that were missed under our chosen approach.
+were planned or used for analysing study outcomes.
 
@@
 For the statistical methods section, researchers are asked to provide a
-``brief description'' of the sample size calculations, statistical
+brief description of all sample size calculations, statistical
 methods and planned analyses, although this section is not compulsory
@@
 All studies available on ANZCTR were downloaded on 1 February 2020 in
-XML format. We used all the text available in the ``Statistical
+XML format. For our analysis, we used all text available in the ``Statistical
 methods'' section. We also collated basic information about the study
@@
 chosen as we believed they might influence the completeness of the
-statistical methods section, because we expected larger studies and
-those with funding to be more complete, and we also were interested in
+statistical methods section. For example, we hypothesised that larger studies and
+those with funding to be more complete. We were also interested in
 changes over time.
@@
 
+\subsubsection{Public Library of Science (PLOS ONE)}
+\label{sec:methodsPLOS}
+
+\emph{PLOS ONE} is a open access mega-journal that publishes original
+research across a wide range of scientific fields. Article submissions
+are handled by an academic editor who selects peer reviewers based on
+their self-nominated area(s) of expertise. Currently there are 324 academic editors out of 9,648 (3\%) 
+with the keywords of "statistics (mathematics)" or "statistical methods" in their expertise list (web search on 25-May-2021,
+ https://journals.plos.org/plosone/static/editorial-board). Submissions do not undergo
+formal statistical review. Instead, reviewers are required to assess
+submissions against several publication criteria, including whether:
+``Experiments, statistics, and other analyses are performed to a high
+technical standard and are described in sufficient detail''
+\citep{PLOS}. All reviewers are asked the question: ``Has the
+statistical analysis been performed appropriately and rigorously?'',
+with the possible responses of ``Yes'', ``No'' and ``I don't know''.
+
+Authors are encouraged to follow published reporting guidelines such as
+EQUATOR, to ensure that chosen statistical methods are appropriate for
+the study design, and adequate details are provided to enable
+independent replication of results.
+
+All \emph{PLOS ONE} articles are freely accessible via the PLOS
+Application Programming Interface (API). This enabled us to conduct
+searches of full-text articles and analyse data on articles' text
+content and general attributes such as publication date and field(s) of
+research. Since \emph{PLOS ONE} does not prescribe a set article format, we applied a two-step
+approach to identify statistical methods sections:
+
+\emph{Step 1}: Targeted API searches were completed using the R package
+`rplos' \citep{rplos}. Search queries targeted analysis-related terms,
+combining the words ``data'' or ``statistical'' with one of:
+``analysis'', ``analyses'', ``method'', ``methodology'' or
+``model(l)ing''. Terms could appear anywhere within the main body of the
+article, to account for the placement of relevant text in different
+sections, for example, in the \emph{Material and Methods} section versus
+\emph{Results}. Search results were indexed by a unique Digital Object
+Identifier (DOI). Attribute data collected per DOI included journal
+volume and subject classification(s).
+
+\emph{Step 2}: \emph{PLOS ONE} does not use standardised headings
+to preface statistical methods sections. To address this, we performed
+partial matching on available headings against frequently used terms in
+initial search results: `Statistical analysis', `Statistical analyses',
+`Statistical method', `Statistics', `Data analysis' and `Data analyses'.
+
+Data were downloaded on 3 July 2020. For
+records that did not pass Step 2, we reviewed where initial search terms
+appeared in the full-text, to estimate the proportion of
+statistical methods sections that were missed under our chosen approach.
+
 \subsection{Full-text processing}
@@
 in-text references (e.g.~``{[}42{]}'') centred equations, and other
-non-ASCII characters. Text contained inside brackets was retained to
-maximise content for analysis, with brackets removed. Common stop words
+non-ASCII characters. Bracketed text was retained with brackets removed to
+maximise content for analysis. Common stop words
 including pronouns, contractions and selected prepositions were removed.
@@
 \citep{Diggle2013,Bland2015,Dobson2018}. Plurals (e.g.,
-`chi-squares') unhyphenated (e.g., `chi square') and combined
-(e.g.~`chisquare') terms were transformed to singular, hyphenated form
+`chi-squares') unhyphenated terms (e.g., `chi square') and combined terms
+(e.g.~`chisquare') were transformed to singular, hyphenated form
 (e.g., `chi-square'). Common statistical tests were also hyphenated
@@
 when dealing with high-dimensional data
-\citep[\citet{luong2019clustering}]{kim2014algorithms}. In this section,
-we outline key details of NMF and its strengths compared with other
-text-based clustering methods.
+\citep[\citet{luong2019clustering}]{kim2014algorithms}.
+%In this section,
+%we outline key details of NMF and its strengths compared with other
+%text-based clustering methods.
 
-For \(N\) studies, let \(P \in R^{M \times N}\) denote a content matrix of text from statistical methods sections as \(M\) unique
+For \(N\) studies, let \(P \in R^{M \times N}\) denote a content matrix of text from statistical methods sections, comprising of \(M\) unique
 terms. Text clustering algorithms for identifying common topics across
@@
 and \(H \in R ^{N \times g}\) \citep{aggarwal2012mining}, such that
-\(P \approx WH^{T}\). The number of subgroups of common topic inferred
+\(P \approx WH^{T}\). The number of subgroups of common topics inferred
 from the data is given by \(g\).
@@
 
-Clustering results were visualised by word clouds to summarise
-frequently occurring terms associated with topic membership. Whilst
-these results were useful for inferring common themes across statistical
-methods sections, they did not indicate evidence of boilerplate text
-beyond the use of common terms. Follow up analysis therefore considered
-similarities in text between sections assigned to the same topic.
+Results were visualised by word clouds to summarise
+frequently occurring terms associated with topic membership. Follow up analysis considered 
+similarities in text between statistical methods sections assign to the same topic.
 
-We used the Jaccard similarity index to compare the statistical methods
+The Jaccard similarity index was used to compare the statistical methods
 section with the highest topic coherence score with all other sections
@@
 
+\subsection{ANZCTR}
+
+We downloaded 28,008 studies. The numbers of excluded studies are shown
+in Figure \ref{fig:consort-diagrams}. Of the 12,700 included studies,
+9,523 (75\%) had a statistical methods section. The median length of
+sections was 129 words with an inter-quartile range of 71 to 219 words.
+
+\begin{figure}[htbp]
+
+{\centering \includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}  \includegraphics[width=0.49\linewidth]{figures/excluded_plosone}
+
+}
+
+\caption{\label{fig:consort-diagrams}Search results for ANZCTR (left) and PLOS ONE (right).}
+\end{figure}
+
+Odds ratios and 95\% credible intervals for study characteristics
+associated with missing statistical methods sections are in Table
+\ref{tab:anzctr-missing-odds}. Observational studies were less likely to
+have a missing statistical methods section compared with interventional
+studies. Missing sections became less likely over time. Studies with
+more funders and a larger target sample size were less likely to have a
+missing statistical methods section.
+
+\begin{table}[]
+\centering
+\caption{Logistic regression results for study characteristics associated with missing statistical methods sections in ANZCTR}
+\label{tab:anzctr-missing-odds}
+\begin{tabular}{lcc}
+\hline
+Variable & Odds ratio & 95$\%$ CI \\
+\hline
+Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
+Date (per year) & 0.90 & (0.88, 0.91) \\
+Number of funders & 0.80 & (0.74, 0.86) \\
+Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
+
+}
+
+\caption{Word clouds for ten topics for statistical methods sections published in ANZCTR}\label{fig:anzctr-wordclouds}
+\end{figure}
+
+Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. 
+Frequent words related to the largest topic (topic 2, n = 1834) reflected descriptions of sample size calculations
+ and represented 19$\%$ of studies analysed. Other topics indicated pilot studies (topic 5, n = 834),
+safety/tolerability studies (topic 6, n = 524), descriptive statistics (topic 7, n = 333),
+intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852). A review of word counts by topic 
+identified statistical methods sections that were only one word, including ``ANOVA'',
+``t-test'', ``SPSS'' and even ``SSPS''.
+
+There were cases where the exact
+same method section had been re-used in a different study. For example,
+in topic 7 (n = 333), 232 sections stated `descriptive statistics' or
+`descriptive statistics used' with no additional details provided. Other
+instances with expanded descriptions of methods included topic 3 (n =
+1277) outlining descriptive analyses only, and topic 6 (n = 524), for
+hypothesis testing methods, statistical significance and software. 
+In other cases, text had been slightly modified to account for changes
+in primary and secondary outcomes. Examples of these text changes were
+found in topics 2 and 4; identified instances
+related to sample size calculations for patient recruitment to different
+studies.
+
+Examples of boilerplate text are provided in Table
+\ref{tab:anzctr-example-boilerplate}.
+
+Since studies in this dataset described planned analyses, we
+hypothesised that some studies did not specify statistical methods because
+they had yet to consult with a statistician. Targeted searches for ``statistician'' across all topics returned 1,277 matching studies,
+with examples including ``A statistician employed by hospital was used''
+and ``Pilot study at this point will use a statistician professionally
+to determine sample size calculations as required''.
+
+\begin{landscape}
+\begin{table}[]
+\centering
+\caption{Example boilerplate text from ANZCTR dataset}
+\label{tab:anzctr-example-boilerplate}
+\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\hline
+\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
+3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
+Two sided p values of less than 0.05 will be considered statistically significant.\\
+ &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
+The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate.
+\\ \hline
+5 & Pilot study \\
+ &  No formal sample size calculation was performed \\ \hline
+7 & Descriptive statistics \\
+ &  Descriptive statistics used \\ \hline
+9 & Linear mixed models will be used to analyse the data. \\ \hline
+10 & Repeated measures of ANOVA \\
+ &  Pre-, during, post- and follow-up variables will be subjected to mixed methods and repeated measures analyses to determine significant changes over (group and) time. \\ \hline
+\end{tabular}
+\end{table}
+\end{landscape}
+
 \subsection{PLOS ONE}\label{plos-one}
@@
 analyses' and 13,627 for `data analysis'.
-
-\begin{figure}[htbp]
-
-{\centering \includegraphics[width=0.49\linewidth]{figures/excluded_plosone} \includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}
-
-}
-
-\caption{\label{fig:consort-diagrams}Search results for PLOS ONE (left) and ANZCTR (right).}
-\end{figure}
 
@@
 
-\subsection{ANZCTR}
-
-We downloaded 28,008 studies. The numbers of excluded studies are shown
-in Figure \ref{fig:consort-diagrams}. Of the 12,700 included studies,
-9,523 (75\%) had a statistical methods section. The median length of
-sections was 129 words with an inter-quartile range of 71 to 219 words.
-
-Odds ratios and 95\% credible intervals for study characteristics
-associated with missing statistical methods sections are in Table
-\ref{tab:anzctr-missing-odds}. Observational studies were less likely to
-have a missing statistical methods section compared with interventional
-studies. Missing sections became less likely over time. Studies with
-more funders and a larger target sample size were less likely to have a
-missing statistical methods section.
-
-\begin{table}[]
-\centering
-\caption{Logistic regression results for study characteristics associated with missing statistical methods sections in ANZCTR}
-\label{tab:anzctr-missing-odds}
-\begin{tabular}{lcc}
-\hline
-Variable & Odds ratio & 95$\%$ CI \\
-\hline
-Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
-Date (per year) & 0.90 & (0.88, 0.91) \\
-Number of funders & 0.80 & (0.74, 0.86) \\
-Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
-
-}
-
-\caption{Word clouds for ten topics for statistical methods sections published in ANZCTR}\label{fig:unnamed-chunk-9}
-\end{figure}
-
-The text clustering algorithm found topics that were purely sample size
-calculations (topic 2, n = 1834), pilot studies (topic 5, n = 834),
-safety/tolerability studies (topic 6, n = 524), intervention studies
-(topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852).
-Examples of boilerplate text are provided in Table
-\ref{tab:anzctr-example-boilerplate} .
-
-Some methods sections were only one word, including ``ANOVA'',
-``t-test'', ``SPSS'' and even ``SSPS''. There were cases where the exact
-same method section had been re-used in a different study. For example,
-in topic 7 (n = 333), 232 sections stated `descriptive statistics' or
-`descriptive statistics used' with no additional details provided. Other
-instances with expanded descriptions of methods included topic 3 (n =
-1277) outlining descriptive analyses only, and topic 6 (n = 524), for
-hypothesis testing methods, statistical significance and software. 
-In other cases, text had been slightly modified to account for changes
-in primary and secondary outcomes. Examples of these text changes were
-found in topics 2 and 4; identified instances
-related to sample size calculations for patient recruitment to different
-studies.
-
-Since studies in this dataset described planned analyses, we
-hypothesised that some studies did not specify statistical methods because
-they had yet to consult with a statistician. Targeted searches for ``statistician'' across all topics returned 1,277 matching studies,
-with examples including ``A statistician employed by hospital was used''
-and ``Pilot study at this point will use a statistician professionally
-to determine sample size calculations as required''.
-
-\begin{landscape}
-\begin{table}[]
-\centering
-\caption{Example boilerplate text from ANZCTR dataset}
-\label{tab:anzctr-example-boilerplate}
-\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
-\hline
-\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
-3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
-Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
-Two sided p values of less than 0.05 will be considered statistically significant.\\
- &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
-The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate.
-\\ \hline
-5 & Pilot study \\
- &  No formal sample size calculation was performed \\ \hline
-7 & Descriptive statistics \\
- &  Descriptive statistics used \\ \hline
-9 & Linear mixed models will be used to analyse the data. \\ \hline
-10 & Repeated measures of ANOVA \\
- &  Pre-, during, post- and follow-up variables will be subjected to mixed methods and repeated measures analyses to determine significant changes over (group and) time. \\ \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
 \hypertarget{discussion}{%

### 2021-06-29 manuscript/asa_template/asa_template_almostdone.tex
update manuscript files

@@
 content and general attributes such as publication date and field(s) of
-research. Since \emph{PLOS ONE} does not prescribe a set article format, we applied a two-step
+research. We applied a two-step
 approach to identify statistical methods sections:
@@
 initial search results: `Statistical analysis', `Statistical analyses',
-`Statistical method', `Statistics', `Data analysis' and `Data analyses'.
-
-Data were downloaded on 3 July 2020. For
-records that did not pass Step 2, we reviewed where initial search terms
-appeared in the full-text, to estimate the proportion of
-statistical methods sections that were missed under our chosen approach.
+`Statistical method', `Statistics', `Data analysis' and `Data analyses'. Data were downloaded on 3 July 2020.
 
@@
 Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. 
-Frequent words related to the largest topic (topic 2, n = 1834) reflected descriptions of sample size calculations
- and represented 19$\%$ of studies analysed. Other topics indicated pilot studies (topic 5, n = 834),
+Words related to the largest topic (topic 2, n = 1,834) reflected descriptions of sample size calculations, for example, ``sample'',``power'' and ``80'' (i.e.
+$80\%$ power) .
+Other topics indicated pilot studies (topic 5, n = 834),
 safety/tolerability studies (topic 6, n = 524), descriptive statistics (topic 7, n = 333),
@@
 total number of API search results peaked at volumes 8 (n = 19,045) and
-9 (n = 19,045) in years 2013 and 2014. This trend aligned with the total
-number of papers published in \emph{PLOS ONE} over the same period. The
+9 (n = 19,045) in years 2013 and 2014.  The
 percentage of records that included a statistical methods section varied between 64\%
@@
 
-For studies excluded based on section headings (n = 20,116), analysis-related search terms for 2,136 DOIs appeared in the general text of the Materials and Methods section.
-For studies without a matching Materials and Methods section,
- search terms oftern appeared in the Introduction (n = 728) and Results and/or Discussion (n = 981) sections.
+%For studies excluded based on section headings (n = 20,116), analysis-related search terms for 2,136 DOIs appeared in the general text of the Materials and Methods section. In other cases,
+ %search terms oftern appeared in the Introduction (n = 728) and Results and/or Discussion (n = 981) sections.
 4,010 DOIs included at least 1 section heading that contained the terms ``analysis'' or ``analyses'' without direct reference to statistical method(s) (e.g.
@@
 
-\caption{\label{fig:plos-clouds}Word clouds for ten topics for statistical methods sections published in PLOS ONE}\label{fig:unnamed-chunk-6}
+\caption{\label{fig:plos-clouds}Word clouds for ten topics for statistical methods sections published in PLOS ONE}\label{fig:plos-clouds}
 \end{figure}

### 2021-06-29 code/7_jaccard_sentence_anzctr.R
update exported matches for anzctr

@@ -10,19 +10,23 @@ library(openxlsx)
 
 ## data
 ## a) get the example data
-load('data/processed_examples_anzctr.RData') # from 7_example_sentences.R
+#load('data/processed_examples_anzctr.RData') # from 7_example_sentences.R
 ## b) get all the data
-load(file='data/anzctr.results.10topics.rda')
+load(file='results/anzctr.results.10topics.rda')
 # data management 
 matches = ungroup(matches) %>%
   mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic '))) # topic number
 
 # split examples into sentences
+# take top ranking result for each topic
+examples <- matches %>% filter(rank==1) %>% select(topic_num,number,words,text_data_clean) %>%
+  rename(topic=topic_num,example = text_data_clean)
+
 examples_in_sentences = NULL
 for (t in 1:nrow(examples)){
   # get the example and split into sentences
   to_compare = examples[t, ]
-  example_sentences = str_split(to_compare$example, pattern='\\. ')[[1]]
+  example_sentences = str_split(to_compare$example, pattern='\\.\\s+')[[1]]
   f = data.frame(topic=to_compare$topic, sentences = example_sentences)
   examples_in_sentences = bind_rows(examples_in_sentences, f)
 }  
@@ -69,6 +73,8 @@ almost_exact = filter(results,
   tally() %>%
   ungroup()
 
+
+
 # merge with original data
 to_export = full_join(examples_in_sentences, almost_exact, by='row') %>%
   select(row, topic, sentences, n) %>%
@@ -82,4 +88,4 @@ writeDataTable(wb, sheet = 1, x = to_export,
                colNames = TRUE, rowNames = FALSE,
                tableStyle = "TableStyleLight9")
 setColWidths(wb, sheet = 1, cols = 1:4, widths = c(5,5,70,5))
-saveWorkbook(wb, file = "results/jaccard_matches_anzctr.xlsx", overwrite = TRUE)
+saveWorkbook(wb, file = "results/jaccard_matches_anzctr_v2.xlsx", overwrite = TRUE)

### 2021-06-29 code/7_jaccard_sentence_anzctr.R
update R script

@@ -74,7 +74,6 @@ almost_exact = filter(results,
   ungroup()
 
 
-
 # merge with original data
 to_export = full_join(examples_in_sentences, almost_exact, by='row') %>%
   select(row, topic, sentences, n) %>%

### 2021-07-12 manuscript/asa_template/asa_template_almostdone.tex
update tex file

@@
 Other topics indicated pilot studies (topic 5, n = 834),
-safety/tolerability studies (topic 6, n = 524), descriptive statistics (topic 7, n = 333),
+safety/tolerability studies (topic 6, n = 524), descriptive analysis (topic 7, n = 333),
 intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852). A review of word counts by topic 

### 2021-07-12 manuscript/asa_template/asa_template_almostdone.tex
update jaccard scripts to compare word tokens

@@
 the few methods they know regardless of their appropriateness, and then
-going through the motions of the analysis \citep{Stark2018}. Acceptance of methodological illiteracy in favour of 
-research quantitiy and perceived novelty is at odds
+going through the motions of the analysis \citep{Stark2018}. Accepting methodological illiteracy in favour of 
+research quantity and novelty is at odds
 with principles of scientific inquiry, yet continues to pervade published scientific research \citep{VanCalster2021}. 
@@
 \citep{Wikipedia}. Use of these methods sections indicates that little
-thought has gone into the statistical analysis.
+thought has gone into the conduct and transparent reporting of statistical analysis.
 
@@
 
-There were cases where the exact
-same method section had been re-used in a different study. For example,
-in topic 7 (n = 333), 232 sections stated `descriptive statistics' or
-`descriptive statistics used' with no additional details provided. Other
-instances with expanded descriptions of methods included topic 3 (n =
-1277) outlining descriptive analyses only, and topic 6 (n = 524), for
-hypothesis testing methods, statistical significance and software. 
+Common statistical methods for data analysis were t-tests (topic 3, n = 1,277), descriptive statistics (topic 7, n = 333), linear models (topic 9, n = 1,195) and repeated measure ANOVA (topic 10, n = 852).
+Instances of boilerplate text were concentrated in Topic 3, which emphasised the use of parametric versus non-parametric hypothesis tests (Table \ref{tab:anzctr-example-boilerplate}).
+Identified matches across topics 7, 9 and 10 were
+relatively low, however, sensitiivty in the Jaccard index to shorter sentence lengths may have been a contributing factor.
+For example, the section with the highest coherence score in topic 7 simply stated `descriptive statistics', with 19 matches.
+Further review of
+sections assigned to this topic identified a further 7 records with `descriptive analysis' as the entire methods section (Jaccard score = 0.79).
+
 In other cases, text had been slightly modified to account for changes
@@
 studies.
-
-Examples of boilerplate text are provided in Table
-\ref{tab:anzctr-example-boilerplate}.
 
@@
 regression (topic 2) and meta-analysis (topic 7) were also identified.
-Examples of boilerplate text for selected topics are in Table
-\ref{tab:plos-example-boilerplate}.
 
@@
 Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
-Diego, CA) was used for all analyses'' was matched in 1,941 papers based on our definition of boilerplate text.
-Similarly, for topic 5, ``SPSS software version
-17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 papers.
+Diego, CA) was used for all analyses'' was matched in 1,941 papers based on our definition of boilerplate text (Table \ref{tab:plos-example-boilerplate}).
+Similarly, ``SPSS software version
+17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 papers in topic 5.
 
@@
 text descriptions varying based on method(s) used. One-way analysis of
-variance also featured strongly in topic 4 (n = 10212), combined
+variance featured strongly in topic 4 (n = 10212), combined
 with common methods for performing post-hoc multiple comparisons. Based
-on the Jaccard index, 1 in 5 studies were matched to the sentence
+on our boilerplate criteria, 1 in 5 studies were matched to the sentence
 ``Significant differences were determined using analysis of variance

### 2021-07-12 code/3_plot_excluded.plos.R

@@ -6,6 +6,11 @@ library(diagram)
 
 #load metadata
 load('results/total.records.plos.rda')
+
+#load excluded sections
+load('data/stats_section_excluded_plos.rda')
+stats_section_excluded = bind_rows(stats_section_excluded)
+
 #number downloaded through target search queries
 n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
 #number of stats sections included following partial matching
@@ -46,7 +51,7 @@ count_exclude = tibble(reason=c('Materials & Methods','Introduction','Results &/
 make_diagram_plos = function(){
   par(mai=c(0,0,0,0))
   labels = c(paste('Downloaded\n n = ', format(n.api, big.mark = ','), sep=''),
-             paste('Excluded\n- ', count_exclude$reason[1], ', n = ', format(count_exclude$n[1],big.mark=','),
+             paste('Excluded by second stage filter\n- ', count_exclude$reason[1], ', n = ', format(count_exclude$n[1],big.mark=','),
                    '\n- ', count_exclude$reason[2], ', n = ', count_exclude$n[2],
                    '\n- ', count_exclude$reason[3], ', n = ', count_exclude$n[3],
                    '\n- ', count_exclude$reason[4], ', n = ', format(count_exclude$n[4],big.mark=','), sep=''),

### 2021-07-12 code/7_jaccard_sentence_anzctr.R

@@ -6,6 +6,7 @@ library(tidyverse)
 library(tidytext)
 library(stringr)
 library(stringdist) # for similarity scores
+library(textreuse)
 library(openxlsx)
 
 ## data
@@ -16,6 +17,29 @@ load(file='results/anzctr.results.10topics.rda')
 # data management 
 matches = ungroup(matches) %>%
   mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic '))) # topic number
+
+
+#tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
+#single number followed by full stop (e.g. [space]1.[space]); reomve number
+matches = mutate(matches,
+                  text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:digit:]{1})(\\.)(\\s{1})",replacement="\\3\\4"))
+
+#a single [a-z] character, predeeced and followed by a full stop[space]
+matches = mutate(matches,
+                  text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:alpha:]{1})(\\.)(\\s{1})",replacement="\\1\\2\\4"))
+#instances where the entire sentence is a number e.g ver. 23.0
+matches = mutate(matches,
+                  text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})([:digit:]+)(\\.)",replacement="\\2\\3\\4"))
+
+#remove final full stop
+matches = mutate(matches,
+                  text_data_clean = str_remove_all(text_data_clean,pattern='\\.$'))
+
+#replace e.g. and i.e.
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='e.g.\\s+|i.e.\\s+')
+                 )
+
 
 # split examples into sentences
 # take top ranking result for each topic
@@ -41,34 +65,47 @@ for (t in 1:nrow(examples_in_sentences)){
                       topic_num == examples_in_sentences$topic[t]) 
   
   # split text_data_clean into sentence as alternative for boilerplate text (takes a while); in list form
-  dat.sentences = lapply(1:nrow(this_topic), function(i) str_split(this_topic[i,]$text_data_clean,"\\. ",simplify=F) %>% unlist() %>% 
-                           tibble(text_data_clean_s=.) %>% add_column(doi = this_topic[i,] %>% pull(number)))
[…]

### 2021-07-12 code/7_jaccard_sentence_plos_updated.R

@@ -8,69 +8,96 @@ library(tidytext)
 library(stringr)
 library(stringdist) # for similarity scores
 library(openxlsx)
+library(textreuse)
 
 ## data
 ## a) get the example data
-load('data/processed_examples.RData') # from 7_example_sentences.R
+#load('data/processed_examples.RData') # from 7_example_sentences.R
 ## b) get all the data
 load(file='results/plos.results.10topics.rda')
+
+
 # data management 
 matches = ungroup(matches) %>%
   mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic ')), # topic number
          doi_num = str_remove(doi, pattern='10.1371/journal\\.pone\\.')) # just get the number from the DOI
+
+#tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
+#single number followed by full stop (e.g. [space]1.[space]); reomve number
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:digit:]{1})(\\.)(\\s{1})",replacement="\\3\\4"))
+
+#a single [a-z] character, predeeced and followed by a full stop[space]
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:alpha:]{1})(\\.)(\\s{1})",replacement="\\1\\2\\4"))
+#instances where the entire sentence is a number e.g ver. 23.0
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})([:digit:]+)(\\.)",replacement="\\2\\3\\4"))
+
+#remove final full stop
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='\\.$'))
+
+#replace e.g. and i.e.
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='e.g.\\s+|i.e.\\s+'))
+
+
+
 # combine paragraphs from the same paper (with the same DOI)
-combined = group_by(matches, doi_num, topic_id, topic_num) %>%
-  summarise(text = toString(text_data_clean )) %>%
+combined = group_by(matches, doi_num, topic_id, topic_num,rank) %>%
+  summarise(text = toString(text_data_clean)) %>%
   ungroup()
 
 # split examples into sentences
+# take top ranking result for each topic
[…]

### 2021-07-12 code/7_jaccard_sentence_anzctr.R
update anzctr jaccard

@@ -1,7 +1,7 @@
 # 7_jaccard_sentence_anzctr.R
 # look for almost identical sentences using the Jaccard score
 # uses the top example sentences from the paper
-# May 2021
+# Last updated July 12 2021
 library(tidyverse)
 library(tidytext)
 library(stringr)
@@ -105,7 +105,7 @@ for (t in 1:nrow(examples_in_sentences)){
 
 # count almost exact matches
 almost_exact = filter(results, 
-                      score >= 0.5) %>% # high similarity score. changed from 0.9
+                      score >= 0.7) %>% # high similarity score. changed from 0.9
   group_by(row) %>%
   tally() %>%
   ungroup()

### 2021-07-13 manuscript/asa_template/asa_template_almostdone.tex
jaccard bug resolved

@@
 \label{tab:anzctr-example-boilerplate}
-\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
+\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
 \hline
-\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
+Topic & Statistical methods text & Matching studies\\
 3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
 Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
-Two sided p values of less than 0.05 will be considered statistically significant.\\
+Two sided p values of less than 0.05 will be considered statistically significant.
+& \\
  &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
 The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate.
-\\ \hline
-5 & Pilot study \\
- &  No formal sample size calculation was performed \\ \hline
-7 & Descriptive statistics \\
- &  Descriptive statistics used \\ \hline
-9 & Linear mixed models will be used to analyse the data. \\ \hline
-10 & Repeated measures of ANOVA \\
- &  Pre-, during, post- and follow-up variables will be subjected to mixed methods and repeated measures analyses to determine significant changes over (group and) time. \\ \hline
+& \\ \hline
+5 & Pilot study & \\
+ &  No formal sample size calculation was performed & \\ \hline
+7 & Descriptive statistics & \\ \hline
+9 & Linear mixed models will be used to analyse the data. & \\ \hline
+10 & Repeated measures of ANOVA & \\ \hline
 \end{tabular}

### 2021-07-13 code/7_jaccard_sentence_plos_updated.R

@@ -4,7 +4,6 @@
 # May 2021
 library(tidyverse)
 library(tidytext)
-#library(openxlsx)
 library(stringr)
 library(stringdist) # for similarity scores
 library(openxlsx)
@@ -51,8 +50,8 @@ combined = group_by(matches, doi_num, topic_id, topic_num,rank) %>%
 
 # split examples into sentences
 # take top ranking result for each topic
-examples <- matches %>% filter(rank==1) %>% select(topic_num,doi_num,words,text_data_clean) %>%
-  rename(topic=topic_num,example = text_data_clean)
+examples <- combined %>% filter(rank==1) %>% select(topic_num,doi_num,text) %>%
+  rename(topic=topic_num,example = text)
 
 
 examples_in_sentences = NULL
@@ -60,46 +59,51 @@ for (t in 1:nrow(examples)){
   # get the example and split into sentences
   to_compare = examples[t, ]
   example_sentences = str_split(to_compare$example, pattern='\\.\\s+')[[1]]
-  f = data.frame(topic=to_compare$topic, sentences = example_sentences)
+  f = data.frame(doi_num=to_compare$doi_num,topic=to_compare$topic, sentences = example_sentences)
   examples_in_sentences = bind_rows(examples_in_sentences, f)
 }  
 examples_in_sentences = mutate(examples_in_sentences, row = 1:n())
 
 
+# split text_data_clean into sentence as alternative for boilerplate text (takes a while); in list form
+dat.sentences = combined %>% mutate(text_data_clean_s = str_split(text,"\\.\\s+",simplify=F)) %>%
+  mutate(n_words = lapply(text_data_clean_s,function(x) str_count(x,"\\w+")))
+
 
 # loop through sentences that were used in the paper
 results = NULL
+results_byrow = NULL
+
 for (t in 1:nrow(examples_in_sentences)){
   #number of words in target sentence
   n_words_example = str_count(examples_in_sentences$sentence[t], "\\w+")
-  ## just text in the same topic 
-  this_topic = filter(combined, 
-                      topic_num == examples_in_sentences$topic[t]) 
   
-  # split text_data_clean into sentence as alternative for boilerplate text (takes a while); in list form
-  dat.sentences = lapply(1:nrow(this_topic), function(i) str_split(this_topic[i,]$text,"\\.\\s+",simplify=F) %>% unlist() %>% 
[…]

### 2021-07-13 code/8_sandpit_topdown_jaccard_updated.R

@@ -1,0 +1,125 @@
+# look for almost identical sentences using the Jaccard score
+# uses the top example sentences from the paper
+# May 2021
+library(tidyverse)
+library(tidytext)
+library(stringdist) # for similarity scores
+library(openxlsx)
+library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
+
+## data
+## a) get the example data
+#load('data/processed_examples.RData') # from 7_example_sentences.R
+## b) get all the data
+load(file='results/plos.results.10topics.rda')
+
+# data management 
+matches = ungroup(matches) %>%
+  mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic ')), # topic number
+         doi_num = str_remove(doi, pattern='10.1371/journal\\.pone\\.')) # just get the number from the DOI
+
+#tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
+#single number followed by full stop (e.g. [space]1.[space]); reomve number
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:digit:]{1})(\\.)(\\s{1})",replacement="\\3\\4"))
+
+#a single [a-z] character, predeeced and followed by a full stop[space]
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:alpha:]{1})(\\.)(\\s{1})",replacement="\\1\\2\\4"))
+#instances where the entire sentence is a number e.g ver. 23.0
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})([:digit:]+)(\\.)",replacement="\\2\\3\\4"))
+
+#remove final full stop
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='\\.$'))
+
+#replace e.g. and i.e.
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern='e.g.\\s+|i\\.e\\.\\s+|\\s+ver\\.\\s+',replacement = ' '))
+#statcorp
+matches =mutate(matches,
+                text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})(college station tx)",replacement="\\2\\3"))
+#inc. or ver.
+  matches =mutate(matches,
[…]

### 2021-07-13 code/8_sandpit_topdown_jaccard_updated.R
resolve jaccard bug

@@ -8,7 +8,7 @@ library(openxlsx)
 library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
 
 ## data
-## a) get the example data
+## a) get the example data  
 #load('data/processed_examples.RData') # from 7_example_sentences.R
 ## b) get all the data
 load(file='results/plos.results.10topics.rda')

### 2021-07-14 code/8_sandpit_topdown_jaccard_updated.R
add updated jaccard results plos

@@ -41,7 +41,7 @@ matches = mutate(matches,
 matches =mutate(matches,
                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})(college station tx)",replacement="\\2\\3"))
 #inc. or ver.
-  matches =mutate(matches,
+matches =mutate(matches,
                   text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
 
 
@@ -59,6 +59,7 @@ dat.sentences = combined %>% mutate(text_data_clean_s = str_split(text,"\\.\\s+"
 #loop over topics - start with 1 topic
 combined_export = list()
 cutoff = 100 #top 100 papers per topic
+
 for (choose.topic in 1:10){
   {cat('Reviewing Topic',choose.topic,'\r')}
   

### 2021-07-15 code/8_sandpit_topdown_jaccard_updated.R
add updated plos boilerplate

@@ -124,3 +124,77 @@ for (choose.topic in 1:10){
   combined_export[[choose.topic]] = to_export
   
 } #end of topic loop
+
+
+save(combined_export,file='results/jaccard_plos_top100.rda')
+
+#summarise the results
+load('results/jaccard_plos_top100.rda')
+
+results = bind_rows(combined_export,.id='topic_num') %>%
+  mutate(topic_num = as.numeric(topic_num))
+
+#rank ==1 matches
+top_ranked_bytopic = results %>% filter(rank==1) %>% arrange(topic_num)
+
+#most common sentence within each topic (rank 1 to 100)
+most_frequent_bytopic = results %>% group_by(topic_num) %>% arrange(-matches) %>% slice(1)
+
+
+#top ranked excluding p-value statements
+results_nop = results %>% filter(!grepl('p-value|p value|\\bp\\b',sentence))
+most_frequent_bytopic_nop = results_nop %>% group_by(topic_num) %>% arrange(-matches) %>% slice(1)
+
+#top 10 sentences per topic
+top10_frequent_bytopic = results %>% group_by(topic_num) %>% arrange(-matches) %>% slice(1:10)
+
+#top 10 excluding p value statments
+top10_frequent_bytopic_nop = results_nop %>% group_by(topic_num) %>% arrange(-matches) %>% slice(1:10)
+
+
+
+#
+# export to Excel
+wb <- createWorkbook()
+#Matches for top ranked paper in each topic
+addWorksheet(wb, sheetName = "Matches rank = 1 by topic", gridLines = FALSE)
+freezePane(wb, sheet = 1, firstRow = TRUE)
+writeDataTable(wb, sheet = 1, x = top_ranked_bytopic,
+               colNames = TRUE, rowNames = FALSE,
+               tableStyle = "TableStyleLight9")
+setColWidths(wb, sheet = 1, cols = 1:5, widths = c(5,5,70,5,5))
+
+#most common sentence within each topic (rank 1 to 100)
+addWorksheet(wb, sheetName = "Matches Most frequent", gridLines = FALSE)
+freezePane(wb, sheet = 2, firstRow = TRUE)
+writeDataTable(wb, sheet = 2, x = most_frequent_bytopic,
+               colNames = TRUE, rowNames = FALSE,
+               tableStyle = "TableStyleLight9")
[…]

### 2021-07-15 manuscript/asa_template/asa_template_almostdone.tex
add updated jaccard anzctr

@@
 number of words that appeared in either sentence (union). Instances of
-boilerplate text were defined by a Jaccard index of 0.9 or higher and a
+boilerplate text were defined by a Jaccard index of 0.75 or higher and a
 difference in word count of plus or minus three words.
@@
 
-\begin{landscape}
+%\begin{landscape}
+%\begin{table}
+%
+%\caption{\label{tab:plos-example-boilerplate} Examples of boilerplate text from PLOS ONE papers with the highest coherence score per topic (sentence level).
+The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words}
+%\centering
+%\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
+%\hline
+%Topic & Statistical methods text & Matching papers\\
+%\hline
+%Topic 1 & Students t-test was used for statistical analysis. & 17\\
+%\cline{2-3}
+% & A p value of $<$ 0.05 was considered statistically significant. & 664\\
+%\hline
+%Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
+%\hline
+%Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
+%\hline
+%Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
+%\hline
+%Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
+%\hline
+%Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
+%\cline{2-3}
+%&  $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
+%\hline
+%\end{tabular}
+%\end{table}
+%\end{landscape}
+
+\pagebreak
+
 \begin{table}
 
-\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE papers (sentence level).
-The number of matching to each sentence was based on a Jaccard score of 0.9 or higher and a difference in word count of $\pm$ 3 words}
+\caption{\label{tab:plos-example-boilerplate} Example boilerplate text from PLOS ONE papers with the highest number of matches per topic (sentence level).
+The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words; $^{*}$ Most frequent match after excluding statements of statistical significance}
 \centering
-\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
+\begin{tabular}[t]{p{0.05\textwidth}p{0.8\textwidth}p{0.1\textwidth}}
 \hline
@@
 \hline
-Topic 1 & Students t-test was used for statistical analysis. & 29\\
-\cline{2-3}
- & A p value of $<$ 0.05 was considered statistically significant. & 1883\\
+1 & A p-value of less-than 0.05 was considered statistically significant & 664\\
+ &  Statistical analysis was performed using student t-test$^{*}$ & 53\\
 \hline
-Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
+2 & A p-value less-than 0.05 was considered to be statistically significant & 1,138\\
+ & All statistical analyses were performed using sas version 9.4 sas institute cary nc usa$^{*}$ & 140\\
 \hline
-Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
+3 & A p-value less-than 0.05 was considered significant & 914\\
+ & Data are expressed as mean $\pm$ SEM$^{*}$ & 350\\
 \hline
-Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
+4 & P less-than 0.05 was considered statistically significant & 580\\
+ &  Data are presented as the mean $\pm$ SEM$^{*}$ & 359\\
 \hline
-Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
+5 & A p-value less-than 0.05 was considered a significant difference & 739\\
+ & All data are presented as mean $\pm$ standard deviation$^{*}$ & 149\\
 \hline
-Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
-\cline{2-3}
-&  $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
+6 & A p-value of less-than 0.05 was considered significant & 519\\
+&  Data are presented as mean $\pm$ SEM$^{*}$ & 223\\
+\hline
+7 & A p-value less-than 0.05 was considered statistically significant & 305\\
+ &  Otherwise the fixed-effects model was used$^{*}$ & 45\\
+\hline
+8 & Data were analysed using R version 2.15.2 & 10\\
+\hline
+9 & Significance was set at p less-than 0.05 & 195\\
+& All data are presented as mean $\pm$ SEM$^{*}$ & 115\\
+\hline
+10 & A p-value less-than 0.05 was considered significant & 397\\
+& All the experiments were repeated at least three times$^{*}$ & 119\\
+
 \hline

### 2021-07-15 code/8_top100_jaccard_anzctr.R

@@ -1,6 +1,5 @@
-# look for almost identical sentences using the Jaccard score
-# uses the top example sentences from the paper
-# May 2021
+#8_top100_jaccard_anzctr.R
+
 library(tidyverse)
 library(tidytext)
 library(stringdist) # for similarity scores
@@ -8,15 +7,14 @@ library(openxlsx)
 library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
 
 ## data
-## a) get the example data  
-#load('data/processed_examples.RData') # from 7_example_sentences.R
+## a) get the example data
+#load('data/processed_examples_anzctr.RData') # from 7_example_sentences.R
 ## b) get all the data
-load(file='results/plos.results.10topics.rda')
-
+load(file='results/anzctr.results.10topics.rda')
 # data management 
 matches = ungroup(matches) %>%
-  mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic ')), # topic number
-         doi_num = str_remove(doi, pattern='10.1371/journal\\.pone\\.')) # just get the number from the DOI
+  mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic '))) # topic number
+
 
 #tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
 #single number followed by full stop (e.g. [space]1.[space]); reomve number
@@ -42,23 +40,17 @@ matches =mutate(matches,
                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})(college station tx)",replacement="\\2\\3"))
 #inc. or ver.
 matches =mutate(matches,
-                  text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
-
-
-# combine paragraphs from the same paper (with the same DOI)
-combined = group_by(matches, doi_num, topic_id, topic_num,rank) %>%
-  summarise(text = toString(text_data_clean)) %>%
-  ungroup()
+                text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
 
 # end data management
 
 # split text_data_clean into sentence as alternative for boilerplate text (takes a while); in list form
[…]

### 2021-07-15 code/8_top100_jaccard_plos.R

@@ -1,0 +1,205 @@
+# look for almost identical sentences using the Jaccard score
+# uses the top example sentences from the paper
+# May 2021
+library(tidyverse)
+library(tidytext)
+library(stringdist) # for similarity scores
+library(openxlsx)
+library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
+
+## data
+## a) get the example data  
+#load('data/processed_examples.RData') # from 7_example_sentences.R
+## b) get all the data
+load(file='results/plos.results.10topics.rda')
+
+# data management 
+matches = ungroup(matches) %>%
+  mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic ')), # topic number
+         doi_num = str_remove(doi, pattern='10.1371/journal\\.pone\\.')) # just get the number from the DOI
+
+#tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
+#single number followed by full stop (e.g. [space]1.[space]); reomve number
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:digit:]{1})(\\.)(\\s{1})",replacement="\\3\\4"))
+
+#a single [a-z] character, predeeced and followed by a full stop[space]
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:alpha:]{1})(\\.)(\\s{1})",replacement="\\1\\2\\4"))
+#instances where the entire sentence is a number e.g ver. 23.0
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})([:digit:]+)(\\.)",replacement="\\2\\3\\4"))
+
+#remove final full stop
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='\\.$'))
+
+#replace e.g. and i.e.
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern='e.g.\\s+|i\\.e\\.\\s+|\\s+ver\\.\\s+',replacement = ' '))
+#statcorp
+matches =mutate(matches,
[…]

### 2021-07-18 code/8_jaccard_plos_keywords.R
add jaccard keywords

@@ -1,0 +1,84 @@
+# look for almost identical sentences using the Jaccard score
+#  targeted searches of keywords
+# May 2021
+library(tidyverse)
+library(tidytext)
+library(stringdist) # for similarity scores
+library(openxlsx)
+library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
+
+## data
+## a) get the example data  
+#load('data/processed_examples.RData') # from 7_example_sentences.R
+## b) get all the data
+load(file='results/plos.results.10topics.rda')
+
+# data management 
+matches = ungroup(matches) %>%
+  mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic ')), # topic number
+         doi_num = str_remove(doi, pattern='10.1371/journal\\.pone\\.')) # just get the number from the DOI
+
+#tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
+#single number followed by full stop (e.g. [space]1.[space]); reomve number
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:digit:]{1})(\\.)(\\s{1})",replacement="\\3\\4"))
+
+#a single [a-z] character, predeeced and followed by a full stop[space]
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:alpha:]{1})(\\.)(\\s{1})",replacement="\\1\\2\\4"))
+#instances where the entire sentence is a number e.g ver. 23.0
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})([:digit:]+)(\\.)",replacement="\\2\\3\\4"))
+
+#remove final full stop
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='\\.$'))
+
+#replace e.g. and i.e.
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern='e.g.\\s+|i\\.e\\.\\s+|\\s+ver\\.\\s+',replacement = ' '))
+#statcorp
+matches =mutate(matches,
+                text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})(college station tx)",replacement="\\2\\3"))
+
+# sas ibm corp
+matches =mutate(matches,
[…]

### 2021-07-19 manuscript/asa_template/asa_template_almostdone.tex
manuscript and code edits

@@
 
-The Jaccard similarity index was used to compare the statistical methods
-section with the highest topic coherence score with all other sections
-assigned to the same topic. Indices were calculated at the sentence
-level, to account for instances of boilerplate text within larger
-methods sections. We chose the Jaccard index as an easy to interpret
-measure, which summarises the similarity between two sentences by the
-number of words common to each sentence (intersection), divided by the
-number of words that appeared in either sentence (union). Instances of
-boilerplate text were defined by a Jaccard index of 0.75 or higher and a
-difference in word count of plus or minus three words.
+Evidence of boilerplate text was assessed at the section and sentences levels using the Jaccard similarity index. 
+We chose the Jaccard index as an easy to interpret measure, which summarises the similarity between two pieces of text by the
+number of words common to each text (intersection), divided by the
+number of words that appeared in either text (union). For both section-level and sentence-level analyses, text was tokenised at the word level. 
+Instances of boilerplate text were defined by a Jaccard index of 0.75 or higher and a difference in word count of plus or minus three words.
+
+At the document level, we applied minhash and locality-sensitive hashing using the R package `textreuse' \citep{textreuseR}. The combination of these approaches 
+allowed us to reduce the total number of pairwise comparisons within a topic by ignoring unlikely matches. For all topics, locality-sensitive hashing was applied using the same random seed, 
+with 1,000 randomly selected minhashes broken into 100 equal-sized bands.
+
+A similar approach was applied at the sentence level.
+In this case, we aimed to identify common sentences within statistical methods sections that were most representative of each topic.
+Analysis therefore focussed
+on identifying the reuse of sentences from the top 100 sections assigned to each topic, based on estimated coherence scores.
+For each example sentence, text was compared with all sections assigned to the same topic.
 
@@
 
-The word clouds for ten topics are in Figure \ref{fig:plos-clouds}.
-Frequently occurring words reflected the use of statistical software
+Frequently occurring words by topic reflected the use of statistical software
 (topics 3 and 5), descriptive statistics (topic 6), group based
 hypothesis testing (topics 1 and 4) and definitions of statistical
-significance (topics 1 and 9). Topics related to
+significance (topics 1 and 9) (\ref{fig:plos-clouds}). Topics related to
 regression (topic 2) and meta-analysis (topic 7) were also identified.
 
-Topics related to statistical software differentiated between Prism
+Statistical software topics differentiated between Prism
 GraphPad (topic 3: n = 9974) and SPSS (topic 5: n = 9648).
-Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
-Diego, CA) was used for all analyses'' was matched in 1,941 papers based on our definition of boilerplate text (Table \ref{tab:plos-example-boilerplate}).
-Similarly, ``SPSS software version
-17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 papers in topic 5.
+Within topic 3, targeted searches for ``GraphPad Prism'' identified 7,141 sentences from 6,844 papers. Of these,
+1,045 papers contained matching sentences based on our definition of boilerplate text, with instances varying based on version and citation details (Table \ref{tab:plos-example-boilerplate-sentence}).
+Similar examples of boilerplate text were identified in topic 5, with ``SPSS'' returning 1,104 matching papers.
 
-%\begin{landscape}
-%\begin{table}
-%
-%\caption{\label{tab:plos-example-boilerplate} Examples of boilerplate text from PLOS ONE papers with the highest coherence score per topic (sentence level).
-The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words}
-%\centering
-%\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
-%\hline
-%Topic & Statistical methods text & Matching papers\\
-%\hline
-%Topic 1 & Students t-test was used for statistical analysis. & 17\\
-%\cline{2-3}
-% & A p value of $<$ 0.05 was considered statistically significant. & 664\\
-%\hline
-%Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
-%\hline
-%Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
-%\hline
-%Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
-%\hline
-%Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
-%\hline
-%Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
-%\cline{2-3}
-%&  $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
-%\hline
-%\end{tabular}
-%\end{table}
-%\end{landscape}
+\begin{landscape}
+\begin{table}
+
+\caption{\label{tab:plos-example-boilerplate-sentence} Examples of boilerplate text from PLOS ONE papers based on targeted keyword searches (sentence level).
+The number of matching papers was based on a Jaccard score of 0.75 or higher}
+\centering
+\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
+\hline
+Topic & Search term &  Matching papers &  Example statistical methods text\\
+\hline
+\multicolumn{4}{l}{Hypothesis testing}\\
+\hline
+1 & ``t-test'' & 531 & Statistical analysis was performed using Student's t-test.\\
+4 & ``ANOVA'' & 270 & Statistical differences were analysed by one-way ANOVA followed by multiple comparisons performed with post-hoc Bonferroni test (SPSS version 16).\\
+9 & ``p $<$'' & 1,962 & $^{*}$p$<$0.05, $^{**}$p$<$0.01, $^{***}$p$<$0.001.\\
+\hline
+\multicolumn{4}{l}{Statistical software}\\
+\hline
+3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
+5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
+\hline
+\multicolumn{4}{l}{Descriptive statistics}\\
+\hline
+6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
+\hline
+\multicolumn{4}{l}{Experiments}\\
+\hline
+10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
+\hline
+\end{tabular}
+\end{table}
+\end{landscape}
 
 \pagebreak
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.8\linewidth]{figures/plosjaccardsectionlevel} 
+%
+%\caption{\label{fig:plos-clouds}Total matches by topic for statistical methods sections published in PLOS ONE (section level)}\label{fig:plos-jaccard-section}
+%\end{figure}
 
@@
 namely means with standard deviations or standard errors.
+
+Statements of statistical significant was the most frequenty occuring text across nine out of ten topics (\ref{tab:plos-example-boilerplate}).
+In all cases, statistical significance was defined at p$<$0.05.
 

### 2021-07-19 code/8_jaccard_plos_keywords.R

@@ -45,40 +45,201 @@ matches =mutate(matches,
 matches =mutate(matches,
                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(ibm corp)(\\.)",replacement="\\1\\2"))
 
-
-
 #inc. or ver.
 matches =mutate(matches,
                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
 
-
 # combine paragraphs from the same paper (with the same DOI)
 combined = group_by(matches, doi_num, topic_id, topic_num,rank) %>%
-  summarise(text = toString(text_data_clean),groups='drop')
+  summarise(text = toString(text_data_clean),.groups='drop')
 
 # end data management
+
+# Document-level analysis, all topics
+n.minhash = 1000
+n.bands = 100
+
+random.seed = TeachingDemos::char2seed('jaccard',set=F)
+
+minhash <- minhash_generator(n = n.minhash, seed = random.seed)
+
+#review chosen number of minhashes and bands
+lsh_threshold(h=n.minhash,b=n.bands) #expected jaccard index
+lsh_probability(h=n.minhash,b=n.bands,s=0.75) #probability of a match based on chosen score cutoff
+
+calc_document_jaccard <- function(indata=combined,choose.topic,minhash,n.bands){
+  dat.topic = indata %>% filter(topic_num==choose.topic) %>% distinct(doi_num,.keep_all=T) %>%
+    mutate(id=row_number())
+
+  text.corpus = TextReuseCorpus(text=dat.topic$text,meta=list(id=dat.topic$id),
+                                tokenizer = tokenize_words,minhash_func = minhash)
+  buckets <- lsh(text.corpus, bands = n.bands)
+  candidates <- lsh_candidates(buckets)
+  
+  jacsim = lsh_compare(candidates, 
+                       text.corpus, 
+                       jaccard_similarity) %>% 
+    arrange(desc(score))  
+  
+  
+  return(list(similarities=jacsim,dat=dat.topic))
+}
+
+#loop over topics
+jaccard_doc = lapply(1:10, function(x) calc_document_jaccard(indata=combined,choose.topic=x,minhash = minhash,n.bands = n.bands))
+
+#for a given topic, find total exact matches, total with score >=0.75
+#then find most frequently matched section as example of boilerplate text
+
[…]

### 2021-07-19 code/8_top100_jaccard_plos.R

@@ -202,4 +202,4 @@ writeDataTable(wb, sheet = 5, x = top10_frequent_bytopic_nop,
 setColWidths(wb, sheet = 5, cols = 1:5, widths = c(5,5,70,5,5))
 
 
-saveWorkbook(wb, file = "results/jaccard_matches_anzctr_v3.xlsx", overwrite = TRUE)
+saveWorkbook(wb, file = "results/jaccard_matches_plos_v3.xlsx", overwrite = TRUE)

### 2021-07-20 manuscript/asa_template/asa_template_almostdone.tex
final run of anzctr boilerplate

@@
 
-A similar approach was applied at the sentence level.
-In this case, we aimed to identify common sentences within statistical methods sections that were most representative of each topic.
-Analysis therefore focussed
-on identifying the reuse of sentences from the top 100 sections assigned to each topic, based on estimated coherence scores.
-For each example sentence, text was compared with all sections assigned to the same topic.
+A similar approach was applied at the sentence level. In this case, we aimed to identify common sentences or phrases within statistical methods sections that were representative of each topic. 
+First, keyword searches using dominant words from each topic cloud were completed to identify common variations on sentences involving a keyword.
+Second, we analysed the reuse of sentences from the top 100 sections assigned to each topic,
+based on estimated coherence scores. For each example sentence, text was compared with all sections assigned to the same topic.
 
@@
 \begin{landscape}
-\begin{table}[]
+\begin{table}
+
+\caption{\label{tab:anzctr-example-boilerplate-sentence} Examples of boilerplate text from ANZCTR studies based on targeted keyword searches (sentence level).
+The number of matching papers was based on a Jaccard score of 0.75 or higher}
 \centering
-\caption{Example boilerplate text from ANZCTR dataset}
-\label{tab:anzctr-example-boilerplate}
-\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
+\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
 \hline
-Topic & Statistical methods text & Matching studies\\
-3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
+Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
+\hline
+\multicolumn{4}{l}{Statistical methods}\\
+\hline
+3 & ``t-test'' & 131 & Continuous Normally distributed variables will be compared using Student's t-test and reported as means standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range\\
+9 & ``model'' &  339 & Linear mixed models will be used to analyse the data.\\
+10 & ``ANOVA'' & 86 & Post-hoc paired comparisons corrected for multiple comparisons will be performed if ANOVA reveal significant effects.\\
+\hline
+\multicolumn{4}{l}{Study designs}\\
+4 & Patient & 375 & At a confidence level of 95\% and a precision around the estimate of 5\% a minimum of 73 patients must be included\\
+5 & ``Pilot study'' & & \\
+6 & ``Subject' & & \\
+8 & ``'group'
+
+\multicolumn{4}{l}{Qualitative analysis}\\
+\hline
+1 & & &\\
+\hline
+\multicolumn{4}{l}{Descriptive statistics}\\
+\hline
+6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
+\hline
+
+3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
+5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
+\hline
+
+\multicolumn{4}{l}{Experiments}\\
+\hline
+10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
+\hline
+\end{tabular}
+\end{table}
+\end{landscape}
+
+\begin{table}
+\caption{\label{tab:anzctr-example-boilerplate} Example boilerplate text from ANZCTR studies with the highest number of matches per topic (sentence level).
+The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words; $^{*}$ Most frequent match after excluding statements of statistical significance}
+\centering
+\begin{tabular}[t]{p{0.05\textwidth}p{0.8\textwidth}p{0.1\textwidth}}
+\hline
+Topic & Statistical methods text & Matching papers\\
+\hline
+1 & Significance will be set at p $<$ 0.05 & 13\\
+ &  All analyses will be conducted on an intention-to-treat basis$^{*}$ & 8\\
+\hline
+2 & Analysis will be by intention-to-treat & 15\\
+\hline
+3 & A p-value $<$ 0.05 is considered statistically significant & 39\\
+ & Continuous normally distributed variables will be compared using Student's t-test and reported as means \& standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range$^{*}$ & 8\\
+\hline
+4 & Data will be analysed by two methods the Kaplan Meier method and per-protocol analysis & 27\\
+\hline
+5 & This is a pilot study only & 5\\
+\hline
+6 & Due to the exploratory nature of this first in human study no power or sample size calculations have been performed & 7\\
+\hline
+7 & Descriptive statistics & 7\\
+\hline
+8 & The level of statistical significance will be set at p $<$ 0.05 & 11\\
+& Relationships between observations at different occasions will be modelled with an unstructured covariance matrix$^{*}$ & 9\\
+\hline
+9 & All analyses will be on an intention-to-treat basis & 24\\
+\hline
+10 & However allowing for a 50 percent attrition rate we will be required to recruit n=70 participants & 10\\
+
+\hline
+\end{tabular}
+\end{table}
+
+%\begin{landscape}
+%\begin{table}[]
+%\centering
+%\caption{Example boilerplate text from ANZCTR dataset}
+%\label{tab:anzctr-example-boilerplate}
+%\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
+%\hline
+%Topic & Statistical methods text & Matching studies\\
+%3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test.
 Continuous data will be compared using the Student’s t-test or Mann-Whitney U test.
@@
 & \\
- &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
+% &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant.
 The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate.
 & \\ \hline
-5 & Pilot study & \\
- &  No formal sample size calculation was performed & \\ \hline
-7 & Descriptive statistics & \\ \hline
-9 & Linear mixed models will be used to analyse the data. & \\ \hline
-10 & Repeated measures of ANOVA & \\ \hline
-\end{tabular}
-\end{table}
-\end{landscape}
+%5 & Pilot study & \\
+% &  No formal sample size calculation was performed & \\ \hline
+%7 & Descriptive statistics & \\ \hline
+%9 & Linear mixed models will be used to analyse the data. & \\ \hline
+%10 & Repeated measures of ANOVA & \\ \hline
+%\end{tabular}
+%\end{table}
+%\end{landscape}
 
@@
 
-Statistical software topics differentiated between Prism
-GraphPad (topic 3: n = 9974) and SPSS (topic 5: n = 9648).
-Within topic 3, targeted searches for ``GraphPad Prism'' identified 7,141 sentences from 6,844 papers. Of these,
-1,045 papers contained matching sentences based on our definition of boilerplate text, with instances varying based on version and citation details (Table \ref{tab:plos-example-boilerplate-sentence}).
+Statistical software topics differentiated between
+GraphPad Prism (topic 3: n = 9,974) and SPSS (topic 5: n = 9,648).
+Within topic 3, targeted searches for ``GraphPad'' identified 7,141 sentences from 6,844 papers. Of these,
+1,306 papers met our definition of boilerplate text, with examples varying based on version and citation details (Table \ref{tab:plos-example-boilerplate-sentence}).
 Similar examples of boilerplate text were identified in topic 5, with ``SPSS'' returning 1,104 matching papers.
@@
 \hline
-Topic & Search term &  Matching papers &  Example statistical methods text\\
+Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
 \hline
@@
 4 & ``ANOVA'' & 270 & Statistical differences were analysed by one-way ANOVA followed by multiple comparisons performed with post-hoc Bonferroni test (SPSS version 16).\\
-9 & ``p $<$'' & 1,962 & $^{*}$p$<$0.05, $^{**}$p$<$0.01, $^{***}$p$<$0.001.\\
+9 & ``p $<$'' & 2,088 & $^{*}$p$<$0.05, $^{**}$p$<$0.01, $^{***}$p$<$0.001.\\
 \hline
@@
 Definitions of statistical significance featured strongly in topic 1 (n
-= 3784) and topic 9 (n = 6195). Topic 1 reflected
-applications of two-tailed and unpaired Student's t-tests, however,
-instances of boilerplate text emphasised the 5\% level of statistical
-significance (Table \ref{tab:plos-example-boilerplate}). In contrast, Topic 9
-focused on multiple thresholds for declaring statistical significance by
-asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
-practice that has been criticised \citep{Wasserstein2019}.
+= 3,784) and topic 9 (n = 6,195). Topic 1 reflected applications of Student's t-tests, however,
+instances of boilerplate text emphasised the 5\% level of statistical significance (Table \ref{tab:plos-example-boilerplate}). Topic 9
+focused on multiple thresholds for declaring statistical significance by asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
+practice that has been criticised \citep{Wasserstein2019}. Variations of this phrase were identfied in 1,676 papers.
 
@@
 text descriptions varying based on method(s) used. One-way analysis of
-variance featured strongly in topic 4 (n = 10212), combined
-with common methods for performing post-hoc multiple comparisons. Based
-on our boilerplate criteria, 1 in 5 studies were matched to the sentence
-``Significant differences were determined using analysis of variance
-(ANOVA) followed by Tukey post-hoc tests for multiple comparisons''.
+variance featured strongly in topic 4 (n = 10,212), combined
+with common methods for performing post-hoc multiple comparisons (Table \ref{tab:plos-example-boilerplate-sentence}). 
 Frequently occurring words in topic 6 (n = 4764) reflected
 mentions of descriptive statistics for summarising continuous variables,
-namely means with standard deviations or standard errors.
+namely means with standard deviations or standard errors. The sentence starting with ``data was (were) expressed as (the) mean $\pm$'' returned exact matches in 523 papers.
 
-Statements of statistical significant was the most frequenty occuring text across nine out of ten topics (\ref{tab:plos-example-boilerplate}).
+Statements of statistical significant were the most common form of boilerplate text across nine out of ten topics (\ref{tab:plos-example-boilerplate}).
 In all cases, statistical significance was defined at p$<$0.05.

### 2021-07-20 code/8_jaccard_anzctr_keywords.R

@@ -1,0 +1,236 @@
+# look for almost identical sentences using the Jaccard score
+#  targeted searches of keywords
+# May 2021
+library(tidyverse)
+library(tidytext)
+library(stringdist) # for similarity scores
+library(openxlsx)
+library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
+
+## data
+## a) get the example data  
+#load('data/processed_examples.RData') # from 7_example_sentences.R
+## b) get all the data
+load(file='results/anzctr.results.10topics.rda')
+
+# data management 
+matches = ungroup(matches) %>%
+  mutate(topic_num = as.numeric(str_remove(topic_id, pattern='Topic '))) # topic number
+
+
+#tidy up a bit more to account for overuse of full stops (affects breakdown by sentence)
+#single number followed by full stop (e.g. [space]1.[space]); reomve number
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:digit:]{1})(\\.)(\\s{1})",replacement="\\3\\4"))
+
+#a single [a-z] character, predeeced and followed by a full stop[space]
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})([:alpha:]{1})(\\.)(\\s{1})",replacement="\\1\\2\\4"))
+#instances where the entire sentence is a number e.g ver. 23.0
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})([:digit:]+)(\\.)",replacement="\\2\\3\\4"))
+
+#remove final full stop
+matches = mutate(matches,
+                 text_data_clean = str_remove_all(text_data_clean,pattern='\\.$'))
+
+#replace e.g. and i.e.
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern='e.g.\\s+|i\\.e\\.\\s+|\\s+ver\\.\\s+',replacement = ' '))
+#statcorp
+matches =mutate(matches,
+                text_data_clean = str_replace_all(text_data_clean,pattern="(\\.)(\\s{1})(college station tx)",replacement="\\2\\3"))
+#inc. or ver.
+matches =mutate(matches,
+                text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
+
[…]

### 2021-08-03 manuscript/asa_template/asa_template_almostdone.tex
updated results

@@
 the few methods they know regardless of their appropriateness, and then
-going through the motions of the analysis \citep{Stark2018}. Accepting methodological illiteracy in favour of 
-research quantity and novelty is at odds
+going through the motions of the analysis \citep{Stark2018}. Accepting this form of methodological illiteracy is at odds
 with principles of scientific inquiry, yet continues to pervade published scientific research \citep{VanCalster2021}. 
@@
 research outputs that include descriptions of statistical methods that
-were planned or used for analysing study outcomes.
+were planned or used for analysing research outcomes.
 
@@
 
-All \emph{PLOS ONE} articles are freely accessible via the PLOS
+Data on all \emph{PLOS ONE} articles can be accessed via the PLOS
 Application Programming Interface (API). This enabled us to conduct
@@
 number of words that appeared in either text (union). For both section-level and sentence-level analyses, text was tokenised at the word level. 
-Instances of boilerplate text were defined by a Jaccard index of 0.75 or higher and a difference in word count of plus or minus three words.
+Instances of boilerplate text were defined by a Jaccard index of 0.8 or higher.
 
-At the document level, we applied minhash and locality-sensitive hashing using the R package `textreuse' \citep{textreuseR}. The combination of these approaches 
-allowed us to reduce the total number of pairwise comparisons within a topic by ignoring unlikely matches. For all topics, locality-sensitive hashing was applied using the same random seed, 
-with 1,000 randomly selected minhashes broken into 100 equal-sized bands.
+To reduce the number of pairwise comparisions, locality-sensitive hashing was implementing using the R package `textreuse' \citep{textreuseR}.
+For all topics, locality-sensitive hashing was applied using the same random seed,
+with 1,000 randomly selected minhashes broken into 200 equal-sized bands.
 
 A similar approach was applied at the sentence level. In this case, we aimed to identify common sentences or phrases within statistical methods sections that were representative of each topic. 
-First, keyword searches using dominant words from each topic cloud were completed to identify common variations on sentences involving a keyword.
-Second, we analysed the reuse of sentences from the top 100 sections assigned to each topic,
-based on estimated coherence scores. For each example sentence, text was compared with all sections assigned to the same topic.
+Analyses considered keyword searches using dominant terms from word clouds and the reuse of sentences within the top 1,000 sections assigned to each topic, based on estimated coherence scores.
 
@@
 sections was 129 words with an inter-quartile range of 71 to 219 words. 
+Sections for TODO studies were only one word, including ``ANOVA'',
+``t-test'', ``SPSS'' and even ``SSPS''.
 
 \begin{figure}[htbp]
-
-{\centering \includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}  \includegraphics[width=0.49\linewidth]{figures/excluded_plosone}
-
-}
-
+\centering 
+\includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}  \includegraphics[width=0.49\linewidth]{figures/excluded_plosone}
 \caption{\label{fig:consort-diagrams}Search results for ANZCTR (left) and PLOS ONE (right).}
@@
 \begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
-
-}
-
+\centering 
+\includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
 \caption{Word clouds for ten topics for statistical methods sections published in ANZCTR}\label{fig:anzctr-wordclouds}
@@
 
-Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. 
-Words related to the largest topic (topic 2, n = 1,834) reflected descriptions of sample size calculations, for example, ``sample'',``power'' and ``80'' (i.e.
-$80\%$ power) .
-Other topics indicated pilot studies (topic 5, n = 834),
-safety/tolerability studies (topic 6, n = 524), descriptive analysis (topic 7, n = 333),
-intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852). A review of word counts by topic 
-identified statistical methods sections that were only one word, including ``ANOVA'',
-``t-test'', ``SPSS'' and even ``SSPS''.
+Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. Topics reflected sample size calculations (topics 2 and 5), 
+study designs (topics 4, 5, 6 and 8),
+quantitative methods (topics 3, 7, 9 and 10) and qualitative methods (topic 1). 
+Common statistical methods for data analysis were t-testing, descriptive statistics, linear models and repeated measures ANOVA.
 
-Common statistical methods for data analysis were t-tests (topic 3, n = 1,277), descriptive statistics (topic 7, n = 333), linear models (topic 9, n = 1,195) and repeated measure ANOVA (topic 10, n = 852).
-Instances of boilerplate text were concentrated in Topic 3, which emphasised the use of parametric versus non-parametric hypothesis tests (Table \ref{tab:anzctr-example-boilerplate}).
-Identified matches across topics 7, 9 and 10 were
-relatively low, however, sensitiivty in the Jaccard index to shorter sentence lengths may have been a contributing factor.
-For example, the section with the highest coherence score in topic 7 simply stated `descriptive statistics', with 19 matches.
-Further review of
-sections assigned to this topic identified a further 7 records with `descriptive analysis' as the entire methods section (Jaccard score = 0.79).
+\begin{table}[htbp]
+\centering
+\caption{Results of boilerplate analysis applied to the ANZCTR dataset.
+The number of studies with Jaccard scores greater than or equal to 0.8 are presented.
+The number of exact matches is provided in brackets.}
+\label{tab:anzctr-boilerplate-matches}
+\begin{tabular}{lccc}
+\hline
+ &  & \multicolumn{2}{c}{Boilerplate text} \\
+\cline{3-4}
+Topic & Total studies & Section-level & {Sentence-level}\\
+\hline
+1 (Qualitative methods)& 854 & 53 (24) & 288 (206)\\
+2 (Sample size calculations) & 1,834 & 68 (26) & 252 (172)\\
+3 (Student's t-test) & 1,277 & 105 (45) & 552 (379)\\
+4 (Efficacy and safety studies) & 909 & 85 (11) & 261 (203)\\
+5 (Pilot studies) & 834 &  53 (34) & 148 (114)\\
+6 (Safety and tolerability studies) & 524 & 51 (29) & 241 (186)\\
+7 (Descriptive analysis) & 333 & 44 (39) & 69 (60)\\
+8 (Intervention studies) & 1,020 & 33 (8) & 300 (203)\\
+9 (Linear models) & 1,195 & 86 (33) & 436 (333)\\
+10 (ANOVA) & 852 &  64 (29) & 270 (208)\\
+\hline
+\end{tabular}
+\end{table}
 
-In other cases, text had been slightly modified to account for changes
-in primary and secondary outcomes. Examples of these text changes were
-found in topics 2 and 4; identified instances
-related to sample size calculations for patient recruitment to different
-studies.
+Approximately 7$\%$ of all study sections were classified as boilerplate text (n = 642).
+One in ten sections associated with descriptive analysis were exact matches (Table \ref{tab:anzctr-boilerplate-matches}).
+Evidence of boilerplate text was stronger at the sentence level, with 29$\%$ of sections having at least one sentence flagged (n = 2,817).
+
+Boilerplate text for selected topics emphasised definitions of statistical significance, combined with planned statistical methods.
+In topic 3 [TODO] Other topics simply stated the main analysis method, including ``descriptive statistics/descriptive analysis'' (topic 7, n = TODO), ``linear (mixed) models'' (topic 9, n = TODO) and ``repeated measures ANOVA'' (topic 10, n = TODO).
+Examples of common boilerplate sentences for describing statistical methods are in Table \ref{tab:anzctr-boilerplate-examples}.
+Full results for all topics are provided as Supplementary material.
+
+In other cases, text had been slightly modified to account for changes in primary and secondary outcomes for a given study design.
+This behaviour was reflected in the high numbers of studies with boilerplate text at the section and sentence levels (Table \ref{tab:anzctr-boilerplate-matches}).
+Evidence of boilerplate text was strongest for topic 6, where matches aligned with the structure:
+
+\begin{center}
+here is some \underline{text} that i changed
+\end{center}
+
+Similar examples were identified for sample size calculations:
+
+\begin{center}
+here is some \underline{text} that i changed
+\end{center}
 
@@
 hypothesised that some studies did not specify statistical methods because
-they had yet to consult with a statistician. Targeted searches for ``statistician'' across all topics returned 1,277 matching studies,
+they had yet to consult with a statistician. Targeted searches for ``statistician'' across all topics returned TODO matching studies,
 with examples including ``A statistician employed by hospital was used''
@@
 
-\begin{landscape}
-\begin{table}
-
-\caption{\label{tab:anzctr-example-boilerplate-sentence} Examples of boilerplate text from ANZCTR studies based on targeted keyword searches (sentence level).
+%\begin{landscape}
+%\begin{table}
+%
+%\caption{\label{tab:anzctr-example-boilerplate-sentence} Examples of boilerplate text from ANZCTR studies based on targeted keyword searches (sentence level).
 The number of matching papers was based on a Jaccard score of 0.75 or higher}
-\centering
-\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
-\hline
-Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
-\hline
-\multicolumn{4}{l}{Statistical methods}\\
-\hline
-3 & ``t-test'' & 131 & Continuous Normally distributed variables will be compared using Student's t-test and reported as means standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range\\
-9 & ``model'' &  339 & Linear mixed models will be used to analyse the data.\\
-10 & ``ANOVA'' & 86 & Post-hoc paired comparisons corrected for multiple comparisons will be performed if ANOVA reveal significant effects.\\
-\hline
-\multicolumn{4}{l}{Study designs}\\
-4 & Patient & 375 & At a confidence level of 95\% and a precision around the estimate of 5\% a minimum of 73 patients must be included\\
-5 & ``Pilot study'' & & \\
-6 & ``Subject' & & \\
-8 & ``'group'
-
-\multicolumn{4}{l}{Qualitative analysis}\\
-\hline
-1 & & &\\
-\hline
-\multicolumn{4}{l}{Descriptive statistics}\\
-\hline
-6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
-\hline
-
-3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
-5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
-\hline
-
-\multicolumn{4}{l}{Experiments}\\
-\hline
-10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
-\hline
-\end{tabular}
-\end{table}
-\end{landscape}
+%\centering
+%\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
+%\hline
+%Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
+%\hline
+%\multicolumn{4}{l}{Statistical methods}\\
+%\hline
+%3 & ``t-test'' & 131 & Continuous Normally distributed variables will be compared using Student's t-test and reported as means standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range\\
+%9 & ``model'' &  339 & Linear mixed models will be used to analyse the data.\\
+%10 & ``ANOVA'' & 86 & Post-hoc paired comparisons corrected for multiple comparisons will be performed if ANOVA reveal significant effects.\\
+%\hline
+%\multicolumn{4}{l}{Study designs}\\
+%4 & ``Patient'' & 375 & At a confidence level of 95\% and a precision around the estimate of 5\% a minimum of 73 patients must be included.\\
+%5 & ``Pilot study'' & 16 & This is a pilot study.\\
+%6 & ``Subject' & & \\
+%8 & ``'group' & & \\
+%\hline
+%\multicolumn{4}{l}{Study designs}\\
+%2 & ``'80 percent power'
+%\hline
+%\multicolumn{4}{l}{Qualitative analysis}\\
+%\hline
+%1 & & &\\
+%\hline
+%\multicolumn{4}{l}{Descriptive statistics}\\
+%\hline
+%7 & ``Descriptive'' & 33 & Analyses will be of a descriptive nature.\\
+%\hline
+%
+%\multicolumn{4}{l}{Experiments}\\
+%\hline
+%10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
+%\hline
+%\end{tabular}
+%\end{table}
+%\end{landscape}
 
@@
 \hline
-7 & Descriptive statistics & 7\\
+7 & Descriptive statistics & 15\\
 \hline
@@
 Similar examples of boilerplate text were identified in topic 5, with ``SPSS'' returning 1,104 matching papers.
+
+\begin{table}[htbp]
+\centering
+\caption{Results of boilerplate analysis applied to the PLOS ONE dataset.
+The number of studies with Jaccard scores greater than or equal to 0.8 are presented.
+The number of exact matches is provided in brackets.}
+\label{tab:plosr-boilerplate-matches}
+\begin{tabular}{lccc}
+\hline
+ &  & \multicolumn{2}{c}{Boilerplate text} \\
+\cline{3-4}
+Topic & Total studies & Section-level & {Sentence-level}\\
+\hline
+1 (Student's t-test) & 3,784 & 344 () & )\\
+2 () & 19,649 & 135 () &)\\
+3 (GraphPad Prism) & 9,974 & 405 () &)\\
+4 (ANOVA) & 10,212 & 452 () &)\\
+5 (SPSS) & 9,648 & 287 () &)\\
+6 (Descriptive analysis) & 4,764 & 331 () & \\
+7 (Meta analysis) & 5,278&  65 () & \\
+8 () & 42,261 &  116 () & \\
+9 () & 6,195 & 260 () & \\
+10 () & 4,693 & 336 () & \\
+\hline
+\end{tabular}
+\end{table}
 

### 2021-08-03 code/8_jaccard_anzctr_keywords.R

@@ -44,16 +44,21 @@ matches =mutate(matches,
 matches =mutate(matches,
                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
 
+#et al. or '..'
+matches = mutate(matches,
+                text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(et al)(\\.)",replacement="\\1\\2"))
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="\\.+",replacement="."))
+
 combined = matches %>% rename(text=text_data_clean) %>%
   mutate(topic_num = str_remove_all(topic_id,'Topic '))
 # end data management
 
 # Document-level analysis, all topics
 n.minhash = 1000
-n.bands = 100
+n.bands = 200
 
 random.seed = TeachingDemos::char2seed('jaccard',set=F)
-
 minhash <- minhash_generator(n = n.minhash, seed = random.seed)
 
 #review chosen number of minhashes and bands
@@ -65,15 +70,16 @@ calc_document_jaccard <- function(indata=combined,choose.topic,minhash,n.bands){
     mutate(id=row_number())
   
   text.corpus = TextReuseCorpus(text=dat.topic$text,meta=list(id=dat.topic$id),
-                                tokenizer = tokenize_words,minhash_func = minhash)
+                                tokenizer = tokenize_words,minhash_func = minhash,skip_short = F)
   buckets <- lsh(text.corpus, bands = n.bands)
   candidates <- lsh_candidates(buckets)
   
   jacsim = lsh_compare(candidates, 
                        text.corpus, 
                        jaccard_similarity) %>% 
-    arrange(desc(score))  
+    arrange(desc(score)) 
   
+  jacsim = jacsim %>% mutate_at(c('a','b'),~as.numeric(str_remove_all(.,pattern='doc-')))
   
   return(list(similarities=jacsim,dat=dat.topic))
 }
@@ -81,52 +87,22 @@ calc_document_jaccard <- function(indata=combined,choose.topic,minhash,n.bands){
 #loop over topics
 jaccard_doc = lapply(1:10, function(x) calc_document_jaccard(indata=combined,choose.topic=x,minhash = minhash,n.bands = n.bands))
 
[…]

### 2021-08-03 code/8_jaccard_plos_keywords.R

@@ -2,15 +2,9 @@
 #  targeted searches of keywords
 # May 2021
 library(tidyverse)
-library(tidytext)
-library(stringdist) # for similarity scores
-library(openxlsx)
 library(textreuse) #updated function for jaccard_similarity (compare to stringdist)
 
 ## data
-## a) get the example data  
-#load('data/processed_examples.RData') # from 7_example_sentences.R
-## b) get all the data
 load(file='results/plos.results.10topics.rda')
 
 # data management 
@@ -49,38 +43,57 @@ matches =mutate(matches,
 matches =mutate(matches,
                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(inc|ver)(\\.)",replacement="\\1\\2"))
 
+#et al. or vs. or fig. [:lower:]{1}[:digit:]{1}
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(et al)(\\.)",replacement="\\1\\2"))
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(vs)(\\.)",replacement="\\1\\2"))
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(fig|figs)(\\.)",replacement="\\1\\2"))
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern="(\\s{1})(no)(\\.)",replacement="\\1\\2"))
+
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern='([:digit:][:lower:])(\\.)',replacement='\\1'))
+matches = mutate(matches,
+                 text_data_clean = str_replace_all(text_data_clean,pattern='([:lower:][:digit:])(\\.)',replacement='\\1'))
+
+#remove excess white space
+matches = mutate(matches,text_data_clean=textclean::replace_white(text_data_clean))
+
 # combine paragraphs from the same paper (with the same DOI)
 combined = group_by(matches, doi_num, topic_id, topic_num,rank) %>%
   summarise(text = toString(text_data_clean),.groups='drop')
 
-# end data management
+combined = combined %>% arrange(topic_num,rank)
 
[…]
